### 2021-09-06 D0/draft.tex
M10

% document layout
\documentclass{article}
\usepackage[margin=1.4in]{geometry}
\usepackage[utf8]{inputenc}

% packages
\usepackage{graphicx}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{syntax}
\usepackage{stmaryrd}
\usepackage{mathtools}
\usepackage{mathpartir}
\usepackage{listings}

\newcommand{\LNLD}{\text{LNL}_\text{D}}

% title and author
\title{Dependent Linear Type Theory for Resource Aware Certified Programming}
\author{Qiancheng Fu}

% document body
\begin{document}
  \maketitle 
  \begin{abstract}
    TBD
  \end{abstract}

  \section{Introduction}
  In the field of certified programming, it is of paramount importance for the specification language to precisely characterize the behavior of programs. For this reason, dependent type theory enjoys great success as the type level language of specifications is exactly the same as the term level programming language, giving tremendous power and control to specification writing. The verification achievements carried out by proof assistants such as Coq or F* is a testament how much dependent type theory has accomplished.

  Dependent type theory is however, not without its flaws. Due to its origins as a logical foundation for mathematics, dependent type theory lacks facilities to reason about resource, a concept that is largely absent from mathematics but ever present in computer science. Objects in dependent type theory can be freely duplicated or deleted by appealing to the structural contraction and weakening rules, this goes against the conservation principle of resource. Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource. As language designers, we hope to alleviate this shortcoming of dependent type theory.

  In the disparate world of substructural logic, Girard introduces Linear Logic in his seminal work. Linear Logic restricts weakening and contraction rules of classical logic, giving rise to an elegant formal foundation for reasoning about resource. Abramsky and Wadler notice the opportunities presented by Linear Logic for resource aware programming, pioneering the development of simple linear type theory. Linear type theory enforces linearity on variable usage, which conservatively subsumes resource usage. This reduces the problem of resource reasoning to linear type checking. The successful extraction of type theory from Linear Logic offers a tantalizing hint to overcoming the challenges of resource reasoning in dependent type theory: integrate linear types.

  The earliest work on integrating dependent type theory and linear type theory was carried out by Cervesato and Pfenning, extending the Edinburgh LF with linearity. Xi extends DML dependent types with linear types in the ATS programming language. Krishnaswami et al. develop $\LNLD$ based on Benton's LNL calculus for simple linear types, allowing term dependency on its non-linear fragment. The authors of $\LNLD$ demonstrate its capacity for certified imperative programming in the style of L3 and Hoare Type Theory.  Luo et al. introduce the notion of essential linearity, developing the first type theory that allows types to depend on linear terms. Atkey's QTT, based on initial ideas of McBride, neatly reduces dependency and linearity checking to checking constraints defined on semi-rings. Idris 2 is a full blown programming language that implements QTT as its core type system.

  The lack of weakening and contraction rules for linear types present unique difficulties when integrating full spectrum dependent types, often making dependent linear types challenging or unsatisfactory for practical use. $\LNLD$ requires use of explicit modality maps F and G to alternate between its linear and non-linear fragment, computation requiring back and forth applications of F and G quickly explode in complexity. Terms in Luo's type theory are not intrinsically linear, linearity is solely determined by annotations. The unfortunate side-effect of this design choice is that linear computations never return non-linear values, contradicting the intuition of retrieving a non-linear int from a linear reference. QTT faces the same drawbacks as Luo due to linearity being determined by similar semi-ring annotations.

  We present a full spectrum dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators or annotations. We believe that linearity is an intrinsic property completely determined by the type of a term. To accomplish this, our type system uses two class of type universes in the style of Krishnaswami, $L$ for linear and $U$ for non-linear. Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $A \multimap B$ for linear functions in the style of Pfenning. We formalize an operational semantics and prove the type soundness of our system. To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for internalized certified imperative programming. We augment our soundness theorem to account for these extensions, showing that imperative resources are properly managed. We have implemented a prototype of our language extended with other features for practicality and ergonomics.

  \section{Core Type Theory}

\end{document}

### 2021-09-08 writing/draft.tex
M10

--- 2021-09-06 D0/draft.tex
@@
 \usepackage{listings}
+\usepackage{float}
+
+\floatstyle{boxed} 
+\restylefloat{figure}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+
+\renewcommand{\syntleft}{}
+\renewcommand{\syntright}{}
+\setlength{\grammarparsep}{10pt}
+\setlength{\grammarindent}{16em}
 
 \newcommand{\LNLD}{\text{LNL}_\text{D}}
+\newcommand{\ok}{\textbf{ok}}
+\newcommand{\pure}{\textbf{pure}}
+\newcommand{\sort}{\textbf{sort}}
+\newcommand{\type}{\textbf{type}}
+\newcommand{\U}{\textbf{U}}
+\renewcommand{\L}{\textbf{L}}
+\newcommand{\pstep}{\leadsto_p}
 
@@
 To accomplish this, our type system uses two class of type universes in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
-Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $A \multimap B$ for linear functions in the style of Pfenning.
+Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $(x : A) \multimap B$ for linear functions in the style of Pfenning.
 We formalize an operational semantics and prove the type soundness of our system.
@@
 
+  \begin{figure}[h]
+    \caption{Syntax}
+    \begin{grammar}
+      <$x, y, z, f, g, h$> $\in$ variables
+
+      <$m, n, A, B, C$> $\in$ expressions
+
+      <$v$> $\in$ values
+
+      <expressions> ::= $\U$ | $\L$ | $x$
+      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
+      \alt $(x : A) \multimap B$ | $A \multimap B$
+      \alt $\lambda x. n$ | $m\ n$
+
+      <values> ::= $\U$ | $\L$ | $x$
+      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
+      \alt $(x : A) \multimap B$ | $A \multimap B$
+      \alt $\lambda x. n$
+    \end{grammar}
+  \end{figure}
+
+  \begin{figure}[h]
+    \caption{Structural Judgments}
+    \begin{mathpar}
+      \inferrule
+      { }
+      { \cdot\ ; \cdot \ok }
+
+      \inferrule
+      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \U }
+      { \Gamma, x : A ; \Delta\ \ok }
+
+      \inferrule
+      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \L }
+      { \Gamma ; \Delta, x : A\ \ok }
+
+    \end{mathpar}
+  \end{figure}
+
+  \begin{figure}[h]
+    \caption{Type Formation} 
+    \begin{mathpar}
+      \inferrule
+      { }
+      { \Gamma ; \cdot \vdash \U : \U }
+
+      \inferrule
+      { }
+      { \Gamma ; \cdot \vdash \L : \U }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \U }
+      { \Gamma ; \cdot \vdash A\ \type }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \L }
+      { \Gamma ; \cdot \vdash A\ \type }
+      \\
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \U \\ \Gamma, x : A ; \cdot \vdash B\ \type }
+      { \Gamma ; \cdot \vdash (x : A) \rightarrow B : \U }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \L \\ \Gamma ; \cdot \vdash B\ \type }
+      { \Gamma ; \cdot \vdash A \rightarrow B : \U } 
+      \\
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \U \\ \Gamma, x : A ; \cdot \vdash B\ \type }
+      { \Gamma ; \cdot \vdash (x : A) \multimap B : \L }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A : \L \\ \Gamma ; \cdot \vdash B\ \type }
+      { \Gamma ; \cdot \vdash A \multimap B : \L } 
+    \end{mathpar}
+  \end{figure}
+
+  \begin{figure}[h]
+    \caption{Term Formation}
+    \begin{mathpar}
+      \inferrule
+      { }
+      { \Gamma, x : A ; \cdot \vdash x : A }
+
+      \inferrule
+      { }
+      { \Gamma ; x : A \vdash x : A }
+
+      \inferrule
+      { \Gamma ; \Delta \vdash m : A \\ \Gamma ; \Delta \vdash A \equiv B }
+      { \Gamma ; \Delta \vdash m : B } 
+      \\
+
+      \inferrule
+      { \Gamma ; \cdot \vdash (x : A) \rightarrow B : \U \\ 
+        \Gamma, x : A ; \cdot \vdash n : B }
+      { \Gamma ; \cdot \vdash \lambda x . n : (x : A) \rightarrow B }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A \rightarrow B : \U \\
+        \Gamma ; x : A \vdash n : B }
+      { \Gamma ; \cdot \vdash \lambda x . n : A \rightarrow B }
+      \\
+
+      \inferrule
+      { \Gamma ; \cdot \vdash (x : A) \multimap B : \L \\ 
+        \Gamma, x : A ; \Delta \vdash n : B }
+      { \Gamma ; \Delta \vdash \lambda x . n : (x : A) \multimap B }
+
+      \inferrule
+      { \Gamma ; \cdot \vdash A \multimap B : \L \\ 
+        \Gamma ; \Delta, x : A \vdash n : B }
+      { \Gamma ; \Delta \vdash \lambda x . n : A \multimap B }
+      \\
+
+      \inferrule
+      { \Gamma ; \Delta_1 \vdash m : (x : A) \rightarrow B \\
+        \Gamma ; \Delta_2 \vdash n : A }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
+
+      \inferrule
+      { \Gamma ; \Delta_1 \vdash m : A \rightarrow B \\
+        \Gamma ; \Delta_2 \vdash n : A }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
+      \\
+
+      \inferrule
+      { \Gamma ; \Delta_1 \vdash m : (x : A) \multimap B \\
+        \Gamma ; \Delta_2 \vdash n : A }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
+
+      \inferrule
+      { \Gamma ; \Delta_1 \vdash m : A \multimap B \\
+        \Gamma ; \Delta_2 \vdash n : A }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
+    \end{mathpar}
+  \end{figure}
+
+  \begin{figure}[h]
+    \caption{Equality}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma ; \Delta \vdash m_1 : A \\ \Gamma ; \Delta \vdash m_2 : A \\ 
+        m_1 \pstep^* n \\ m_2 \pstep^* n }
+      { \Gamma ; \Delta \vdash m_1 \equiv m_2 }
+    \end{mathpar}
+  \end{figure}
+
+  \begin{figure}[h]
+    \caption{Parallel Reduction}
+    \begin{mathpar}
+      \inferrule
+      { }
+      { n \pstep n }
+
+      \inferrule
+      { n \pstep n' }
+      { \lambda x.n \pstep \lambda x.n' }
+
+      \inferrule
+      { m \pstep m' \\ n \pstep n' }
+      { m\ n \pstep m'\ n' }
+
+      \inferrule
+      { n \pstep n' \\ v \pstep v' }
+      { (\lambda x.n)\ v \pstep [v'/x]n' }
+      \\
+
+      \inferrule
+      { A \pstep A' \\ B \pstep B' }
+      { (x : A) \rightarrow B \pstep (x : A') \rightarrow B' }
+
+      \inferrule
+      { A \pstep A' \\ B \pstep B' }
+      { A \rightarrow B \pstep A' \rightarrow B' }
+      \\
+
+      \inferrule
+      { A \pstep A' \\ B \pstep B' }
+      { (x : A) \multimap B \pstep (x : A') \multimap B' }
+
+      \inferrule
+      { A \pstep A' \\ B \pstep B' }
+      { A \multimap B \pstep A' \multimap B' }
+    \end{mathpar}
+  \end{figure}
+
 \end{document}

### 2021-09-09 writing/draft.tex
M10

@@
 \usepackage{amssymb}
+\usepackage{mdwtab}
 \usepackage{syntax}
@@
 \setlength{\grammarparsep}{10pt}
-\setlength{\grammarindent}{16em}
-
+\setlength{\grammarindent}{10em}
+\newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
+\newcommand{\rname}[1]{\textsc{\footnotesize #1}}
 \newcommand{\LNLD}{\text{LNL}_\text{D}}
@@
 \renewcommand{\L}{\textbf{L}}
+\newcommand{\lambdah}{\hat{\lambda}}
+\newcommand{\betah}{\hat{\beta}}
 \newcommand{\pstep}{\leadsto_p}
@@
 This reduces the problem of resource reasoning to linear type checking.
+Languages based on linear type systems have shown the capacity to safely manipulate imperative data structures.
 The successful extraction of type theory from Linear Logic offers a tantalizing hint to overcoming the challenges of resource reasoning in dependent type theory: integrate linear types.
@@
 To accomplish this, our type system uses two class of type universes in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
-Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $(x : A) \multimap B$ for linear functions in the style of Pfenning.
-We formalize an operational semantics and prove the type soundness of our system.
-To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for internalized certified imperative programming.
+Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $(x : A) \multimap B$ for linear functions.
+We formalize a call-by-value operational semantics and prove the type soundness of our system.
+To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for certified imperative programming.
 We augment our soundness theorem to account for these extensions, showing that imperative resources are properly managed.
@@
   \section{Core Type Theory}
+  In the efforts of integrating dependency and linearity, there are two predominant schools of thought: \textit{Types can only depended on non-linear terms} and \textit{Types can depend on all terms}.
+
+  Advocates of restricted dependency argue that the restriction is both intuitive and a necessity.
+A natural example of linear types depending on a non-linear terms is that of the length indexed random access array.
+Term dependency on array length statically prevents out-of-bounds array access whilst linearity ensures proper memory management.
+In the converse situation of types depending on linear terms, one immediately faces a dilemma: \textit{Do linear terms at the type level possess weakening and contraction?}.
+
+  If one answers yes, then the linear terms at the type level are defacto non-linear.
+In this case, a language with restricted dependency can simulate linear dependency by depending on non-linear terms that reflect the shape of linear terms, erasing the advantage of increased type level expressivity.
+All prior works featuring linear dependency fall into this category, for good reason: \textit{the alternative is worse}.
+
+  If type level linear terms do not possess weakening and contraction, types with linear dependencies themselves are linear.
+This breaks the very notion of the typing judgment as even $\alpha$-equivalent terms may not possess the same type.
+Though the idea of such a type system is deeply interesting, the philosophical burden is prohibitive.
+
+  Our language firmly falls in the former category.
+Besides the reasons listed above, restricted dependency gives a clear distinction between linear and non-linear terms, a property we believe will aid in the compilation to high performance code.
+
+  \subsection{Syntax}
+  The syntax of the core type theory is presented in Figure \ref{syntax}.
+Our type theory contains two universes, $\U$ and $\L$ for denoting the universe of non-linear types and linear types respectively.
+An important fact to note is that our language possess the "Type-in-Type" axiom in the form of $\Gamma ; \cdot \vdash \U : \U$ which is well known to be logically unsound.
+We chose this design deliberately as we are willing to sacrifice logical soundness for increased expressivity and convenience.
+We forsee no issue in extending the core language with a hierarchy of universes to enforce logical soundness.
 
@@
     \caption{Syntax}
+    \centering
+    \begin{minipage}{0.8\linewidth}
     \begin{grammar}
-      <$x, y, z, f, g, h$> $\in$ variables
+      <$m, n, A, B, C$> ::= $\U$ | $\L$ | $x$  \hspace*{8em} expressions
+      \indalt $(x : A) \rightarrow B$ | $A \rightarrow B$
+      \indalt $(x : A) \multimap B$ | $A \multimap B$
+      \indalt $\lambda x. n$ | $\lambdah x.n$ | $m\ n$
 
-      <$m, n, A, B, C$> $\in$ expressions
+      <$v$> ::= $\U$ | $\L$ | $x$ \hspace*{8em} values
+      \indalt $(x : A) \rightarrow B$ | $A \rightarrow B$
+      \indalt $(x : A) \multimap B$ | $A \multimap B$
+      \indalt $\lambda x. n$ | $\lambdah x.n$
+    \end{grammar}
+    \end{minipage}
+    \label{syntax}
+  \end{figure}
 
-      <$v$> $\in$ values
+  A clear departure of our language from standard presentations of type theory is the presence of four function types: $(x : A) \rightarrow B$, $A \rightarrow B$, $(x : A) \multimap B$, $A \multimap B$.
+The reason for these variants is the decoupling of function linearity from the linearity of function domain and co-domain.
+Instead, the linearity of functions is determined by the linearity of the closure it forms.
+For example, a function with linear input and output type can itself be non-linear if its free variables have non-linear type, indicating this function can be applied repeatedly without contraction of linear variables.
+In practice, these function types can be consolidated to just $(x : A) \rightarrow B$ and $(x : A) \multimap B$ by typechecking using Luo's mixed context and $merge$ techniques.
+For the sake of clarity, we leave them as four.
 
-      <expressions> ::= $\U$ | $\L$ | $x$
-      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
-      \alt $(x : A) \multimap B$ | $A \multimap B$
-      \alt $\lambda x. n$ | $m\ n$
+  \subsection{Context and Structural Judgments}
+  The context of our type theory is divided into two partitions, $\Gamma$ for variables of non-linear type and $\Delta$ for variables of linear type.
+Weakening and contraction structural rules are permitted on $\Gamma$ but forbidden on $\Delta$.
+Structural rules for well-founded contexts are shown in Figure \ref{structural}.
 
-      <values> ::= $\U$ | $\L$ | $x$
-      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
-      \alt $(x : A) \multimap B$ | $A \multimap B$
-      \alt $\lambda x. n$
-    \end{grammar}
-  \end{figure}
+  Types in our language are determined to be linear or non-linear by the universe they belong to.
+The typing judgment $\Gamma; \Delta \vdash A : \U$ asserts that $A$ is a non-linear type under context $\Gamma; \Delta$, whereas the judgment $\Gamma; \Delta \vdash B : \L$ asserts that $B$ is a linear type under context $\Gamma; \Delta$.
+In the rules for constructing well-founded contexts, care is taken to prevent $\Delta$ from leaking into typing judgments, restricting type dependency to $\Gamma$.
 
@@
       { \cdot\ ; \cdot \ok }
+      \rname{Empty-Ok}
 
       \inferrule
-      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \U }
-      { \Gamma, x : A ; \Delta\ \ok }
+      { \Gamma ; \cdot\ \ok \\ \Gamma ; \cdot \vdash A : \U }
+      { \Gamma, x : A ; \cdot\ \ok }
+      \rname{U-Ok}
 
       \inferrule
-      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \L }
+      { \Gamma ; \Delta\ \ok \\ \Gamma ; \cdot \vdash A : \L }
       { \Gamma ; \Delta, x : A\ \ok } 
+      \rname{L-Ok}
+    \end{mathpar}
+    \label{structural}
+  \end{figure}
 
-    \end{mathpar}
-  \end{figure}
+  \subsection{Type Formation}
+  Type formation rules are presented in Figure \ref{type}. Rules worth detailed discussion are the \rname{L-Axiom} and the rules for constructing various function types.
+
+  For \rname{L-Axiom}, the universe of linear types itself belongs to the universe of non-linear types.
+This means that a linear type is itself a non-linear term, avoiding the philosophical trappings discussed above.
+
+  The function types $(x : A) \rightarrow B$ and $A \rightarrow B$ represent non-linear functions.
+Terms of these types contain no linear free variables, thus can be applied repeatedly without duplication of linear variables.
+The difference between the two is that $(x : A) \rightarrow B$ allows co-domain type $B$ to depended on the input term $x$ of non-linear type $A$.
+For $A \rightarrow B$, the domain type $A$ is linear so $B$ is not allowed to depend on the function input.
+
+  The variants $(x : A) \multimap B$ and $A \multimap B$ represent linear functions.
+Terms of these types may contain linear free variables, thus cannot be applied repeated without duplication of linear variables.
+Similar to the non-linear versions, the difference between $(x : A) \multimap B$ and $A \multimap B$ lies in the allowance of dependency on function input for non-linear domain types.
 
@@
       { \Gamma ; \cdot \vdash \U : \U } 
+      \rname{U-Axiom}
 
@@
       { \Gamma ; \cdot \vdash \L : \U } 
+      \rname{L-Axiom}
 
@@
       { \Gamma ; \cdot \vdash A\ \type }
+      \rname{U-Type}
 
@@
       { \Gamma ; \cdot \vdash A\ \type }
+      \rname{L-Type}
       \\
@@
       { \Gamma ; \cdot \vdash (x : A) \rightarrow B : \U } 
+      \rname{U-Prod}
 
@@
       { \Gamma ; \cdot \vdash A \rightarrow B : \U } 
+      \rname{Arrow}
       \\
@@
       { \Gamma ; \cdot \vdash (x : A) \multimap B : \L } 
+      \rname{L-Prod}
 
@@
       { \Gamma ; \cdot \vdash A \multimap B : \L } 
+      \rname{Lolli}
     \end{mathpar}
+    \label{type}
   \end{figure}
+
+  \subsection{Term Formation}
+  Term formation rules are presented in Figure \ref{term}.
 
@@
       { \Gamma, x : A ; \cdot \vdash x : A } 
+      \rname{U-Var}
 
@@
       { \Gamma ; x : A \vdash x : A } 
+      \rname{L-Var}
+      \\
 
       \inferrule
-      { \Gamma ; \Delta \vdash m : A \\ \Gamma ; \Delta \vdash A \equiv B }
+      { \Gamma ; \Delta \vdash m : A \\ 
+        \Gamma ; \cdot \vdash A \equiv B : \U }
       { \Gamma ; \Delta \vdash m : B } 
+      \rname{U-Conv}
+
+      \inferrule
+      { \Gamma ; \Delta \vdash m : A \\ 
+        \Gamma ; \cdot \vdash A \equiv B : \L }
+      { \Gamma ; \Delta \vdash m : B } 
+      \rname{L-Conv}
       \\
@@
       { \Gamma ; \cdot \vdash \lambda x . n : (x : A) \rightarrow B }
+      \rname{U-$\lambda_1$}
+      \\
 
@@
       { \Gamma ; \cdot \vdash \lambda x . n : A \rightarrow B }
+      \rname{U-$\lambda_2$}
       \\
@@
         \Gamma, x : A ; \Delta \vdash n : B }
-      { \Gamma ; \Delta \vdash \lambda x . n : (x : A) \multimap B }
+      { \Gamma ; \Delta \vdash \lambdah x . n : (x : A) \multimap B }
+      \rname{L-$\lambda_1$}
 
@@
         \Gamma ; \Delta, x : A \vdash n : B }
-      { \Gamma ; \Delta \vdash \lambda x . n : A \multimap B }
+      { \Gamma ; \Delta \vdash \lambdah x . n : A \multimap B }
+      \rname{L-$\lambda_2$}
       \\
@@
         \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : (\lambda x.B)\ n }
+      \rname{U-App-1}
 
@@
       { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
+      \rname{U-App-2}
       \\
@@
         \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
+      { \Gamma; \Delta_1, \Delta_2 : m\ n : (\lambda x.B)\ n }
+      \rname{L-App-1}
 
@@
       { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
+      \rname{L-App-2}
     \end{mathpar}
+    \label{term}
   \end{figure}
+
+  \subsection{Equality and Reduction}
 
@@
         m_1 \pstep^* n \\ m_2 \pstep^* n }
-      { \Gamma ; \Delta \vdash m_1 \equiv m_2 }
+      { \Gamma ; \Delta \vdash m_1 \equiv m_2 : A }
     \end{mathpar}
@@
       { }
-      { n \pstep n }
+      { x \pstep x }
+      \rname{PR-Var}
+
+      \inferrule
+      { }
+      { \U \pstep \U }
+      \rname{PR-U}
+
+      \inferrule
+      { }
+      { \L \pstep \L }
+      \rname{PR-L}
 
@@
       { \lambda x.n \pstep \lambda x.n' }
+      \rname{PR-$\lambda$}
+
+      \inferrule
+      { n \pstep n' }
+      { \lambdah x.n \pstep \lambdah x.n' }
+      \rname{PR-$\lambdah$}
 
@@
       { m\ n \pstep m'\ n' }
+      \rname{PR-App}
 
@@
       { (\lambda x.n)\ v \pstep [v'/x]n' }
-      \\
+      \rname{PR-$\beta$}
+
+      \inferrule
+      { n \pstep n' \\ v \pstep v' }
+      { (\lambdah x.n)\ v \pstep [v'/x]n' }
+      \rname{PR-$\betah$}
 
@@
       { (x : A) \rightarrow B \pstep (x : A') \rightarrow B' }
+      \rname{PR-U-Prod}
 
@@
       { A \rightarrow B \pstep A' \rightarrow B' }
-      \\
+      \rname{PR-Arrow}
 
@@
       { (x : A) \multimap B \pstep (x : A') \multimap B' }
+      \rname{PR-L-Prod}
 
@@
       { A \multimap B \pstep A' \multimap B' }
+      \rname{PR-L-Lolli}
     \end{mathpar}
@@
 
+  \subsection{Meta Theory}
+
+  \begin{lemma}
+    Parallel reduction satisfies the diamond property. If $m \pstep m_1$ and $m \pstep m_2$ then there exists $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+  \end{lemma}
+
+  \begin{lemma}
+    The transitive reflexive closure of parallel reduction is confluent. If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
+  \end{lemma}
+
 \end{document}

### 2021-09-10 writing/draft.tex
M10

@@
 \newcommand{\betah}{\hat{\beta}}
+\newcommand{\step}{\leadsto}
 \newcommand{\pstep}{\leadsto_p}
@@
 
-  \subsection{Type Formation}
+  \subsection{Type Formation} \label{tyformation}
   Type formation rules are presented in Figure \ref{type}. Rules worth detailed discussion are the \rname{L-Axiom} and the rules for constructing various function types.
@@
   Term formation rules are presented in Figure \ref{term}.
+As outlined previously, terms constructed by $\lambda$ for non-linear function types $(x : A) \rightarrow B$ and $A \rightarrow B$ are only well type if their function bodies contain only non-linear variables listed in $\Gamma$.
+Conversely, terms constructed by $\lambdah$ for linear types $(x : A) \multimap B$ and $A \multimap B$ are allowed to contain both linear and non-linear variables listed in $\Gamma; \Delta$.
+
+  In classical dependent type theories such as Martin L\"{o}f Type Theory or Calculus of Constructions, strong normalization and confluence ensure that term normalization is agnostic to reduction strategy.
+Furthermore, the preservation theorem shows that reduction preserves the typing judgment of well-typed terms.
+The management of substructural contexts for dependent linear types complicates the direct use of these theorems.
+In the application rules $\rname{U-App-1}$ and $\rname{L-App-1}$, the argument $n$ is a non-linear term that may contain free variables in linear context $\Delta$.
+If a call-by-name reduction strategy is used, substitution of $n$ into the function body may cause duplication of its linear free variables.
+Some existing literature ban non-linear arguments from containing linear free variables.
+We find this approach to be overly restrictive as it excludes the pattern of freely using retrieved data (non-linear) from references (linear) ubiquitous in imperative programs.
+Instead, we adjust the operational semantics to favor call-by-value reduction strategy.
+Our value soundness theorem guarantees non-linear values to contain only non-linear free variables.
+Thus, substitution will no longer duplicate the linear free variables found in arguments.
 
@@
 
-  \subsection{Equality and Reduction}
+  \subsection{Reduction and Equality}
+  Figure \ref{single} presents the call-by-value operational semantics that we have eluded to. The rules defining the single step relation $\step$ are completely standard.
+
+  For full spectrum dependently typed languages, terms can appear at the type level.
+A definitional equality judgment is required to identify types beyond simple $\alpha$-equivalence.
+This is usually accomplished by normalization and comparing normal forms.
+Due to the complications brought on by linearity outlined in \ref{tyformation}, standard normalization techniques cannot be directly applied.
+Unfortunately, the single step relation $\step$ defined previously is not sufficient either as binders block full normalization.
+Following in the footsteps of the Trellys project, we define a call-by-value parallel step relation $\pstep$.
+The $\pstep$ relation is essentially the same as $\step$ except that it normalizes under binders.
+We use the transitive reflexive closure of $\pstep$ to define definitional equality.
+
+  \begin{figure}[h]
+    \caption{Single Step Reduction}
+    \begin{mathpar}
+      \inferrule
+      { }
+      { (\lambda x.n)\ v \step [v/x]n }
+      \rname{S-$\beta$}
+
+      \inferrule
+      { }
+      { (\lambdah x.n)\ v \step [v/x]n }
+      \rname{S-$\betah$}
+
+      \inferrule
+      { m \step m' }
+      { m\ n \step m'\ n }
+      \rname{S-App-L}
+
+      \inferrule
+      { n \step n' }
+      { v\ n \step v\ n' }
+      \rname{S-App-R}
+    \end{mathpar}
+    \label{single}
+  \end{figure}
 
@@
       { x \pstep x }
-      \rname{PR-Var}
+      \rname{P-Var}
 
@@
       { \U \pstep \U }
-      \rname{PR-U}
+      \rname{P-U}
 
@@
       { \L \pstep \L }
-      \rname{PR-L}
+      \rname{P-L}
 
@@
       { \lambda x.n \pstep \lambda x.n' }
-      \rname{PR-$\lambda$}
+      \rname{P-$\lambda$}
 
@@
       { \lambdah x.n \pstep \lambdah x.n' }
-      \rname{PR-$\lambdah$}
+      \rname{P-$\lambdah$}
 
@@
       { m\ n \pstep m'\ n' }
-      \rname{PR-App}
+      \rname{P-App}
 
@@
       { (\lambda x.n)\ v \pstep [v'/x]n' }
-      \rname{PR-$\beta$}
+      \rname{P-$\beta$}
 
@@
       { (\lambdah x.n)\ v \pstep [v'/x]n' }
-      \rname{PR-$\betah$}
+      \rname{P-$\betah$}
 
@@
       { (x : A) \rightarrow B \pstep (x : A') \rightarrow B' }
-      \rname{PR-U-Prod}
+      \rname{P-U-Prod}
 
@@
       { A \rightarrow B \pstep A' \rightarrow B' }
-      \rname{PR-Arrow}
+      \rname{P-Arrow}
 
@@
       { (x : A) \multimap B \pstep (x : A') \multimap B' }
-      \rname{PR-L-Prod}
+      \rname{P-L-Prod}
 
@@
       { A \multimap B \pstep A' \multimap B' }
-      \rname{PR-L-Lolli}
+      \rname{P-L-Lolli}
     \end{mathpar}
+    \label{parallel}
   \end{figure}
@@
   \subsection{Meta Theory}
+  \begin{lemma}
+    Single step reduction implies parallel step reduction. If $m \step m'$, then $m \pstep m'$.
+  \end{lemma}
 
@@
 
-  \begin{lemma}
+  \begin{corollary}
     The transitive reflexive closure of parallel reduction is confluent. If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
-  \end{lemma}
+  \end{corollary}
+
+  \begin{theorem}
+    Types are non-linear. If $\Gamma; \Delta \vdash n : A$ then either $\Gamma; \cdot \vdash A : \U$ or $\Gamma; \cdot \vdash A : \L$.
+  \end{theorem}
+
+  \begin{theorem}
+    Value soundness. If $\Gamma; \Delta \vdash v : A$ then $\Delta = \cdot$.
+  \end{theorem}
 

### 2021-09-22 writing/draft.tex
M10
writing

--- 2021-09-08 writing/draft.tex
@@
 \newcommand{\type}{\textbf{type}}
-\newcommand{\U}{\textbf{U}}
-\renewcommand{\L}{\textbf{L}}
+\newcommand{\utype}{\overset{U}{:}}
+\newcommand{\ltype}{\overset{L}{:}}
+\newcommand{\stype}[1]{\overset{#1}{:}}
+\newcommand{\step}{\leadsto}
 \newcommand{\pstep}{\leadsto_p}
@@
 For this reason, dependent type theory enjoys great success as the type level language of specifications is exactly the same as the term level programming language, giving tremendous power and control to specification writing.
-The verification achievements carried out by proof assistants such as Coq or F* is a testament how much dependent type theory has accomplished.
+The verification achievements\cite{compcert,deepspec,everest} carried out by proof assistants such as Coq\cite{coq} or F*\cite{fstar} is a testament how much dependent type theory has accomplished.
 
   Dependent type theory is however, not without its flaws.
-Due to its origins as a logical foundation for mathematics, dependent type theory lacks facilities to reason about resource, a concept that is largely absent from mathematics but ever present in computer science.
-Objects in dependent type theory can be freely duplicated or deleted by appealing to the structural contraction and weakening rules, this goes against the conservation principle of resource.
-Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource.
+Due to its origins as a logical foundation for mathematics\cite{martinlof,cc}, dependent type theory lacks facilities to reason about resource, a concept that is largely absent from mathematics but ever present in computer science.
+Objects in dependent type theory can be freely duplicated or deleted by appealing to the contraction and weakening rules, this goes against the conservation principle of resource.
+Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource\cite{bedrock}.
 As language designers, we hope to alleviate this shortcoming of dependent type theory.
 
-  In the disparate world of substructural logic, Girard introduces Linear Logic in his seminal work.
+  In the disparate world of substructural logic, Girard introduces Linear Logic in his seminal work\cite{girard}.
 Linear Logic restricts weakening and contraction rules of classical logic, giving rise to an elegant formal foundation for reasoning about resource.
-Abramsky and Wadler notice the opportunities presented by Linear Logic for resource aware programming, pioneering the development of simple linear type theory.
-Linear type theory enforces linearity on variable usage, which conservatively subsumes resource usage.
+Wadler\cite{wadler1990,wadler1991} and Abramsky\cite{abramsky1993} notice the opportunities presented by Linear Logic for resource aware programming, pioneering the development of simple linear type theory as a programming language.
+Linear type theory enforces safe variable usage, which conservatively subsumes resource usage.
 This reduces the problem of resource reasoning to linear type checking.
+Languages based on linear type systems have shown the capacity to safely manipulate imperative data structures\cite{l3,ats}.
 The successful extraction of type theory from Linear Logic offers a tantalizing hint to overcoming the challenges of resource reasoning in dependent type theory: integrate linear types.
 
-  The earliest work on integrating dependent type theory and linear type theory was carried out by Cervesato and Pfenning, extending the Edinburgh LF with linearity.
-Xi extends DML dependent types with linear types in the ATS programming language.
+  The earliest work on integrating dependent type theory and linear type theory was carried out by Cervesato and Pfenning\cite{llf}, extending the Edinburgh LF with linear types.
+Xi extends DML dependent types with linear types in the ATS programming language\cite{ats}.
 Krishnaswami et al.
-develop $\LNLD$ based on Benton's LNL calculus for simple linear types, allowing term dependency on its non-linear fragment.
-The authors of $\LNLD$ demonstrate its capacity for certified imperative programming in the style of L3 and Hoare Type Theory.
+develop $\LNLD$\cite{neel15} based on Benton's LNL\cite{benton1994} calculus for simple linear types, allowing term dependency on its non-linear fragment.
+The authors of $\LNLD$ demonstrate its capacity for certified imperative programming in the style of L3\cite{l3} and Hoare Type Theory\cite{htt}.
 Luo et al.
-introduce the notion of essential linearity, developing the first type theory that allows types to depend on linear terms.
-Atkey's QTT, based on initial ideas of McBride, neatly reduces dependency and linearity checking to checking constraints defined on semi-rings.
-Idris 2 is a full blown programming language that implements QTT as its core type system.
+introduce the notion of essential linearity\cite{luo}, developing the first type theory that allows types to depend on linear terms.
+Atkey's QTT\cite{qtt}, based on initial ideas of McBride\cite{nothing}, neatly reduces dependency and linearity checking to checking constraints defined on semi-rings.
+Idris 2\cite{idris2} is a full featured programming language that implements QTT as its core type system.
 
   The lack of weakening and contraction rules for linear types present unique difficulties when integrating full spectrum dependent types, often making dependent linear types challenging or unsatisfactory for practical use.
-$\LNLD$ requires use of explicit modality maps F and G to alternate between its linear and non-linear fragment, computation requiring back and forth applications of F and G quickly explode in complexity.
+$\LNLD$ requires use of explicit modality maps F and G to alternate between its linear and non-linear fragment.
+Computation requiring back and forth applications of F and G quickly explode in complexity.
 Terms in Luo's type theory are not intrinsically linear, linearity is solely determined by annotations.
-The unfortunate side-effect of this design choice is that linear computations never return non-linear values, contradicting the intuition of retrieving a non-linear int from a linear reference.
-QTT faces the same drawbacks as Luo due to linearity being determined by similar semi-ring annotations.
+The unfortunate side-effect of this design choice is that linear computations never return non-linear values, eliminating the common pattern of retrieving a non-linear values from a linear references.
+QTT faces the similar drawbacks to Luo due to linearity being determined by semi-ring annotations.
 
-  We present a full spectrum dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators or annotations.
-We believe that linearity is an intrinsic property completely determined by the type of a term.
-To accomplish this, our type system uses two class of type universes in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
-Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow B$ for non-linear functions and $(x : A) \multimap B$ for linear functions in the style of Pfenning.
-We formalize an operational semantics and prove the type soundness of our system.
-To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for internalized certified imperative programming.
+  We present a full spectrum dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators.
+We believe that linearity is an intrinsic property that can be determined by the type of a term.
+To accomplish this, our type system uses two class of type sorts in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
+Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow_s B$ for non-linear functions and $(x : A) \multimap_s B$ for linear functions.
+We formalize a call-by-value operational semantics and prove the type soundness of our system.
+To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for certified imperative programming.
 We augment our soundness theorem to account for these extensions, showing that imperative resources are properly managed.
@@
   \section{Core Type Theory}
+  In the efforts of integrating dependency and linearity, there are two predominant schools of thought: \textit{Types can only depended on non-linear terms} and \textit{Types can depend on all terms}.
 
-  \begin{figure}[h]
+  Advocates of restricted dependency argue that the restriction is both intuitive and a necessity.
+A natural example of linear types depending on a non-linear terms is that of the length indexed random access array.
+Term dependency on array length statically prevents out-of-bounds array access whilst linearity ensures proper memory management.
+In the converse situation of types depending on linear terms, one immediately faces a dilemma: \textit{Do linear terms at the type level possess weakening and contraction?}.
+
+  If one answers yes, then the linear terms at the type level are defacto non-linear.
+In this case, a language with restricted dependency can simulate linear dependency by depending on non-linear terms that reflect the shape of linear terms\cite{ldqt}, erasing the advantage of increased type level expressivity.
+All prior works featuring linear dependency fall into this category, for good reason: \textit{the alternative is worse}.
+
+  If type level linear terms do not possess weakening and contraction, types with linear dependencies themselves are linear.
+This breaks the very notion of the typing judgment as even $\alpha$-equivalent terms are not allowed to possess the same type.
+Though the idea of such a type system is deeply interesting, the philosophical burden is prohibitive.
+
+  Our language firmly falls in the former category.
+Besides the reasons listed above, restricted dependency gives a clear distinction between linear and non-linear terms, a property we believe will aid in the compilation to high performance code.
+
+  \subsection{Syntax}
+  The syntax of the core type theory is presented in Figure \ref{syntax}.
+Our type theory contains two sorts, $U$ and $L$ for denoting the sort of non-linear types and linear types respectively.
+An important fact to note is that our language possess the "Type-in-Type" axiom in the form of $\Gamma \vdash U \utype U$ which is well known to be logically unsound.
+We chose this design deliberately as we are willing to sacrifice logical soundness for increased expressivity and convenience.
+We forsee no issue in extending the core language with a hierarchy of sorts to enforce logical soundness.
+
+  \begin{figure}[H]
     \caption{Syntax}
+    \centering
+    \begin{minipage}{0.8\linewidth}
     \begin{grammar}
-      <$x, y, z, f, g, h$> $\in$ variables
+      <$s, t$> ::= $U$ | $L$ \phantom{| $x$} \hspace*{8em} sorts
 
-      <$m, n, A, B, C$> $\in$ expressions
+      <$m, n, A, B, C$> ::= $U$ | $L$ | $x$  \hspace*{8em} expressions
+      \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
+      \indalt $(x : A) \multimap_s B$ | $A \multimap_s B$
+      \indalt $\lambda x. n$ | $m\ n$
 
-      <$v$> $\in$ values
-
-      <expressions> ::= $\U$ | $\L$ | $x$
-      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
-      \alt $(x : A) \multimap B$ | $A \multimap B$
-      \alt $\lambda x. n$ | $m\ n$
-
-      <values> ::= $\U$ | $\L$ | $x$
-      \alt $(x : A) \rightarrow B$ | $A \rightarrow B$
-      \alt $(x : A) \multimap B$ | $A \multimap B$
-      \alt $\lambda x. n$
+      <$v$> ::= $U$ | $L$ | $x$ \hspace*{8em} values
+      \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
+      \indalt $(x : A) \multimap_s B$ | $A \multimap_s B$
+      \indalt $\lambda x. n$
     \end{grammar}
+    \end{minipage}
+    \label{syntax}
   \end{figure}
 
-  \begin{figure}[h]
+  A clear departure of our language from standard presentations of type theory is the presence of four function types: $(x : A) \rightarrow_s B$, $A \rightarrow_s B$, $(x : A) \multimap_s B$, $A \multimap_s B$.
+The reason for these variants is the decoupling of function linearity from the linearity of function domain and co-domain.
+Instead, the linearity of functions is determined by the linearity of the closure it forms.
+For example, a function with linear input and output type can itself be non-linear if its free variables have non-linear type, indicating this function can be applied repeatedly without contraction of linear variables.
+In practice, these function types can be consolidated to just $(x : A) \rightarrow_s B$ and $(x : A) \multimap_s B$ by viewing $A \rightarrow_s B$ and $A \multimap_s B$ as their non-binding special cases.
+For the sake of clarity, we leave them as four.
+
+  \subsection{Context and Structural Judgments}
+  The context of our language employs a mixed linear/non-linear representation introduced by Luo\cite{luo}.
+Variables in the context are annotated to indicate whether they are linear or non-linear.
+A non-linear variable is annotated as $\Gamma, x \utype A$, whereas a linear variable is annotated as $\Gamma, x \ltype A$.
+
+  We define a $merge$ relation to combine two mixed contexts $\Gamma_1$ and $\Gamma_2$ by performing contraction on shared non-linear variables.
+For linear variables, the $merge$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
+The definition of $merge$ allows contraction for non-linear variables whilst forbidding it for linear ones.
+
+  An auxiliary judgment $\pure$ is defined to assert that a context $\Gamma$ do not contain linear variables.
+In other words, all variables found in a pure context are of the form $\Gamma, x \utype A$.
+A restriction function $\overline{\Gamma}$ is defined that removes all linear variables from context $\Gamma$, the result of restriction is a pure subset of the original context.
+
+  \begin{figure}[H]
     \caption{Structural Judgments}
@@
       { }
-      { \cdot\ ; \cdot \ok }
+      { \epsilon \ \ok }
+      \rname{Ok-$\epsilon$}
 
       \inferrule
-      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \U }
-      { \Gamma, x : A ; \Delta\ \ok }
+      { \Gamma\ \ok \\ 
+        \overline{\Gamma} \vdash A \utype U }
+      { \Gamma, x \utype A\ \ok }
+      \rname{Ok-U}
 
       \inferrule
-      { \Gamma ; \Delta\ \ok \\ \Gamma ; \Delta \vdash A : \L }
-      { \Gamma ; \Delta, x : A\ \ok }
+      { \Gamma\ \ok \\ 
+        \overline{\Gamma} \vdash A \utype L }
+      { \Gamma, x \ltype A\ \ok } 
+      \rname{Ok-L}
+      \\
 
+      \inferrule
+      { }
+      { \epsilon\ \pure }
+      \rname{Pure-$\epsilon$}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype U }
+      { \Gamma, x \utype A\ \pure }
+      \rname{Pure-U}
+      \\
+
+      \inferrule
+      { }
+      { merge\ \epsilon\ \epsilon\ \epsilon }
+      \rname{Merge-$\epsilon$}
+
+      \inferrule
+      { merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+      { merge\ (\Gamma_1, x \utype A)
+             \ (\Gamma_2, x \utype A)
+             \ (\Gamma, x \utype A) }
+      \rname{Merge-U}
+
+      \inferrule
+      { merge\ \Gamma_1\ \Gamma_2\ \Gamma \\
+        x \notin \Gamma_2 }
+      { merge\ (\Gamma_1, x \ltype A)
+             \ \Gamma_2
+             \ (\Gamma, x \ltype A) }
+      \rname{Merge-L1}
+
+      \inferrule
+      { merge\ \Gamma_1\ \Gamma_2\ \Gamma \\
+        x \notin \Gamma_1 }
+      { merge\ \Gamma_1
+             \ (\Gamma_2, x \ltype A)
+             \ (\Gamma, x \ltype A) }
+      \rname{Merge-L2} 
     \end{mathpar}
+    \begin{align*}
+      \overline{\epsilon} = \epsilon
+      \hspace*{4em}
+      \overline{\Gamma, x \utype A} = \overline{\Gamma}, x \utype A 
+      \hspace*{4em}
+      \overline{\Gamma, x \ltype A} = \overline{\Gamma}
+    \end{align*}
+    \label{structural}
   \end{figure}
 
-  \begin{figure}[h]
+  \subsection{Type Formation} \label{tyformation}
+  Typing judgments in our language take on the form $\Gamma \vdash m \stype{s} A$, $s$ is an indexing sort that is either $U$ or $L$.
+Intuitively, this judgment states that expression $m$ has type $A$ and sort $s$ under context $\Gamma$.
+An expression of sort $L$ is linear, and sort $U$ is non-linear.
+Surprisingly, duplication of non-linear expression is not immediately allowed, only non-linear \textit{values} can be safely duplicated.
+We will discuss this subtlety later in more detail.
+
+  Type formation rules are presented in Figure \ref{type}.
+Rules worth discussing in detail are the \rname{L-Axiom} and the rules for constructing various function types.
+For \rname{L-Axiom}, the sort of all linear types $L$ itself is of non-linear sort $U$.
+This means that a linear type can be freely used, avoiding the philosophical trappings discussed earlier.
+
+  The function types $(x : A) \rightarrow_s B$ and $A \rightarrow_s B$ represent non-linear functions, where the subscript $s$ is the sort of co-domain $B$.
+Terms of these types contain no linear free variables, thus can be applied repeatedly without duplication of linear resources.
+The difference between the two is that $(x : A) \rightarrow_s B$ allows co-domain $B$ to depended on input $x$ of non-linear domain $A$.
+For $A \rightarrow_s B$, the domain $A$ is linear so $B$ is not allowed to depend on the function input.
+
+  The variants $(x : A) \multimap_s B$ and $A \multimap_s B$ represent linear functions.
+Terms of these types may contain linear free variables, thus cannot be applied repeated without duplication of linear variables.
+Similar to the non-linear versions, the difference between $(x : A) \multimap_s B$ and $A \multimap_s B$ lies in the allowance of dependency on function input for non-linear domain types.
+
+  \begin{figure}[H]
     \caption{Type Formation} 
@@
       \inferrule
+      { \Gamma\ \pure }
+      { \Gamma \vdash U \utype U } 
+      \rname{U-Axiom}
+
+      \inferrule
+      { \Gamma\ \pure }
+      { \Gamma \vdash L \utype U } 
+      \rname{L-Axiom}
+      \\
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype U \\ 
+        \Gamma, x \utype A \vdash B \utype s }
+      { \Gamma \vdash (x : A) \rightarrow_s B \utype U } 
+      \rname{U-Prod}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype L \\ 
+        \Gamma \vdash B \utype s }
+      { \Gamma \vdash A \rightarrow_s B \utype U } 
+      \rname{Arrow}
+      \\
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype U \\ 
+        \Gamma, x \utype A \vdash B\ \utype s }
+      { \Gamma \vdash (x : A) \multimap_s B \utype L } 
+      \rname{L-Prod}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype L \\ 
+        \Gamma \vdash B \utype s }
+      { \Gamma \vdash A \multimap_s B \utype L } 
+      \rname{Lolli}
+    \end{mathpar}
+    \label{type}
+  \end{figure}
+
+  \subsection{Term Formation}
+  In classical dependent type theories such as Martin L\"{o}f Type Theory or Calculus of Constructions, strong normalization and confluence ensure that term normalization is agnostic to reduction strategy.
+Furthermore, the preservation theorem shows that reduction preserves the typing judgment of well-typed terms.
+The management of substructural contexts for dependent linear types complicates the direct transplant of these theorems.
+In the application rules $\rname{U-App-1}$ and $\rname{L-App-1}$, the argument $n$ is a non-linear expression that could contain free variables.
+If a call-by-name reduction strategy is used, substitution of $n$ into the function body may cause duplication of these variables.
+Some existing literature ban non-linear arguments from containing linear free variables.
+We find this approach to be overly restrictive as it excludes the pattern of freely using retrieved data (non-linear) from references (linear) ubiquitous in imperative programs.
+Instead, we adjust the operational semantics to favor call-by-value reduction strategy.
+Our value soundness theorem guarantee non-linear values to contain only non-linear free variables.
+Thus, substitution will no longer duplicate linear variables found in argument expressions.
+
+  Furthermore, in the \rname{U-App-1} and \rname{L-App-1} rules, the input expression $n$ is not directly substituted into the dependent co-domain $B$ as this may introduce linear variables into types.
+Instead, a lambda abstract is formed around $B$ and applied to $n$.
+This delays $\beta$-reduction until $n$ can be resolved into value-form.
+
+  \begin{figure}[H]
+    \caption{Term Formation}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma\ \pure}
+      { \Gamma, x \utype A \vdash x \utype A } 
+      \rname{U-Var}
+
+      \inferrule
+      { \Gamma\ \pure }
+      { \Gamma, x \ltype A \vdash x \ltype A } 
+      \rname{L-Var}
+
+      \inferrule
+      { \Gamma \vdash m \stype{s} A \\ A \equiv B }
+      { \Gamma \vdash m \stype{s} B } 
+      \rname{Conv}
+      \\
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash (x : A) \rightarrow_s B \utype U \\ 
+        \Gamma, x \utype A \vdash n \stype{s} B }
+      { \Gamma \vdash \lambda x . n \utype (x : A) \rightarrow_s B }
+      \rname{U-$\lambda_1$}
+      \\
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \rightarrow_s B \utype U \\
+        \Gamma, x \ltype A \vdash n \stype{s} B }
+      { \Gamma \vdash \lambda x . n \utype A \rightarrow_s B }
+      \rname{U-$\lambda_2$}
+      \\
+
+      \inferrule
+      { \overline{\Gamma} \vdash (x : A) \multimap_s B \utype L \\ 
+        \Gamma, x \utype A \vdash n \stype{s} B }
+      { \Gamma \vdash \lambda x . n \ltype (x : A) \multimap_s B }
+      \rname{L-$\lambda_1$}
+      \\
+
+      \inferrule
+      { \overline{\Gamma} \vdash A \multimap_s B \utype L \\ 
+        \Gamma, x \ltype A \vdash n \stype{s} B }
+      { \Gamma \vdash \lambda x . n \ltype A \multimap_s B }
+      \rname{L-$\lambda_2$}
+      \\
+
+      \inferrule
+      { \Gamma_1 \vdash m \utype (x : A) \rightarrow_s B \\
+        \Gamma_2 \vdash n \utype A \\
+        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+      { \Gamma \vdash m\ n \stype{s} (\lambda x.B)\ n }
+      \rname{U-App-1}
+      \\
+
+      \inferrule
+      { \Gamma_1 \vdash m \utype A \rightarrow_s B \\
+        \Gamma_2 \vdash n \ltype A \\
+        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+      { \Gamma \vdash m\ n \stype{s} B }
+      \rname{U-App-2}
+      \\
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype (x : A) \multimap_s B \\
+        \Gamma_2 \vdash n \utype A \\
+        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+      { \Gamma \vdash m\ n \stype{s} (\lambda x.B)\ n }
+      \rname{L-App-1}
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype A \multimap_s B \\
+        \Gamma_2 \vdash n \ltype A \\
+        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+      { \Gamma \vdash m\ n \stype{s} B }
+      \rname{L-App-2}
+    \end{mathpar}
+    \label{term}
+  \end{figure}
+
+  \subsection{Reduction and Equality}
+  Figure \ref{single} presents the call-by-value operational semantics that we have eluded to. The rules defining the single step relation $\step$ are completely standard.
+
+  For full spectrum dependently typed languages, terms can appear at the type level.
+A definitional equality judgment is required to identify types beyond simple $\alpha$-equivalence.
+This is usually accomplished by normalization and comparing normal forms.
+Due to the complications brought on by linearity outlined in \ref{tyformation}, standard normalization techniques cannot be directly applied.
+Unfortunately, the single step relation $\step$ defined previously is not sufficient either as binders block evaluation.
+Following the footsteps of the Trellys project\cite{trellys}, we define a call-by-value parallel step relation $\pstep$ that evaluates under binders.
+We use the transitive reflexive closure of $\pstep$ to define definitional equality.
+
+  \begin{figure}[H]
+    \caption{Single Step Reduction}
+    \begin{mathpar}
+      \inferrule
       { }
-      { \Gamma ; \cdot \vdash \U : \U }
+      { (\lambda x.n)\ v \step [v/x]n }
+      \rname{S-U-$\beta$}
 
@@
       { }
-      { \Gamma ; \cdot \vdash \L : \U }
+      { (\lambda x.n)\ v \step [v/x]n }
+      \rname{S-L-$\beta$}
 
       \inferrule
-      { \Gamma ; \cdot \vdash A : \U }
-      { \Gamma ; \cdot \vdash A\ \type }
+      { m \step m' }
+      { m\ n \step m'\ n }
+      \rname{S-App-L}
 
       \inferrule
-      { \Gamma ; \cdot \vdash A : \L }
-      { \Gamma ; \cdot \vdash A\ \type }
-      \\
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A : \U \\ \Gamma, x : A ; \cdot \vdash B\ \type }
-      { \Gamma ; \cdot \vdash (x : A) \rightarrow B : \U }
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A : \L \\ \Gamma ; \cdot \vdash B\ \type }
-      { \Gamma ; \cdot \vdash A \rightarrow B : \U } 
-      \\
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A : \U \\ \Gamma, x : A ; \cdot \vdash B\ \type }
-      { \Gamma ; \cdot \vdash (x : A) \multimap B : \L }
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A : \L \\ \Gamma ; \cdot \vdash B\ \type }
-      { \Gamma ; \cdot \vdash A \multimap B : \L } 
+      { n \step n' }
+      { v\ n \step v\ n' }
+      \rname{S-App-R}
     \end{mathpar}
+    \label{single}
   \end{figure}
 
-  \begin{figure}[h]
-    \caption{Term Formation}
+  \begin{figure}[H]
+    \caption{Equality and Parallel Reduction}
     \begin{mathpar}
       \inferrule
-      { }
-      { \Gamma, x : A ; \cdot \vdash x : A }
+      { m_1 \pstep^* n \\ m_2 \pstep^* n }
+      { m_1 \equiv m_2 : A }
+      \rname{Join}
 
@@
       { }
-      { \Gamma ; x : A \vdash x : A }
+      { x \pstep x }
+      \rname{P-Var}
 
       \inferrule
-      { \Gamma ; \Delta \vdash m : A \\ \Gamma ; \Delta \vdash A \equiv B }
-      { \Gamma ; \Delta \vdash m : B } 
-      \\
+      { }
+      { U \pstep U }
+      \rname{P-U}
 
       \inferrule
-      { \Gamma ; \cdot \vdash (x : A) \rightarrow B : \U \\ 
-        \Gamma, x : A ; \cdot \vdash n : B }
-      { \Gamma ; \cdot \vdash \lambda x . n : (x : A) \rightarrow B }
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A \rightarrow B : \U \\
-        \Gamma ; x : A \vdash n : B }
-      { \Gamma ; \cdot \vdash \lambda x . n : A \rightarrow B }
-      \\
-
-      \inferrule
-      { \Gamma ; \cdot \vdash (x : A) \multimap B : \L \\ 
-        \Gamma, x : A ; \Delta \vdash n : B }
-      { \Gamma ; \Delta \vdash \lambda x . n : (x : A) \multimap B }
-
-      \inferrule
-      { \Gamma ; \cdot \vdash A \multimap B : \L \\ 
-        \Gamma ; \Delta, x : A \vdash n : B }
-      { \Gamma ; \Delta \vdash \lambda x . n : A \multimap B }
-      \\
-
-      \inferrule
-      { \Gamma ; \Delta_1 \vdash m : (x : A) \rightarrow B \\
-        \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
-
-      \inferrule
-      { \Gamma ; \Delta_1 \vdash m : A \rightarrow B \\
-        \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
-      \\
-
-      \inferrule
-      { \Gamma ; \Delta_1 \vdash m : (x : A) \multimap B \\
-        \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : [n/x]B }
-
-      \inferrule
-      { \Gamma ; \Delta_1 \vdash m : A \multimap B \\
-        \Gamma ; \Delta_2 \vdash n : A }
-      { \Gamma; \Delta_1, \Delta_2 : m\ n : B }
-    \end{mathpar}
-  \end{figure}
-
-  \begin{figure}[h]
-    \caption{Equality}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma ; \Delta \vdash m_1 : A \\ \Gamma ; \Delta \vdash m_2 : A \\ 
-        m_1 \pstep^* n \\ m_2 \pstep^* n }
-      { \Gamma ; \Delta \vdash m_1 \equiv m_2 }
-    \end{mathpar}
-  \end{figure}
-
-  \begin{figure}[h]
-    \caption{Parallel Reduction}
-    \begin{mathpar}
-      \inferrule
       { }
-      { n \pstep n }
+      { L \pstep L }
+      \rname{P-L}
 
@@
       { \lambda x.n \pstep \lambda x.n' }
+      \rname{P-$\lambda$}
+
+      \inferrule
+      { n \pstep n' }
+      { \lambda x.n \pstep \lambda x.n' }
+      \rname{P-$\lambda$}
 
@@
       { m\ n \pstep m'\ n' }
+      \rname{P-App}
 
@@
       { (\lambda x.n)\ v \pstep [v'/x]n' }
-      \\
+      \rname{P-U-$\beta$}
+
+      \inferrule
+      { n \pstep n' \\ v \pstep v' }
+      { (\lambda x.n)\ v \pstep [v'/x]n' }
+      \rname{P-L-$\beta$}
 
@@
       { A \pstep A' \\ B \pstep B' }
-      { (x : A) \rightarrow B \pstep (x : A') \rightarrow B' }
+      { (x : A) \rightarrow_s B \pstep (x : A') \rightarrow_s B' }
+      \rname{P-U-Prod}
 
@@
       { A \pstep A' \\ B \pstep B' }
-      { A \rightarrow B \pstep A' \rightarrow B' }
-      \\
+      { A \rightarrow_s B \pstep A' \rightarrow_s B' }
+      \rname{P-Arrow}
 
@@
       { A \pstep A' \\ B \pstep B' }
-      { (x : A) \multimap B \pstep (x : A') \multimap B' }
+      { (x : A) \multimap_s B \pstep (x : A') \multimap_s B' }
+      \rname{P-L-Prod}
 
@@
       { A \pstep A' \\ B \pstep B' }
-      { A \multimap B \pstep A' \multimap B' }
+      { A \multimap_s B \pstep A' \multimap_s B' }
+      \rname{P-Lolli}
     \end{mathpar}
+    \label{parallel}
   \end{figure}
 
+  \subsection{Meta Theory}
+  We have proven the type soundness of our language in the form of \textit{progress} and \textit{preservation} theorems.
+The proofs have been formalized in Coq with help from the Autosubst\cite{autosubst} library.
+
+  \subsubsection{Step and Parallel Step}
+
+  The following lemmas and proofs are entirely standard. The restriction to value-form arguments for $\beta$-reduction does not pose any complications.
+
+  \begin{lemma}
+    Single step reduction implies parallel step reduction. If $m \step m'$, then $m \pstep m'$.
+  \end{lemma}
+
+  \begin{lemma}
+    Parallel reduction satisfies the diamond property. If $m \pstep m_1$ and $m \pstep m_2$ then there exists $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+  \end{lemma}
+
+  \begin{corollary}
+    The transitive reflexive closure of parallel reduction is confluent. If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
+  \end{corollary}
+
+  \begin{corollary}
+    The definitional equality relation is an equivalence relation.
+  \end{corollary}
+
+  \subsubsection{Substitution}
+  Though the \textit{substitution} lemma is generally considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the \textit{substitution} lemma is admissible.
+Much of this difficulty arises during the substitution of non-linear expressions.
+Perhaps the most famous example detailing the issues of substitution is due to Wadler\cite{substitute}.
+Since computation arises as a consequence of substitution, it is paramount to get it right.
+
+  \begin{lemma}
+    Value soundness. If $\Gamma \vdash v \utype A$ then $\Gamma\ \pure$.
+  \end{lemma}
+
+  \begin{lemma}
+    Substitution.
+For $\Gamma_1, x \stype{s} A \vdash m \stype{t} B$ and $\Gamma_2 \vdash v \stype{s} A$, if there exists $\Gamma$ such that $merge\ \Gamma_1\ \Gamma_2\ \Gamma$ is defined, then $\Gamma \vdash [v/x]m \stype{t} [v/x]B$.
+  \end{lemma}
+
+  \begin{theorem}
+    Partial Weakening. If $\Gamma \vdash m \stype{s} A$ then $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+  \end{theorem}
+
+  \begin{theorem}
+    Progress. For $\epsilon \vdash m \stype{s} A$, either $m$ is a value or there exists $n$ such that $m \step n$.
+  \end{theorem}
+
+  \begin{theorem}
+    Preservation. For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
+  \end{theorem}
+
+  \subsection{Data Types}
+
+  \subsection{Future Work}
+  All of the theorems we have developed so far are purely syntactic in nature.
+But the natural correspondence between linear types and call-by-value semantics seem to hint at a deeper connection.
+Indeed, past works\cite{cbvsemantics} have provided insight for the simply typed case.
+We would like to extend these results our dependent linear type theory.
+
+  \subsection{Conclusion}
+
+\bibliographystyle{acm}
+\bibliography{ref}
+
 \end{document}

### 2021-09-22 writing/draft.tex
writing

@@
   \subsubsection{Substitution}
-  Though the \textit{substitution} lemma is generally considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the \textit{substitution} lemma is admissible.
+  Though the \textit{substitution} lemma is widely considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the \textit{substitution} lemma is admissible.
 Much of this difficulty arises during the substitution of non-linear expressions.
-Perhaps the most famous example detailing the issues of substitution is due to Wadler\cite{substitute}.
+Perhaps the most famous work detailing the issues of substitution is due to Wadler\cite{substitute}.
 Since computation arises as a consequence of substitution, it is paramount to get it right.
+
+  Generally, the application rule looks similar to the following for languages with linear types.
+  \begin{mathpar}
+    \inferrule
+    { \Gamma \vdash m : A \multimap B \\ 
+      \Delta \vdash n : A }
+    { \Gamma, \Delta \vdash m\ n : B } 
+  \end{mathpar}
+
+  If type $A$ is non-linear, then $n$ ought to be used freely.
+However, it is possible for $n$ to contain linear variables in $\Delta$.
+If $m$ is a lambda abstraction $\lambda x.m'$ where $x$ occurs multiple times within $m'$, substitution of $n$ for $x$ will cause duplication of linear variable.
+A common approach for solving this issue is to wrap non-linear values in an explicit modality, unpacking the internal value only when needed\cite{substitute,neel15}.
+Another is to outright ban non-linear expressions from containing linear variables\cite{llf,luo,qtt}.
+
+  Our call-by-value semantics neatly solves the substitution problem without imposing any of the modifications mentioned previously to the typing of application.
+The crucial realization is the following \textit{value soundness} lemma: non-linear \textit{values} contain no linear variables.
 
@@
   \end{lemma}
+
+  From the \textit{value soundness} lemma, the standard rule for application is admissible.
+Intuitively, a single copy of each linear resource is used to create a single non-linear value.
+This value can then be freely duplicated without needing the original resources to generate fresh copies.
 
@@
 
+  \subsubsection{Type Soundness}
+  The following theorems are a direct result of the \textit{substitution} lemma and various canonical-form lemmas.
+
   \begin{theorem}
-    Partial Weakening. If $\Gamma \vdash m \stype{s} A$ then $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+    Non-linear Weakening. If $\Gamma \vdash m \stype{s} A$ then $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
   \end{theorem}

### 2021-10-07 writing/draft0.tex
draft

--- 2021-09-22 writing/draft.tex
@@
 \documentclass{article}
-\usepackage[margin=1.4in]{geometry}
+\usepackage[margin=1.3in]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \newcommand{\type}{\textbf{type}}
+\newcommand{\refl}{\text{refl}}
+\newcommand{\letin}[3]{$\text{let }#1\text{ := }#2\text{ in }#3$}
+\newcommand{\new}{\text{new}}
+\newcommand{\free}{\text{free}}
+\newcommand{\get}{\text{get}}
+\newcommand{\set}{\text{set}}
+\newcommand{\subst}{\text{subst}}
 \newcommand{\utype}{\overset{U}{:}}
@@
 Objects in dependent type theory can be freely duplicated or deleted by appealing to the contraction and weakening rules, this goes against the conservation principle of resource.
-Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource\cite{bedrock}.
+Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource.
 As language designers, we hope to alleviate this shortcoming of dependent type theory.
@@
 
-  The lack of weakening and contraction rules for linear types present unique difficulties when integrating full spectrum dependent types, often making dependent linear types challenging or unsatisfactory for practical use.
+  The lack of weakening and contraction rules for linear types present unique difficulties when integrating dependent types, often making dependent linear types challenging for practical use.
 $\LNLD$ requires use of explicit modality maps F and G to alternate between its linear and non-linear fragment.
@@
 The unfortunate side-effect of this design choice is that linear computations never return non-linear values, eliminating the common pattern of retrieving a non-linear values from a linear references.
-QTT faces the similar drawbacks to Luo due to linearity being determined by semi-ring annotations.
+QTT faces similar drawbacks to Luo due to linearity being determined by semi-ring annotations.
 
-  We present a full spectrum dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators.
+  We present a dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators.
 We believe that linearity is an intrinsic property that can be determined by the type of a term.
-To accomplish this, our type system uses two class of type sorts in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
+To accomplish this, our type system uses two sorts of types in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
 Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow_s B$ for non-linear functions and $(x : A) \multimap_s B$ for linear functions.
@@
 To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for certified imperative programming.
-We augment our soundness theorem to account for these extensions, showing that imperative resources are properly managed.
 We have implemented a prototype of our language extended with other features for practicality and ergonomics.
@@
 
-  Our language firmly falls in the former category.
+  Our language falls firmly in the former category.
 Besides the reasons listed above, restricted dependency gives a clear distinction between linear and non-linear terms, a property we believe will aid in the compilation to high performance code.
@@
 Our type theory contains two sorts, $U$ and $L$ for denoting the sort of non-linear types and linear types respectively.
-An important fact to note is that our language possess the "Type-in-Type" axiom in the form of $\Gamma \vdash U \utype U$ which is well known to be logically unsound.
+An important fact to note is that our language possess the ``Type-in-Type" axiom in the form of $\Gamma \vdash U \utype U$ which is well known to be logically unsound.
 We chose this design deliberately as we are willing to sacrifice logical soundness for increased expressivity and convenience.
@@
 Instead, the linearity of functions is determined by the linearity of the closure it forms.
-For example, a function with linear input and output type can itself be non-linear if its free variables have non-linear type, indicating this function can be applied repeatedly without contraction of linear variables.
-In practice, these function types can be consolidated to just $(x : A) \rightarrow_s B$ and $(x : A) \multimap_s B$ by viewing $A \rightarrow_s B$ and $A \multimap_s B$ as their non-binding special cases.
+For example, a function with linear input and output type can itself be non-linear if its free variables have non-linear type, indicating this function can be applied repeatedly without duplication of linear variables.
+In practice, these function types can be consolidated to just $(x : A) \rightarrow_s B$ and $(x : A) \multimap_s B$ by treating $A \rightarrow_s B$ and $A \multimap_s B$ as non-binding special cases.
 For the sake of clarity, we leave them as four.
@@
   \subsection{Context and Structural Judgments}
-  The context of our language employs a mixed linear/non-linear representation introduced by Luo\cite{luo}.
+  The context of our language employs a mixed linear/non-linear representation in the style of Luo\cite{luo}.
 Variables in the context are annotated to indicate whether they are linear or non-linear.
@@
 
-  An auxiliary judgment $\pure$ is defined to assert that a context $\Gamma$ do not contain linear variables.
+  An auxiliary judgment $\pure$ is defined to assert that a context $\Gamma$ does not contain linear variables.
 In other words, all variables found in a pure context are of the form $\Gamma, x \utype A$.
-A restriction function $\overline{\Gamma}$ is defined that removes all linear variables from context $\Gamma$, the result of restriction is a pure subset of the original context.
+A restriction function $\overline{\Gamma}$ is defined that removes all linear variables from context $\Gamma$.
+The result of restriction is a pure subset of the original context.
 
@@
   \subsection{Type Formation} \label{tyformation}
-  Typing judgments in our language take on the form $\Gamma \vdash m \stype{s} A$, $s$ is an indexing sort that is either $U$ or $L$.
+  Typing judgments in our language take on the form $\Gamma \vdash m \stype{s} A$ where $s$ is an indexing sort that is either $U$ or $L$.
 Intuitively, this judgment states that expression $m$ has type $A$ and sort $s$ under context $\Gamma$.
@@
 Surprisingly, duplication of non-linear expression is not immediately allowed, only non-linear \textit{values} can be safely duplicated.
-We will discuss this subtlety later in more detail.
+We will discuss this subtlety later in Section \ref{subst}.
 
@@
   The function types $(x : A) \rightarrow_s B$ and $A \rightarrow_s B$ represent non-linear functions, where the subscript $s$ is the sort of co-domain $B$.
-Terms of these types contain no linear free variables, thus can be applied repeatedly without duplication of linear resources.
+Functions of these types contain no linear free variables, thus can be applied repeatedly without duplication of linear resources.
 The difference between the two is that $(x : A) \rightarrow_s B$ allows co-domain $B$ to depended on input $x$ of non-linear domain $A$.
@@
   The variants $(x : A) \multimap_s B$ and $A \multimap_s B$ represent linear functions.
-Terms of these types may contain linear free variables, thus cannot be applied repeated without duplication of linear variables.
+Functions of these types may contain linear free variables, thus cannot be applied repeated without duplication of linear variables.
 Similar to the non-linear versions, the difference between $(x : A) \multimap_s B$ and $A \multimap_s B$ lies in the allowance of dependency on function input for non-linear domain types.
@@
 
-  \subsection{Term Formation}
-  In classical dependent type theories such as Martin L\"{o}f Type Theory or Calculus of Constructions, strong normalization and confluence ensure that term normalization is agnostic to reduction strategy.
-Furthermore, the preservation theorem shows that reduction preserves the typing judgment of well-typed terms.
-The management of substructural contexts for dependent linear types complicates the direct transplant of these theorems.
-In the application rules $\rname{U-App-1}$ and $\rname{L-App-1}$, the argument $n$ is a non-linear expression that could contain free variables.
-If a call-by-name reduction strategy is used, substitution of $n$ into the function body may cause duplication of these variables.
-Some existing literature ban non-linear arguments from containing linear free variables.
-We find this approach to be overly restrictive as it excludes the pattern of freely using retrieved data (non-linear) from references (linear) ubiquitous in imperative programs.
-Instead, we adjust the operational semantics to favor call-by-value reduction strategy.
-Our value soundness theorem guarantee non-linear values to contain only non-linear free variables.
-Thus, substitution will no longer duplicate linear variables found in argument expressions.
+  \subsection{Term Formation} \label{teformation}
+  The term formations rules are presented in Figure \ref{term}.
+
+  The rules \rname{U-Var} and \rname{L-Var} state that a variable's type is determined by its context.
+In the case of \rname{L-Var}, the linear variable $x$ must be the only linear variable within its context.
+This enforcement of uniqueness eliminates the weakening rule for linear variables as redundant linear variables in the context will prevent the usage of \rname{L-Var}.
+
+  The \rname{Conv} rule is standard with regards to prior dependent type theory literature.
+Two types $A$ and $B$ are treated equivalently if they are $\beta\eta$-convertible.
+Our treatment of the convertibility relation $A \equiv B$ differs from standard due to our call-by-value semantics.
+We give a detailed account of this in the next section.
+
+  For the non-linear function formation rules \rname{U-$\lambda_1$} and \rname{U-$\lambda_2$}, the context $\Gamma$ is asserted to be pure.
+The purity of $\Gamma$ ensures that no linear variables occur within function body $n$, allowing the function to be freely used without duplication of linear variables.
+For the linear function formation rules \rname{L-$\lambda_1$} and \rname{L-$\lambda_2$}, the restriction on $\Gamma$'s purity is lifted.
+Linear variables within $\Gamma$ are allowed to occur freely within function body $n$.
+However, the tradeoff is that these linear functions may only be applied once, as multiple uses may result in duplication of its linear free variables.
+
+  In the application rules $\rname{U-App-1}$ and $\rname{L-App-1}$, the argument $n$ is a non-linear expression that may contain linear free variables.
+If reduction is performed naively, substitution of $n$ into the function body may cause duplication of these variables.
+Our operational semantics and value soundness lemma guarantee that substitution will not duplicate linear variables.
 
   Furthermore, in the \rname{U-App-1} and \rname{L-App-1} rules, the input expression $n$ is not directly substituted into the dependent co-domain $B$ as this may introduce linear variables into types.
-Instead, a lambda abstract is formed around $B$ and applied to $n$.
-This delays $\beta$-reduction until $n$ can be resolved into value-form.
+Instead, a $\lambda$-abstraction is formed around $B$ and applied to $n$.
+This delays $\beta$-reduction until $n$ can be evaluated into a value.
 
@@
 
-  \subsection{Reduction and Equality}
+  \subsection{Reduction and Equality} \label{reduction}
   Figure \ref{single} presents the call-by-value operational semantics that we have eluded to. The rules defining the single step relation $\step$ are completely standard.
 
-  For full spectrum dependently typed languages, terms can appear at the type level.
+  For dependently typed languages, terms can appear at the type level.
 A definitional equality judgment is required to identify types beyond simple $\alpha$-equivalence.
 This is usually accomplished by normalization and comparing normal forms.
-Due to the complications brought on by linearity outlined in \ref{tyformation}, standard normalization techniques cannot be directly applied.
+Due to the complications brought on by linearity and substitution hinted at in \ref{teformation}, standard normalization techniques cannot be directly applied.
 Unfortunately, the single step relation $\step$ defined previously is not sufficient either as binders block evaluation.
@@
 
-  \subsubsection{Substitution}
+  \subsubsection{Substitution} \label{subst}
   Though the \textit{substitution} lemma is widely considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the \textit{substitution} lemma is admissible.
-Much of this difficulty arises during the substitution of non-linear expressions.
+Much of this difficulty arise during the substitution of non-linear expressions.
 Perhaps the most famous work detailing the issues of substitution is due to Wadler\cite{substitute}.
-Since computation arises as a consequence of substitution, it is paramount to get it right.
+Since computation arise as a consequence of substitution, it is imperative to get it right.
 
@@
 
-  If type $A$ is non-linear, then $n$ ought to be used freely.
-However, it is possible for $n$ to contain linear variables in $\Delta$.
-If $m$ is a lambda abstraction $\lambda x.m'$ where $x$ occurs multiple times within $m'$, substitution of $n$ for $x$ will cause duplication of linear variable.
-A common approach for solving this issue is to wrap non-linear values in an explicit modality, unpacking the internal value only when needed\cite{substitute,neel15}.
-Another is to outright ban non-linear expressions from containing linear variables\cite{llf,luo,qtt}.
+  If type $A$ is a non-linear type, then $n$ ought to be used freely.
+However, it is possible for $n$ to contain linear variables present in $\Delta$.
+If $m$ is a lambda abstraction $\lambda x.m'$ where $x$ occurs multiple times within $m'$, substitution of $n$ for $x$ will cause duplication of linear variables.
+One approach for solving this issue is to wrap non-linear values in an explicit modality, unpacking the internal value only when needed\cite{substitute,neel15}.
+Another is to ban non-linear expressions from containing linear variables\cite{llf,luo,qtt}.
 
-  Our call-by-value semantics neatly solves the substitution problem without imposing any of the modifications mentioned previously to the typing of application.
+  Our call-by-value semantics resolves the substitution problem without imposing any of the modifications mentioned previously to the typing of application.
 The crucial realization is the following \textit{value soundness} lemma: non-linear \textit{values} contain no linear variables.
@@
 This value can then be freely duplicated without needing the original resources to generate fresh copies.
+This is consistent with the common practice of retrieving non-linear values from linear references.
 
@@
   \begin{theorem}
-    Non-linear Weakening. If $\Gamma \vdash m \stype{s} A$ then $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
-  \end{theorem}
-
-  \begin{theorem}
     Progress. For $\epsilon \vdash m \stype{s} A$, either $m$ is a value or there exists $n$ such that $m \step n$.
@@
 
+  \section{Extensions}
+
   \subsection{Data Types}
+  Though it is possible to encode data and propositions directly in the core language using Church-encodings, it is incredibly inconvenient.
+To address this, our implementation allows users to define inductive data types\cite{inductive} similar to Coq or Agda\cite{agda}.
+A pattern matching construct\cite{pattern} is defined on inductive types for data elimination.
+Checking is performed to ensure that non-linear inductive types do not carry linear terms and linear types cannot be used as type indices or parameters.
 
-  \subsection{Future Work}
+  Commonly used inductive types are formalized in Figures \ref{dataform} with their Introduction and Elimination rules formalized in Figure \ref{dataintro} and Figure \ref{dataelim} respectively.
+Reduction for data types obey the same call-by-value semantics as outlined in \ref{reduction}.
+Other standard data types that are not covered here are $\mathbb{N}$ for natural numbers and $\top$ for the unit type.
+
+  \begin{figure}[H]
+    \caption{Data Formation}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma\ \pure \\ 
+        \Gamma \vdash A \utype U \\ 
+        \Gamma, x \utype A \vdash B \utype U }
+      { \Gamma \vdash \Sigma x:A.B \utype U }
+      \rname{$\Sigma$}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype U \\ 
+        \Gamma, x \utype A \vdash B \utype L }
+      { \Gamma \vdash F x:A.B \utype L }
+      \rname{$F$}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype L \\ 
+        \Gamma \vdash B \utype L }
+      { \Gamma \vdash A \otimes B \utype L }
+      \rname{$\otimes$}
+
+      \inferrule
+      { \Gamma\ \pure \\ 
+        \Gamma \vdash m \utype A \\
+        \Gamma \vdash n \utype A }
+      { \Gamma \vdash m =_A n \utype U }
+    \end{mathpar}
+    \label{dataform}
+  \end{figure}
+
+  \begin{figure}[H]
+    \caption{Data Introduction}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma_1 \vdash m \utype A \\
+        \Gamma_2 \vdash n \utype (\lambda x.B)\ m \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash (m, n) \utype \Sigma x:A.B }
+      \rname{$\Sigma$-intro}
+
+      \inferrule
+      { \Gamma_1 \vdash m \utype A \\
+        \Gamma_2 \vdash n \ltype (\lambda x.B)\ m \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash [m, n] \ltype F x:A.B }
+      \rname{$F$-intro}
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype A \\
+        \Gamma_2 \vdash n \ltype B \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash \langle m, n \rangle \ltype A \otimes B }
+      \rname{$\otimes$-intro}
+
+      \inferrule
+      { \Gamma\ \pure \\ \Gamma \vdash n \utype A }
+      { \Gamma \vdash \refl\ n \utype n =_A n }
+    \end{mathpar}
+    \label{dataintro}
+  \end{figure}
+
+  \begin{figure}[H]
+    \caption{Data Elimination}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma_1 \vdash m \utype \Sigma x.A.B \\
+        \Gamma_2, x \utype A, y \utype B \vdash n \stype{s} C \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash \letin{(x , y)}{m}{n} \stype{s} C }
+      \rname{$\Sigma$-elim}
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype F x.A.B \\
+        \Gamma_2, x \utype A, y \ltype B \vdash n \stype{s} C \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash \letin{[x , y]}{m}{n} \stype{s} C }
+      \rname{$F$-elim}
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype A \otimes B \\
+        \Gamma_2, x \ltype A, y \ltype B \vdash n \stype{s} C \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash \letin{$\langle x , y\rangle$}{m}{n} \stype{s} C }
+      \rname{$\otimes$-elim}
+
+      \inferrule
+      { \Gamma_1 \vdash p \utype m =_A n \\
+        \Gamma_2 \vdash q \stype{s} B[m] \\
+        \overline{\Gamma}, x \utype A \vdash B[x] \utype s \\
+        merge\ \Gamma_1 \Gamma_2 \Gamma }
+      { \Gamma \vdash \subst(\lambda x.B, p, q) \stype{s} B[n] }
+    \end{mathpar}
+    \label{dataelim}
+  \end{figure}
+
+  \subsection{Imperative Programming}
+  With the addition of data types, axioms for imperative programming can be added to the language.
+In Figure \ref{state} we give a set of axioms for stateful programs.
+Note that these axioms are not associated with any reductions.
+This unsurprisingly breaks the \textit{Progress Theorem} as axioms are neither values nor can they reduce.
+Despite this obvious shortcoming, the axiomatic treatment of state provides an interface for extraction of imperative code with safe manual memory management.
+
+  \subsubsection{State Programs}
+  \begin{figure}[H]
+    \caption{State Programs}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash A \utype U \\ 
+        \Gamma \vdash m \utype A \\
+        \Gamma \vdash l \utype \mathbb{N} }
+      { \Gamma \vdash l \mapsto_A m \utype L }
+      \rname{Capability}
+
+      \inferrule
+      { \Gamma\ \pure \\ 
+        \Gamma \vdash A \utype U \\ 
+        \Gamma \vdash m \utype A }
+      { \Gamma \vdash \new(m) \ltype F l:\mathbb{N}. l \mapsto_A m }
+      \rname{New}
+
+      \inferrule
+      { \Gamma\ \pure \\
+        \Gamma \vdash c \ltype l \mapsto_A m }
+      { \Gamma \vdash \free(c) \utype \top }
+      \rname{Free}
+
+      \inferrule
+      { \Gamma \vdash c \ltype l \mapsto_A m }
+      { \Gamma \vdash \get(l, c) \ltype
+        F x:A. F e: (x =_A m). l \mapsto_A m }
+      \rname{Get}
+
+      \inferrule
+      { \Gamma \vdash c \ltype l \mapsto_A m \\
+        \overline{\Gamma} \vdash n \utype A }
+      { \Gamma \vdash \set(l, c, n) \ltype l \mapsto_A n }
+      \rname{Set}
+    \end{mathpar}
+    \label{state}
+  \end{figure}
+
+  \subsubsection{Laws for Free}
+
+  \section{Future Work}
   All of the theorems we have developed so far are purely syntactic in nature.
@@
 
-  \subsection{Conclusion}
+  \section{Conclusion}
 

### 2021-10-07 writing/draft1.tex
draft

--- 2021-10-07 writing/draft0.tex
@@
 % title and author
-\title{Dependent Linear Type Theory for Resource Aware Certified Programming}
+\title{The Calculus of Linear Constructions}
 \author{Qiancheng Fu}
@@
   \begin{abstract}
-    TBD
+    The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
+Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
+We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource, with proofs formalized in the Coq Proof Assistant.
+We extend CLC with linear inductive types and show that it is straightforward to use CLC as a programming language for the safe manipulation of mutable data structures, with applications to verified functional programming in GC-less environments.
   \end{abstract}
@@
   \section{Introduction}
-  In the field of certified programming, it is of paramount importance for the specification language to precisely characterize the behavior of programs.
-For this reason, dependent type theory enjoys great success as the type level language of specifications is exactly the same as the term level programming language, giving tremendous power and control to specification writing.
-The verification achievements\cite{compcert,deepspec,everest} carried out by proof assistants such as Coq\cite{coq} or F*\cite{fstar} is a testament how much dependent type theory has accomplished.
+  The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
+In CC types can depend on terms, allowing one to write fine-grain propositions as types.
+Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean}, and others.
+These theorem provers have found great success in the fields of software verification \cite{compcert,bedrock}, and constructive mathematics \cite{four-color,schemes}.
+However, due to its origins as a logical framework for constructive mathematics, it is quite difficult for CC to encode and reason about resources.
+Intuitively, a mathematical theorem can be applied an unrestricted number of times.
+On the other hand, resource must respect the conservation principle: resources are constant through change.
+Users of proof assistants based on CC often embed external logics into CC to provide additional reasoning principles to prove theorems concerning resources.
+The embedding of these logics is a difficult problem in its own right, requiring additional proofs to justify its soundness.
+We propose an alternative solution: extend CC with linear types.
 
-  Dependent type theory is however, not without its flaws.
-Due to its origins as a logical foundation for mathematics\cite{martinlof,cc}, dependent type theory lacks facilities to reason about resource, a concept that is largely absent from mathematics but ever present in computer science.
-Objects in dependent type theory can be freely duplicated or deleted by appealing to the contraction and weakening rules, this goes against the conservation principle of resource.
-Users of dependent type theories are often required to laboriously embed memory models and program logics into type theory to effectively reason about resource.
-As language designers, we hope to alleviate this shortcoming of dependent type theory.
+  Linear Logic is a substructural logic introduced by Girard in his seminal work \cite{girard}.
+Girard notice that the Weakening and Contraction rules of Classical Logic when restricted carefully, gives rise to a new logical foundation for reasoning about resource.
+Wadler \cite{wadler1990,wadler1991} first notice that an analogous restriction to variable usage in simple type theory leads to a linear type theory, where terms respect resources.
+A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
+Benton \cite{benton1994} investigates the ramifications of the !
+exponential in linear term calculi, decomposing it to adjoint connectives F and G that map between linear and non-linear judgments.
+Programming languages \cite{l3,ats,linear-haskell} featuring linear types have been implemented in practice, allowing programmers to write memory safe software.
+The success of integrating Linear Logic with simple type theory exposes a tantalizing new frontier of integrating linearity with a richer type theory.
 
-  In the disparate world of substructural logic, Girard introduces Linear Logic in his seminal work\cite{girard}.
-Linear Logic restricts weakening and contraction rules of classical logic, giving rise to an elegant formal foundation for reasoning about resource.
-Wadler\cite{wadler1990,wadler1991} and Abramsky\cite{abramsky1993} notice the opportunities presented by Linear Logic for resource aware programming, pioneering the development of simple linear type theory as a programming language.
-Linear type theory enforces safe variable usage, which conservatively subsumes resource usage.
-This reduces the problem of resource reasoning to linear type checking.
-Languages based on linear type systems have shown the capacity to safely manipulate imperative data structures\cite{l3,ats}.
-The successful extraction of type theory from Linear Logic offers a tantalizing hint to overcoming the challenges of resource reasoning in dependent type theory: integrate linear types.
-
-  The earliest work on integrating dependent type theory and linear type theory was carried out by Cervesato and Pfenning\cite{llf}, extending the Edinburgh LF with linear types.
-Xi extends DML dependent types with linear types in the ATS programming language\cite{ats}.
+  Cervesato and Pfenning extends the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r gives a categorical semantics for linear dependent types.
 Krishnaswami et al.
-develop $\LNLD$\cite{neel15} based on Benton's LNL\cite{benton1994} calculus for simple linear types, allowing term dependency on its non-linear fragment.
-The authors of $\LNLD$ demonstrate its capacity for certified imperative programming in the style of L3\cite{l3} and Hoare Type Theory\cite{htt}.
+present a dependent linear type theory \cite{neel15} based on Benton's early work of mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming the style of Hoare Type Theory \cite{htt}.
 Luo et al.
-introduce the notion of essential linearity\cite{luo}, developing the first type theory that allows types to depend on linear terms.
-Atkey's QTT\cite{qtt}, based on initial ideas of McBride\cite{nothing}, neatly reduces dependency and linearity checking to checking constraints defined on semi-rings.
-Idris 2\cite{idris2} is a full featured programming language that implements QTT as its core type system.
-
-  The lack of weakening and contraction rules for linear types present unique difficulties when integrating dependent types, often making dependent linear types challenging for practical use.
-$\LNLD$ requires use of explicit modality maps F and G to alternate between its linear and non-linear fragment.
-Computation requiring back and forth applications of F and G quickly explode in complexity.
-Terms in Luo's type theory are not intrinsically linear, linearity is solely determined by annotations.
-The unfortunate side-effect of this design choice is that linear computations never return non-linear values, eliminating the common pattern of retrieving a non-linear values from a linear references.
-QTT faces similar drawbacks to Luo due to linearity being determined by semi-ring annotations.
-
-  We present a dependent linear type theory that provides a more direct approach to enforce linearity, eschewing the use of modality operators.
-We believe that linearity is an intrinsic property that can be determined by the type of a term.
-To accomplish this, our type system uses two sorts of types in the style of Krishnaswami, $L$ for linear and $U$ for non-linear.
-Due to the abandonment of modalities, we use two distinct functions types $(x : A) \rightarrow_s B$ for non-linear functions and $(x : A) \multimap_s B$ for linear functions.
-We formalize a call-by-value operational semantics and prove the type soundness of our system.
-To further demonstrate the applications of dependent linear types, we extend our type system with imperative programming features, leveraging both dependency and linearity for certified imperative programming.
-We have implemented a prototype of our language extended with other features for practicality and ergonomics.
+introduce the property of essential linearity, and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+Based on initial ideas of McBride \cite{nothing}, Atkey \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear, and affine types within a unified framework.
 

### 2021-10-08 writing/draft1.tex
draft

@@
     The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource, with proofs formalized in the Coq Proof Assistant.
-We extend CLC with linear inductive types and show that it is straightforward to use CLC as a programming language for the safe manipulation of mutable data structures, with applications to verified functional programming in GC-less environments.
+Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the introduction and elimination of its term level inhabitants.
+We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource.
+CLC is highly backwards compatible with CC, allowing CLC to enjoy the decades of CC research.
+We have formalized all major results in the Coq Proof Assistant.
+We extend CLC with linear inductive types and show that CLC as a programming language enables the manipulation of mutable data structures in a principled way.
   \end{abstract}
@@
 Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean}, and others.
-These theorem provers have found great success in the fields of software verification \cite{compcert,bedrock}, and constructive mathematics \cite{four-color,schemes}.
+These theorem provers have found great success in the fields of software verification \cite{compcert,deepspec}, and constructive mathematics \cite{four-color,schemes}.
+
   However, due to its origins as a logical framework for constructive mathematics, it is quite difficult for CC to encode and reason about resources.
 Intuitively, a mathematical theorem can be applied an unrestricted number of times.
-On the other hand, resource must respect the conservation principle: resources are constant through change.
-Users of proof assistants based on CC often embed external logics into CC to provide additional reasoning principles to prove theorems concerning resources.
-The embedding of these logics is a difficult problem in its own right, requiring additional proofs to justify its soundness.
+Comparatively, the usage of resources is more limited.
+For example, if we encode Girard's classical example \cite{girard95} of purchasing cigarettes literally into CC as a function of type:
+
+  \begin{equation*}
+    Money \rightarrow Camels \times Marlboro
+  \end{equation*}
+
+  \noindent The customer will still maintain full ownership of their money after paying the vendor.
+Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
+Users of proof assistants based on CC often need to embed external logics into CC \cite{bedrock} to provide additional reasoning principles for dealing with resource.
+The design and embedding of these logics is a difficult problem in its own right, requiring additional proofs to justify its soundness.
 We propose an alternative solution: extend CC with linear types.
@@
 Benton \cite{benton1994} investigates the ramifications of the !
-exponential in linear term calculi, decomposing it to adjoint connectives F and G that map between linear and non-linear judgments.
-Programming languages \cite{l3,ats,linear-haskell} featuring linear types have been implemented in practice, allowing programmers to write memory safe software.
-The success of integrating Linear Logic with simple type theory exposes a tantalizing new frontier of integrating linearity with a richer type theory.
+exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
+Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
+The success of integrating Linear Logic with simple type theory exposes a tantalizing new frontier of integrating linearity with richer type theories.
 
+  Work have been done to extend dependent type theories with linear types.
 Cervesato and Pfenning extends the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r gives a categorical semantics for linear dependent types.
+V\'{a}k\'{a}r \cite{vakar14} gives a categorical semantics for linear dependent types.
 Krishnaswami et al.
@@
 introduce the property of essential linearity, and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear, and affine types within a unified framework.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear, and affine types within a unified framework.
+The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
+
+  We propose a new linear dependent type system - The Calculus of Linear Constructions (CLC).
+Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``minimally invasive'' approach to extending CC with linear types.
+This allows for a straightforward embedding of CLC into CC, and lifting of CC into CLC, endowing CLC with the fruits of decades of CC research.
+We accomplish this by adding extra universes $L$ of linear types with cumulativity parallel to the universes $U$ of non-linear types.
+Universe information is propagated by an indexed typing judgment down to the term level, controlling the usage of Weakening and Contraction rules.
+We further extend CLC with inductive types, showing that as a programming language it can manipulate mutable data structures in a principled way.
+We have formalized all major results in Coq, and implemented a prototype in OCaml.
+
+  \medskip
+
+  \noindent \textbf{\textit{Contributions}}: 
+  Our contributions can be summarized as follows.
+  \begin{itemize}
+    \item Fist, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types.
+The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
+    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, progress, regularity, and preservation. 
+    \item We observe that CLC is highly backwards compatible with CC. We construct a reduction preserving embedding of CLC into CC, showing that CLC is consistent. 
+    \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
+To the best of our knowledge, our development is the first machine checked formalization of a linear dependent type theory.
+    \item Furthermore, we extend CLC with linear inductive data, demonstrating that as a programming language, CLC can safely manipulate mutable data structures.
+    \item Finally, we give a implementation extended with user definable linear and non-linear data types.
+Algorithmic type checking employed by the implementation streamlines the process of writing CLC.
+  \end{itemize}
 

### 2021-10-09 writing/draft1.tex
coq

@@
     The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the introduction and elimination of its term level inhabitants.
+Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
 We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource.
 CLC is highly backwards compatible with CC, allowing CLC to enjoy the decades of CC research.
-We have formalized all major results in the Coq Proof Assistant.
+We have formalized and proven correct all major results of the core calculus in the Coq Proof Assistant.
 We extend CLC with linear inductive types and show that CLC as a programming language enables the manipulation of mutable data structures in a principled way.
@@
 For example, if we encode Girard's classical example \cite{girard95} of purchasing cigarettes literally into CC as a function of type:
-
   \begin{equation*}
-    Money \rightarrow Camels \times Marlboro
+    Money \rightarrow Camels + Marlboro
   \end{equation*}
-
-  \noindent The customer will still maintain full ownership of their money after paying the vendor.
+  The customer will still maintain full ownership of their money after paying the vendor.
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
@@
   Linear Logic is a substructural logic introduced by Girard in his seminal work \cite{girard}.
-Girard notice that the Weakening and Contraction rules of Classical Logic when restricted carefully, gives rise to a new logical foundation for reasoning about resource.
+Girard notice that the weakening and contraction rules of Classical Logic when restricted carefully, gives rise to a new logical foundation for reasoning about resource.
 Wadler \cite{wadler1990,wadler1991} first notice that an analogous restriction to variable usage in simple type theory leads to a linear type theory, where terms respect resources.
@@
   We propose a new linear dependent type system - The Calculus of Linear Constructions (CLC).
-Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``minimally invasive'' approach to extending CC with linear types.
+CLC extends CC$\omega$ with linear types.
+CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
+We add extra universes $L$ of linear types with cumulativity parallel to the universes $U$ of non-linear types.
+Universe information is propagated by an indexed typing judgment down to the term level, controlling the usage of weakening and contraction rules.
+This ultimately results in the \textit{linearity} theorem, stating that all resources are used exactly once.
+
+  The presence of both linear and dependent types enables CLC to write specifications that faithfully encodes the usage of resource.
+The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$, and a linear arrow $\multimap_L$ as follows.
+  \begin{equation*}
+    Money\ 5 \multimap_L Camels + Marlboro
+  \end{equation*}
+  This new specification for transaction states that it requires a payment of 5 units of money, the customer is relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
+
+  Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``minimally invasive'' approach to extending CC with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
 This allows for a straightforward embedding of CLC into CC, and lifting of CC into CLC, endowing CLC with the fruits of decades of CC research.
-We accomplish this by adding extra universes $L$ of linear types with cumulativity parallel to the universes $U$ of non-linear types.
-Universe information is propagated by an indexed typing judgment down to the term level, controlling the usage of Weakening and Contraction rules.
 We further extend CLC with inductive types, showing that as a programming language it can manipulate mutable data structures in a principled way.

### 2021-10-09 writing/draft1.tex
draft

@@
 \usepackage{float}
+\usepackage{tikz-cd}
 
-\floatstyle{boxed} 
+\floatstyle{plain} 
 \restylefloat{figure}
@@
 \newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
 
@@
 \newcommand{\ok}{\textbf{ok}}
-\newcommand{\pure}{\textbf{pure}}
+\newcommand{\pure}[1]{|#1|}
 \newcommand{\sort}{\textbf{sort}}
@@
 \newcommand{\subst}{\text{subst}}
-\newcommand{\utype}{\overset{U}{:}}
-\newcommand{\ltype}{\overset{L}{:}}
+\newcommand{\utype}{:_U}
+\newcommand{\ltype}{:_L}
 \newcommand{\stype}[1]{\overset{#1}{:}}
@@
 \newcommand{\pstep}{\leadsto_p}
+\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
 
@@
 We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource.
-CLC is highly backwards compatible with CC, allowing CLC to enjoy the decades of CC research.
+CLC is backwards compatible with CC, allowing CLC to enjoy the decades of CC research.
 We have formalized and proven correct all major results of the core calculus in the Coq Proof Assistant.
@@
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, progress, regularity, and preservation. 
+    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, regularity, and subject reduction. 
     \item We observe that CLC is highly backwards compatible with CC. We construct a reduction preserving embedding of CLC into CC, showing that CLC is consistent. 
@@
     \item Furthermore, we extend CLC with linear inductive data, demonstrating that as a programming language, CLC can safely manipulate mutable data structures.
-    \item Finally, we give a implementation extended with user definable linear and non-linear data types.
+    \item Finally, we give an implementation extended with user definable linear and non-linear inductive types.
 Algorithmic type checking employed by the implementation streamlines the process of writing CLC.
@@
 
-  \section{Core Type Theory}
-  In the efforts of integrating dependency and linearity, there are two predominant schools of thought: \textit{Types can only depended on non-linear terms} and \textit{Types can depend on all terms}.
-
-  Advocates of restricted dependency argue that the restriction is both intuitive and a necessity.
-A natural example of linear types depending on a non-linear terms is that of the length indexed random access array.
-Term dependency on array length statically prevents out-of-bounds array access whilst linearity ensures proper memory management.
-In the converse situation of types depending on linear terms, one immediately faces a dilemma: \textit{Do linear terms at the type level possess weakening and contraction?}.
-
-  If one answers yes, then the linear terms at the type level are defacto non-linear.
-In this case, a language with restricted dependency can simulate linear dependency by depending on non-linear terms that reflect the shape of linear terms\cite{ldqt}, erasing the advantage of increased type level expressivity.
-All prior works featuring linear dependency fall into this category, for good reason: \textit{the alternative is worse}.
-
-  If type level linear terms do not possess weakening and contraction, types with linear dependencies themselves are linear.
-This breaks the very notion of the typing judgment as even $\alpha$-equivalent terms are not allowed to possess the same type.
-Though the idea of such a type system is deeply interesting, the philosophical burden is prohibitive.
-
-  Our language falls firmly in the former category.
-Besides the reasons listed above, restricted dependency gives a clear distinction between linear and non-linear terms, a property we believe will aid in the compilation to high performance code.
+  \section{The Language of CLC}
 
@@
   The syntax of the core type theory is presented in Figure \ref{syntax}.
-Our type theory contains two sorts, $U$ and $L$ for denoting the sort of non-linear types and linear types respectively.
-An important fact to note is that our language possess the ``Type-in-Type" axiom in the form of $\Gamma \vdash U \utype U$ which is well known to be logically unsound.
-We chose this design deliberately as we are willing to sacrifice logical soundness for increased expressivity and convenience.
-We forsee no issue in extending the core language with a hierarchy of sorts to enforce logical soundness.
+Our type theory contains two sorts of predicative universes $U$ and $L$, being the types of non-linear types and linear types respectively.
+An additional impredicative universe $Prop$ is the type of propositions in the style of CC.
 
@@
 
-      <$m, n, A, B, C$> ::= $U$ | $L$ | $x$  \hspace*{8em} expressions
+      <$m, n, A, B, C$> ::= $U_i$ | $L_i$ | $Prop$ | $x$  \hspace*{4.3em} expressions
       \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
@@
       \indalt $\lambda x. n$ | $m\ n$
-
-      <$v$> ::= $U$ | $L$ | $x$ \hspace*{8em} values
-      \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
-      \indalt $(x : A) \multimap_s B$ | $A \multimap_s B$
-      \indalt $\lambda x. n$
     \end{grammar}
@@
 
-  A clear departure of our language from standard presentations of type theory is the presence of four function types: $(x : A) \rightarrow_s B$, $A \rightarrow_s B$, $(x : A) \multimap_s B$, $A \multimap_s B$.
-The reason for these variants is the decoupling of function linearity from the linearity of function domain and co-domain.
-Instead, the linearity of functions is determined by the linearity of the closure it forms.
-For example, a function with linear input and output type can itself be non-linear if its free variables have non-linear type, indicating this function can be applied repeatedly without duplication of linear variables.
-In practice, these function types can be consolidated to just $(x : A) \rightarrow_s B$ and $(x : A) \multimap_s B$ by treating $A \rightarrow_s B$ and $A \multimap_s B$ as non-binding special cases.
-For the sake of clarity, we leave them as four.
+  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the presence of four function types: $(x : A) \rightarrow_s B$, $A \rightarrow_s B$, $(x : A) \multimap_s B$, $A \multimap_s B$.
+The reason for these variants is that we have built the !
+exponential of linear logic directly into the syntax of function types.
+The behavior of !
+is difficult to account for even in simple linear type theory.
+Subtle issues arise if !!
+is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
+By integrating the exponential directly into function syntax, we give canonicity to ways that !
+can be used.
+This allows us to derive substitution, and a direct embedding into CC$\omega$ without needing extra machinery for manipulating exponential.
+  \begin{figure}[H]
+    \caption{Correspondence of CLC types and MELL implications}
+    \begin{align}
+      (\_ : A) \rightarrow_U B \quad &\equiv \quad !(!A \multimap !B) \\
+      (\_ : A) \rightarrow_L B \quad &\equiv \quad !(!A \multimap B) \\
+      A \rightarrow_U B \quad &\equiv \quad !(A \multimap !B) \\
+      A \rightarrow_L B \quad &\equiv \quad !(A \multimap B) \\
+      (\_ : A) \multimap_U B \quad &\equiv \quad !A \multimap !B \\
+      (\_ : A) \multimap_L B \quad &\equiv \quad !A \multimap B \\
+      A \multimap_U B \quad &\equiv \quad A \multimap !B \\
+      A \multimap_L B \quad &\equiv \quad A \multimap B
+    \end{align}
+    \label{correspondence}
+  \end{figure}
+  Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
+MELL lacks counterparts for the cases (1), (2), (5), (6) if the co-domain $B$ is dependent on arguments of the domain $A$.
+
+  \subsection{Universes and Cumulativity}
+  CLC features three sorts of universes, $Prop$, $U$, and $L$.
+$Prop$ is the impredicative universe propositions.
+$U$, and $L$ are the predicative universe of non-linear types, and linear types respectively.
+The main mechanism that CLC uses to distinguish between linear and non-linear types is by the universe which they belong.
+Intuitively, terms with types that occur within $Prop$, and $U$ are unrestricted in their usage.
+Terms with types that occur within $L$ are restricted to being used exactly once.
+
+  In order to lift terms from lower universes to higher ones, there exists cumulativity between universe levels of the same sort. We define cumulativity as follows.
+
+  \begin{definition}
+    The cumulativity relation ($\preceq$) is the smallest binary relation over terms such that
+    \begin{enumerate}
+      \item $\preceq$ is a partial order with respect to definitional equality.
+        \begin{enumerate}
+          \item If $A \equiv B$, then $A \preceq B$.
+          \item If $A \preceq B$ and $B \preceq A$, then $A \equiv B$.
+          \item If $A \preceq B$ and $B \preceq C$, then $A \preceq B$.
+        \end{enumerate}
+      \item $Prop \preceq U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
+      \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
+      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
+        \begin{enumerate}
+          \item $(x : A_1) \rightarrow B_1 \preceq (x : A_2) \rightarrow B_2$
+          \item $(x : A_1) \multimap B_1 \preceq (x : A_2) \multimap B_2$
+          \item $A_1 \rightarrow B_1 \preceq A_2 \rightarrow B_2$
+          \item $A_1 \multimap B_1 \preceq A_2 \multimap B_2$
+        \end{enumerate}
+    \end{enumerate}
+  \end{definition}
+
+  Figure \ref{universe} illustrates the structure of our universe hierarchy.
+Each linear universes $L_i$ has $U_{i+1}$ as its type, allowing functions to dependently quantify over linear \textit{types}.
+However, $L_i$ cumulates to $L_{i+1}$.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of linearity.
+
+  \begin{figure}[H]
+    \caption{The Universe Hierarchy}
+    \centering
+    \begin{tikzcd}
+      Prop \arrow[r, ":"', "\preceq", dash] 
+      & U_0 \arrow[r, ":"', "\preceq", dash] 
+      & U_1 \arrow[r, ":"', "\preceq", dash] 
+      & U_2 \arrow[r, ":"', "\preceq", dash] 
+      & \cdots \\
+      & L_0 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash] 
+      & L_1 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash] 
+      & L_2 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash] 
+      & \cdots
+    \end{tikzcd}
+    \label{universe}
+  \end{figure}
 
@@
 
-  We define a $merge$ relation to combine two mixed contexts $\Gamma_1$ and $\Gamma_2$ by performing contraction on shared non-linear variables.
-For linear variables, the $merge$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
-The definition of $merge$ allows contraction for non-linear variables whilst forbidding it for linear ones.
+  Next, we define a $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ relation that merges two mixed contexts $\Gamma_1$, and $\Gamma_2$ into $\Gamma$, by performing contraction on shared non-linear variables.
+For linear variables, the $\mrg{\_}{\_}{\_}$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
+This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for non-linear variables whilst forbidding it for linear ones.
 
-  An auxiliary judgment $\pure$ is defined to assert that a context $\Gamma$ does not contain linear variables.
-In other words, all variables found in a pure context are of the form $\Gamma, x \utype A$.
-A restriction function $\overline{\Gamma}$ is defined that removes all linear variables from context $\Gamma$.
-The result of restriction is a pure subset of the original context.
+  An auxiliary judgment $\pure{\Gamma}$ is defined to assert that a context $\Gamma$ does not contain linear variables.
+In other words, all variables found in $\pure{\Gamma}$ are annotated of the form $x \utype A$.
+The full rules for structural judgments are presented in Figure \ref{structural}.
 
@@
       { }
-      { \epsilon \ \ok }
-      \rname{Ok-$\epsilon$}
+      { \epsilon \vdash }
+      \rname{Wf-$\epsilon$}
 
       \inferrule
-      { \Gamma\ \ok \\ 
+      { \Gamma\ \vdash \\ 
         \overline{\Gamma} \vdash A \utype U }
-      { \Gamma, x \utype A\ \ok }
-      \rname{Ok-U}
+      { \Gamma, x \utype A \vdash }
+      \rname{Wf-U}
 
       \inferrule
-      { \Gamma\ \ok \\ 
+      { \Gamma\ \vdash \\ 
         \overline{\Gamma} \vdash A \utype L }
-      { \Gamma, x \ltype A\ \ok } 
-      \rname{Ok-L}
+      { \Gamma, x \ltype A\ \vdash } 
+      \rname{Wf-L}
       \\
@@
       { }
-      { \epsilon\ \pure }
+      { \pure{\epsilon} }
       \rname{Pure-$\epsilon$}
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \pure{\Gamma} \\
         \Gamma \vdash A \utype U }
-      { \Gamma, x \utype A\ \pure }
+      { \pure{\Gamma, x \utype A} }
       \rname{Pure-U}
@@
       { }
-      { merge\ \epsilon\ \epsilon\ \epsilon }
+      { \mrg{\epsilon}{\epsilon}{\epsilon} }
       \rname{Merge-$\epsilon$}
@@
       \inferrule
-      { merge\ \Gamma_1\ \Gamma_2\ \Gamma }
-      { merge\ (\Gamma_1, x \utype A)
-             \ (\Gamma_2, x \utype A)
-             \ (\Gamma, x \utype A) }
+      { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \mrg{\Gamma_1, x \utype A}
+            {\Gamma_2, x \utype A}
+            {\Gamma, x \utype A} }
       \rname{Merge-U}
@@
       \inferrule
-      { merge\ \Gamma_1\ \Gamma_2\ \Gamma \\
+      { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
         x \notin \Gamma_2 }
-      { merge\ (\Gamma_1, x \ltype A)
-             \ \Gamma_2
-             \ (\Gamma, x \ltype A) }
+      { \mrg{\Gamma_1, x \ltype A}
+            {\Gamma_2}
+            {\Gamma, x \ltype A} }
       \rname{Merge-L1}
@@
       \inferrule
-      { merge\ \Gamma_1\ \Gamma_2\ \Gamma \\
+      { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
         x \notin \Gamma_1 }
-      { merge\ \Gamma_1
-             \ (\Gamma_2, x \ltype A)
-             \ (\Gamma, x \ltype A) }
+      { \mrg{\Gamma_1}
+            {\Gamma_2, x \ltype A}
+            {\Gamma, x \ltype A} }
       \rname{Merge-L2} 
     \end{mathpar}
+    \label{structural}
+  \end{figure}
+
+  \begin{definition}
+    The context restriction function $\overline{\Gamma}$ is defined as a recursive filter over $\Gamma$ as follows.
+All linear variables are removed from context $\Gamma$.
+The result of context restriction is the non-linear subset of the original context.
     \begin{align*}
@@
     \end{align*}
-    \label{structural}
-  \end{figure}
+  \end{definition}
 
@@
       \inferrule
-      { \Gamma\ \pure }
+      { \Gamma }
       { \Gamma \vdash U \utype U } 
@@
       \inferrule
-      { \Gamma\ \pure }
+      { \Gamma }
       { \Gamma \vdash L \utype U } 
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \Gamma \\
         \Gamma \vdash A \utype U \\ 
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \Gamma \\
         \Gamma \vdash A \utype L \\ 
@@
       \inferrule
-      { \Gamma\ \pure}
+      { \Gamma }
       { \Gamma, x \utype A \vdash x \utype A } 
@@
       \inferrule
-      { \Gamma\ \pure }
+      { \Gamma }
       { \Gamma, x \ltype A \vdash x \ltype A } 
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \Gamma \\
         \Gamma \vdash (x : A) \rightarrow_s B \utype U \\ 
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \Gamma \\
         \Gamma \vdash A \rightarrow_s B \utype U \\
@@
   \begin{lemma}
-    Value soundness. If $\Gamma \vdash v \utype A$ then $\Gamma\ \pure$.
+    Value soundness. If $\Gamma \vdash v \utype A$ then $\Gamma$.
   \end{lemma}
@@
       \inferrule
-      { \Gamma\ \pure \\ 
+      { \Gamma \\ 
         \Gamma \vdash A \utype U \\ 
@@
       \inferrule
-      { \Gamma\ \pure \\ 
+      { \Gamma \\ 
         \Gamma \vdash m \utype A \\
@@
       \inferrule
-      { \Gamma\ \pure \\ \Gamma \vdash n \utype A }
+      { \Gamma \\ \Gamma \vdash n \utype A }
       { \Gamma \vdash \refl\ n \utype n =_A n }
@@
       \inferrule
-      { \Gamma\ \pure \\
+      { \Gamma \\
         \Gamma \vdash c \ltype l \mapsto_A m }

### 2021-10-09 writing/draft1.tex
draft

@@
 \documentclass{article}
-\usepackage[margin=1.3in]{geometry}
+\usepackage[margin=1.2in]{geometry}
 \usepackage[utf8]{inputenc}
@@
 \newcommand{\ltype}{:_L}
-\newcommand{\stype}[1]{\overset{#1}{:}}
+\newcommand{\stype}[1]{:_#1}
 \newcommand{\step}{\leadsto}
-\newcommand{\pstep}{\leadsto_p}
+\newcommand{\pstep}{\leadsto}
 \newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
@@
 
-  Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``minimally invasive'' approach to extending CC with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
+  Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach to extending CC$\omega$ with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
 This allows for a straightforward embedding of CLC into CC, and lifting of CC into CLC, endowing CLC with the fruits of decades of CC research.
@@
 Our type theory contains two sorts of predicative universes $U$ and $L$, being the types of non-linear types and linear types respectively.
-An additional impredicative universe $Prop$ is the type of propositions in the style of CC.
+An additional impredicative universe $U_*$ is the type of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
 
@@
     \begin{grammar}
+      <$k$> := 0 | 1 | 3 $\cdots$ \hspace*{6.8em} predicative levels
+
+      <$i$> := * | 0 | 1 | 3 $\cdots$ \hspace*{5.3em} (im)predicative levels
+
       <$s, t$> ::= $U$ | $L$ \phantom{| $x$} \hspace*{8em} sorts
 
-      <$m, n, A, B, C$> ::= $U_i$ | $L_i$ | $Prop$ | $x$  \hspace*{4.3em} expressions
+      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
       \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
@@
   \end{figure}
-
   A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the presence of four function types: $(x : A) \rightarrow_s B$, $A \rightarrow_s B$, $(x : A) \multimap_s B$, $A \multimap_s B$.
@@
 is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into function syntax, we give canonicity to ways that !
-can be used.
-This allows us to derive substitution, and a direct embedding into CC$\omega$ without needing extra machinery for manipulating exponential.
+By integrating the exponential directly into function syntax, we give implicitly limit !
+to only be used canonically.
+This allows us to derive substitution, and a direct embedding of CLC into CC$\omega$ without needing extra machinery for manipulating exponential.
   \begin{figure}[H]
@@
   \subsection{Type Formation} \label{tyformation}
-  Typing judgments in our language take on the form $\Gamma \vdash m \stype{s} A$ where $s$ is an indexing sort that is either $U$ or $L$.
-Intuitively, this judgment states that expression $m$ has type $A$ and sort $s$ under context $\Gamma$.
-An expression of sort $L$ is linear, and sort $U$ is non-linear.
-Surprisingly, duplication of non-linear expression is not immediately allowed, only non-linear \textit{values} can be safely duplicated.
-We will discuss this subtlety later in Section \ref{subst}.
-
-  Type formation rules are presented in Figure \ref{type}.
-Rules worth discussing in detail are the \rname{L-Axiom} and the rules for constructing various function types.
-For \rname{L-Axiom}, the sort of all linear types $L$ itself is of non-linear sort $U$.
-This means that a linear type can be freely used, avoiding the philosophical trappings discussed earlier.
-
-  The function types $(x : A) \rightarrow_s B$ and $A \rightarrow_s B$ represent non-linear functions, where the subscript $s$ is the sort of co-domain $B$.
-Functions of these types contain no linear free variables, thus can be applied repeatedly without duplication of linear resources.
-The difference between the two is that $(x : A) \rightarrow_s B$ allows co-domain $B$ to depended on input $x$ of non-linear domain $A$.
-For $A \rightarrow_s B$, the domain $A$ is linear so $B$ is not allowed to depend on the function input.
-
-  The variants $(x : A) \multimap_s B$ and $A \multimap_s B$ represent linear functions.
-Functions of these types may contain linear free variables, thus cannot be applied repeated without duplication of linear variables.
-Similar to the non-linear versions, the difference between $(x : A) \multimap_s B$ and $A \multimap_s B$ lies in the allowance of dependency on function input for non-linear domain types.
 
@@
       \inferrule
-      { \Gamma }
-      { \Gamma \vdash U \utype U } 
+      { \pure{\Gamma} }
+      { \Gamma \vdash U_* \utype U_0 } 
+      \rname{Prop-Axiom}
+
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash U_k \utype U_{k+1} } 
       \rname{U-Axiom}
@@
       \inferrule
-      { \Gamma }
-      { \Gamma \vdash L \utype U } 
+      { \pure{\Gamma} }
+      { \Gamma \vdash L_k \utype U_{k+1} } 
       \rname{L-Axiom}
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype U \\ 
-        \Gamma, x \utype A \vdash B \utype s }
-      { \Gamma \vdash (x : A) \rightarrow_s B \utype U } 
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype U_i \\ 
+        \Gamma, x \utype A \vdash B \utype U_* }
+      { \Gamma \vdash (x : A) \rightarrow_U B \utype U_* } 
+      \rname{U-Prop}
+      \\
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype U_k \\ 
+        \Gamma, x \utype A \vdash B \utype s_k }
+      { \Gamma \vdash (x : A) \rightarrow_s B \utype U_k } 
       \rname{U-Prod}
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype L \\ 
-        \Gamma \vdash B \utype s }
-      { \Gamma \vdash A \rightarrow_s B \utype U } 
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype L_k \\ 
+        \Gamma \vdash B \utype s_k }
+      { \Gamma \vdash A \rightarrow_s B \utype U_k } 
       \rname{Arrow}
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype U \\ 
-        \Gamma, x \utype A \vdash B\ \utype s }
-      { \Gamma \vdash (x : A) \multimap_s B \utype L } 
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype U_k \\ 
+        \Gamma, x \utype A \vdash B \utype s_k }
+      { \Gamma \vdash (x : A) \multimap_s B \utype L_k } 
       \rname{L-Prod}
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype L \\ 
-        \Gamma \vdash B \utype s }
-      { \Gamma \vdash A \multimap_s B \utype L } 
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype L_k \\ 
+        \Gamma \vdash B \utype s_k }
+      { \Gamma \vdash A \multimap_s B \utype L_k } 
       \rname{Lolli}
@@
   \subsection{Term Formation} \label{teformation}
-  The term formations rules are presented in Figure \ref{term}.
-
-  The rules \rname{U-Var} and \rname{L-Var} state that a variable's type is determined by its context.
-In the case of \rname{L-Var}, the linear variable $x$ must be the only linear variable within its context.
-This enforcement of uniqueness eliminates the weakening rule for linear variables as redundant linear variables in the context will prevent the usage of \rname{L-Var}.
-
-  The \rname{Conv} rule is standard with regards to prior dependent type theory literature.
-Two types $A$ and $B$ are treated equivalently if they are $\beta\eta$-convertible.
-Our treatment of the convertibility relation $A \equiv B$ differs from standard due to our call-by-value semantics.
-We give a detailed account of this in the next section.
-
-  For the non-linear function formation rules \rname{U-$\lambda_1$} and \rname{U-$\lambda_2$}, the context $\Gamma$ is asserted to be pure.
-The purity of $\Gamma$ ensures that no linear variables occur within function body $n$, allowing the function to be freely used without duplication of linear variables.
-For the linear function formation rules \rname{L-$\lambda_1$} and \rname{L-$\lambda_2$}, the restriction on $\Gamma$'s purity is lifted.
-Linear variables within $\Gamma$ are allowed to occur freely within function body $n$.
-However, the tradeoff is that these linear functions may only be applied once, as multiple uses may result in duplication of its linear free variables.
-
-  In the application rules $\rname{U-App-1}$ and $\rname{L-App-1}$, the argument $n$ is a non-linear expression that may contain linear free variables.
-If reduction is performed naively, substitution of $n$ into the function body may cause duplication of these variables.
-Our operational semantics and value soundness lemma guarantee that substitution will not duplicate linear variables.
-
-  Furthermore, in the \rname{U-App-1} and \rname{L-App-1} rules, the input expression $n$ is not directly substituted into the dependent co-domain $B$ as this may introduce linear variables into types.
-Instead, a $\lambda$-abstraction is formed around $B$ and applied to $n$.
-This delays $\beta$-reduction until $n$ can be evaluated into a value.
 
@@
       \inferrule
-      { \Gamma }
+      { \pure{\Gamma} }
       { \Gamma, x \utype A \vdash x \utype A } 
@@
       \inferrule
-      { \Gamma }
+      { \pure{\Gamma} }
       { \Gamma, x \ltype A \vdash x \ltype A } 
       \rname{L-Var}
-
-      \inferrule
-      { \Gamma \vdash m \stype{s} A \\ A \equiv B }
-      { \Gamma \vdash m \stype{s} B } 
-      \rname{Conv}
       \\
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash (x : A) \rightarrow_s B \utype U \\ 
+      { \pure{\Gamma} \\
+        \Gamma \vdash (x : A) \rightarrow_s B \utype U_i \\ 
         \Gamma, x \utype A \vdash n \stype{s} B }
@@
       \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \rightarrow_s B \utype U \\
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \rightarrow_s B \utype U_k \\
         \Gamma, x \ltype A \vdash n \stype{s} B }
@@
       \inferrule
-      { \overline{\Gamma} \vdash (x : A) \multimap_s B \utype L \\ 
+      { \overline{\Gamma} \vdash (x : A) \multimap_s B \utype L_k \\ 
         \Gamma, x \utype A \vdash n \stype{s} B }
@@
       \inferrule
-      { \overline{\Gamma} \vdash A \multimap_s B \utype L \\ 
+      { \overline{\Gamma} \vdash A \multimap_s B \utype L_k \\ 
         \Gamma, x \ltype A \vdash n \stype{s} B }
@@
       { \Gamma_1 \vdash m \utype (x : A) \rightarrow_s B \\
-        \Gamma_2 \vdash n \utype A \\
-        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
-      { \Gamma \vdash m\ n \stype{s} (\lambda x.B)\ n }
+        \overline{\Gamma_2} \vdash n \utype A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash m\ n \stype{s} B[n/x] }
       \rname{U-App-1}
@@
         \Gamma_2 \vdash n \ltype A \\
-        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
       { \Gamma \vdash m\ n \stype{s} B }
@@
       { \Gamma_1 \vdash m \ltype (x : A) \multimap_s B \\
-        \Gamma_2 \vdash n \utype A \\
-        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
-      { \Gamma \vdash m\ n \stype{s} (\lambda x.B)\ n }
+        \overline{\Gamma_2} \vdash n \utype A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash m\ n \stype{s} B[n/x] }
       \rname{L-App-1}
@@
         \Gamma_2 \vdash n \ltype A \\
-        merge\ \Gamma_1\ \Gamma_2\ \Gamma }
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
       { \Gamma \vdash m\ n \stype{s} B }
       \rname{L-App-2} 
+      \\
+
+      \inferrule
+      { \Gamma \vdash m \stype{s} A \\
+        \overline{\Gamma} \vdash B \utype s_i \\ A \preceq B }
+      { \Gamma \vdash m \stype{s} B } 
+      \rname{Conv}
     \end{mathpar}
@@
   \subsection{Reduction and Equality} \label{reduction}
-  Figure \ref{single} presents the call-by-value operational semantics that we have eluded to. The rules defining the single step relation $\step$ are completely standard.
-
-  For dependently typed languages, terms can appear at the type level.
-A definitional equality judgment is required to identify types beyond simple $\alpha$-equivalence.
-This is usually accomplished by normalization and comparing normal forms.
-Due to the complications brought on by linearity and substitution hinted at in \ref{teformation}, standard normalization techniques cannot be directly applied.
-Unfortunately, the single step relation $\step$ defined previously is not sufficient either as binders block evaluation.
-Following the footsteps of the Trellys project\cite{trellys}, we define a call-by-value parallel step relation $\pstep$ that evaluates under binders.
-We use the transitive reflexive closure of $\pstep$ to define definitional equality.
-
-  \begin{figure}[H]
-    \caption{Single Step Reduction}
-    \begin{mathpar}
-      \inferrule
-      { }
-      { (\lambda x.n)\ v \step [v/x]n }
-      \rname{S-U-$\beta$}
-
-      \inferrule
-      { }
-      { (\lambda x.n)\ v \step [v/x]n }
-      \rname{S-L-$\beta$}
-
-      \inferrule
-      { m \step m' }
-      { m\ n \step m'\ n }
-      \rname{S-App-L}
-
-      \inferrule
-      { n \step n' }
-      { v\ n \step v\ n' }
-      \rname{S-App-R}
-    \end{mathpar}
-    \label{single}
-  \end{figure}
 
@@
       \inferrule
-      { n \pstep n' \\ v \pstep v' }
-      { (\lambda x.n)\ v \pstep [v'/x]n' }
-      \rname{P-U-$\beta$}
-
-      \inferrule
-      { n \pstep n' \\ v \pstep v' }
-      { (\lambda x.n)\ v \pstep [v'/x]n' }
-      \rname{P-L-$\beta$}
+      { m \pstep m' \\ n \pstep n' }
+      { (\lambda x.m)\ n \pstep m'[n'/x] }
+      \rname{P-$\beta$}
 

### 2021-10-10 writing/draft1.tex
draft

@@
     The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, proving that it is a sound logical framework for reasoning about resource.
-CLC is backwards compatible with CC, allowing CLC to enjoy the decades of CC research.
+Specifically, CLC extends CC with a hierarchy of linear universes, and an indexed typing judgment that precisely controls the weakening and contraction of its term level inhabitants.
+We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
+CLC is backwards compatible with CC, allowing CLC to enjoy the fruits of decades of CC research.
 We have formalized and proven correct all major results of the core calculus in the Coq Proof Assistant.
@@
   Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach to extending CC$\omega$ with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
-This allows for a straightforward embedding of CLC into CC, and lifting of CC into CLC, endowing CLC with the fruits of decades of CC research.
+This allows for a straightforward modeling of CLC in CC$\omega$, and lifting of CC$\omega$ into CLC, endowing CLC with the fruits of decades of CC research.
 We further extend CLC with inductive types, showing that as a programming language it can manipulate mutable data structures in a principled way.
@@
     \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, regularity, and subject reduction. 
-    \item We observe that CLC is highly backwards compatible with CC. We construct a reduction preserving embedding of CLC into CC, showing that CLC is consistent. 
+    \item We observe that CLC is highly backwards compatible with CC$\omega$. We construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent. 
     \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
@@
 An additional impredicative universe $U_*$ is the type of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
+We use the meta variable $k$ to specifically quantify over levels $0, 1, 2, ...$ that correspond to the predicative universes.
+We use the meta variable $i$ to quantify over all levels $*, 0, 1, 2, ...$.
 
@@
     \begin{grammar}
-      <$k$> := 0 | 1 | 3 $\cdots$ \hspace*{6.8em} predicative levels
+      <$k$> := 0 | 1 | 2 ... | \phantom{*} \hspace*{5.8em} predicative levels
 
-      <$i$> := * | 0 | 1 | 3 $\cdots$ \hspace*{5.3em} (im)predicative levels
+      <$i$> := * | 0 | 1 | 2 ... \hspace*{5.8em} all levels
 
@@
 is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into function syntax, we give implicitly limit !
+By integrating the exponential directly into function syntax, we implicitly limit !
 to only be used canonically.
-This allows us to derive substitution, and a direct embedding of CLC into CC$\omega$ without needing extra machinery for manipulating exponential.
+This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing extra machinery for manipulating exponential.
   \begin{figure}[H]
@@
   \subsection{Universes and Cumulativity}
-  CLC features three sorts of universes, $Prop$, $U$, and $L$.
-$Prop$ is the impredicative universe propositions.
-$U$, and $L$ are the predicative universe of non-linear types, and linear types respectively.
-The main mechanism that CLC uses to distinguish between linear and non-linear types is by the universe which they belong.
-Intuitively, terms with types that occur within $Prop$, and $U$ are unrestricted in their usage.
-Terms with types that occur within $L$ are restricted to being used exactly once.
+  CLC features two sorts of universes $U$, and $L$ with level indices $*, 0, 1, 2, \cdots$.
+$U_*$ is the impredicative universe of propositions.
+$U_k$, and $L_k$ are the predicative universe of non-linear types, and linear types respectively.
+The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
+Basically, terms with types that occur within $U_i$ are unrestricted in their usage.
+Terms with types that occur within $L_k$ are restricted to being used exactly once.
 
@@
         \end{enumerate}
-      \item $Prop \preceq U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
+      \item $U_* \preceq U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
       \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
@@
   Figure \ref{universe} illustrates the structure of our universe hierarchy.
-Each linear universes $L_i$ has $U_{i+1}$ as its type, allowing functions to dependently quantify over linear \textit{types}.
-However, $L_i$ cumulates to $L_{i+1}$.
-These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of linearity.
+Each linear universe $L_k$ has $U_{k+1}$ as its type, allowing functions to dependently quantify over linear \textit{types}.
+However, $L_k$ cumulates to $L_{k+1}$.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of its occupants' linearity.
 
@@
     \begin{tikzcd}
-      Prop \arrow[r, ":"', "\preceq", dash] 
+      U_* \arrow[r, ":"', "\preceq", dash] 
       & U_0 \arrow[r, ":"', "\preceq", dash] 
@@
       { \Gamma\ \vdash \\ 
-        \overline{\Gamma} \vdash A \utype U }
+        \overline{\Gamma} \vdash A \utype U_i }
       { \Gamma, x \utype A \vdash }
@@
       { \Gamma\ \vdash \\ 
-        \overline{\Gamma} \vdash A \utype L }
+        \overline{\Gamma} \vdash A \utype L_k }
       { \Gamma, x \ltype A\ \vdash } 
@@
 
+  \subsection{Typing Judgment}
+  Typing judgments in CLC take on the form of $\Gamma \vdash m \stype{s} A$.
+Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$.
+The sort index $s$ tells us the linearity of $m$.
+Specifically, if $s = U$, then $m$ has unrestricted usage.
+Conversely, if $s = L$, then $m$ must be used exactly once.
+In Section \ref{meta}, we show through the regularity theorem that $s$ corresponds exactly to the sort of type $A$'s universe.
+
   \subsection{Type Formation} \label{tyformation}
+  The rules for forming types are presented in Figure \ref{type}.
 
@@
   \end{figure}
+
+  The axiom rules \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
+Variables not pertaining to actual proofs could be introduced this way, thus giving rise to the admissibility of weakening.
+To support linear types, we must restrict weakening to non-linear variables.
+This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom}.
+
+  Recall from Figure \ref{correspondence} where we presented the correspondence between CLC function types and MELL implications.
+There exists 8 canonical combinations of !
+and $\multimap$ that we must encode.
+
+  The rules \rname{U-Prop} and \rname{U-Prod} are dependent function types for $\lambda$-abstractions that could be applied an unrestricted number of times.
 
@@
 
-  \subsection{Meta Theory}
+  \subsection{Meta Theory} \label{meta}
   We have proven the type soundness of our language in the form of \textit{progress} and \textit{preservation} theorems.

### 2021-10-10 writing/draft1.tex
wip

@@
   \section{Introduction}
-  The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
-In CC types can depend on terms, allowing one to write fine-grain propositions as types.
+  The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand, and Huet in their landmark work \cite{cc}.
+In CC types can depend on terms, allowing one to write precise specifications as types.
 Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean}, and others.
@@
   \end{equation*}
-  The customer will still maintain full ownership of their money after paying the vendor.
+  If viewed as an propositional implication, the customer will still maintain full ownership of their money after paying the vendor, because implication does not diminish the validity of its antecedent.
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
-Users of proof assistants based on CC often need to embed external logics into CC \cite{bedrock} to provide additional reasoning principles for dealing with resource.
+Users of proof assistants based on CC often need to embed external logics \cite{bedrock} to provide additional reasoning principles for dealing with resource.
 The design and embedding of these logics is a difficult problem in its own right, requiring additional proofs to justify its soundness.
@@
   \section{The Language of CLC}
-
   \subsection{Syntax}
@@
 MELL lacks counterparts for the cases (1), (2), (5), (6) if the co-domain $B$ is dependent on arguments of the domain $A$.
+We will discuss the detail of these function types in Section \ref{tyformation}.
 
@@
   Typing judgments in CLC take on the form of $\Gamma \vdash m \stype{s} A$.
-Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$.
+Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$, with free variables typed in $\Gamma$.
 The sort index $s$ tells us the linearity of $m$.
@@
   The rules for forming types are presented in Figure \ref{type}.
+In CLC, we forbid types from depending on linear terms similar to \cite{llf,neel15} for the same reason of avoiding philosophical pitfalls.
 
@@
 This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom}.
+From \rname{L-Axiom} we can see that the universe of linear types $L_k$ is an inhabitant of $U_{k+1}$.
+This is reminiscent of Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
 
@@
 and $\multimap$ that we must encode.
+Note that the function types $(x : A) \rightarrow_s B$, and $A \rightarrow_s B$ are intrinsically different entities, even if the variable $x$ is unbinding in the former.
+So $A \rightarrow_s B$ should not be viewed as a shorthand for $(\_ : A) \rightarrow_s B$.
+Likewise for the linear variants, $A \multimap_s B$ is not a shorthand for $(\_ : A) \multimap_s B$.
 
-  The rules \rname{U-Prop} and \rname{U-Prod} are dependent function types for $\lambda$-abstractions that could be applied an unrestricted number of times.
+  The \rname{U-Prod} rule is used for forming non-linear dependent function types.
+$\lambda$-abstractions of this type could be applied an unrestricted number of times, with the argument being allowed unrestricted usage within the function body as well.
+Notice that the domain $A$ must be of type $U_k$, this entails that the co-domain $B$ may depend on an unrestricted term of type $A$.
+However, $B$ itself may chose to be of any universe sort with an appropriate predicative level.
+The sort of $B$'s universe is used to annotate the function arrow.
+The side-condition $\pure{\Gamma}$ ensures that all components used to form $(x : A)\rightarrow_s B$ do not dependent on linear variables, thus allowing the \textit{type} itself to used as an unrestricted \textit{term}.
+
+  The \rname{Arrow} rule is also used for forming non-linear function types.
+However, as we have alluded to previously, the formation rule \rname{Arrow} differs from \rname{U-Prod}.
+Namely, the domain $A$ must be of type $L_k$.
+This means that if the co-domain $B$ is dependent on a term of type $A$, $B$ will depend on a linear term.
+To prevent this situation from ever occurring, we explicitly forbid the binding of a dependent variable $x$ within $B$.
+This leaves us with a type of the shape $A \rightarrow_s B$, where $s$ is the sort of $B$'s universe.
+$\lambda$-abstractions of this type could be applied an unrestricted number of times, but the argument must be used within the body of the function once.
+
+  The \rname{L-Prod} rule is used for forming linear function types.
+$\lambda$-abstractions of this type must be applied exactly once, with the argument being allowed unrestricted usage within the function.
+\rname{L-Prod} shares the same
 

### 2021-10-10 writing/draft1.tex
draft1

@@
 \newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
+\newcommand{\arw}[2]
+{\hspace*{0.2em}{\scriptstyle #1}\hspace*{-0.2em}
+\rightarrow
+\hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
 
@@
   The presence of both linear and dependent types enables CLC to write specifications that faithfully encodes the usage of resource.
-The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$, and a linear arrow $\multimap_L$ as follows.
+The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$ as follows.
   \begin{equation*}
-    Money\ 5 \multimap_L Camels + Marlboro
+    (Money\ 5) \arw{L}{L} (Camels + Marlboro)
   \end{equation*}
@@
       <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
-      \indalt $(x : A) \rightarrow_s B$ | $A \rightarrow_s B$
-      \indalt $(x : A) \multimap_s B$ | $A \multimap_s B$
+      \indalt $(x : A) \arw{s}{t} B$
       \indalt $\lambda x. n$ | $m\ n$
@@
   \end{figure}
-  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the presence of four function types: $(x : A) \rightarrow_s B$, $A \rightarrow_s B$, $(x : A) \multimap_s B$, $A \multimap_s B$.
-The reason for these variants is that we have built the !
-exponential of linear logic directly into the syntax of function types.
+  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the indexed function type: $(x : A) \arw{s}{t} B$.
+The reason for these indices is that we have built the !
+exponential of linear logic directly into universe sorts.
+The indices $s$, and $t$ annotate the domain, and co-domain's respective universe sort, hence annotating linearity.
 The behavior of !
@@
 is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into function syntax, we implicitly limit !
+By integrating the exponential directly into type formation rules, we implicitly limit !
 to only be used canonically.
@@
     \begin{align}
-      (\_ : A) \rightarrow_U B \quad &\equiv \quad !(!A \multimap !B) \\
-      (\_ : A) \rightarrow_L B \quad &\equiv \quad !(!A \multimap B) \\
-      A \rightarrow_U B \quad &\equiv \quad !(A \multimap !B) \\
-      A \rightarrow_L B \quad &\equiv \quad !(A \multimap B) \\
-      (\_ : A) \multimap_U B \quad &\equiv \quad !A \multimap !B \\
-      (\_ : A) \multimap_L B \quad &\equiv \quad !A \multimap B \\
-      A \multimap_U B \quad &\equiv \quad A \multimap !B \\
-      A \multimap_L B \quad &\equiv \quad A \multimap B
+      (\_ : A) \arw{U}{U} B \quad &\equiv \quad !A \multimap !B \\
+      (\_ : A) \arw{U}{L} B \quad &\equiv \quad !A \multimap B \\
+      (\_ : A) \arw{L}{U} B \quad &\equiv \quad A \multimap !B \\
+      (\_ : A) \arw{L}{L} B \quad &\equiv \quad A \multimap B
     \end{align}
@@
   Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
-MELL lacks counterparts for the cases (1), (2), (5), (6) if the co-domain $B$ is dependent on arguments of the domain $A$.
+MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of the domain $A$.
 We will discuss the detail of these function types in Section \ref{tyformation}.
@@
       \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
-        \begin{enumerate}
-          \item $(x : A_1) \rightarrow B_1 \preceq (x : A_2) \rightarrow B_2$
-          \item $(x : A_1) \multimap B_1 \preceq (x : A_2) \multimap B_2$
-          \item $A_1 \rightarrow B_1 \preceq A_2 \rightarrow B_2$
-          \item $A_1 \multimap B_1 \preceq A_2 \multimap B_2$
-        \end{enumerate}
+        $(x : A_1) \arw{s}{t} B_1 \preceq (x : A_2) \arw{s}{t} B_2$
     \end{enumerate}
@@
   Figure \ref{universe} illustrates the structure of our universe hierarchy.
-Each linear universe $L_k$ has $U_{k+1}$ as its type, allowing functions to dependently quantify over linear \textit{types}.
+Each linear universe $L_k$ has $U_{k+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
 However, $L_k$ cumulates to $L_{k+1}$.
@@
         \Gamma, x \utype A \vdash B \utype U_* }
-      { \Gamma \vdash (x : A) \rightarrow_U B \utype U_* } 
+      { \Gamma \vdash (x : A) \arw{U}{U} B \utype U_* } 
       \rname{U-Prop}
@@
         \Gamma, x \utype A \vdash B \utype s_k }
-      { \Gamma \vdash (x : A) \rightarrow_s B \utype U_k } 
+      { \Gamma \vdash (x : A) \arw{U}{s} B \utype t_k } 
       \rname{U-Prod}
@@
         \Gamma \vdash B \utype s_k }
-      { \Gamma \vdash A \rightarrow_s B \utype U_k } 
-      \rname{Arrow}
-      \\
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A \utype U_k \\ 
-        \Gamma, x \utype A \vdash B \utype s_k }
-      { \Gamma \vdash (x : A) \multimap_s B \utype L_k } 
+      { \Gamma \vdash (\_ : A) \arw{L}{s} B \utype t_k } 
       \rname{L-Prod}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A \utype L_k \\ 
-        \Gamma \vdash B \utype s_k }
-      { \Gamma \vdash A \multimap_s B \utype L_k } 
-      \rname{Lolli}
     \end{mathpar}

### 2021-10-11 writing/draft1.tex
wip

@@
 \newcommand{\subst}{\text{subst}}
-\newcommand{\utype}{:_U}
-\newcommand{\ltype}{:_L}
+\newcommand{\utype}{:_{\scriptscriptstyle U}}
+\newcommand{\ltype}{:_{\scriptscriptstyle L}}
 \newcommand{\stype}[1]{:_#1}
@@
 MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of the domain $A$.
-We will discuss the detail of these function types in Section \ref{tyformation}.
+Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort in their formation judgment.
+The encoding for this example would be $\Gamma \vdash A \arw{L}{L} B \utype U_i $.
+We will discuss function type formation in Section \ref{tyformation} in more detail.
 
@@
 Conversely, if $s = L$, then $m$ must be used exactly once.
-In Section \ref{meta}, we show through the regularity theorem that $s$ corresponds exactly to the sort of type $A$'s universe.
+In Section \ref{meta}, we show through the regularity theorem that $s$ corresponds exactly to the sort of $A$'s universe.
+
+  \begin{definition} We formally define the terms \textit{non-linear}, \textit{linear}, \textit{unrestricted}, and \textit{restricted}.
+    \begin{enumerate}
+      \item A \textit{type} $A$ is \textit{non-linear} under context $\Gamma$ if $\Gamma \vdash A \utype U_i$.
+      \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A \utype L_k$.
+      \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \utype A$.
+      \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \ltype A$
+    \end{enumerate}
+  \end{definition}
 
@@
   The rules for forming types are presented in Figure \ref{type}.
-In CLC, we forbid types from depending on linear terms similar to \cite{llf,neel15} for the same reason of avoiding philosophical pitfalls.
+In CLC, we forbid types from depending on linear terms similar to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
 
@@
 
-  Recall from Figure \ref{correspondence} where we presented the correspondence between CLC function types and MELL implications.
-There exists 8 canonical combinations of !
-and $\multimap$ that we must encode.
-Note that the function types $(x : A) \rightarrow_s B$, and $A \rightarrow_s B$ are intrinsically different entities, even if the variable $x$ is unbinding in the former.
-So $A \rightarrow_s B$ should not be viewed as a shorthand for $(\_ : A) \rightarrow_s B$.
-Likewise for the linear variants, $A \multimap_s B$ is not a shorthand for $(\_ : A) \multimap_s B$.
+  The \rname{U-Prop} rule is used for forming propositions.
+From the judgment $\Gamma \vdash A \utype U_i$ we can see that \rname{U-Prop} allows for quantification over non-linear types of arbitrary level, hence impredicative.
+The judgment $\Gamma, x \utype A \vdash B \utype U_*$ asserts that the co-domain must be in the impredicative universe $U_*$.
+The final resulting judgment $\Gamma \vdash (x : A) \arw{U}{U} B \utype U_*$ is indexed by $U$, indicating that the \textit{type} $(x : A) \arw{U}{U} B$ can be use unrestrictedly as a \textit{term}.
+Since $(x : A) \arw{U}{U} B$ belongs to universe $U_*$, its \textit{terms} also enjoys unrestricted usage.
 
-  The \rname{U-Prod} rule is used for forming non-linear dependent function types.
-$\lambda$-abstractions of this type could be applied an unrestricted number of times, with the argument being allowed unrestricted usage within the function body as well.
-Notice that the domain $A$ must be of type $U_k$, this entails that the co-domain $B$ may depend on an unrestricted term of type $A$.
-However, $B$ itself may chose to be of any universe sort with an appropriate predicative level.
-The sort of $B$'s universe is used to annotate the function arrow.
-The side-condition $\pure{\Gamma}$ ensures that all components used to form $(x : A)\rightarrow_s B$ do not dependent on linear variables, thus allowing the \textit{type} itself to used as an unrestricted \textit{term}.
+  The \rname{U-Prod} rule is used for forming function types with non-linear domains.
+This is evident from the judgment $\Gamma \vdash A \utype U_k$.
+Since $A$ is in the predicative non-linear universe $U_k$, terms of type $A$ have unrestricted usage.
+The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B \utype s_k$.
+The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
+From the resulting judgment $\Gamma \vdash (x : A) \arw{U}{s} B \utype t_k$, we see that $A$, and $B$'s universe sorts are used as indices to the arrow.
+Interestingly, $(x : A) \arw{U}{s} B$ is assigned to a universe $t_k$ of level $k$ and arbitrary sort $t$.
+If $t$ is chosen to be $U$, then $\Gamma \vdash (x : A) \arw{U}{s} B \utype U_k$ tells us $\lambda$-abstractions of this type have unrestricted usage.
+Conversely, if $t$ is chose to be $L$, then $\Gamma \vdash (x : A) \arw{U}{s} B \utype L_k$ tells us $\lambda$-abstractions of this type have restricted usage.
+This is how we encode the missing $!(A \multimap B)$ forms from MELL shown in Figure \ref{correspondence}.
 
-  The \rname{Arrow} rule is also used for forming non-linear function types.
-However, as we have alluded to previously, the formation rule \rname{Arrow} differs from \rname{U-Prod}.
-Namely, the domain $A$ must be of type $L_k$.
-This means that if the co-domain $B$ is dependent on a term of type $A$, $B$ will depend on a linear term.
-To prevent this situation from ever occurring, we explicitly forbid the binding of a dependent variable $x$ within $B$.
-This leaves us with a type of the shape $A \rightarrow_s B$, where $s$ is the sort of $B$'s universe.
-$\lambda$-abstractions of this type could be applied an unrestricted number of times, but the argument must be used within the body of the function once.
-
-  The \rname{L-Prod} rule is used for forming linear function types.
-$\lambda$-abstractions of this type must be applied exactly once, with the argument being allowed unrestricted usage within the function.
-\rname{L-Prod} shares the same
+  The \rname{L-Prod} rule is used for forming function types with linear domains. From the judgment $\Gamma \vdash A \utype L_k$,
 

### 2021-10-11 writing/draft1.tex
draft

@@
 \hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
+\newcommand{\erase}[1]{\llbracket #1 \rrbracket}
 
@@
   The syntax of the core type theory is presented in Figure \ref{syntax}.
-Our type theory contains two sorts of predicative universes $U$ and $L$, being the types of non-linear types and linear types respectively.
-An additional impredicative universe $U_*$ is the type of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
+Our type theory contains two sorts of universes $U$ and $L$.
 We use the meta variable $k$ to specifically quantify over levels $0, 1, 2, ...$ that correspond to the predicative universes.
 We use the meta variable $i$ to quantify over all levels $*, 0, 1, 2, ...$.
+Here, $U_*$ is the impredicative universe of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
+$U_k$, and $L_k$ are the predicative universes of non-linear, and linear types respectively.
 
@@
 is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into type formation rules, we implicitly limit !
+By integrating the exponential directly into universe sorts, we implicitly limit !
 to only be used canonically.
-This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing extra machinery for manipulating exponential.
+This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing any machinery for manipulating exponential.
   \begin{figure}[H]
@@
   Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
-MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of the domain $A$.
-Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort in their formation judgment.
+MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of domain $A$.
+Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort through their formation judgment.
 The encoding for this example would be $\Gamma \vdash A \arw{L}{L} B \utype U_i $.
-We will discuss function type formation in Section \ref{tyformation} in more detail.
+We will discuss function type formation in Section \ref{tyformation} in greater detail.
+
+  In practice, algorithmic type checking techniques such as bi-directional typing allow users to omit writing most of these sort indices.
 
@@
 Conversely, if $s = L$, then $m$ must be used exactly once.
-In Section \ref{meta}, we show through the regularity theorem that $s$ corresponds exactly to the sort of $A$'s universe.
+Theorem \ref{regularity} (regularity) shows that $s$ corresponds exactly to the sort of $A$'s universe.
 
@@
       \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A \utype L_k$.
-      \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \utype A$.
-      \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \ltype A$
+      \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \utype A$. A unrestricted term may be used an arbitrary number of times.
+      \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \ltype A$ A restricted term must be used exactly once.
     \end{enumerate}
@@
         \Gamma \vdash A \utype L_k \\ 
-        \Gamma \vdash B \utype s_k }
-      { \Gamma \vdash (\_ : A) \arw{L}{s} B \utype t_k } 
+        \Gamma \vdash B \utype s_k \\
+        x \notin \Gamma }
+      { \Gamma \vdash (x : A) \arw{L}{s} B \utype t_k } 
       \rname{L-Prod}
@@
 
-  The \rname{L-Prod} rule is used for forming function types with linear domains. From the judgment $\Gamma \vdash A \utype L_k$,
+  The \rname{L-Prod} rule is used for forming function types with linear domains.
+From the judgment $\Gamma \vdash A \utype L_k$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
+Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B \utype s_k$.
+Like \rname{U-Prod}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
+The final resulting judgment is $\Gamma \vdash (x : A) \arw{L}{s} B \utype t_k$.
+Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
+Again, we may chose $t$ to be either $U$ or $L$, which form function types for $\lambda$-abstractions with unrestricted or restricted usage respectively.
+
+  For rules \rname{U-Prod}, and \rname{L-Prod}, it may seem as if the final choice of $t_k$ does not matter.
+Indeed, the choice of $t = U$ or $t = L$ has no effect on type formation.
+But the differences will become apparent when typing $\lambda$-abstractions in Section \ref{teformation}.
 
   \subsection{Term Formation} \label{teformation}
+  The rules for term formation are presented in Figure \ref{term}.
 
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash (x : A) \rightarrow_s B \utype U_i \\ 
-        \Gamma, x \utype A \vdash n \stype{s} B }
-      { \Gamma \vdash \lambda x . n \utype (x : A) \rightarrow_s B }
-      \rname{U-$\lambda_1$}
+        \Gamma \vdash (x : A) \arw{s}{t} B \utype U_i \\ 
+        \Gamma, x \stype{s} A \vdash n \stype{t} B }
+      { \Gamma \vdash \lambda x . n \utype (x : A) \arw{s}{t} B }
+      \rname{U-$\lambda$}
       \\
@@
       \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A \rightarrow_s B \utype U_k \\
-        \Gamma, x \ltype A \vdash n \stype{s} B }
-      { \Gamma \vdash \lambda x . n \utype A \rightarrow_s B }
-      \rname{U-$\lambda_2$}
+      { \overline{\Gamma} \vdash (x : A) \arw{s}{t} B \utype L_k \\ 
+        \Gamma, x \stype{s} A \vdash n \stype{t} B }
+      { \Gamma \vdash \lambda x . n \ltype (x : A) \arw{s}{t} B }
+      \rname{L-$\lambda$}
       \\
@@
       \inferrule
-      { \overline{\Gamma} \vdash (x : A) \multimap_s B \utype L_k \\ 
-        \Gamma, x \utype A \vdash n \stype{s} B }
-      { \Gamma \vdash \lambda x . n \ltype (x : A) \multimap_s B }
-      \rname{L-$\lambda_1$}
+      { \Gamma_1 \vdash m \stype{t} (x : A) \arw{U}{s} B \\
+        \pure{\Gamma_2} \\
+        \Gamma_2 \vdash n \utype A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      \rname{U-App}
       \\
@@
       \inferrule
-      { \overline{\Gamma} \vdash A \multimap_s B \utype L_k \\ 
-        \Gamma, x \ltype A \vdash n \stype{s} B }
-      { \Gamma \vdash \lambda x . n \ltype A \multimap_s B }
-      \rname{L-$\lambda_2$}
-      \\
-
-      \inferrule
-      { \Gamma_1 \vdash m \utype (x : A) \rightarrow_s B \\
-        \overline{\Gamma_2} \vdash n \utype A \\
+      { \Gamma_1 \vdash m \stype{t} (x : A) \arw{L}{s} B \\
+        \Gamma_2 \vdash n \ltype A \\
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
       { \Gamma \vdash m\ n \stype{s} B[n/x] }
-      \rname{U-App-1}
-      \\
-
-      \inferrule
-      { \Gamma_1 \vdash m \utype A \rightarrow_s B \\
-        \Gamma_2 \vdash n \ltype A \\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B }
-      \rname{U-App-2}
-      \\
-
-      \inferrule
-      { \Gamma_1 \vdash m \ltype (x : A) \multimap_s B \\
-        \overline{\Gamma_2} \vdash n \utype A \\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B[n/x] }
-      \rname{L-App-1}
-
-      \inferrule
-      { \Gamma_1 \vdash m \ltype A \multimap_s B \\
-        \Gamma_2 \vdash n \ltype A \\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B }
-      \rname{L-App-2} 
+      \rname{L-App} 
       \\
@@
 
-  \subsection{Reduction and Equality} \label{reduction}
+  The rules \rname{U-Var}, and \rname{L-Var} are used for typing free variables.
+The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma, x \utype A$ with a non-linear type $A$.
+The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma, x \ltype A$ with linear type $A$.
+For both rules, the side condition $\pure{\Gamma}$ forbids irrelevant variables with linear type from occurring within $\Gamma$.
+This prevents weakening variables with linear type.
 
+  As promised in Section \ref{tyformation}, the choice of $t_k$ for function type formation comes into play during the typing of $\lambda$-abstractions.
+Namely, in the pair of rules \rname{U-$\lambda$}, and \rname{L-$\lambda$}.
+
+  By design, types with universe sort $U$ are non-linear, terms with these types enjoy unrestricted usage.
+In \rname{U-$\lambda$}, the function type being addressed has universe $U_i$, as seen by the judgment $\Gamma \vdash (x : A) \arw{s}{t} \utype U_i$.
+$\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage.
+This consideration is realized by the side condition $\pure{\Gamma}$, asserting all variables in context $\Gamma$ are unrestricted.
+Next, the body of the abstraction $n$ is typed as $\Gamma, x \stype{s} A \vdash n \stype{t} B$, where $s$ is the sort of $A$ and $t$ is the sort of $B$.
+Finally, the resulting judgment $\Gamma \vdash \lambda x.n \utype (x : A) \arw{s}{t} B$ with index $U$, asserts that the $\lambda$-abstraction can be used unrestrictedly.
+
+  In contrast to \rname{U-$\lambda$}, the \rname{L-$\lambda$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
+Since these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n \stype{t} B$, and lack of side condition $\pure{\Gamma}$.
+However, in judgment $\overline{\Gamma} \vdash (x : A) \arw{s}{t} B \utype L_k$ the context must be filtered, since types are not allowed to depend on restricted variables.
+The final resulting judgment $\Gamma \vdash \lambda x.n \ltype (x : A) \arw{s}{t} B$ with index $L$, asserts that the $\lambda$-abstraction must be used exactly once.
+
+  For \rname{U-App}, domain $A$ is a non-linear type, as seen by its arrow index in $(x : A) \arw{U}{s} B$.
+Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
+Thus, the supplied argument $n$ must not depend on restricted variables in context $\Gamma_2$.
+Otherwise, substitution may put multiple copies of $n$ into $m$ during $\beta$-reduction, duplicating variables that should have been restricted.
+This justifies the side condition of $\pure{\Gamma_2}$.
+The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
+
+  Now for \rname{L-App}, domain $A$ is a linear type, as seen by its arrow index in $(x : A) \arw{L}{s} B$.
+Intuitively, this tells us that $x$ must be used once within the body of $m$.
+During $\beta$-reduction, substitution will only put a single copy of $n$ into the body of $m$, so $n$ can depend on restricted variables within $\Gamma_2$ without fear of duplicating them.
+This justifies the lack of side condition $\pure{\Gamma_2}$.
+The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
+
+  Finally, the \rname{Conv} rule allows judgment $\Gamma \vdash m \stype{s} A$ to convert to judgment $\Gamma \vdash m \stype{s} B$, if $B$ is a valid type in context $\overline{\Gamma}$, and of the same sort $s$ as $A$.
+Futhermore, $A$ must be a subtype of $B$ satisfying the cumulativity relation $A \preceq B$.
+This rule gives rise to large eliminations, as computations embedded at the type level can convert to canonical types.
+
+  \subsection{Equality and Parallel Reduction} \label{reduction}
+  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi}.
+The reduction, and equality rules for CLC are presented in Figure \ref{parallel}, and are entirely standard.
   \begin{figure}[H]
@@
       \inferrule
-      { n \pstep n' }
-      { \lambda x.n \pstep \lambda x.n' }
-      \rname{P-$\lambda$}
-
-      \inferrule
       { m \pstep m' \\ n \pstep n' }
@@
       { A \pstep A' \\ B \pstep B' }
-      { (x : A) \rightarrow_s B \pstep (x : A') \rightarrow_s B' }
-      \rname{P-U-Prod}
-
-      \inferrule
-      { A \pstep A' \\ B \pstep B' }
-      { A \rightarrow_s B \pstep A' \rightarrow_s B' }
-      \rname{P-Arrow}
-
-      \inferrule
-      { A \pstep A' \\ B \pstep B' }
-      { (x : A) \multimap_s B \pstep (x : A') \multimap_s B' }
-      \rname{P-L-Prod}
-
-      \inferrule
-      { A \pstep A' \\ B \pstep B' }
-      { A \multimap_s B \pstep A' \multimap_s B' }
-      \rname{P-Lolli}
+      { (x : A) \arw{s}{t} B \pstep (x : A') \arw{s}{t} B' }
+      \rname{P-Prod}
     \end{mathpar}
@@
   \end{figure}
+  As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain a simple operational semantics, whose $\beta$-reductions are very well behaved.
 
   \subsection{Meta Theory} \label{meta}
-  We have proven the type soundness of our language in the form of \textit{progress} and \textit{preservation} theorems.
-The proofs have been formalized in Coq with help from the Autosubst\cite{autosubst} library.
+  In this section, we focus our discussion on the meta-properties of CLC.
+First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
+Next, we show that CLC is a valid linear type theory through the \textit{linearity} theorem.
+Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, showing that CLC is strongly normalizing.
 
-  \subsubsection{Step and Parallel Step}
+  All proofs have been formalized in Coq with help from the Autosubst \cite{autosubst} library.
+The Coq development is publicly available on the first author's Github repository.
+To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
+We given a hand written version of the proof in the appendix as well.
 
-  The following lemmas and proofs are entirely standard. The restriction to value-form arguments for $\beta$-reduction does not pose any complications.
+  \subsubsection{Reduction and Confluence}
 
-  \begin{lemma}
-    Single step reduction implies parallel step reduction. If $m \step m'$, then $m \pstep m'$.
-  \end{lemma}
+  The following lemmas and proofs are entirely standard. The presence of linear types do not pose any complications as reductions are untyped.
 
@@
   \begin{corollary}
-    The definitional equality relation is an equivalence relation.
+    The definitional equality relation $\equiv$ is an equivalence relation.
   \end{corollary}
 
+  \subsubsection{Weakening} \label{subst}
+  CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.
+
+  \begin{theorem}
+    Weakening. If $\Gamma \vdash m \stype{s} A$ is a valid, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+  \end{theorem}
+
   \subsubsection{Substitution} \label{subst}
-  Though the \textit{substitution} lemma is widely considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the \textit{substitution} lemma is admissible.
-Much of this difficulty arise during the substitution of non-linear expressions.
-Perhaps the most famous work detailing the issues of substitution is due to Wadler\cite{substitute}.
-Since computation arise as a consequence of substitution, it is imperative to get it right.
+  Though the substitution lemma is widely considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
+Much of this difficulty arise during the substitution of arguments containing !
+exponential.
+Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
+He defines additional syntax, and semantics for the intricate unboxing of !
+terms, solving the lack of substitute in Abramsky's term calculus \cite{abramsky1993}.
 
-  Generally, the application rule looks similar to the following for languages with linear types.
-  \begin{mathpar}
-    \inferrule
-    { \Gamma \vdash m : A \multimap B \\ 
-      \Delta \vdash n : A }
-    { \Gamma, \Delta \vdash m\ n : B } 
-  \end{mathpar}
-
-  If type $A$ is a non-linear type, then $n$ ought to be used freely.
-However, it is possible for $n$ to contain linear variables present in $\Delta$.
-If $m$ is a lambda abstraction $\lambda x.m'$ where $x$ occurs multiple times within $m'$, substitution of $n$ for $x$ will cause duplication of linear variables.
-One approach for solving this issue is to wrap non-linear values in an explicit modality, unpacking the internal value only when needed\cite{substitute,neel15}.
-Another is to ban non-linear expressions from containing linear variables\cite{llf,luo,qtt}.
-
-  Our call-by-value semantics resolves the substitution problem without imposing any of the modifications mentioned previously to the typing of application.
-The crucial realization is the following \textit{value soundness} lemma: non-linear \textit{values} contain no linear variables.
-
-  \begin{lemma}
-    Value soundness. If $\Gamma \vdash v \utype A$ then $\Gamma$.
-  \end{lemma}
-
-  From the \textit{value soundness} lemma, the standard rule for application is admissible.
-Intuitively, a single copy of each linear resource is used to create a single non-linear value.
-This value can then be freely duplicated without needing the original resources to generate fresh copies.
-This is consistent with the common practice of retrieving non-linear values from linear references.
+  Our design of integrating ! into universe sorts removes the need for ! manipulating syntax, and semantics. The substitution lemma is directly proved by induction on typing derivations.
 
@@
     Substitution.
-For $\Gamma_1, x \stype{s} A \vdash m \stype{t} B$ and $\Gamma_2 \vdash v \stype{s} A$, if there exists $\Gamma$ such that $merge\ \Gamma_1\ \Gamma_2\ \Gamma$ is defined, then $\Gamma \vdash [v/x]m \stype{t} [v/x]B$.
+For $\Gamma_1, x \stype{s} A \vdash m \stype{t} B$ and $\Gamma_2 \vdash v \stype{s} A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is defined for some $\Gamma$, then $\Gamma \vdash [v/x]m \stype{t} [v/x]B$.
   \end{lemma}
@@
   \subsubsection{Type Soundness}
-  The following theorems are a direct result of the \textit{substitution} lemma and various canonical-form lemmas.
+  In order to prove subject reduction, we first prove the regularity theorem.
+The main purpose of regularity is to lower types down to the term level, enabling the application of various inversion lemmas.
+Regularity also serves a second purpose, it tells us the judgment index $s$ of a term $\Gamma \vdash m \stype{s} A$ is exactly the same sort as its type's universe $\overline{\Gamma} \vdash A \utype s_i$.
 
   \begin{theorem}
-    Progress. For $\epsilon \vdash m \stype{s} A$, either $m$ is a value or there exists $n$ such that $m \step n$.
+    Regularity.
+For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid judgment, then there exists some level $i$ such that $\overline{\Gamma} \vdash A \utype s_i$.
+    \label{regularity}
   \end{theorem}
 
+  With weakening, substitution, regularity, and various inversion lemmas proven, subject reduction is proved by induction on typing derivation.
+
   \begin{theorem}
-    Preservation. For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
+    Subject reduction. For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
   \end{theorem}
 
-  \section{Extensions}
+  \subsubsection{Linearity}
+  At this point, we have proven that CLC is type sound.
+However, we still need to prove that the removal of weakening, and contraction for restricted variables yield tangible impact on the structure of terms.
+For this purpose, we define a binding aware recursive function $occurs(x, m)$ that counts the number of times variable $x$ appears within term $m$.
+The linearity theorem asserts that restricted variables are used exactly once within a term, subsuming safe resource usage.
 
-  \subsection{Data Types}
-  Though it is possible to encode data and propositions directly in the core language using Church-encodings, it is incredibly inconvenient.
-To address this, our implementation allows users to define inductive data types\cite{inductive} similar to Coq or Agda\cite{agda}.
-A pattern matching construct\cite{pattern} is defined on inductive types for data elimination.
-Checking is performed to ensure that non-linear inductive types do not carry linear terms and linear types cannot be used as type indices or parameters.
+  Before we proceed, we first prove the seemingly obvious narity lemma.
 
-  Commonly used inductive types are formalized in Figures \ref{dataform} with their Introduction and Elimination rules formalized in Figure \ref{dataintro} and Figure \ref{dataelim} respectively.
-Reduction for data types obey the same call-by-value semantics as outlined in \ref{reduction}.
-Other standard data types that are not covered here are $\mathbb{N}$ for natural numbers and $\top$ for the unit type.
+  \begin{lemma}
+    Narity. If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
+  \end{lemma}
 
-  \begin{figure}[H]
-    \caption{Data Formation}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma \\ 
-        \Gamma \vdash A \utype U \\ 
-        \Gamma, x \utype A \vdash B \utype U }
-      { \Gamma \vdash \Sigma x:A.B \utype U }
-      \rname{$\Sigma$}
+  For cases with branched syntax in the linearity theorem, such as application $(m\ n)$, the $occurs$ function sums up the occurrences of variable $x$ in branches $m$, and $n$.
+The narity lemma is used to prove that either $occurs(x, m) = 1 \wedge occurs(x, n) = 0$ or $occurs(x, m) = 0 \wedge occurs(x, n) = 1$ is true.
 
-      \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype U \\ 
-        \Gamma, x \utype A \vdash B \utype L }
-      { \Gamma \vdash F x:A.B \utype L }
-      \rname{$F$}
+  \begin{theorem}
+    Linearity. If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+  \end{theorem}
 
-      \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype L \\ 
-        \Gamma \vdash B \utype L }
-      { \Gamma \vdash A \otimes B \utype L }
-      \rname{$\otimes$}
+  \subsubsection{Strong Normalization}
+  The strong normalization theorem of CLC is proven by construction of a typing, and reduction preserving erasure function to CC$\omega$.
+We assume familiarity with CC$\omega$ syntax here, and define the erasure function as follows.
 
-      \inferrule
-      { \Gamma \\ 
-        \Gamma \vdash m \utype A \\
-        \Gamma \vdash n \utype A }
-      { \Gamma \vdash m =_A n \utype U }
-    \end{mathpar}
-    \label{dataform}
-  \end{figure}
-
-  \begin{figure}[H]
-    \caption{Data Introduction}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma_1 \vdash m \utype A \\
-        \Gamma_2 \vdash n \utype (\lambda x.B)\ m \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash (m, n) \utype \Sigma x:A.B }
-      \rname{$\Sigma$-intro}
-
-      \inferrule
-      { \Gamma_1 \vdash m \utype A \\
-        \Gamma_2 \vdash n \ltype (\lambda x.B)\ m \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash [m, n] \ltype F x:A.B }
-      \rname{$F$-intro}
-
-      \inferrule
-      { \Gamma_1 \vdash m \ltype A \\
-        \Gamma_2 \vdash n \ltype B \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash \langle m, n \rangle \ltype A \otimes B }
-      \rname{$\otimes$-intro}
-
-      \inferrule
-      { \Gamma \\ \Gamma \vdash n \utype A }
-      { \Gamma \vdash \refl\ n \utype n =_A n }
-    \end{mathpar}
-    \label{dataintro}
-  \end{figure}
-
-  \begin{figure}[H]
-    \caption{Data Elimination}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma_1 \vdash m \utype \Sigma x.A.B \\
-        \Gamma_2, x \utype A, y \utype B \vdash n \stype{s} C \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash \letin{(x , y)}{m}{n} \stype{s} C }
-      \rname{$\Sigma$-elim}
-
-      \inferrule
-      { \Gamma_1 \vdash m \ltype F x.A.B \\
-        \Gamma_2, x \utype A, y \ltype B \vdash n \stype{s} C \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash \letin{[x , y]}{m}{n} \stype{s} C }
-      \rname{$F$-elim}
-
-      \inferrule
-      { \Gamma_1 \vdash m \ltype A \otimes B \\
-        \Gamma_2, x \ltype A, y \ltype B \vdash n \stype{s} C \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash \letin{$\langle x , y\rangle$}{m}{n} \stype{s} C }
-      \rname{$\otimes$-elim}
-
-      \inferrule
-      { \Gamma_1 \vdash p \utype m =_A n \\
-        \Gamma_2 \vdash q \stype{s} B[m] \\
-        \overline{\Gamma}, x \utype A \vdash B[x] \utype s \\
-        merge\ \Gamma_1 \Gamma_2 \Gamma }
-      { \Gamma \vdash \subst(\lambda x.B, p, q) \stype{s} B[n] }
-    \end{mathpar}
-    \label{dataelim}
-  \end{figure}
-
-  \subsection{Imperative Programming}
-  With the addition of data types, axioms for imperative programming can be added to the language.
-In Figure \ref{state} we give a set of axioms for stateful programs.
-Note that these axioms are not associated with any reductions.
-This unsurprisingly breaks the \textit{Progress Theorem} as axioms are neither values nor can they reduce.
-Despite this obvious shortcoming, the axiomatic treatment of state provides an interface for extraction of imperative code with safe manual memory management.
-
-  \subsubsection{State Programs}
-  \begin{figure}[H]
-    \caption{State Programs}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma \\
-        \Gamma \vdash A \utype U \\ 
-        \Gamma \vdash m \utype A \\
-        \Gamma \vdash l \utype \mathbb{N} }
-      { \Gamma \vdash l \mapsto_A m \utype L }
-      \rname{Capability}
-
-      \inferrule
-      { \Gamma \\ 
-        \Gamma \vdash A \utype U \\ 
-        \Gamma \vdash m \utype A }
-      { \Gamma \vdash \new(m) \ltype F l:\mathbb{N}. l \mapsto_A m }
-      \rname{New}
-
-      \inferrule
-      { \Gamma \\
-        \Gamma \vdash c \ltype l \mapsto_A m }
-      { \Gamma \vdash \free(c) \utype \top }
-      \rname{Free}
-
-      \inferrule
-      { \Gamma \vdash c \ltype l \mapsto_A m }
-      { \Gamma \vdash \get(l, c) \ltype
-        F x:A. F e: (x =_A m). l \mapsto_A m }
-      \rname{Get}
-
-      \inferrule
-      { \Gamma \vdash c \ltype l \mapsto_A m \\
-        \overline{\Gamma} \vdash n \utype A }
-      { \Gamma \vdash \set(l, c, n) \ltype l \mapsto_A n }
-      \rname{Set}
-    \end{mathpar}
-    \label{state}
-  \end{figure}
-
-  \subsubsection{Laws for Free}
+  \begin{definition}
+    \begin{align*}
+      \erase{x} &= x \\
+      \erase{U_*} &= Prop \\
+      \erase{U_k} &= Type_k \\
+      \erase{L_k} &= Type_k \\
+      \erase{(x : A) \arw{s}{t} B} &= (x : \erase{A}) \rightarrow \erase{B} \\
+      \erase{\lambda x.n} &= \lambda x.\erase{n} \\
+      \erase{m\ n} &= \erase{m}\ \erase{n}
+    \end{align*}
+  \end{definition}
 
@@
 We would like to extend these results our dependent linear type theory.
-
   \section{Conclusion}

### 2021-10-12 writing/draft1.tex
draft

@@
 
-  \begin{lemma}
-    Parallel reduction satisfies the diamond property. If $m \pstep m_1$ and $m \pstep m_2$ then there exists $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+  \begin{lemma} Parallel reduction satisfies the diamond property. \\
+    If $m \pstep m_1$ and $m \pstep m_2$ then there exists $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
   \end{lemma}
 
-  \begin{corollary}
-    The transitive reflexive closure of parallel reduction is confluent. If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
+  \begin{corollary} The transitive reflexive closure of parallel reduction is confluent. \\ 
+    If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
   \end{corollary}
@@
 
-  \begin{theorem}
-    Weakening. If $\Gamma \vdash m \stype{s} A$ is a valid, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
-  \end{theorem}
+  \begin{lemma} Weakening. \\
+    If $\Gamma \vdash m \stype{s} A$ is a valid, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+  \end{lemma}
 
@@
 
-  \begin{lemma}
-    Substitution.
+  \begin{lemma} Substitution. \\
     For $\Gamma_1, x \stype{s} A \vdash m \stype{t} B$ and $\Gamma_2 \vdash v \stype{s} A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is defined for some $\Gamma$, then $\Gamma \vdash [v/x]m \stype{t} [v/x]B$.
@@
 
-  \begin{theorem}
-    Regularity.
+  \begin{theorem} Regularity. \\
      For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid judgment, then there exists some level $i$ such that $\overline{\Gamma} \vdash A \utype s_i$.
@@
 
-  \begin{theorem}
-    Subject reduction. For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
+  \begin{theorem} Subject reduction. \\
+    For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
   \end{theorem}
@@
 
-  \begin{lemma}
-    Narity. If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
+  \begin{lemma} Narity. \\
+    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
   \end{lemma}
@@
 
-  \begin{theorem}
-    Linearity. If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+  \begin{theorem} Linearity. \\
+    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
   \end{theorem}
@@
 
+  We prove the following lemma to commute erasure and substitution when needed.
+  \begin{lemma} Erasure commutes with substitution. \\ 
+    For any CLC terms $m$, $n$, and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
+  \end{lemma}
+
+  For the following theorem, we refer to the reductions of CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions of CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
+  \begin{theorem} \label{preserve} Erasure preserves reduction. \\
+    For any CLC terms $m$, and $n$, if there is $m \pstep_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \pstep_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  \end{theorem}
+
+  \begin{theorem} \label{embedding} Embedding. \\
+    For any CLC context $\Gamma$, terms $m$, $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid typing judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid a typing judgment in CC$\omega$.
+  \end{theorem}
+
+  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions.
+This is contradictory to the strong normalization property of CC$\omega$, thus this hypothetical term does not exist in CLC.
+
+  \begin{theorem}
+    Well-typed CLC terms are strongly normalizing.
+  \end{theorem}
+
   \section{Future Work}

### 2021-10-12 writing/draft1.tex
draft

@@
 
-  We prove the following lemma to commute erasure and substitution when needed.
+  With slight overloading of notation, we define erasure for CLC contexts recursively.
+
+  \begin{definition}
+    \begin{align*}
+      \erase{\epsilon} &= \epsilon \\
+      \erase{\Gamma, x \stype{s} A} &= \erase{\Gamma}, x : \erase{A}
+    \end{align*}
+  \end{definition}
+
+  We prove the following lemma to commute erasure and substitution whenever needed.
   \begin{lemma} Erasure commutes with substitution. \\ 
@@
 
-  For the following theorem, we refer to the reductions of CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions of CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
+  For the following theorem, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
   \begin{theorem} \label{preserve} Erasure preserves reduction. \\
@@
   \section{Future Work}
-  All of the theorems we have developed so far are purely syntactic in nature.
-But the natural correspondence between linear types and call-by-value semantics seem to hint at a deeper connection.
-Indeed, past works\cite{cbvsemantics} have provided insight for the simply typed case.
-We would like to extend these results our dependent linear type theory.
   \section{Conclusion}

### 2021-10-12 writing/draft1.tex
draft

@@
     \begin{align}
-      (\_ : A) \arw{U}{U} B \quad &\equiv \quad !A \multimap !B \\
-      (\_ : A) \arw{U}{L} B \quad &\equiv \quad !A \multimap B \\
-      (\_ : A) \arw{L}{U} B \quad &\equiv \quad A \multimap !B \\
-      (\_ : A) \arw{L}{L} B \quad &\equiv \quad A \multimap B
+      (\_ : A) \arw{U}{U} B \quad &\Leftrightarrow \quad !A \multimap !B \\
+      (\_ : A) \arw{U}{L} B \quad &\Leftrightarrow \quad !A \multimap B \\
+      (\_ : A) \arw{L}{U} B \quad &\Leftrightarrow \quad A \multimap !B \\
+      (\_ : A) \arw{L}{L} B \quad &\Leftrightarrow \quad A \multimap B
     \end{align}
@@
   \subsection{Equality and Parallel Reduction} \label{reduction}
-  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi}.
-The reduction, and equality rules for CLC are presented in Figure \ref{parallel}, and are entirely standard.
+  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi} as presented in Figure \ref{parallel}, all of which are entirely standard.
   \begin{figure}[H]
@@
   If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions.
-This is contradictory to the strong normalization property of CC$\omega$, thus this hypothetical term does not exist in CLC.
+This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
 

### 2021-10-12 writing/draft1.tex
draft

@@
 For linear variables, the $\mrg{\_}{\_}{\_}$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
-This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for non-linear variables whilst forbidding it for linear ones.
+This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for unrestricted variables whilst forbidding it for restricted.
 
@@
 
-  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions.
+  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}.
 This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.

### 2021-10-12 writing/draft1.tex
draft

@@
   \end{equation*}
-  If viewed as an propositional implication, the customer will still maintain full ownership of their money after paying the vendor, because implication does not diminish the validity of its antecedent.
+  If viewed as a propositional implication, the customer will still maintain full ownership of their money after paying the vendor, because implication does not diminish the validity of its antecedent.
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.

### 2021-10-14 writing/draft2.tex
draft

--- 2021-10-12 writing/draft1.tex
@@
 \rightarrow
+\hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
+\newcommand{\lrw}[2]
+{\hspace*{0.2em}{\scriptstyle #1}\hspace*{-0.2em}
+\multimap
 \hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
@@
 
-  \subsection{Equality and Parallel Reduction} \label{reduction}
-  The operational semantics, and definitional equality of CLC is defined by parallel reductions \cite{takahashi} as presented in Figure \ref{parallel}, all of which are entirely standard.
+  \subsection{Equality and Reduction} \label{reduction}
+  The operational semantics, and definitional $\beta$-equality of CLC is defined reductions as presented in Figure \ref{parallel}, all of which are entirely standard.
   \begin{figure}[H]
@@
       \inferrule
-      { m_1 \pstep^* n \\ m_2 \pstep^* n }
+      { m_1 \step^* n \\ m_2 \step^* n }
       { m_1 \equiv m_2 : A }
@@
       {  }
-      { x \pstep x }
-      \rname{P-Var}
-
-      \inferrule
-      { }
-      { U \pstep U }
-      \rname{P-U}
-
-      \inferrule
-      { }
-      { L \pstep L }
-      \rname{P-L}
-
-      \inferrule
-      { n \pstep n' }
-      { \lambda x.n \pstep \lambda x.n' }
-      \rname{P-$\lambda$}
-
-      \inferrule
-      { m \pstep m' \\ n \pstep n' }
-      { m\ n \pstep m'\ n' }
-      \rname{P-App}
-
-      \inferrule
-      { m \pstep m' \\ n \pstep n' }
-      { (\lambda x.m)\ n \pstep m'[n'/x] }
+      { (\lambda x.m)\ n \pstep m[n/x] }
       \rname{P-$\beta$}
@@
       \inferrule
-      { A \pstep A' \\ B \pstep B' }
-      { (x : A) \arw{s}{t} B \pstep (x : A') \arw{s}{t} B' }
-      \rname{P-Prod}
+      { m \step m' }
+      { \lambda x.m \step \lambda x.m' }
+      \rname{Step-$\lambda$}
+
+      \inferrule
+      { A \pstep A' }
+      { (x : A) \arw{s}{t} B \pstep (x : A') \arw{s}{t} B }
+      \rname{Step-ProdL}
+
+      \inferrule
+      { B \pstep B' }
+      { (x : A) \arw{s}{t} B \pstep (x : A) \arw{s}{t} B' }
+      \rname{Step-ProdR}
+
+      \inferrule
+      { m \step m' }
+      { m\ n \step m'\ n }
+      \rname{Step-AppL}
+
+      \inferrule
+      { n \step n' }
+      { m\ n \step m\ n' }
+      \rname{Step-AppR}
     \end{mathpar}
@@
   \begin{lemma} Parallel reduction satisfies the diamond property. \\
-    If $m \pstep m_1$ and $m \pstep m_2$ then there exists $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+    If $m \pstep_p m_1$ and $m \pstep_p m_2$ then there exists $m'$ such that $m_1 \pstep_p m'$ and $m_2 \pstep_p m'$.
   \end{lemma}
 
-  \begin{corollary} The transitive reflexive closure of parallel reduction is confluent. \\ 
-    If $m \pstep^* m_1$ and $m \pstep^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
+  \begin{corollary} The transitive reflexive closure of  reduction is confluent. \\ 
+    If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
   \end{corollary}
@@
   \begin{lemma} Weakening. \\
-    If $\Gamma \vdash m \stype{s} A$ is a valid, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
   \end{lemma}

### 2021-10-14 writing/draft2.tex
draft

@@
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
-Users of proof assistants based on CC often need to embed external logics \cite{bedrock} to provide additional reasoning principles for dealing with resource.
-The design and embedding of these logics is a difficult problem in its own right, requiring additional proofs to justify its soundness.
+Users of proof assistants based on CC often need to embed external logics \cite{iris} to provide additional reasoning principles for dealing with resource.
+The design and embedding of these logics are a difficult problem in their own right, requiring additional proofs to justify its soundness.
 We propose an alternative solution: extend CC with linear types.
@@
   \subsection{Equality and Reduction} \label{reduction}
-  The operational semantics, and definitional $\beta$-equality of CLC is defined reductions as presented in Figure \ref{parallel}, all of which are entirely standard.
+  The operational semantics, and definitional $\beta$-equality of CLC are presented in Figure \ref{parallel}, all of which are entirely standard.
   \begin{figure}[H]

### 2021-10-14 writing/draft2.tex
draft
clc

@@
       \indalt $(x : A) \arw{s}{t} B$
+      \indalt $(x : A) \lrw{s}{t} B$
       \indalt $\lambda x. n$ | $m\ n$
@@
   \end{figure}
-  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory is the indexed function type: $(x : A) \arw{s}{t} B$.
+  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory are the indexed function types: $(x : A) \arw{s}{t} B$, $(x : A) \lrw{s}{t} B$.
 The reason for these indices is that we have built the !
@@
     \begin{align}
-      (\_ : A) \arw{U}{U} B \quad &\Leftrightarrow \quad !A \multimap !B \\
-      (\_ : A) \arw{U}{L} B \quad &\Leftrightarrow \quad !A \multimap B \\
-      (\_ : A) \arw{L}{U} B \quad &\Leftrightarrow \quad A \multimap !B \\
-      (\_ : A) \arw{L}{L} B \quad &\Leftrightarrow \quad A \multimap B
+      (\_ : A) \arw{U}{U} B \quad &\Leftrightarrow \quad !(!A \multimap\ !B) \\
+      (\_ : A) \arw{U}{L} B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
+      (\_ : A) \arw{L}{U} B \quad &\Leftrightarrow \quad !(A \multimap\ !B) \\
+      (\_ : A) \arw{L}{L} B \quad &\Leftrightarrow \quad !(A \multimap B) \\
+      (\_ : A) \lrw{U}{U} B \quad &\Leftrightarrow \quad !A \multimap\ !B \\
+      (\_ : A) \lrw{U}{L} B \quad &\Leftrightarrow \quad !A \multimap B \\
+      (\_ : A) \lrw{L}{U} B \quad &\Leftrightarrow \quad A \multimap\ !B \\
+      (\_ : A) \lrw{L}{L} B \quad &\Leftrightarrow \quad A \multimap B
     \end{align}
@@
   Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
-MELL lacks counterparts for the cases (1), (2) if the co-domain $B$ is dependent on arguments of domain $A$.
-Though it may seem as if implications of the $!(A \multimap B)$ form are left out, CLC encodes these by endowing each function type with a universe sort through their formation judgment.
-The encoding for this example would be $\Gamma \vdash A \arw{L}{L} B \utype U_i $.
+MELL lacks counterparts for the cases (1), (2), (5), (6) if the co-domain $B$ is dependent on arguments of domain $A$.
 We will discuss function type formation in Section \ref{tyformation} in greater detail.

### 2021-10-14 writing/draft2.tex
draft

@@
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, regularity, and subject reduction. 
+    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation, and subject reduction. 
     \item We observe that CLC is highly backwards compatible with CC$\omega$. We construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent. 
@@
 
-  In practice, algorithmic type checking techniques such as bi-directional typing allow users to omit writing most of these sort indices.
+  In practice, algorithmic type checking techniques allow users to omit writing most of these sort indices.
+Our implementation employs bi-directional type checking and users never interact with indices in the surface syntax.
 
@@
         $(x : A_1) \arw{s}{t} B_1 \preceq (x : A_2) \arw{s}{t} B_2$
+      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
+        $(x : A_1) \lrw{s}{t} B_1 \preceq (x : A_2) \arw{s}{t} B_2$
     \end{enumerate}
@@
 Conversely, if $s = L$, then $m$ must be used exactly once.
-Theorem \ref{regularity} (regularity) shows that $s$ corresponds exactly to the sort of $A$'s universe.
+Theorem \ref{propagation} (propagation) shows that $s$ corresponds exactly to the sort of $A$'s universe.
 
@@
         \Gamma, x \utype A \vdash B \utype s_k }
-      { \Gamma \vdash (x : A) \arw{U}{s} B \utype t_k } 
+      { \Gamma \vdash (x : A) \arw{U}{s} B \utype U_k } 
       \rname{U-Prod}
@@
         x \notin \Gamma }
-      { \Gamma \vdash (x : A) \arw{L}{s} B \utype t_k } 
+      { \Gamma \vdash (x : A) \arw{L}{s} B \utype U_k } 
       \rname{L-Prod}
+      \\
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype U_k \\ 
+        \Gamma, x \utype A \vdash B \utype s_k }
+      { \Gamma \vdash (x : A) \lrw{U}{s} B \utype L_k } 
+      \rname{U-Lolli}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A \utype L_k \\ 
+        \Gamma \vdash B \utype s_k \\
+        x \notin \Gamma }
+      { \Gamma \vdash (x : A) \lrw{L}{s} B \utype L_k } 
+      \rname{L-Lolli}
     \end{mathpar}
@@
 
-  The \rname{U-Prod} rule is used for forming function types with non-linear domains.
+  The \rname{U-Prod} rule is used for forming non-linear function types with non-linear domains.
 This is evident from the judgment $\Gamma \vdash A \utype U_k$.
@@
 The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-From the resulting judgment $\Gamma \vdash (x : A) \arw{U}{s} B \utype t_k$, we see that $A$, and $B$'s universe sorts are used as indices to the arrow.
-Interestingly, $(x : A) \arw{U}{s} B$ is assigned to a universe $t_k$ of level $k$ and arbitrary sort $t$.
-If $t$ is chosen to be $U$, then $\Gamma \vdash (x : A) \arw{U}{s} B \utype U_k$ tells us $\lambda$-abstractions of this type have unrestricted usage.
-Conversely, if $t$ is chose to be $L$, then $\Gamma \vdash (x : A) \arw{U}{s} B \utype L_k$ tells us $\lambda$-abstractions of this type have restricted usage.
-This is how we encode the missing $!(A \multimap B)$ forms from MELL shown in Figure \ref{correspondence}.
+From the resulting judgment $\Gamma \vdash (x : A) \arw{U}{s} B \utype U_k$, we see that $A$ and $B$'s universe sorts are used as indices to the arrow, and the overall type is a non-linear type.
+The term level $\lambda$-abstractions of these types can be used unrestrictedly.
 
-  The \rname{L-Prod} rule is used for forming function types with linear domains.
+  The \rname{L-Prod} rule is used for non-linear forming function types with linear domains.
 From the judgment $\Gamma \vdash A \utype L_k$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
@@
 Like \rname{U-Prod}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-The final resulting judgment is $\Gamma \vdash (x : A) \arw{L}{s} B \utype t_k$.
+The final resulting judgment is $\Gamma \vdash (x : A) \arw{L}{s} B \utype U_k$.
 Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
-Again, we may chose $t$ to be either $U$ or $L$, which form function types for $\lambda$-abstractions with unrestricted or restricted usage respectively.
+Again, the overall resulting type a non-linear type, and the term level $\lambda$-abstractions of these types can be used unrestrictedly.
 
-  For rules \rname{U-Prod}, and \rname{L-Prod}, it may seem as if the final choice of $t_k$ does not matter.
-Indeed, the choice of $t = U$ or $t = L$ has no effect on type formation.
-But the differences will become apparent when typing $\lambda$-abstractions in Section \ref{teformation}.
+  The \rname{U-Lolli}, \rname{L-Lolli} rules are similar to \rname{U-Prod}, \rname{L-Prod} in spirit.
+The dependency considerations for domain $A$ and co-domain $B$ are exactly the same.
+The main difference between \rname{U-Lolli}, \rname{L-Lolli} and \rname{U-Prod}, \rname{L-Prod} is the universe sort of the resulting type.
+The sorts of the function types formed by \rname{U-Lolli}, \rname{L-Lolli} are $L$, meaning that they are linear function types.
+The term level $\lambda$-abstractions of these types must be used exactly once.
 
@@
       { \Gamma \vdash \lambda x . n \utype (x : A) \arw{s}{t} B }
-      \rname{U-$\lambda$}
+      \rname{Prod-$\lambda$}
       \\
@@
       \inferrule
-      { \overline{\Gamma} \vdash (x : A) \arw{s}{t} B \utype L_k \\ 
+      { \overline{\Gamma} \vdash (x : A) \lrw{s}{t} B \utype L_k \\ 
         \Gamma, x \stype{s} A \vdash n \stype{t} B }
-      { \Gamma \vdash \lambda x . n \ltype (x : A) \arw{s}{t} B }
-      \rname{L-$\lambda$}
+      { \Gamma \vdash \lambda x . n \ltype (x : A) \lrw{s}{t} B }
+      \rname{Lolli-$\lambda$}
       \\
@@
       \inferrule
-      { \Gamma_1 \vdash m \stype{t} (x : A) \arw{U}{s} B \\
+      { \Gamma_1 \vdash m \utype (x : A) \arw{U}{s} B \\
         \pure{\Gamma_2} \\
@@
       { \Gamma \vdash m\ n \stype{s} B[n/x] }
-      \rname{U-App}
+      \rname{U-Prod-App}
       \\
@@
       \inferrule
-      { \Gamma_1 \vdash m \stype{t} (x : A) \arw{L}{s} B \\
+      { \Gamma_1 \vdash m \utype (x : A) \arw{L}{s} B \\
         \Gamma_2 \vdash n \ltype A \\
@@
       { \Gamma \vdash m\ n \stype{s} B[n/x] }
-      \rname{L-App} 
+      \rname{L-Prod-App} 
+      \\
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype (x : A) \lrw{U}{s} B \\
+        \pure{\Gamma_2} \\
+        \Gamma_2 \vdash n \utype A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      \rname{U-Lolli-App}
+      \\
+
+      \inferrule
+      { \Gamma_1 \vdash m \ltype (x : A) \lrw{L}{s} B \\
+        \Gamma_2 \vdash n \ltype A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      \rname{L-Lolli-App} 
       \\
@@
 
-  As promised in Section \ref{tyformation}, the choice of $t_k$ for function type formation comes into play during the typing of $\lambda$-abstractions.
-Namely, in the pair of rules \rname{U-$\lambda$}, and \rname{L-$\lambda$}.
-
   By design, types with universe sort $U$ are non-linear, terms with these types enjoy unrestricted usage.
-In \rname{U-$\lambda$}, the function type being addressed has universe $U_i$, as seen by the judgment $\Gamma \vdash (x : A) \arw{s}{t} \utype U_i$.
+In \rname{Prod-$\lambda$}, the function type being addressed has universe $U_i$, as seen by the judgment $\Gamma \vdash (x : A) \arw{s}{t} \utype U_i$.
 $\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage.
@@
 
-  In contrast to \rname{U-$\lambda$}, the \rname{L-$\lambda$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
+  In contrast to \rname{Prod-$\lambda$}, the \rname{Lolli-$\lambda$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
 Since these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n \stype{t} B$, and lack of side condition $\pure{\Gamma}$.
 However, in judgment $\overline{\Gamma} \vdash (x : A) \arw{s}{t} B \utype L_k$ the context must be filtered, since types are not allowed to depend on restricted variables.
-The final resulting judgment $\Gamma \vdash \lambda x.n \ltype (x : A) \arw{s}{t} B$ with index $L$, asserts that the $\lambda$-abstraction must be used exactly once.
+The final resulting judgment $\Gamma \vdash \lambda x.n \ltype (x : A) \lrw{s}{t} B$ with index $L$, asserts that the $\lambda$-abstraction must be used exactly once.
 
-  For \rname{U-App}, domain $A$ is a non-linear type, as seen by its arrow index in $(x : A) \arw{U}{s} B$.
+  For \rname{U-Prod-App}, domain $A$ is a non-linear type, as seen by its arrow index in $(x : A) \arw{U}{s} B$.
 Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
@@
 
-  Now for \rname{L-App}, domain $A$ is a linear type, as seen by its arrow index in $(x : A) \arw{L}{s} B$.
+  Now for \rname{L-Prod-App}, domain $A$ is a linear type, as seen by its arrow index in $(x : A) \arw{L}{s} B$.
 Intuitively, this tells us that $x$ must be used once within the body of $m$.
@@
 The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
+
+  The rules \rname{U-Lolli-App}, \rname{L-Lolli-App} follow the same considerations as \rname{U-Prod-App}, \rname{L-Prod-App}. Additional conditions are not required for these two rules to be sound.
 
@@
       \inferrule
+      { A \pstep A' }
+      { (x : A) \lrw{s}{t} B \pstep (x : A') \lrw{s}{t} B }
+      \rname{Step-LolliL}
+
+      \inferrule
+      { B \pstep B' }
+      { (x : A) \lrw{s}{t} B \pstep (x : A) \lrw{s}{t} B' }
+      \rname{Step-LolliR}
+
+      \inferrule
       { m \step m' }
@@
 
-  The following lemmas and proofs are entirely standard. The presence of linear types do not pose any complications as reductions are untyped.
+  The following lemmas and proofs are entirely standard using parallel step technique \cite{takahashi}. The presence of linear types do not pose any complications as reductions are untyped.
 
@@
 
-  \begin{corollary} The transitive reflexive closure of  reduction is confluent. \\ 
+  \begin{lemma} Each $\pstep_p$ is reachable with $\step^*$. \\
+    If $m \pstep_p m'$ then $m \step^* m'$.
+  \end{lemma}
+
+  \begin{theorem} The transitive reflexive closure of reduction is confluent. \\ 
     If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
-  \end{corollary}
+  \end{theorem}
 
@@
   \subsubsection{Type Soundness}
-  In order to prove subject reduction, we first prove the regularity theorem.
-The main purpose of regularity is to lower types down to the term level, enabling the application of various inversion lemmas.
-Regularity also serves a second purpose, it tells us the judgment index $s$ of a term $\Gamma \vdash m \stype{s} A$ is exactly the same sort as its type's universe $\overline{\Gamma} \vdash A \utype s_i$.
+  In order to prove subject reduction, we first prove the propagation theorem.
+The main purpose of propagation is to lower types down to the term level, enabling the application of various inversion lemmas.
+Propagation also serves a second purpose, it tells us the judgment index $s$ of a term $\Gamma \vdash m \stype{s} A$ is exactly the same sort as its type's universe $\overline{\Gamma} \vdash A \utype s_i$.
 
-  \begin{theorem} Regularity. \\
+  \begin{theorem} Propagation. \\
      For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid judgment, then there exists some level $i$ such that $\overline{\Gamma} \vdash A \utype s_i$.
-    \label{regularity}
+    \label{propagation}
   \end{theorem}
 
-  With weakening, substitution, regularity, and various inversion lemmas proven, subject reduction is proved by induction on typing derivation.
+  With weakening, substitution, propagation, and various inversion lemmas proven, subject reduction can now be proved by induction on typing derivation.
 
@@
       \erase{(x : A) \arw{s}{t} B} &= (x : \erase{A}) \rightarrow \erase{B} \\
+      \erase{(x : A) \lrw{s}{t} B} &= (x : \erase{A}) \rightarrow \erase{B} \\
       \erase{\lambda x.n} &= \lambda x.\erase{n} \\

### 2021-10-15 writing/draft2.tex
wip

@@
 
+  \section{CLC in Action}
+
+  \subsection{Inductive Types}
+  A core design goal of CLC is to support an ergonomic interface for users to define both non-linear and linear inductive types.
+Inductive data should be both easy to construct, and to eliminate.
+The !
+exponential, or its adjoint alternatives often get in the way of directly using data.
+Due to CLC's usage of universe sorts instead of !, data construction and elimination becomes much.
+
+  \subsection{Resource Reasoning}
+  \subsection{Parametric Reasoning}
+  \subsection{Mutable Data Structures}
+  \subsection{Continuations}
+  \subsection{Program Synthesis}
+
   \section{Future Work}

### 2021-10-15 writing/draft3.tex
draft

--- 2021-10-15 writing/draft2.tex
@@
     The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends CC with a hierarchy of linear universes, and an indexed typing judgment that precisely controls the weakening and contraction of its term level inhabitants.
+Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
 We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
@@
 We have formalized and proven correct all major results of the core calculus in the Coq Proof Assistant.
-We extend CLC with linear inductive types and show that CLC as a programming language enables the manipulation of mutable data structures in a principled way.
+We extend CLC with linear inductive types, showing that CLC facilitates correct by construction imperative programming.
   \end{abstract}
@@
 Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean}, and others.
-These theorem provers have found great success in the fields of software verification \cite{compcert,deepspec}, and constructive mathematics \cite{four-color,schemes}.
+These theorem provers have found great success in the fields of software verification, and constructive mathematics.
 
@@
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
-Users of proof assistants based on CC often need to embed external logics \cite{iris} to provide additional reasoning principles for dealing with resource.
-The design and embedding of these logics are a difficult problem in their own right, requiring additional proofs to justify its soundness.
+Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide additional reasoning principles for dealing with resource.
+The design and embedding of these logics are a difficult problem in their own right, requiring additional proofs to justify their soundness, and effectiveness.
 We propose an alternative solution: extend CC with linear types.
@@
 
-  Work have been done to extend dependent type theories with linear types.
-Cervesato and Pfenning extends the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r \cite{vakar14} gives a categorical semantics for linear dependent types.
+  Work has been done to extend dependent type theories with linear types.
+Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
 Krishnaswami et al.
-present a dependent linear type theory \cite{neel15} based on Benton's early work of mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming the style of Hoare Type Theory \cite{htt}.
+present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
 Luo et al.
@@
 We add extra universes $L$ of linear types with cumulativity parallel to the universes $U$ of non-linear types.
-Universe information is propagated by an indexed typing judgment down to the term level, controlling the usage of weakening and contraction rules.
-This ultimately results in the \textit{linearity} theorem, stating that all resources are used exactly once.
+This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
 
-  The presence of both linear and dependent types enables CLC to write specifications that faithfully encodes the usage of resource.
+  The presence of both linear and dependent types enables CLC to write specifications that faithfully encode the usage of resources.
 The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$ as follows.
   \begin{equation*}
-    (Money\ 5) \arw{L}{L} (Camels + Marlboro)
+    (x \ltype Money\ 5) \rightarrow (Camels + Marlboro)
   \end{equation*}
@@
 This allows for a straightforward modeling of CLC in CC$\omega$, and lifting of CC$\omega$ into CLC, endowing CLC with the fruits of decades of CC research.
-We further extend CLC with inductive types, showing that as a programming language it can manipulate mutable data structures in a principled way.
+We further extend CLC with inductive types, showing that CLC facilitates correct by construction imperative programming.
 We have formalized all major results in Coq, and implemented a prototype in OCaml.
@@
     \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation, and subject reduction. 
-    \item We observe that CLC is highly backwards compatible with CC$\omega$. We construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent. 
+    \item We observe that CLC is highly backwards compatible with CC$\omega$, enabling us to construct a reduction preserving model of CLC in CC$\omega$. This shows that CLC is consistent. 
     \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
 To the best of our knowledge, our development is the first machine checked formalization of a linear dependent type theory.
-    \item Furthermore, we extend CLC with linear inductive data, demonstrating that as a programming language, CLC can safely manipulate mutable data structures.
+    \item Furthermore, we extend CLC with linear inductive data, demonstrating CLC capabilities for correct by construction imperative programming.
     \item Finally, we give an implementation extended with user definable linear and non-linear inductive types.
@@
       <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
-      \indalt $(x : A) \arw{s}{t} B$
-      \indalt $(x : A) \lrw{s}{t} B$
+      \indalt $(x :_s A) \rightarrow B$
+      \indalt $(x :_s A) \multimap B$
       \indalt $\lambda x. n$ | $m\ n$
@@
   \end{figure}
-  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory are the indexed function types: $(x : A) \arw{s}{t} B$, $(x : A) \lrw{s}{t} B$.
-The reason for these indices is that we have built the !
+  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory are the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
+The reason for these function types is that we have built the !
 exponential of linear logic directly into universe sorts.
-The indices $s$, and $t$ annotate the domain, and co-domain's respective universe sort, hence annotating linearity.
+The sort annotation $s$ here records the universe sort of the function domain.
 The behavior of !
@@
 By integrating the exponential directly into universe sorts, we implicitly limit !
-to only be used canonically.
+to only be canonically usable.
 This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing any machinery for manipulating exponential.
@@
     \begin{align}
-      (\_ : A) \arw{U}{U} B \quad &\Leftrightarrow \quad !(!A \multimap\ !B) \\
-      (\_ : A) \arw{U}{L} B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
-      (\_ : A) \arw{L}{U} B \quad &\Leftrightarrow \quad !(A \multimap\ !B) \\
-      (\_ : A) \arw{L}{L} B \quad &\Leftrightarrow \quad !(A \multimap B) \\
-      (\_ : A) \lrw{U}{U} B \quad &\Leftrightarrow \quad !A \multimap\ !B \\
-      (\_ : A) \lrw{U}{L} B \quad &\Leftrightarrow \quad !A \multimap B \\
-      (\_ : A) \lrw{L}{U} B \quad &\Leftrightarrow \quad A \multimap\ !B \\
-      (\_ : A) \lrw{L}{L} B \quad &\Leftrightarrow \quad A \multimap B
+      (\_ \utype A) \rightarrow B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
+      (\_ \ltype A) \rightarrow B \quad &\Leftrightarrow \quad !(A \multimap B) \\
+      (\_ \utype A) \multimap B \quad &\Leftrightarrow \quad !A \multimap B \\
+      (\_ \ltype A) \multimap B \quad &\Leftrightarrow \quad A \multimap B
     \end{align}
@@
   Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
-MELL lacks counterparts for the cases (1), (2), (5), (6) if the co-domain $B$ is dependent on arguments of domain $A$.
+MELL lacks counterparts for the cases (1), (3) if the co-domain $B$ is dependent on arguments of domain $A$.
 We will discuss function type formation in Section \ref{tyformation} in greater detail.
 
-  In practice, algorithmic type checking techniques allow users to omit writing most of these sort indices.
+  In practice, algorithmic type checking techniques allow users to omit writing sort indices.
 Our implementation employs bi-directional type checking and users never interact with indices in the surface syntax.
@@
       \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
-        $(x : A_1) \arw{s}{t} B_1 \preceq (x : A_2) \arw{s}{t} B_2$
+        $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{s} A_2) \rightarrow B_2$
       \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
-        $(x : A_1) \lrw{s}{t} B_1 \preceq (x : A_2) \arw{s}{t} B_2$
+        $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{s} A_2) \multimap B_2$
     \end{enumerate}
@@
   \subsection{Typing Judgment}
-  Typing judgments in CLC take on the form of $\Gamma \vdash m \stype{s} A$.
-Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$, with free variables typed in $\Gamma$.
-The sort index $s$ tells us the linearity of $m$.
-Specifically, if $s = U$, then $m$ has unrestricted usage.
-Conversely, if $s = L$, then $m$ must be used exactly once.
-Theorem \ref{propagation} (propagation) shows that $s$ corresponds exactly to the sort of $A$'s universe.
+  Typing judgments in CLC take on the form of $\Gamma \vdash m : A$. Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$, with free variables typed in $\Gamma$.
 
@@
     \begin{enumerate}
-      \item A \textit{type} $A$ is \textit{non-linear} under context $\Gamma$ if $\Gamma \vdash A \utype U_i$.
-      \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A \utype L_k$.
-      \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \utype A$. A unrestricted term may be used an arbitrary number of times.
-      \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if it has type $A$, and $\Gamma \vdash m \ltype A$ A restricted term must be used exactly once.
+      \item A \textit{type} $A$ is \textit{non-linear} under context $\Gamma$ if $\Gamma \vdash A : U_i$.
+      \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A : L_k$.
+      \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$, such that $\Gamma \vdash m : A$.
+A unrestricted term may be used an arbitrary number of times.
+      \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$, such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
     \end{enumerate}

### 2021-10-15 writing/draft3.tex
draft

@@
       { \pure{\Gamma} }
-      { \Gamma \vdash U_* \utype U_0 } 
+      { \Gamma \vdash U_* : U_0 } 
       \rname{Prop-Axiom}
@@
       { \pure{\Gamma} }
-      { \Gamma \vdash U_k \utype U_{k+1} } 
-      \rname{U-Axiom}
+      { \Gamma \vdash s_k : U_{k+1} } 
+      \rname{Sort-Axiom}
 
       \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash L_k \utype U_{k+1} } 
-      \rname{L-Axiom}
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_i \\ 
+        \Gamma, x \utype A \vdash B : U_* }
+      { \Gamma \vdash (x \utype A) \rightarrow B : U_* } 
+      \rname{Prop}
       \\
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash A \utype U_i \\ 
-        \Gamma, x \utype A \vdash B \utype U_* }
-      { \Gamma \vdash (x : A) \arw{U}{U} B \utype U_* } 
-      \rname{U-Prop}
-      \\
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A \utype U_k \\ 
-        \Gamma, x \utype A \vdash B \utype s_k }
-      { \Gamma \vdash (x : A) \arw{U}{s} B \utype U_k } 
+        \Gamma \vdash A : U_k \\ 
+        \Gamma, x \utype A \vdash B : s_k }
+      { \Gamma \vdash (x \utype A) \rightarrow B : U_k } 
       \rname{U-Prod}
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash A \utype L_k \\ 
-        \Gamma \vdash B \utype s_k \\
+        \Gamma \vdash A : L_k \\ 
+        \Gamma \vdash B : s_k \\
         x \notin \Gamma }
-      { \Gamma \vdash (x : A) \arw{L}{s} B \utype U_k } 
+      { \Gamma \vdash (x \ltype A) \rightarrow B : U_k } 
       \rname{L-Prod}
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash A \utype U_k \\ 
-        \Gamma, x \utype A \vdash B \utype s_k }
-      { \Gamma \vdash (x : A) \lrw{U}{s} B \utype L_k } 
+        \Gamma \vdash A : U_k \\ 
+        \Gamma, x \utype A \vdash B : s_k }
+      { \Gamma \vdash (x \utype A) \multimap B : L_k } 
       \rname{U-Lolli}
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash A \utype L_k \\ 
-        \Gamma \vdash B \utype s_k \\
+        \Gamma \vdash A : L_k \\ 
+        \Gamma \vdash B : s_k \\
         x \notin \Gamma }
-      { \Gamma \vdash (x : A) \lrw{L}{s} B \utype L_k } 
+      { \Gamma \vdash (x \ltype A) \multimap B : L_k } 
       \rname{L-Lolli}
@@
 From \rname{L-Axiom} we can see that the universe of linear types $L_k$ is an inhabitant of $U_{k+1}$.
-This is reminiscent of Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
+This is inspired by Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
 
-  The \rname{U-Prop} rule is used for forming propositions.
-From the judgment $\Gamma \vdash A \utype U_i$ we can see that \rname{U-Prop} allows for quantification over non-linear types of arbitrary level, hence impredicative.
-The judgment $\Gamma, x \utype A \vdash B \utype U_*$ asserts that the co-domain must be in the impredicative universe $U_*$.
-The final resulting judgment $\Gamma \vdash (x : A) \arw{U}{U} B \utype U_*$ is indexed by $U$, indicating that the \textit{type} $(x : A) \arw{U}{U} B$ can be use unrestrictedly as a \textit{term}.
-Since $(x : A) \arw{U}{U} B$ belongs to universe $U_*$, its \textit{terms} also enjoys unrestricted usage.
+  The \rname{Prop} rule is used for forming propositions.
+From the judgment $\Gamma \vdash A : U_i$ we can see that \rname{Prop} allows impredicative quantification over non-linear types of arbitrary level $i$.
+The judgment $\Gamma, x \utype A \vdash B : U_*$ asserts that the co-domain must be in the impredicative universe $U_*$.
+The final resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_*$ has sort $U$, indicating that terms with type $(x \utype A) \rightarrow B$ enjoy unrestricted usage.
 
   The \rname{U-Prod} rule is used for forming non-linear function types with non-linear domains.
-This is evident from the judgment $\Gamma \vdash A \utype U_k$.
-Since $A$ is in the predicative non-linear universe $U_k$, terms of type $A$ have unrestricted usage.
-The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B \utype s_k$.
+This is evident from the judgment $\Gamma \vdash A : U_k$.
+Due to the fact that $A$ is in the predicative non-linear universe $U_k$, terms of type $A$ have unrestricted usage.
+The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_k$.
 The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-From the resulting judgment $\Gamma \vdash (x : A) \arw{U}{s} B \utype U_k$, we see that $A$ and $B$'s universe sorts are used as indices to the arrow, and the overall type is a non-linear type.
+From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_k$, we see that the overall type is a non-linear type.
 The term level $\lambda$-abstractions of these types can be used unrestrictedly.
 
-  The \rname{L-Prod} rule is used for non-linear forming function types with linear domains.
-From the judgment $\Gamma \vdash A \utype L_k$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
-Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B \utype s_k$.
+  The \rname{L-Prod} rule is used for forming non-linear function types with linear domains.
+From the judgment $\Gamma \vdash A : L_k$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
+Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B : s_k$.
 Like \rname{U-Prod}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-The final resulting judgment is $\Gamma \vdash (x : A) \arw{L}{s} B \utype U_k$.
+The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_k$.
 Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
-Again, the overall resulting type a non-linear type, and the term level $\lambda$-abstractions of these types can be used unrestrictedly.
+Again, the overall resulting type is non-linear, so the term level $\lambda$-abstractions of these types can be used unrestrictedly.
 
@@
 The main difference between \rname{U-Lolli}, \rname{L-Lolli} and \rname{U-Prod}, \rname{L-Prod} is the universe sort of the resulting type.
-The sorts of the function types formed by \rname{U-Lolli}, \rname{L-Lolli} are $L$, meaning that they are linear function types.
+The sorts of the function types formed by \rname{U-Lolli}, \rname{L-Lolli} are $L$, meaning they are linear function types.
 The term level $\lambda$-abstractions of these types must be used exactly once.
@@
       { \pure{\Gamma} }
-      { \Gamma, x \utype A \vdash x \utype A } 
+      { \Gamma, x \utype A \vdash x : A } 
       \rname{U-Var}
@@
       { \pure{\Gamma} }
-      { \Gamma, x \ltype A \vdash x \ltype A } 
+      { \Gamma, x \ltype A \vdash x : A } 
       \rname{L-Var}
@@
       { \pure{\Gamma} \\
-        \Gamma \vdash (x : A) \arw{s}{t} B \utype U_i \\ 
-        \Gamma, x \stype{s} A \vdash n \stype{t} B }
-      { \Gamma \vdash \lambda x . n \utype (x : A) \arw{s}{t} B }
+        \Gamma \vdash (x \stype{s} A) \rightarrow B : t_i \\ 
+        \Gamma, x \stype{s} A \vdash n : B }
+      { \Gamma \vdash \lambda x . n : (x \stype{s} A) \rightarrow B }
       \rname{Prod-$\lambda$}
@@
       \inferrule
-      { \overline{\Gamma} \vdash (x : A) \lrw{s}{t} B \utype L_k \\ 
-        \Gamma, x \stype{s} A \vdash n \stype{t} B }
-      { \Gamma \vdash \lambda x . n \ltype (x : A) \lrw{s}{t} B }
+      { \overline{\Gamma} \vdash (x \stype{s} A) \multimap B : t_i \\ 
+        \Gamma, x \stype{s} A \vdash n : B }
+      { \Gamma \vdash \lambda x . n : (x \stype{s} A) \multimap B }
       \rname{Lolli-$\lambda$}
@@
       \inferrule
-      { \Gamma_1 \vdash m \utype (x : A) \arw{U}{s} B \\
+      { \Gamma_1 \vdash m : (x \utype A) \rightarrow B \\
         \pure{\Gamma_2} \\
-        \Gamma_2 \vdash n \utype A \\
+        \Gamma_2 \vdash n : A \\
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      { \Gamma \vdash m\ n : B[n/x] }
       \rname{U-Prod-App}
@@
       \inferrule
-      { \Gamma_1 \vdash m \utype (x : A) \arw{L}{s} B \\
-        \Gamma_2 \vdash n \ltype A \\
+      { \Gamma_1 \vdash m : (x \ltype A) \rightarrow B \\
+        \Gamma_2 \vdash n : A \\
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      { \Gamma \vdash m\ n : B[n/x] }
       \rname{L-Prod-App} 
@@
       \inferrule
-      { \Gamma_1 \vdash m \ltype (x : A) \lrw{U}{s} B \\
+      { \Gamma_1 \vdash m : (x \utype A) \multimap B \\
         \pure{\Gamma_2} \\
-        \Gamma_2 \vdash n \utype A \\
+        \Gamma_2 \vdash n : A \\
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      { \Gamma \vdash m\ n : B[n/x] }
       \rname{U-Lolli-App}
@@
       \inferrule
-      { \Gamma_1 \vdash m \ltype (x : A) \lrw{L}{s} B \\
-        \Gamma_2 \vdash n \ltype A \\
+      { \Gamma_1 \vdash m : (x \ltype A) \multimap B \\
+        \Gamma_2 \vdash n : A \\
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash m\ n \stype{s} B[n/x] }
+      { \Gamma \vdash m\ n : B[n/x] }
       \rname{L-Lolli-App} 
@@
       \inferrule
-      { \Gamma \vdash m \stype{s} A \\
-        \overline{\Gamma} \vdash B \utype s_i \\ A \preceq B }
-      { \Gamma \vdash m \stype{s} B } 
+      { \Gamma \vdash m : A \\
+        \overline{\Gamma} \vdash B : s_i \\ A \preceq B }
+      { \Gamma \vdash m : B } 
       \rname{Conv}
@@
 The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma, x \ltype A$ with linear type $A$.
-For both rules, the side condition $\pure{\Gamma}$ forbids irrelevant variables with linear type from occurring within $\Gamma$.
-This prevents weakening variables with linear type.
+For both rules, the side condition $\pure{\Gamma}$ forbids irrelevant variables of linear types from occurring within $\Gamma$.
+This prevents another vector of weakening variables with linear type.
 
-  By design, types with universe sort $U$ are non-linear, terms with these types enjoy unrestricted usage.
-In \rname{Prod-$\lambda$}, the function type being addressed has universe $U_i$, as seen by the judgment $\Gamma \vdash (x : A) \arw{s}{t} \utype U_i$.
-$\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage.
+  In \rname{Prod-$\lambda$}, the function type being addressed has form $(x \stype{s} A) \rightarrow B$.
+$\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
 This consideration is realized by the side condition $\pure{\Gamma}$, asserting all variables in context $\Gamma$ are unrestricted.
-Next, the body of the abstraction $n$ is typed as $\Gamma, x \stype{s} A \vdash n \stype{t} B$, where $s$ is the sort of $A$ and $t$ is the sort of $B$.
-Finally, the resulting judgment $\Gamma \vdash \lambda x.n \utype (x : A) \arw{s}{t} B$ with index $U$, asserts that the $\lambda$-abstraction can be used unrestrictedly.
+Next, the body of the abstraction $n$ is typed as $\Gamma, x \stype{s} A \vdash n : B$, where $s$ is the sort of $A$.
+Finally, the resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \rightarrow B$ asserts that the $\lambda$-abstraction can be used unrestrictedly.
 
   In contrast to \rname{Prod-$\lambda$}, the \rname{Lolli-$\lambda$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
-Since these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n \stype{t} B$, and lack of side condition $\pure{\Gamma}$.
-However, in judgment $\overline{\Gamma} \vdash (x : A) \arw{s}{t} B \utype L_k$ the context must be filtered, since types are not allowed to depend on restricted variables.
-The final resulting judgment $\Gamma \vdash \lambda x.n \ltype (x : A) \lrw{s}{t} B$ with index $L$, asserts that the $\lambda$-abstraction must be used exactly once.
+Due to the fact these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
+However, in judgment $\overline{\Gamma} \vdash (x \stype{s} A) \multimap B : L_k$ the context must be filtered, because types are not allowed to depend on restricted variables.
+The final resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \multimap B$ asserts that the $\lambda$-abstraction must be used exactly once.
 
-  For \rname{U-Prod-App}, domain $A$ is a non-linear type, as seen by its arrow index in $(x : A) \arw{U}{s} B$.
+  For \rname{U-Prod-App}, domain $A$ is a non-linear type, as seen by its annotation in $(x \utype A) \rightarrow B$.
 Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
@@
 
-  Now for \rname{L-Prod-App}, domain $A$ is a linear type, as seen by its arrow index in $(x : A) \arw{L}{s} B$.
+  Now for \rname{L-Prod-App}, domain $A$ is a linear type, as seen by its annotation in $(x \ltype A) \rightarrow B$.
 Intuitively, this tells us that $x$ must be used once within the body of $m$.
@@
 
-  Finally, the \rname{Conv} rule allows judgment $\Gamma \vdash m \stype{s} A$ to convert to judgment $\Gamma \vdash m \stype{s} B$, if $B$ is a valid type in context $\overline{\Gamma}$, and of the same sort $s$ as $A$.
+  Finally, the \rname{Conv} rule allows judgment $\Gamma \vdash m : A$ to convert to judgment $\Gamma \vdash m : B$ if $B$ is a valid type in context $\overline{\Gamma}$.
 Futhermore, $A$ must be a subtype of $B$ satisfying the cumulativity relation $A \preceq B$.
-This rule gives rise to large eliminations, as computations embedded at the type level can convert to canonical types.
+This rule gives rise to large eliminations (compute types from terms), as computations embedded at the type level can convert to canonical types.
 
@@
       { A \pstep A' }
-      { (x : A) \arw{s}{t} B \pstep (x : A') \arw{s}{t} B }
+      { (x \stype{s} A) \rightarrow B \step (x \stype{s} A') \rightarrow B }
       \rname{Step-ProdL}
@@
       { B \pstep B' }
-      { (x : A) \arw{s}{t} B \pstep (x : A) \arw{s}{t} B' }
+      { (x \stype{s} A) \rightarrow B \step (x \stype{s} A) \rightarrow B' }
       \rname{Step-ProdR}
@@
       { A \pstep A' }
-      { (x : A) \lrw{s}{t} B \pstep (x : A') \lrw{s}{t} B }
+      { (x \stype{s} A) \multimap B \step (x \stype{s} A') \multimap B }
       \rname{Step-LolliL}
@@
       { B \pstep B' }
-      { (x : A) \lrw{s}{t} B \pstep (x : A) \lrw{s}{t} B' }
+      { (x \stype{s} A) \multimap B \step (x \stype{s} A) \multimap B' }
       \rname{Step-LolliR}
@@
   \end{figure}
-  As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain a simple operational semantics, whose $\beta$-reductions are very well behaved.
+  As we have discussed previously, the elimination of the explicit !
+exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
 
@@
 Next, we show that CLC is a valid linear type theory through the \textit{linearity} theorem.
+Furthermore, we show that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
 Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, showing that CLC is strongly normalizing.
@@
   \begin{lemma} Weakening. \\
-    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, then for any $x \notin \Gamma$, judgment $\Gamma, x \utype B \vdash m \stype{s} A$ is derivable.
+    If $\Gamma \vdash m : A$ is a valid judgment, then for any $x \notin \Gamma$, the judgment $\Gamma, x \utype B \vdash m : A$ is derivable.
   \end{lemma}
@@
 Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
-He defines additional syntax, and semantics for the intricate unboxing of !
+He defines additional syntax and semantics for the intricate unboxing of !
 terms, solving the lack of substitute in Abramsky's term calculus \cite{abramsky1993}.
 
-  Our design of integrating ! into universe sorts removes the need for ! manipulating syntax, and semantics. The substitution lemma is directly proved by induction on typing derivations.
+  Our design of integrating ! into universe sorts removes the need for ! manipulating syntax and semantics. The following substitution lemmas are directly proved by induction on typing derivations.
 
-  \begin{lemma} Substitution. \\
-    For $\Gamma_1, x \stype{s} A \vdash m \stype{t} B$ and $\Gamma_2 \vdash v \stype{s} A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is defined for some $\Gamma$, then $\Gamma \vdash [v/x]m \stype{t} [v/x]B$.
+  \begin{lemma} Non-linear Substitution. \\
+    For $\Gamma_1, x \utype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ are valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
+  \end{lemma}
+
+  \begin{lemma} Linear Substitution. \\
+    For $\Gamma_1, x \ltype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
   \end{lemma}
@@
 The main purpose of propagation is to lower types down to the term level, enabling the application of various inversion lemmas.
-Propagation also serves a second purpose, it tells us the judgment index $s$ of a term $\Gamma \vdash m \stype{s} A$ is exactly the same sort as its type's universe $\overline{\Gamma} \vdash A \utype s_i$.
 
   \begin{theorem} Propagation. \\
-     For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid judgment, then there exists some level $i$ such that $\overline{\Gamma} \vdash A \utype s_i$.
+     For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exists some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
     \label{propagation}
@@
   \begin{theorem} Subject reduction. \\
-    For $\Gamma \vdash m \stype{s} A$, if $m \pstep n$ then $\Gamma \vdash n \stype{s} A$.
+    For $\Gamma \vdash m : A$, if $m \pstep n$ then $\Gamma \vdash n : A$.
   \end{theorem}
@@
   \begin{lemma} Narity. \\
-    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
+    If $\Gamma \vdash m : A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
   \end{lemma}
@@
   \begin{theorem} Linearity. \\
-    If $\Gamma \vdash m \stype{s} A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+    If $\Gamma \vdash m : A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+  \end{theorem}
+
+  \subsubsection{Promotion and Dereliction}
+  In most presentations of linear logic, there is no clear distinction between linear and non-linear objects.
+Non-linear objects are created by packing linear objects under non-linear contexts into the !
+exponential through the \textit{promotion} rule.
+Conversely, the \textit{dereliction} rule unpacks !
+objects so they can be used in place of linear ones.
+Linear type theories \cite{abramsky1993,dill,vakar14} based on !
+will contain introduction and elimination rules for !
+types that serve similar purpose to \textit{promotion} and \textit{dereliction}.
+
+  The linear types of CLC are not designed around packing and unpacking !, so there are no explicit rules for \textit{promotion} and \textit{dereliction}.
+However, \textit{promotion} and \textit{dereliction} for canonical types are derivable as theorems through $\eta$-expansion.
+
+  \begin{theorem} Promotion \\
+    If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
+  \end{theorem}
+
+  \begin{theorem} Dereliction \\
+    If $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$ is a valid judgment, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$ is derivable.
   \end{theorem}
@@
       \erase{L_k} &= Type_k \\
-      \erase{(x : A) \arw{s}{t} B} &= (x : \erase{A}) \rightarrow \erase{B} \\
-      \erase{(x : A) \lrw{s}{t} B} &= (x : \erase{A}) \rightarrow \erase{B} \\
+      \erase{(x \stype{s} A) \rightarrow B} &= (x : \erase{A}) \rightarrow \erase{B} \\
+      \erase{(x \stype{s} A) \multimap B} &= (x : \erase{A}) \rightarrow \erase{B} \\
       \erase{\lambda x.n} &= \lambda x.\erase{n} \\
@@
   \begin{theorem} \label{embedding} Embedding. \\
-    For any CLC context $\Gamma$, terms $m$, $A$, and sort $s$, if $\Gamma \vdash m \stype{s} A$ is a valid typing judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid a typing judgment in CC$\omega$.
+    If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
   \end{theorem}
@@
 
-  \section{CLC in Action}
+  % \section{CLC in Action}
 
-  \subsection{Inductive Types}
-  A core design goal of CLC is to support an ergonomic interface for users to define both non-linear and linear inductive types.
+  % \subsection{Inductive Types}
+  % A core design goal of CLC is to support an ergonomic interface for users to define both non-linear and linear inductive types.
 Inductive data should be both easy to construct, and to eliminate.
@@
 
-  \subsection{Resource Reasoning}
-  \subsection{Parametric Reasoning}
-  \subsection{Mutable Data Structures}
-  \subsection{Continuations}
-  \subsection{Program Synthesis}
+  % \subsection{Resource Reasoning}
+  % \subsection{Parametric Reasoning}
+  % \subsection{Mutable Data Structures}
+  % \subsection{Continuations}
+  % \subsection{Program Synthesis}
 

### 2021-10-16 writing/draft3.tex
draft

@@
 \newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
-\newcommand{\arw}[2]
-{\hspace*{0.2em}{\scriptstyle #1}\hspace*{-0.2em}
-\rightarrow
-\hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
-\newcommand{\lrw}[2]
-{\hspace*{0.2em}{\scriptstyle #1}\hspace*{-0.2em}
-\multimap
-\hspace*{-0.2em}{\scriptstyle #2}\hspace*{0.2em}}
 \newcommand{\erase}[1]{\llbracket #1 \rrbracket}
@@
   \subsubsection{Promotion and Dereliction}
-  In most presentations of linear logic, there is no clear distinction between linear and non-linear objects.
-Non-linear objects are created by packing linear objects under non-linear contexts into the !
+  In most presentations of linear logic, non-linear objects are created by packing linear objects under non-linear contexts into the !
 exponential through the \textit{promotion} rule.
-Conversely, the \textit{dereliction} rule unpacks !
-objects so they can be used in place of linear ones.
-Linear type theories \cite{abramsky1993,dill,vakar14} based on !
-will contain introduction and elimination rules for !
-types that serve similar purpose to \textit{promotion} and \textit{dereliction}.
+Conversely, the \textit{dereliction} rule unpacks non-linear objects so they can be used in place of linear ones.
+Linear type theories \cite{abramsky1993,dill,vakar14} utilizing exponential for non-linearity will contain !
+types and rules that are equivalent to \textit{promotion} and \textit{dereliction}.
 
-  The linear types of CLC are not designed around packing and unpacking !, so there are no explicit rules for \textit{promotion} and \textit{dereliction}.
-However, \textit{promotion} and \textit{dereliction} for canonical types are derivable as theorems through $\eta$-expansion.
+  Linear types of CLC are not designed around packing and unpacking !, so there are no rules for \textit{promotion} and \textit{dereliction}.
+Arbitrary computations existing at the type level also muddle the association between which types can be promoted or derelicted to which.
+Nevertheless, \textit{promotion} and \textit{dereliction} for canonical types are derivable as theorems through $\eta$-expansion.
 
@@
 
-  For the following theorem, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
-  \begin{theorem} \label{preserve} Erasure preserves reduction. \\
+  For the following lemma, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
+  \begin{lemma} \label{preserve} Erasure preserves reduction. \\
     For any CLC terms $m$, and $n$, if there is $m \pstep_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \pstep_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
-  \end{theorem}
+  \end{lemma}
 
-  \begin{theorem} \label{embedding} Embedding. \\
+  \begin{lemma} \label{embedding} Embedding. \\
     If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
-  \end{theorem}
+  \end{lemma}
 
-  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}.
+  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of lemmas \ref{preserve}, and \ref{embedding}.
 This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.

### 2021-10-16 writing/draft3.tex
draft

@@
   \subsubsection{Promotion and Dereliction}
-  In most presentations of linear logic, non-linear objects are created by packing linear objects under non-linear contexts into the !
-exponential through the \textit{promotion} rule.
+  In most presentations of linear logic, non-linear objects are created by packing linear objects into the !
+exponential using the \textit{promotion} rule.
 Conversely, the \textit{dereliction} rule unpacks non-linear objects so they can be used in place of linear ones.

### 2021-10-16 writing/lics22.tex
lics22

--- 2021-10-16 writing/draft3.tex
@@
 % document layout
-\documentclass{article}
-\usepackage[margin=1.2in]{geometry}
+\documentclass[sigplan,screen,review,authordraft]{acmart}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsmath}
-\usepackage{amssymb}
 \usepackage{mdwtab}
@@
 
-\floatstyle{plain} 
-\restylefloat{figure}
-
 \newtheorem{theorem}{Theorem}[section]
@@
 \renewcommand{\syntright}{}
-\setlength{\grammarparsep}{10pt}
-\setlength{\grammarindent}{10em}
+\setlength{\grammarparsep}{7pt}
+\setlength{\grammarindent}{7em}
 \newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
@@
 \author{Qiancheng Fu}
+\affiliation{
+  \institution{Boston University}
+  \city{Boston}
+  \state{MA}
+  \country{USA}
+}
 
@@
 \begin{document}
-  \maketitle 
   \begin{abstract}
@@
   \end{abstract}
+  \maketitle
 
@@
 
-  We propose a new linear dependent type system - The Calculus of Linear Constructions (CLC).
+  We present a new linear dependent type system - The Calculus of Linear Constructions (CLC).
 CLC extends CC$\omega$ with linear types.
 CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
-We add extra universes $L$ of linear types with cumulativity parallel to the universes $U$ of non-linear types.
+We add an extra universe sort $L$ of linear types with cumulativity parallel to the sort $U$ of non-linear types.
 This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
@@
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-    \item Next, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation, and subject reduction. 
-    \item We observe that CLC is highly backwards compatible with CC$\omega$, enabling us to construct a reduction preserving model of CLC in CC$\omega$. This shows that CLC is consistent. 
+    \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation, and subject reduction.
+    \item Next, we prove the linearity theorem, showing that the tangible impact linear types have how terms can be structured.
+We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
+    \item Additionally, we observe that CLC is highly backwards compatible with CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$.
+This shows that CLC is consistent.
     \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
 To the best of our knowledge, our development is the first machine checked formalization of a linear dependent type theory.
-    \item Furthermore, we extend CLC with linear inductive data, demonstrating CLC capabilities for correct by construction imperative programming.
+    \item Furthermore, we extend CLC with linear inductive data, demonstrating CLC's capabilities for correct by construction imperative programming.
     \item Finally, we give an implementation extended with user definable linear and non-linear inductive types.
@@
   \subsection{Syntax}
+  \begin{figure}[h]
+    \vspace{-1em}
+    \caption{Syntax}
+    \vspace{1.5em}
+    \begin{grammar}
+      <$k$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} predicative levels
+
+      <$i$> := * | 0 | 1 | 2 ... \hspace*{3em} all levels
+
+      <$s, t$> ::= $U$ | $L$ \phantom{| $x$} \hspace*{4.6em} sorts
+
+      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
+      \indalt $(x :_s A) \rightarrow B$
+      \indalt $(x :_s A) \multimap B$
+      \indalt $\lambda x. n$ | $m\ n$
+    \end{grammar}
+    \label{syntax}
+  \end{figure}
   The syntax of the core type theory is presented in Figure \ref{syntax}.
@@
 $U_k$, and $L_k$ are the predicative universes of non-linear, and linear types respectively.
+  \begin{figure}[H]
+    \vspace{-1em}
+    \caption{Correspondence of CLC types and MELL implications}
+    \Description{}
+    \vspace{-1em}
+    \begin{align}
+      (\_ \utype A) \rightarrow B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
+      (\_ \ltype A) \rightarrow B \quad &\Leftrightarrow \quad !(A \multimap B) \\
+      (\_ \utype A) \multimap B \quad &\Leftrightarrow \quad !A \multimap B \\
+      (\_ \ltype A) \multimap B \quad &\Leftrightarrow \quad A \multimap B
+    \end{align}
+    \vspace{-1em}
+    \label{correspondence}
+  \end{figure}
 
-  \begin{figure}[H]
-    \caption{Syntax}
-    \centering
-    \begin{minipage}{0.8\linewidth}
-    \begin{grammar}
-      <$k$> := 0 | 1 | 2 ... | \phantom{*} \hspace*{5.8em} predicative levels
-
-      <$i$> := * | 0 | 1 | 2 ... \hspace*{5.8em} all levels
-
-      <$s, t$> ::= $U$ | $L$ \phantom{| $x$} \hspace*{8em} sorts
-
-      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{7.3em} expressions
-      \indalt $(x :_s A) \rightarrow B$
-      \indalt $(x :_s A) \multimap B$
-      \indalt $\lambda x. n$ | $m\ n$
-    \end{grammar}
-    \end{minipage}
-    \label{syntax}
-  \end{figure}
   A clear departure of our language from standard presentations of both linear type theory, and dependent type theory are the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
@@
 to only be canonically usable.
-This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without needing any machinery for manipulating exponential.
-  \begin{figure}[H]
-    \caption{Correspondence of CLC types and MELL implications}
-    \begin{align}
-      (\_ \utype A) \rightarrow B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
-      (\_ \ltype A) \rightarrow B \quad &\Leftrightarrow \quad !(A \multimap B) \\
-      (\_ \utype A) \multimap B \quad &\Leftrightarrow \quad !A \multimap B \\
-      (\_ \ltype A) \multimap B \quad &\Leftrightarrow \quad A \multimap B
-    \end{align}
-    \label{correspondence}
-  \end{figure}
+This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
+
   Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
@@
       \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
-      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
+      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
         $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{s} A_2) \rightarrow B_2$
-      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, then
+      \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
         $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{s} A_2) \multimap B_2$
@@
 However, $L_k$ cumulates to $L_{k+1}$.
-These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of its occupants' linearity.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of its linearity.
 
   \begin{figure}[H]
+    \vspace{-1.6em}
     \caption{The Universe Hierarchy}
+    \vspace{0.8em}
     \centering
@@
     \end{tikzcd}
+    \vspace{-1em}
     \label{universe}
@@
 
-  \begin{figure}[H]
+  \begin{figure}[h]
+    \vspace{-1em}
     \caption{Structural Judgments}
+    \vspace{-0.5em}
     \begin{mathpar}
@@
     \end{mathpar}
+    \vspace{-0.8em}
     \label{structural}
@@
   \begin{figure}[H]
+    \vspace{-1em}
     \caption{Type Formation} 
@@
     \end{mathpar}
+    \vspace{-1em}
     \label{type}
@@
 
-  \begin{figure}[H]
+  \begin{figure}[h]
     \caption{Term Formation}
@@
     \end{mathpar}
+    \vspace{-1em}
     \label{term}
@@
   The operational semantics, and definitional $\beta$-equality of CLC are presented in Figure \ref{parallel}, all of which are entirely standard.
-  \begin{figure}[H]
+  \begin{figure}[h]
     \caption{Equality and Parallel Reduction}
@@
 
-  \begin{lemma} Parallel reduction satisfies the diamond property. \\
-    If $m \pstep_p m_1$ and $m \pstep_p m_2$ then there exists $m'$ such that $m_1 \pstep_p m'$ and $m_2 \pstep_p m'$.
+  \begin{lemma} 
+    Parallel reduction satisfies the diamond property. If $m \pstep_p m_1$ and $m \pstep_p m_2$ then there exists $m'$ such that $m_1 \pstep_p m'$ and $m_2 \pstep_p m'$.
   \end{lemma}
 
-  \begin{lemma} Each $\pstep_p$ is reachable with $\step^*$. \\
-    If $m \pstep_p m'$ then $m \step^* m'$.
+  \begin{lemma} 
+    Each $\pstep_p$ is reachable with $\step^*$. If $m \pstep_p m'$ then $m \step^* m'$.
   \end{lemma}
 
-  \begin{theorem} The transitive reflexive closure of reduction is confluent. \\ 
-    If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
+  \begin{theorem} 
+    The transitive reflexive closure of reduction is confluent. If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
   \end{theorem}
@@
 
-  \subsubsection{Weakening} \label{subst}
+  \subsubsection{Weakening} \label{weakening}
   CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.
 
-  \begin{lemma} Weakening. \\
-    If $\Gamma \vdash m : A$ is a valid judgment, then for any $x \notin \Gamma$, the judgment $\Gamma, x \utype B \vdash m : A$ is derivable.
+  \begin{lemma} 
+    Weakening. If $\Gamma \vdash m : A$ is a valid judgment, then for any $x \notin \Gamma$, the judgment $\Gamma, x \utype B \vdash m : A$ is derivable.
   \end{lemma}
@@
   \subsubsection{Substitution} \label{subst}
-  Though the substitution lemma is widely considered a boring and bureaucratic theorem, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
-Much of this difficulty arise during the substitution of arguments containing !
+  Though the substitution lemma is considered a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
+Much of the difficulty arise during the substitution of arguments containing !
 exponential.
@@
 He defines additional syntax and semantics for the intricate unboxing of !
-terms, solving the lack of substitute in Abramsky's term calculus \cite{abramsky1993}.
+terms, solving the lack of substitute in Abramsky's term calculus.
 
@@
 
-  \begin{lemma} Non-linear Substitution. \\
+  \begin{lemma} 
+    Non-linear Substitution.
 For $\Gamma_1, x \utype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ are valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
@@
 
-  \begin{lemma} Linear Substitution. \\
-    For $\Gamma_1, x \ltype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
+  \begin{lemma} 
+    Linear Substitution. For $\Gamma_1, x \ltype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
   \end{lemma}
@@
 
-  \begin{theorem} Propagation. \\
+  \begin{theorem} 
+    Propagation.
 For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exists some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
@@
 
-  \begin{theorem} Subject reduction. \\
-    For $\Gamma \vdash m : A$, if $m \pstep n$ then $\Gamma \vdash n : A$.
+  \begin{theorem} 
+    Subject Reduction. For $\Gamma \vdash m : A$, if $m \pstep n$ then $\Gamma \vdash n : A$.
   \end{theorem}
@@
 
-  \begin{lemma} Narity. \\
-    If $\Gamma \vdash m : A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
+  \begin{lemma} 
+    Narity. If $\Gamma \vdash m : A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
   \end{lemma}
@@
 
-  \begin{theorem} Linearity. \\
-    If $\Gamma \vdash m : A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+  \begin{theorem} 
+    Linearity. If $\Gamma \vdash m : A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
   \end{theorem}
@@
 
-  Linear types of CLC are not designed around packing and unpacking !, so there are no rules for \textit{promotion} and \textit{dereliction}.
+  Linear types of CLC are not obtained through packing and unpacking !, so there are no rules for \textit{promotion} and \textit{dereliction}.
 Arbitrary computations existing at the type level also muddle the association between which types can be promoted or derelicted to which.
@@
 
-  \begin{theorem} Promotion \\
+  \begin{theorem} 
+    Promotion.
 If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
@@
 
-  \begin{theorem} Dereliction \\
-    If $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$ is a valid judgment, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$ is derivable.
+  \begin{theorem} 
+    Dereliction. If $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$ is a valid judgment, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$ is derivable.
   \end{theorem}
@@
   We prove the following lemma to commute erasure and substitution whenever needed.
-  \begin{lemma} Erasure commutes with substitution. \\ 
-    For any CLC terms $m$, $n$, and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
+  \begin{lemma} 
+    Erasure commutes with substitution. For any CLC terms $m$, $n$, and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
   \end{lemma}
@@
   For the following lemma, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
-  \begin{lemma} \label{preserve} Erasure preserves reduction. \\
+  \begin{lemma} \label{preserve} 
+    Erasure preserves reduction.
 For any CLC terms $m$, and $n$, if there is $m \pstep_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \pstep_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
@@
 
-  \begin{lemma} \label{embedding} Embedding. \\
-    If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
+  \begin{lemma} \label{embedding} 
+    Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
   \end{lemma}

### 2021-10-16 writing/lics22.tex
lics22

@@
   \section{Introduction}
-  The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand, and Huet in their landmark work \cite{cc}.
+  The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
 In CC types can depend on terms, allowing one to write precise specifications as types.
-Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean}, and others.
+Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
 These theorem provers have found great success in the fields of software verification, and constructive mathematics.
@@
   Linear Logic is a substructural logic introduced by Girard in his seminal work \cite{girard}.
-Girard notice that the weakening and contraction rules of Classical Logic when restricted carefully, gives rise to a new logical foundation for reasoning about resource.
-Wadler \cite{wadler1990,wadler1991} first notice that an analogous restriction to variable usage in simple type theory leads to a linear type theory, where terms respect resources.
+Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
+Wadler \cite{wadler1990,wadler1991} first notices that an analogous restriction to variable usage in simple type theory leads to a linear type theory, where terms respect resources.
 A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
@@
 Luo et al.
-introduce the property of essential linearity, and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear, and affine types within a unified framework.
+introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
 The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
 
-  We present a new linear dependent type system - The Calculus of Linear Constructions (CLC).
+  We present a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
 CLC extends CC$\omega$ with linear types.
@@
   \end{equation*}
-  This new specification for transaction states that it requires a payment of 5 units of money, the customer is relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
+  This new specification for transaction demands a payment of 5 units of money, and the customer be relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
 
@@
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-    \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation, and subject reduction.
+    \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation and subject reduction.
     \item Next, we prove the linearity theorem, showing that the tangible impact linear types have how terms can be structured.
@@
   \begin{figure}[h]
-    \vspace{-1em}
+    \vspace{-1.4em}
     \caption{Syntax}
-    \vspace{1.5em}
+    \vspace{1.4em}
     \begin{grammar}
@@
 Here, $U_*$ is the impredicative universe of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
-$U_k$, and $L_k$ are the predicative universes of non-linear, and linear types respectively.
+$U_k$, and $L_k$ are the predicative universes of non-linear and linear types respectively.
   \begin{figure}[H]
-    \vspace{-1em}
+    \vspace{-1.4em}
     \caption{Correspondence of CLC types and MELL implications}
@@
     \end{align}
-    \vspace{-1em}
+    \vspace{-1.6em}
     \label{correspondence}
@@
 
-  A clear departure of our language from standard presentations of both linear type theory, and dependent type theory are the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
+  A clear departure of our language from standard presentations of both linear type theory and dependent type theory is the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
 The reason for these function types is that we have built the !
@@
 The behavior of !
-is difficult to account for even in simple linear type theory.
+is difficult to account for even in simple linear type theories.
 Subtle issues arise if !!
@@
   \subsection{Universes and Cumulativity}
-  CLC features two sorts of universes $U$, and $L$ with level indices $*, 0, 1, 2, \cdots$.
+  CLC features two sorts of universes $U$ and $L$ with level indices $*, 0, 1, 2, \cdots$.
 $U_*$ is the impredicative universe of propositions.
-$U_k$, and $L_k$ are the predicative universe of non-linear types, and linear types respectively.
+$U_k$ and $L_k$ are the predicative universes of non-linear types and linear types respectively.
 The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
@@
     \end{tikzcd}
-    \vspace{-1em}
+    \vspace{-0.5em}
     \label{universe}
@@
 For linear variables, the $\mrg{\_}{\_}{\_}$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
-This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for unrestricted variables whilst forbidding it for restricted.
+This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for unrestricted variables whilst forbidding it for restricted ones.
 
@@
     \end{mathpar}
-    \vspace{-0.8em}
+    \vspace{-1em}
     \label{structural}
@@
 
-  \begin{definition} We formally define the terms \textit{non-linear}, \textit{linear}, \textit{unrestricted}, and \textit{restricted}.
+  \begin{definition} We formally define the terms \textit{non-linear}, \textit{linear}, \textit{unrestricted} and \textit{restricted}.
     \begin{enumerate}
@@
   \subsection{Equality and Reduction} \label{reduction}
-  The operational semantics, and definitional $\beta$-equality of CLC are presented in Figure \ref{parallel}, all of which are entirely standard.
+  The operational semantics and definitional $\beta$-equality of CLC are presented in Figure \ref{parallel}, all of which are entirely standard.
   \begin{figure}[h]
@@
 To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
-We given a hand written version of the proof in the appendix as well.
+We give a hand written version of the proof in the appendix as well.
 
@@
   \subsubsection{Substitution} \label{subst}
-  Though the substitution lemma is considered a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
-Much of the difficulty arise during the substitution of arguments containing !
+  Though the substitution lemma is regarded as a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
+Much of the difficulty arises during the substitution of arguments containing !
 exponential.
@@
     Propagation.
-For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exists some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
+For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exist some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
     \label{propagation}
@@
 
-  With weakening, substitution, propagation, and various inversion lemmas proven, subject reduction can now be proved by induction on typing derivation.
+  With weakening, substitution, propagation and various inversion lemmas proven, subject reduction can now be proved by induction on typing derivation.
 
@@
   At this point, we have proven that CLC is type sound.
-However, we still need to prove that the removal of weakening, and contraction for restricted variables yield tangible impact on the structure of terms.
+However, we still need to prove that the removal of weakening and contraction for restricted variables yields tangible impact on the structure of terms.
 For this purpose, we define a binding aware recursive function $occurs(x, m)$ that counts the number of times variable $x$ appears within term $m$.
@@
 
-  For cases with branched syntax in the linearity theorem, such as application $(m\ n)$, the $occurs$ function sums up the occurrences of variable $x$ in branches $m$, and $n$.
+  For cases with branched syntax in the linearity theorem, such as application $(m\ n)$, the $occurs$ function sums up the occurrences of variable $x$ in branches $m$ and $n$.
 The narity lemma is used to prove that either $occurs(x, m) = 1 \wedge occurs(x, n) = 0$ or $occurs(x, m) = 0 \wedge occurs(x, n) = 1$ is true.
@@
   \subsubsection{Strong Normalization}
-  The strong normalization theorem of CLC is proven by construction of a typing, and reduction preserving erasure function to CC$\omega$.
+  The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function to CC$\omega$.
 We assume familiarity with CC$\omega$ syntax here, and define the erasure function as follows.

### 2021-10-16 writing/lics22.tex
lics22

@@
 }
+\email{qcfu@bu.edu}
 
@@
 
-      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
+      <$m, n, A, B$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
       \indalt $(x :_s A) \rightarrow B$
@@
     \begin{enumerate}
-      \item $\preceq$ is a partial order with respect to definitional equality.
+      \item $\preceq$ is a partial order with respect to equality.
         \begin{enumerate}
@@
 However, $L_k$ cumulates to $L_{k+1}$.
-These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universe, and subsequently losing track of its linearity.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes, and subsequently losing track of linearity.
 
@@
       \rname{U-Prod-App}
-      \\
 
@@
       \rname{U-Lolli-App}
-      \\
 

### 2021-10-17 writing/lics22/lics22.tex
writing
lics22

--- 2021-10-16 writing/lics22.tex
@@
 % document layout
-\documentclass[sigplan,screen,review,authordraft]{acmart}
+%\documentclass[sigplan,screen,review,authordraft]{acmart}
+\documentclass[sigplan,screen]{acmart}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{listings}
+\usepackage{lstlangcoq}
 \usepackage{float}
@@
 
+\lstset{basicstyle=\scriptsize,language=Coq,showstringspaces=false}
 \newtheorem{theorem}{Theorem}[section]
@@
 \newcommand{\rname}[1]{\textsc{\footnotesize #1}}
-\newcommand{\LNLD}{\text{LNL}_\text{D}}
-\newcommand{\ok}{\textbf{ok}}
 \newcommand{\pure}[1]{|#1|}
@@
 \newcommand{\set}{\text{set}}
-\newcommand{\subst}{\text{subst}}
 \newcommand{\utype}{:_{\scriptscriptstyle U}}
@@
 
-  \subsubsection{Weakening} \label{weakening}
+  \subsubsection{Weakening} \label{weakening}ß
   CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.
@@
 
+  \section{Extensions}
+  \subsection{Inductive Types}
+  Readers familiar with the literature on linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
+Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written this way will be incredibly unwieldy.
+
+  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic}.
+The presence of universe sorts $U$ and $L$ allows users to define inductive types to be non-linear or linear.
+The lack of !
+exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching.
+The $\otimes$ connective becomes a definable inductive type.
+
+  \begin{figure}[h]
+  \vspace{-1.4em}
+  \caption{Inductive Type Definition and Elimination}
+  \begin{lstlisting}
+  (* Sigma Definition *)
+  Inductive Sigma (A: Type) (F: A -> Type): Type :=
+  | pair : (x: A) -> F x -> Sigma A F.
+
+  (* Tensor Definition *)
+  Inductive Tensor (A: Linear) (B: Linear): Linear :=
+  | tpair: A -> B -> Tensor A B.
+
+  (* FTensor Definition *)
+  Inductive FTensor (A: Type) (F: A -> Linear): Linear :=
+  | fpair: (x: A) -> F x -> FTensor A F.
+
+  (* Sigma Elimination
+     x and y are unrestricted in clause body *)
+  match m with
+  | pair x y => (* clause body *)
+  end
+
+  (* Tensor Elimination
+     x and y must be used once in clause body *)
+  match m with
+  | tpair x y => (* clause body *)
+  end
+
+  (* FTensor Elimination
+     x is unrestricted in clause body, 
+     y must be used once in clause body *)
+  match m with
+  | fpair x y => (* clause body *)
+  end
+  \end{lstlisting}
+  \label{inductive}
+  \Description{}
+  \end{figure}
+
+  Figure \ref{inductive} is an excerpt taken from our implementation's prelude, with syntax drawing heavily from Coq.
+
+  \begin{itemize}
+    \item \texttt{Sigma} defines the standard dependent sum type.
+Since first component of the pair constructor is of non-linear type, the type of the second component is allowed to depend on the first component.
+    \item \texttt{Tensor} defines the $\otimes$ connective for CLC.
+Both components of the \texttt{tpair} constructor are of linear types.
+When a \texttt{Tensor} term is eliminated through pattern-matching, its first and second components must be used exactly once in the body of its pattern clause.
+    \item \texttt{FTensor} fulfills a role similar to Krishnaswami et al.'s $F$ adjoint connective \cite{neel15}.
+The type of \texttt{FTensor}'s first component is non-linear.
+The linear type of its second component is allowed to depend on the first component.
+Due to the fact that \texttt{FTensor}'s second component is linear, \texttt{FTensor} must be linear as well, otherwise duplication of restricted variables may occur.
+  \end{itemize}
+
+  Due to the complexity of checking soundness for dependent linear type definitions (such as \texttt{FTensor}) and pattern-matching based elimination, we have not verified the algorithm employed by our implementation yet.
+
   % \section{CLC in Action}
@@
 \bibliographystyle{acm}
-\bibliography{ref}
+\bibliography{../ref}
 

### 2021-10-18 writing/lics22/lics22.tex
lics22

@@
 \newcommand{\refl}{\text{refl}}
-\newcommand{\letin}[3]{$\text{let }#1\text{ := }#2\text{ in }#3$}
+\newcommand{\letin}[3]{\ $\text{let }#1\text{ := }#2\text{ in }#3$\ }
+\newcommand{\ind}{\ $\text{ind}$\ }
+\newcommand{\iter}{\ $\text{iter}$\ }
 \newcommand{\new}{\text{new}}
@@
 
-      <$m, n, A, B$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
+      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
       \indalt $(x :_s A) \rightarrow B$
@@
   \section{Extensions}
-  \subsection{Inductive Types}
   Readers familiar with the literature on linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
-Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written this way will be incredibly unwieldy.
+Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written in this style will be incredibly difficult to use.
 
-  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic}.
-The presence of universe sorts $U$ and $L$ allows users to define inductive types to be non-linear or linear.
+  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic} as an extension to the core theory.
+The presence of universe sorts $U$ and $L$ allows users to declare inductive types to be non-linear or linear.
 The lack of !
 exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching.
-The $\otimes$ connective becomes a definable inductive type.
+The $\otimes$ connective becomes definable as an inductive type.
 
   \begin{figure}[h]
-  \vspace{-1.4em}
+  \vspace{-1.2em}
   \caption{Inductive Type Definition and Elimination}
@@
 
-  Due to the complexity of checking soundness for dependent linear type definitions (such as \texttt{FTensor}) and pattern-matching based elimination, we have not verified the algorithm employed by our implementation yet.
+  Due to the complexity of checking soundness for dependent linear type definitions (such as \texttt{FTensor}) and pattern-matching based elimination, we have not verified the algorithm employed by our implementation.
 
-  % \section{CLC in Action}
+  Figure \ref{formation}, Figure \ref{introduction} and Figure \ref{elimination} presents the formation, introduction and elimination rules for select inductive types respectively.
+We will refer to these inductive types during the construction of examples.
 
-  % \subsection{Inductive Types}
-  % A core design goal of CLC is to support an ergonomic interface for users to define both non-linear and linear inductive types.
-Inductive data should be both easy to construct, and to eliminate.
-The !
-exponential, or its adjoint alternatives often get in the way of directly using data.
-Due to CLC's usage of universe sorts instead of !, data construction and elimination becomes much.
+  \begin{figure}[h]
+    \vspace{-0.8em}
+    \caption{Select Formation Rules}
+    \begin{mathpar}
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash \mathbf{1} : U_0 }
+      \rname{$1$-Form}
 
-  % \subsection{Resource Reasoning}
-  % \subsection{Parametric Reasoning}
-  % \subsection{Mutable Data Structures}
-  % \subsection{Continuations}
-  % \subsection{Program Synthesis}
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash \mathbb{N} : U_0 }
+      \rname{$\mathbb{N}$-Form}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_i \\
+        \Gamma \vdash m : A \\
+        \Gamma \vdash n : A }
+      { \Gamma \vdash m =_A n : U_* }
+      \rname{=-Form}
+
+      \inferrule
+      { \pure{\Gamma} \\ 
+        \Gamma \vdash A : U_k \\ 
+        \Gamma x \utype A \vdash B : U_k  }
+      { \Gamma \vdash \Sigma x:A.B : U_k }
+      \rname{$\Sigma$-Form}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : L_k \\
+        \Gamma \vdash B : L_k }
+      { \Gamma \vdash A \otimes B : L_k }
+      \rname{$\otimes$-Form}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_k \\ 
+        \Gamma x \utype A \vdash B : L_k }
+      { \Gamma \vdash F x:A.B : L_k }
+      \rname{$F$-Form}
+    \end{mathpar}
+    \vspace{-0.4em}
+    \label{formation}
+    \Description{}
+  \end{figure}
+
+  Of all the rules presented here, perhaps the most surprising to readers familiar with linear types or dependent types is the \rname{$\mathbb{N}$-L-Elim} rule.
+This rule constructs a term with linear type through primitive recursion.
+The $\multimap$ arrows prevent duplication of the O-case element of linear type $C\ O$, whereas the S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
+
+  \begin{figure}[h]
+    \vspace{-0.8em}
+    \caption{Select Introduction Rules}
+    \begin{mathpar}
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash () : \mathbf{1} }
+      \rname{$1$-Intro}
+
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash O : \mathbb{N} }
+      \rname{O-Intro}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash n : \mathbb{N} }
+      { \Gamma \vdash S\ n : \mathbb{N} }
+      \rname{S-Intro}
+
+      \inferrule
+      { \pure{\Gamma} \\ 
+        \Gamma \vdash A : U_i \\
+        \Gamma \vdash m : A }
+      { \Gamma \vdash \refl_A m : m =_A m }
+      \rname{=-Intro}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash m : A \\
+        \Gamma \vdash n : B[m/x] \\\\
+        \Gamma \vdash \Sigma x:A.B : U_k }
+      { \Gamma \vdash (m, n) : \Sigma x:A.B }
+      \rname{$\Sigma$-Intro}
+
+      \inferrule
+      { \Gamma_1 \vdash m : A \\
+        \Gamma_2 \vdash n : B \\ 
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\\\
+        \Gamma \vdash A \otimes B : L_k }
+      { \Gamma \vdash (m, n) : A \otimes B }
+      \rname{$\otimes$-Intro}
+
+      \inferrule
+      { \pure{\Gamma_1} \\
+        \Gamma_1 \vdash m : A \\
+        \Gamma_2 \vdash n : B[m/x] \\\\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+        \Gamma \vdash F x:A.B : L_k }
+      { \Gamma \vdash (m, n) : F x:A.B }
+      \rname{$F$-Intro}
+    \end{mathpar}
+    \vspace{-2em}
+    \label{introduction}
+    \Description{}
+  \end{figure}
+
+  \begin{figure}[h]
+    \vspace{-0.8em}
+    \caption{Select Elimination Rules}
+    \begin{mathpar}
+      \inferrule
+      { \Gamma_1 \vdash m : \mathbf{1} \\ 
+        \Gamma_2 \vdash n : A \\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash \letin{()}{m}{n} : A }
+      \rname{$1$-Elim}
+
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash \iter_\mathbb{N}^U :
+        (C \utype \mathbb{N} \rightarrow U_i) \rightarrow C\ O \rightarrow \\
+        ((n \utype \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)) \rightarrow
+        (n \utype \mathbb{N}) \rightarrow C\ n }
+      \rname{$\mathbb{N}$-U-Elim}
+
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash \iter_\mathbb{N}^L :
+        (C \utype \mathbb{N} \rightarrow L_k) \rightarrow C\ O \multimap \\
+        ((n \utype \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)) \multimap
+        (n \utype \mathbb{N}) \multimap C\ n }
+      \rname{$\mathbb{N}$-L-Elim}
+
+      \inferrule
+      { \Gamma_1 \vdash m : \Sigma x:A.B \\ 
+        \Gamma_2, x \utype A, y \utype B \vdash n : C \\\\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
+      \rname{$\Sigma$-Elim}
+
+      \inferrule
+      { \Gamma_1 \vdash m : A \otimes B \\ 
+        \Gamma_2, x \ltype A, y \ltype B \vdash n : C \\\\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
+      \rname{$\otimes$-Elim}
+
+      \inferrule
+      { \Gamma_1 \vdash m : F x:A.B \\
+        \Gamma_2, x \utype A, y \ltype B \vdash n : C \\\\
+        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
+      \rname{$F$-Elim}
+    \end{mathpar}
+    \vspace{-0.8em}
+    \label{elimination}
+    \Description{}
+  \end{figure}
+
+  \section{Applications}
+  In this section, we demonstrate some of the possible applications of CLC through examples.
+We collapse the sort annotations on arrows $(x \utype A) \rightarrow B$ and $(x \ltype A) \rightarrow B$ into $(x : A) \rightarrow B$ whenever $A$'s sort is inferable from context.
+
+  \subsection{Stateful Reasoning}
+  We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
+  \begin{definition}
+    State axioms.
+    \begin{itemize}
+      \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U_k) \rightarrow A \rightarrow L_k$ \\
+      The $\mapsto$ is a linear type constructor.
+It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
+Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
+      \item $\new : (A : U_k) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
+      $\new$ is a function that allocates a new memory store for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $l \mapsto_A m$.
+      \item $\free : (A : U_k) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \mathbf{1}$ \\
+      $\free$ is a function that de-allocates the memory store at location $l$ which is currently in use.
+This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the $(l \mapsto_A m)$ component is a proof that location $l$ is currently allocated.
+      \item $\get : (A : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
+      $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
+The type of $\get$'s output is the rather complicated $F \_: (\Sigma x.
+x =_A m).
+(l \mapsto_A m)$.
+This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed.
+The proof $(l \mapsto_A m)$ is returned unchanged as well.
+      \item $\set : (A\ B : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
+      $\set$ has the most complicated type of all the axioms presented here.
+Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original stored term.
+This is the so called "strong update".
+The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ encode this update process.
+    \end{itemize}
+  \end{definition}
 

### 2021-10-19 writing/lics22/lics22.tex
lics22

@@
 \newcommand{\iter}{\ $\text{iter}$\ }
-\newcommand{\new}{\text{new}}
-\newcommand{\free}{\text{free}}
-\newcommand{\get}{\text{get}}
-\newcommand{\set}{\text{set}}
+\newcommand{\unit}{\text{unit}}
+\newcommand{\new}{\textbf{new}}
+\newcommand{\free}{\textbf{free}}
+\newcommand{\get}{\textbf{get}}
+\newcommand{\set}{\textbf{set}}
 \newcommand{\utype}{:_{\scriptscriptstyle U}}
@@
   Readers familiar with the literature on linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
-Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written in this style will be incredibly difficult to use.
+Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, but proofs and programs written in this style will be incredibly difficult to maintain.
 
@@
       { \pure{\Gamma} }
-      { \Gamma \vdash \mathbf{1} : U_0 }
-      \rname{$1$-Form}
+      { \Gamma \vdash \unit : U_0 }
+      \rname{Unit-Form}
 
@@
       { \pure{\Gamma} }
-      { \Gamma \vdash () : \mathbf{1} }
-      \rname{$1$-Intro}
+      { \Gamma \vdash () : \unit }
+      \rname{Unit-Intro}
 
@@
         \Gamma \vdash A \otimes B : L_k }
-      { \Gamma \vdash (m, n) : A \otimes B }
+      { \Gamma \vdash [m, n] : A \otimes B }
       \rname{$\otimes$-Intro}
@@
         \Gamma \vdash F x:A.B : L_k }
-      { \Gamma \vdash (m, n) : F x:A.B }
+      { \Gamma \vdash \{m, n\} : F x:A.B }
       \rname{$F$-Intro}
@@
       \inferrule
-      { \Gamma_1 \vdash m : \mathbf{1} \\ 
+      { \Gamma_1 \vdash m : \unit \\ 
         \Gamma_2 \vdash n : A \\
@@
       { \Gamma \vdash \letin{()}{m}{n} : A }
-      \rname{$1$-Elim}
+      \rname{Unit-Elim}
 
@@
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
+      { \Gamma \vdash \letin{$[x, y]$}{$m$}{$n$} \vdash C }
       \rname{$\otimes$-Elim}
@@
         \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
+      { \Gamma \vdash \letin{$\{x, y\}$}{$m$}{$n$} \vdash C }
       \rname{$F$-Elim}
@@
   \end{definition}
+  Due to the fact that $(l \mapsto_A m)$ is a linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
+This is the same non-sharing principle that empowers Separation Logic.
+Unlike Separation Logic, $(l \mapsto_A m)$ is a first-class object of CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures etc.
+
+  \subsection{Internalization of Effect Laws}
+  The combination of linear types and dependent types within the state axioms allow the internalization of the so called "effect laws" for state, where these laws are provable theorems within the CLC framework.
+
+  \begin{enumerate}
+    \item Performing $\get$ twice on the same address will produce two terms that are provably equal.
+    \item After using $\set$ to update the store at an address $l$ to term $m$, performing $\get$ again on the same address will retrieve $m$.
+  \end{enumerate}
+
+  Figure \ref{law1} is an excerpt taken from our implementation.
+It demonstrates the internalization of Law 1.
+The usage of implicit parameters allows us to omit much of the redundant arguments that could be inferred from the types of others.
+Here, $\new$ allocates a store for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
+When $\get$ is applied to $c$ twice, the returned equality proofs $pf1$ and $pf2$ allow one to create proof $pf$ that the returned values $x$ and $y$ are equal.
+Finally, due to $c$ having linear type, it is required for it to be freed at the very end.
+
+  \begin{figure}[h]
+    \vspace{-1em}
+    \caption{Law 1}
+    \begin{lstlisting}
+    let [ _, c ] := new _ 1 in
+    let [ xeq, c ] := get _ _ _ c in
+    let [ yeq, c ] := get _ _ _ c in
+    let ( x, pf1 ) := xeq in
+    let ( y, pf2 ) := yeq in
+    let pf1 := Eq_sym _ _ _ pf1 in
+    let pf : Eq _ x y := Eq_trans _ _ _ _ pf1 pf2 in
+    free _ _ _ c.
+    \end{lstlisting}
+    \vspace{-1em}
+    \label{law1}
+    \Description{}
+  \end{figure}
+
+  Figure \ref{law2} demonstrates the internalization of Law 2.
+Again, implicit parameters fill in the annoying bureaucratic arguments when they can be inferred.
+Here, a store is allocated by $\new$ for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
+After using $\set$ to update the store at address $l$ to 2 and then performing $\get$ on $l$, one can prove that the value retrieved by $\get$ is exactly equal to 2.
+
+  \begin{figure}[h]
+    \vspace{-1em}
+    \caption{Law 2}
+    \begin{lstlisting}
+    let [ _, c ] := new _ 1 in
+    let c := set _ _ _ _ c 2 in
+    let [ zeq, c ] := get _ _ _ c in
+    let ( z, pf1 ) := zeq in
+    let pf : Eq _ 2 z := pf1 in
+    free _ _ _ c.
+    \end{lstlisting}
+    \vspace{-1em}
+    \label{law2}
+    \Description{}
+  \end{figure}
+
+  Though we have not implemented program extraction for CLC yet, our preliminary experiments hint at the possibility of compiling the state axioms to constant time imperative operations after proof erasure.
+This can greatly increase the runtime efficiency of CLC over pure functional languages when used for programming.
+More examples are available in the first author's Github repository.
+
+  \subsection{Mutable Data}
+  Immutable data in functional programming languages are easy to reason about due their values staying constant over their lifespan.
+However, this convenience comes at the price of efficiency when attempting to perform "updates" on existing data.
+In order to update immutable data, an entirely new copy must be created with new entries substituted in for the original ones.
+So after update, there are two data structures in memory, the original one and the updated one.
+A garbage collector is necessary to reclaim memory allocated for the original copy.
+
+  Due to the non-sharing property of linear types, it is possible to safely update linear data in-place without encountering synchronization issues.
+Garbage collection is no longer necessary as extra copies of data are not created.
+Languages such as ATS that implement mutable linear data have tiny runtime environments when compared to immutable garbage collection based ones such as OCaml or Haskell.
+
+  \begin{figure}[h]
+    \vspace{-1em}
+    \caption{Linear Lists}
+    \begin{lstlisting}
+    Inductive list (A : Type) : Linear :=
+    | nil : list A
+    | cons : A -> list A -> list A.
+
+    Fixpoint append (A : Type) : list A -o list A -o 
+      (list A -o list A) -o list A :=
+      fun ls1 ls2 k =>
+        match ls1 with
+        | nil => k ls2
+        | (cons h t) as _cons_ => 
+          append _ t ls2 (fun res => k (_cons_ h res))
+        end.
+    \end{lstlisting}
+    \label{list}
+    \vspace{-1em}
+    \Description{}
+  \end{figure}
+
+  Figure \ref{list} presents a CPS append function for linear lists.
+In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
+Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
+Intuitively, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
 
   \section{Future Work}
+  The marriage of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
+
+  On the theoretical side, we intend to investigate the semantic models of CLC, with the goal of explaining the seemingly natural cohesion between its rules.
+The impredicative nature of the \textit{Prop}-like universe $U_*$ hints at the possibility of extending term dependency to terms of linear types in a philosophically satisfying way.
+Linear types also appears to augment parametricity, which may produce stronger free theorems.
+
+  On the application side, we intend to bolster our implementation with more features, making it a full-blown theorem prover/programming language.
+We aim to fully verify the typechecking algorithms and extensions employed by the implementation.
+The unification algorithm used to solve implicit parameters is extremely ad-hoc, as the problem of unification for linear types has not been thoroughly explored.
+We hope to improve this situation.
+
   \section{Conclusion}
+  The Calculus of Linear Constructions is a lightweight extension of the Calculus of Constructions with linear types.
+As a type theory, it is a sound and effective framework for reasoning about resources.
+As a programing language, it allows one to write safe and efficient code, alleviating some of the issues faced by standard functional programming.
 

### 2021-10-19 writing/lics22/lics22.tex
lics22

@@
   \section{Extensions}
-  Readers familiar with the literature on linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
-Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, but proofs and programs written in this style will be incredibly difficult to maintain.
+  Readers familiar with the literature of linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
+Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written in this style will be incredibly difficult to maintain.
 
@@
 The linear type of its second component is allowed to depend on the first component.
-Due to the fact that \texttt{FTensor}'s second component is linear, \texttt{FTensor} must be linear as well, otherwise duplication of restricted variables may occur.
+Due to the fact that \texttt{FTensor}'s second component is linear, the entire \texttt{FTensor} must be linear as well, otherwise duplication of restricted variables may occur.
   \end{itemize}
@@
   Figure \ref{formation}, Figure \ref{introduction} and Figure \ref{elimination} presents the formation, introduction and elimination rules for select inductive types respectively.
-We will refer to these inductive types during the construction of examples.
+We will use these inductive types during the construction of examples.
 
@@
   Of all the rules presented here, perhaps the most surprising to readers familiar with linear types or dependent types is the \rname{$\mathbb{N}$-L-Elim} rule.
-This rule constructs a term with linear type through primitive recursion.
-The $\multimap$ arrows prevent duplication of the O-case element of linear type $C\ O$, whereas the S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
+This rule constructs a term with linear type through a primitive recursion principle.
+The $\multimap$ arrows protect the O-case element of linear type $C\ O$ from duplication by currying.
+The S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
 
@@
       \item $\new : (A : U_k) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
-      $\new$ is a function that allocates a new memory store for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $l \mapsto_A m$.
+      $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
       \item $\free : (A : U_k) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \mathbf{1}$ \\
-      $\free$ is a function that de-allocates the memory store at location $l$ which is currently in use.
-This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the $(l \mapsto_A m)$ component is a proof that location $l$ is currently allocated.
+      $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use.
+This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
       \item $\get : (A : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
@@
 (l \mapsto_A m)$.
-This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed.
+This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored.
 The proof $(l \mapsto_A m)$ is returned unchanged as well.
@@
       $\set$ has the most complicated type of all the axioms presented here.
-Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original stored term.
-This is the so called "strong update".
-The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ encode this update process.
+Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term.
+This is the so called "strong update" operation.
+The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
     \end{itemize}
@@
 This is the same non-sharing principle that empowers Separation Logic.
-Unlike Separation Logic, $(l \mapsto_A m)$ is a first-class object of CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures etc.
+Unlike Separation Logic, $(l \mapsto_A m)$ is a first-class object of CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
 
   \subsection{Internalization of Effect Laws}
-  The combination of linear types and dependent types within the state axioms allow the internalization of the so called "effect laws" for state, where these laws are provable theorems within the CLC framework.
+  The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CLC framework.
 
@@
 It demonstrates the internalization of Law 1.
-The usage of implicit parameters allows us to omit much of the redundant arguments that could be inferred from the types of others.
+The usage of implicit parameters allows one to omit writing much of the redundant arguments that could be inferred from context.
 Here, $\new$ allocates a store for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-When $\get$ is applied to $c$ twice, the returned equality proofs $pf1$ and $pf2$ allow one to create proof $pf$ that the returned values $x$ and $y$ are equal.
+When $\get$ is applied to $c$ twice, the equality proofs \textit{pf1} $: (1 =_\mathbb{N} x)$ and \textit{pf2} $: (1 =_\mathbb{N} y)$ generated by $\get$ allow one to create proof \textit{pf} $: (x =_\mathbb{N} y)$ that the retrieved values $x$ and $y$ are equal.
 Finally, due to $c$ having linear type, it is required for it to be freed at the very end.
@@
 Here, a store is allocated by $\new$ for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-After using $\set$ to update the store at address $l$ to 2 and then performing $\get$ on $l$, one can prove that the value retrieved by $\get$ is exactly equal to 2.
+After using $\set$ to update the store at $l$ to 2 and then performing $\get$ on $l$, one can create proof \textit{pf} $: (2 =_\mathbb{N} z)$ that the value $z$ retrieved by $\get$ is exactly equal to 2.
 
@@
 Garbage collection is no longer necessary as extra copies of data are not created.
-Languages such as ATS that implement mutable linear data have tiny runtime environments when compared to immutable garbage collection based ones such as OCaml or Haskell.
+Languages such as ATS that implement mutable linear data have tiny runtimes when compared to immutable garbage collector based ones such as OCaml or Haskell, making them suitable for programming in computationally constrained settings such as embedded programming.
 
@@
 Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
-Intuitively, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
+Basically, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
 
   \section{Future Work}
-  The marriage of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
-
-  On the theoretical side, we intend to investigate the semantic models of CLC, with the goal of explaining the seemingly natural cohesion between its rules.
-The impredicative nature of the \textit{Prop}-like universe $U_*$ hints at the possibility of extending term dependency to terms of linear types in a philosophically satisfying way.
-Linear types also appears to augment parametricity, which may produce stronger free theorems.
-
-  On the application side, we intend to bolster our implementation with more features, making it a full-blown theorem prover/programming language.
-We aim to fully verify the typechecking algorithms and extensions employed by the implementation.
-The unification algorithm used to solve implicit parameters is extremely ad-hoc, as the problem of unification for linear types has not been thoroughly explored.
-We hope to improve this situation.
+  The integration of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
+
+  On the theoretical side, we intend to investigate the semantic models of CLC, with the goal of explaining the natural cohesion between its rules.
+The impredicative nature of the \textit{Prop}-like universe $U_*$ hints at the possibility of extending dependency to terms of linear types in a philosophically satisfying way.
+Linear types also appear to augment parametricity, which can produce stronger free theorems.
+
+  On the application side, we intend to bolster our implementation with more features, making it a full-featured theorem prover/programming language.
+We aim to fully verify the typechecking algorithms and extensions employed by the current implementation.
+The unification algorithm used to resolve implicit parameters is extremely ad-hoc, as the problem of unification for linear types has not thoroughly studied, we hope to improve this situation.
 
@@
 As a type theory, it is a sound and effective framework for reasoning about resources.
-As a programing language, it allows one to write safe and efficient code, alleviating some of the issues faced by standard functional programming.
+As a programing language, it allows one to write safe and efficient code, alleviating some of the common issues faced by standard functional programming.
 

### 2021-10-19 writing/lics22/lics22.tex
lics22

@@
 % document layout
-%\documentclass[sigplan,screen,review,authordraft]{acmart}
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,screen,review,authordraft]{acmart}
+%\documentclass[sigplan,screen]{acmart}
 \usepackage[utf8]{inputenc}
@@
 This is the same non-sharing principle that empowers Separation Logic.
-Unlike Separation Logic, $(l \mapsto_A m)$ is a first-class object of CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
+Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
 

### 2021-10-19 writing/lics22/lics22.tex
lics22

@@
 \newcommand{\erase}[1]{\llbracket #1 \rrbracket}
+\newcommand{\lift}[1]{\llparenthesis #1 \rrparenthesis}
 
@@
   \begin{lemma} 
-    Erasure commutes with substitution. For any CLC terms $m$, $n$, and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
+    Erasure commutes with substitution. For any CLC terms $m$, $n$ and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
   \end{lemma}
 
-  For the following lemma, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
-  \begin{lemma} \label{preserve} 
+  For the following lemma, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
+  \begin{theorem} \label{preserve} 
     Erasure preserves reduction.
-For any CLC terms $m$, and $n$, if there is $m \pstep_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \pstep_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
-  \end{lemma}
-
-  \begin{lemma} \label{embedding} 
+For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  \end{theorem}
+
+  \begin{theorem} \label{embedding} 
     Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
-  \end{lemma}
-
-  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of lemmas \ref{preserve}, and \ref{embedding}.
+  \end{theorem}
+
+  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}.
 This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
@@
     Well-typed CLC terms are strongly normalizing.
+  \end{theorem}
+
+  \subsubsection{Backwards Capability}
+  To show that CLC is backwards compatible with CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
+
+  \begin{definition}
+    \begin{align*}
+      \lift{x} &= x \\
+      \lift{Prop} &= U_* \\
+      \lift{Type_k} &= U_k \\
+      \lift{(x : A) \rightarrow B} &= (x \utype \lift{A}) \rightarrow \lift{B} \\
+      \lift{\lambda x.n} &= \lambda x.\lift{n} \\
+      \lift{m\ n} &= \lift{m}\ \lift{n}
+    \end{align*}
+  \end{definition}
+
+  With slight overloading of notation, we define lifting for CC$\omega$ recursively.
+  \begin{definition}
+    \begin{align*}
+      \lift{\epsilon} &= \epsilon \\
+      \lift{\Gamma, x : A} &= \lift{\Gamma}, x \utype \lift{A}
+    \end{align*}
+  \end{definition}
+
+  The following lemmas and theorems are all proved following a similar procedure to proving the soundness of erasure.
+  \begin{lemma}
+    Lift commutes with substitution. For any CC$\omega$ terms $m$, $n$ and some variable $x$, $\lift{m[n/x]} = \lift{m}[\lift{n}/x]$.
+  \end{lemma}
+
+  \begin{theorem}
+    Lift preserves reduction.
+For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+  \end{theorem}
+
+  \begin{theorem}
+    Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} \utype \lift{A}$ is a valid judgment in CLC.
   \end{theorem}

### 2021-10-19 writing/lics22/lics22.tex
lics22

@@
 
-  \subsubsection{Backwards Capability}
+  \subsubsection{Backwards Compatibility}
   To show that CLC is backwards compatible with CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.

### 2021-10-20 writing/lics22/lics22.tex
lics22

@@
 
-\lstset{basicstyle=\scriptsize,language=Coq,showstringspaces=false}
+\lstset{basicstyle=\small,language=Coq,showstringspaces=false}
 \newtheorem{theorem}{Theorem}[section]
@@
   \begin{figure}[h]
-  \vspace{-1.2em}
+  \vspace{-1em}
   \caption{Inductive Type Definition and Elimination}
@@
   (* Sigma Definition *)
-  Inductive Sigma (A: Type) (F: A -> Type): Type :=
+  Inductive Sigma (A: U) (F: A -> U): U :=
   | pair : (x: A) -> F x -> Sigma A F.
@@
   (* Tensor Definition *)
-  Inductive Tensor (A: Linear) (B: Linear): Linear :=
+  Inductive Tensor (A: L) (B: L): L :=
   | tpair: A -> B -> Tensor A B.
@@
   (* FTensor Definition *)
-  Inductive FTensor (A: Type) (F: A -> Linear): Linear :=
+  Inductive FTensor (A: U) (F: A -> L): L :=
   | fpair: (x: A) -> F x -> FTensor A F.
 
-  (* Sigma Elimination
-     x and y are unrestricted in clause body *)
+  (* Sigma Elimination *)
   match m with
@@
 
-  (* Tensor Elimination
-     x and y must be used once in clause body *)
+  (* Tensor Elimination *)
   match m with
@@
 
-  (* FTensor Elimination
-     x is unrestricted in clause body, 
-     y must be used once in clause body *)
+  (* FTensor Elimination *)
   match m with
@@
   \end{lstlisting}
+  \vspace{-1em}
   \label{inductive}
@@
 
-  Of all the rules presented here, perhaps the most surprising to readers familiar with linear types or dependent types is the \rname{$\mathbb{N}$-L-Elim} rule.
-This rule constructs a term with linear type through a primitive recursion principle.
-The $\multimap$ arrows protect the O-case element of linear type $C\ O$ from duplication by currying.
-The S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
-
   \begin{figure}[h]
@@
   \end{figure}
+
+  Of all the rules presented here, perhaps the most surprising to readers familiar with linear types or dependent types is the \rname{$\mathbb{N}$-L-Elim} rule.
+This rule constructs a term with linear type through a primitive recursion principle.
+The $\multimap$ arrows protect the O-case element of linear type $C\ O$ from duplication by currying.
+The S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
 
@@
   let pf1 := Eq_sym _ _ _ pf1 in
-    let pf : Eq _ x y := Eq_trans _ _ _ _ pf1 pf2 in
-    free _ _ _ c.
+  let pf : Eq _ x y := 
+    Eq_trans _ _ _ _ pf1 pf2 
+  in free _ _ _ c.
   \end{lstlisting}
@@
 
-  Though we have not implemented program extraction for CLC yet, our preliminary experiments hint at the possibility of compiling the state axioms to constant time imperative operations after proof erasure.
-This can greatly increase the runtime efficiency of CLC over pure functional languages when used for programming.
-More examples are available in the first author's Github repository.
+  Though we have not implemented program extraction for CLC yet, our preliminary experiments show the feasibility of compiling the state axioms to constant time imperative operations after proof erasure.
+This can greatly improve the runtime efficiency of CLC over pure functional languages when used for programming.
+More complex examples are available in the first author's Github repository.
 
@@
 
-    Fixpoint append (A : Type) : list A -o list A -o 
-      (list A -o list A) -o list A :=
+  Fixpoint append (A : Type) : list A -o 
+    list A -o (list A -o list A) -o list A 
+  :=
     fun ls1 ls2 k =>
@@
       | (cons h t) as _cons_ => 
-          append _ t ls2 (fun res => k (_cons_ h res))
+        append _ t ls2 
+          (fun res => k (_cons_ h res))
       end.
@@
 Basically, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
+Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
+Using linear types for continuations is surprisingly accurate as control-flow itself is linear.
 

### 2021-10-20 writing/lics22/lics22.tex
lics22

@@
   \begin{figure}[h]
-    \vspace{-0.8em}
+    \vspace{-1em}
     \caption{Select Elimination Rules}
@@
   \begin{figure}[h]
-  \vspace{-1em}
+  \vspace{-0.6em}
   \caption{Law 1}
@@
   \end{lstlisting}
-  \vspace{-1em}
+  \vspace{-0.6em}
   \label{law1}
@@
   \begin{figure}[h]
-  \vspace{-1em}
+  \vspace{-0.6em}
   \caption{Law 2}
@@
   \end{lstlisting}
-  \vspace{-1em}
+  \vspace{-0.6em}
   \label{law2}
@@
   \begin{figure}[h]
-  \vspace{-1em}
+  \vspace{-0.6em}
   \caption{Linear Lists}
   \begin{lstlisting}
-  Inductive list (A : Type) : Linear :=
+  Inductive list (A : U) : L :=
   | nil : list A
@@
 
-  Fixpoint append (A : Type) : list A -o 
+  Fixpoint append (A : U) : list A -o 
     list A -o (list A -o list A) -o list A 
@@
   \label{list}
-  \vspace{-1em}
+  \vspace{-0.6em}
   \Description{}

### 2021-10-26 writing/lics22/lics22.tex
lics22

@@
 
-  Linear Logic is a substructural logic introduced by Girard in his seminal work \cite{girard}.
-Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
-Wadler \cite{wadler1990,wadler1991} first notices that an analogous restriction to variable usage in simple type theory leads to a linear type theory, where terms respect resources.
-A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
-Benton \cite{benton1994} investigates the ramifications of the !
-exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
-Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
-The success of integrating Linear Logic with simple type theory exposes a tantalizing new frontier of integrating linearity with richer type theories.
-
-  Work has been done to extend dependent type theories with linear types.
-Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
-Krishnaswami et al.
-present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
-Luo et al.
-introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
-The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
-
-  We present a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
+  This paper presents a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
 CLC extends CC$\omega$ with linear types.
@@
   \end{equation*}
-  This new specification for transaction demands a payment of 5 units of money, and the customer be relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
+  This new specification for the transaction demands a payment of 5 units of money, and the customer be relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
 
   Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach to extending CC$\omega$ with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
-This allows for a straightforward modeling of CLC in CC$\omega$, and lifting of CC$\omega$ into CLC, endowing CLC with the fruits of decades of CC research.
+Prior works have either utilized a literal !
+exponential or a pair of adjoint connectives to bridge the gap between linear and non-linear types, adding clutter that greatly hinders their practical usage.
+Our method requires neither exponential nor adjoint connectives, allowing for a straightforward modeling of CLC in CC$\omega$, and lifting of CC$\omega$ into CLC with minimal annotations, endowing CLC with the fruits of decades of CC research.
+
   We further extend CLC with inductive types, showing that CLC facilitates correct by construction imperative programming.
 We have formalized all major results in Coq, and implemented a prototype in OCaml.
-
-  \medskip
 
@@
     \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation and subject reduction.
-    \item Next, we prove the linearity theorem, showing that the tangible impact linear types have how terms can be structured.
+    \item Next, we prove the linearity theorem, showing the tangible impact linear types have on the structure of terms.
 We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
-    \item Additionally, we observe that CLC is highly backwards compatible with CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$.
-This shows that CLC is consistent.
+    \item Additionally, we observe that CLC is highly backwards compatible with CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
+We also prove that all valid CC$\omega$ terms after adding minimal amounts of annotation are valid CLC terms.
     \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
@@
   \begin{figure}[h]
-    \vspace{-1.4em}
+    \vspace{-1em}
     \caption{Syntax}
-    \vspace{1.4em}
+    \vspace{1em}
     \begin{grammar}
@@
 $U_k$, and $L_k$ are the predicative universes of non-linear and linear types respectively.
-  \begin{figure}[H]
-    \vspace{-1.4em}
+  \begin{figure}[h]
+    \vspace{-1em}
     \caption{Correspondence of CLC types and MELL implications}
@@
     \end{align}
-    \vspace{-1.6em}
+    \vspace{-1em}
     \label{correspondence}
@@
   \begin{figure}[H]
-    \vspace{-1.6em}
+    \vspace{-1em}
     \caption{The Universe Hierarchy}
@@
     \label{universe}
+    \Description{}
   \end{figure}
@@
     \label{structural}
+    \Description{}
   \end{figure}
@@
 
-  \begin{figure}[H]
-    \vspace{-1em}
-    \caption{Type Formation} 
-    \begin{mathpar}
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash U_* : U_0 } 
-      \rname{Prop-Axiom}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash s_k : U_{k+1} } 
-      \rname{Sort-Axiom}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : U_i \\ 
-        \Gamma, x \utype A \vdash B : U_* }
-      { \Gamma \vdash (x \utype A) \rightarrow B : U_* } 
-      \rname{Prop}
-      \\
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : U_k \\ 
-        \Gamma, x \utype A \vdash B : s_k }
-      { \Gamma \vdash (x \utype A) \rightarrow B : U_k } 
-      \rname{U-Prod}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : L_k \\ 
-        \Gamma \vdash B : s_k \\
-        x \notin \Gamma }
-      { \Gamma \vdash (x \ltype A) \rightarrow B : U_k } 
-      \rname{L-Prod}
-      \\
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : U_k \\ 
-        \Gamma, x \utype A \vdash B : s_k }
-      { \Gamma \vdash (x \utype A) \multimap B : L_k } 
-      \rname{U-Lolli}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : L_k \\ 
-        \Gamma \vdash B : s_k \\
-        x \notin \Gamma }
-      { \Gamma \vdash (x \ltype A) \multimap B : L_k } 
-      \rname{L-Lolli}
-    \end{mathpar}
-    \vspace{-1em}
-    \label{type}
-  \end{figure}
-
   The axiom rules \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
@@
 
+  \begin{figure}[h]
+    \caption{Type Formation} 
+    \begin{mathpar}
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash U_* : U_0 } 
+      \rname{Prop-Axiom}
+
+      \inferrule
+      { \pure{\Gamma} }
+      { \Gamma \vdash s_k : U_{k+1} } 
+      \rname{Sort-Axiom}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_i \\ 
+        \Gamma, x \utype A \vdash B : U_* }
+      { \Gamma \vdash (x \utype A) \rightarrow B : U_* } 
+      \rname{Prop}
+      \\
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_k \\ 
+        \Gamma, x \utype A \vdash B : s_k }
+      { \Gamma \vdash (x \utype A) \rightarrow B : U_k } 
+      \rname{U-Prod}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : L_k \\ 
+        \Gamma \vdash B : s_k \\
+        x \notin \Gamma }
+      { \Gamma \vdash (x \ltype A) \rightarrow B : U_k } 
+      \rname{L-Prod}
+      \\
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : U_k \\ 
+        \Gamma, x \utype A \vdash B : s_k }
+      { \Gamma \vdash (x \utype A) \multimap B : L_k } 
+      \rname{U-Lolli}
+
+      \inferrule
+      { \pure{\Gamma} \\
+        \Gamma \vdash A : L_k \\ 
+        \Gamma \vdash B : s_k \\
+        x \notin \Gamma }
+      { \Gamma \vdash (x \ltype A) \multimap B : L_k } 
+      \rname{L-Lolli}
+    \end{mathpar}
+    \label{type}
+    \Description{}
+  \end{figure}
+
   The \rname{U-Prod} rule is used for forming non-linear function types with non-linear domains.
@@
   \subsection{Term Formation} \label{teformation}
-  The rules for term formation are presented in Figure \ref{term}.
+  The rules for term formation are presented in Figure~\ref{term}.
 
@@
       \inferrule
-      { \pure{\Gamma} }
-      { \Gamma, x \utype A \vdash x : A } 
+      {\pure{\Gamma}}
+      {\Gamma, x \utype A \vdash x : A} 
       \rname{U-Var}
@@
     \end{mathpar}
-    \vspace{-1em}
     \label{term}
+    \Description{}
   \end{figure}
@@
     \label{parallel}
+    \Description{}
   \end{figure}
@@
   \subsection{Meta Theory} \label{meta}
-  In this section, we focus our discussion on the meta-properties of CLC.
+  In this section, we focus our discussion on the properties of CLC.
 First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
@@
   \subsubsection{Promotion and Dereliction}
-  In most presentations of linear logic, non-linear objects are created by packing linear objects into the !
-exponential using the \textit{promotion} rule.
-Conversely, the \textit{dereliction} rule unpacks non-linear objects so they can be used in place of linear ones.
-Linear type theories \cite{abramsky1993,dill,vakar14} utilizing exponential for non-linearity will contain !
-types and rules that are equivalent to \textit{promotion} and \textit{dereliction}.
-
-  Linear types of CLC are not obtained through packing and unpacking !, so there are no rules for \textit{promotion} and \textit{dereliction}.
+  Linear types of CLC are not obtained through packing and unpacking !, so there are no explicit rules for \textit{promotion} and \textit{dereliction} of Linear Logic.
 Arbitrary computations existing at the type level also muddle the association between which types can be promoted or derelicted to which.
@@
   \begin{theorem}
-    Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} \utype \lift{A}$ is a valid judgment in CLC.
+    Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
   \end{theorem}
 
+  The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domain of function types are annotated with sort $U$.
+In practice, even these annotations can be omitted if algorithmic type checking is able to infer them from context.
+
   \section{Extensions}
-  Readers familiar with the literature of linear logic or linear type theory will have noticed at this point that CLC does not possess the famous $\otimes$ connective.
+  Readers familiar with the literature of linear logic or linear type theory will have noticed at this point that CLC does not possess the iconic $\otimes$ connective.
 Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written in this style will be incredibly difficult to maintain.
-
-  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic} as an extension to the core theory.
-The presence of universe sorts $U$ and $L$ allows users to declare inductive types to be non-linear or linear.
-The lack of !
-exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching.
-The $\otimes$ connective becomes definable as an inductive type.
 
@@
 
-  Figure \ref{inductive} is an excerpt taken from our implementation's prelude, with syntax drawing heavily from Coq.
+  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic} as an extension to the core theory.
+The presence of universe sorts $U$ and $L$ allow users to declare inductive types to be non-linear or linear.
+The lack of !
+exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching.
+The $\otimes$ connective becomes definable as an inductive type.
+
+  Figure \ref{inductive} is an excerpt taken from our implementation's prelude, with syntax heavily inspired by Coq.
 
@@
 
-  \begin{figure}[h]
-    \vspace{-0.8em}
+  \begin{figure}[H]
     \caption{Select Formation Rules}
@@
     \end{mathpar}
-    \vspace{-0.4em}
     \label{formation}
@@
 
-  \begin{figure}[h]
+  \begin{figure}[H]
     \vspace{-0.8em}
@@
     \end{mathpar}
-    \vspace{-2em}
+    \vspace{-1em}
     \label{introduction}
@@
 This can greatly improve the runtime efficiency of CLC over pure functional languages when used for programming.
-More complex examples are available in the first author's Github repository.
+More complex examples such as safe array indexing are available in the first author's Github repository.
 
@@
   \begin{figure}[h]
-  \vspace{-0.6em}
   \caption{Linear Lists}
@@
   \label{list}
-  \vspace{-0.6em}
   \Description{}
@@
 Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
-Using linear types for continuations is surprisingly accurate as control-flow itself is linear.
+Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.
+
+  \section{Related Work}
+  Linear types are a class of type systems inspired by Girard' substructural Linear Logic \cite{girard}.
+Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
+Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
+A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
+Benton \cite{benton1994} investigates the ramifications of the !
+exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
+Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
+
+  Over the years, work has been done to enrich linear type theories with dependent types.
+Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
+Krishnaswami et al.
+present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
+Luo et al.
+\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
+The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
 
@@
 
-  \section{Conclusion}
-  The Calculus of Linear Constructions is a lightweight extension of the Calculus of Constructions with linear types.
-As a type theory, it is a sound and effective framework for reasoning about resources.
-As a programing language, it allows one to write safe and efficient code, alleviating some of the common issues faced by standard functional programming.
-
 \bibliographystyle{acm}

### 2021-10-27 writing/lics22/lics22.tex
lics22

@@
 % document layout
-\documentclass[sigplan,screen,review,authordraft]{acmart}
-%\documentclass[sigplan,screen]{acmart}
+%\documentclass[sigplan,screen,review,authordraft]{acmart}
+\documentclass[sigplan,screen]{acmart}
 \usepackage[utf8]{inputenc}
@@
 Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
+We study the meta-theory of CLC, showing that it is a sound and consistent logical framework for reasoning about resource.
 CLC is backwards compatible with CC, allowing CLC to enjoy the fruits of decades of CC research.
@@
       \inferrule
-      {\pure{\Gamma}}
-      {\Gamma, x \utype A \vdash x : A} 
+      { \pure{\Gamma_1, \Gamma_2} }
+      { \Gamma_1, x \utype A, \Gamma_2 \vdash x : A }
       \rname{U-Var}
@@
       \inferrule
-      { \pure{\Gamma} }
-      { \Gamma, x \ltype A \vdash x : A } 
+      { \pure{\Gamma_1, \Gamma_2} }
+      { \Gamma_1, x \ltype A, \Gamma_2 \vdash x : A } 
       \rname{L-Var}
@@
 
-  \subsubsection{Weakening} \label{weakening}ß
+  \subsubsection{Weakening} \label{weakening}
   CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.

### 2021-10-27 writing/lics22/lics22.tex
lics22

@@
 Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, showing that it is a sound and consistent logical framework for reasoning about resource.
+We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
 CLC is backwards compatible with CC, allowing CLC to enjoy the fruits of decades of CC research.
@@
   The rules \rname{U-Var}, and \rname{L-Var} are used for typing free variables.
-The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma, x \utype A$ with a non-linear type $A$.
-The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma, x \ltype A$ with linear type $A$.
-For both rules, the side condition $\pure{\Gamma}$ forbids irrelevant variables of linear types from occurring within $\Gamma$.
+The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \utype A, \Gamma_2$ with a non-linear type $A$.
+The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \ltype A, \Gamma_2$ with linear type $A$.
+For both rules, the side condition $\pure{\Gamma_1, \Gamma_2}$ forbids irrelevant variables of linear types from occurring within the context.
 This prevents another vector of weakening variables with linear type.

### 2021-11-06 notes/scratch1.tex
scratch

\documentclass{article}

\usepackage[margin=0.6in]{geometry}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathpartir}
\pagestyle{empty}

\newcommand{\rname}[1]{\textsc{\footnotesize #1}}
\newcommand{\utype}{:_{\scriptscriptstyle U}}
\newcommand{\ltype}{:_{\scriptscriptstyle L}}
\newcommand{\stype}{:_s}
\newcommand{\ucons}{constructor_{\scriptscriptstyle U}}
\newcommand{\lcons}{constructor_{\scriptscriptstyle L}}
\newcommand{\scons}{constructor_{s}}
\newcommand{\pure}[1]{|#1|}

\begin{document}

\section{Inductive Linear Types}

\subsection{Arity}
For some type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as follows.
\begin{mathpar}
  \inferrule
  { }
  { arity(s,s) }
  \rname{ar-sort}

  \inferrule
  { arity(A, s) }
  { arity((x \utype M) \rightarrow A, s)}
  \rname{ar-arrow}
\end{mathpar}

\subsection{Strict Positivity}
For some type $P$ and type variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as follows.
\begin{mathpar}
  \inferrule
  { (\forall i = 1...n) \\ X \notin m_i }
  { postive(X\ m_1...m_n, X) }
  \rname{pos-x}

  \inferrule
  { X \notin M \\ postive(P, X)}
  { postive((x \stype M) \rightarrow P, X) }
  \rname{pos-arrow}

  \inferrule
  { X \notin M \\ postive(P, X) }
  { postive((x \stype M) \multimap P, X) }
  \rname{pos-lolli}
\end{mathpar}
We say that $X$ occurs strictly postive in $P$ if the judgment $postive(P,X)$ is valid.

\subsection{Non-Linear Constructor}
For some type $C$ and type variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as follows.
\begin{mathpar}
  \inferrule
  { (\forall i = 1...n) \\ X \notin m_i }
  { \ucons(X\ m_i...m_n, X) }
  \rname{u-constr-x}

  \inferrule
  { postive(P, X) \\ \ucons(C, X) }
  { \ucons((\_ \utype P) \rightarrow C, X)}
  \rname{u-constr-pos}

  \inferrule
  { X \notin M \\ \ucons(C, X) }
  { \ucons((x \utype M) \rightarrow C, X)}
  \rname{u-constr-arrow}
\end{mathpar}
Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not applied to linear terms as this may cause duplication of linear variables.

\subsection{Linear Constructor}
For some type $C$ and type variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as follows.
\begin{mathpar}
  \inferrule
  { (\forall i = 1...n) \\ X \notin m_i }
  { \lcons(X\ m_i...m_n, X) }
  \rname{l-constr-x} \\

  \inferrule
  { postive(P, X) \\ \lcons(C, X) }
  { \lcons((\_ \utype P) \rightarrow C, X)}
  \rname{l-constr-pos-1}

  \inferrule
  { X \notin M \\ \lcons(C, X) }
  { \lcons((x \utype M) \rightarrow C, X)}
  \rname{l-constr-arrow-1}

  \inferrule
  { postive(P, X) \\ activation(C, X) }
  { \lcons((\_ \ltype P) \rightarrow C, X)}
  \rname{l-constr-pos-2}

  \inferrule
  { X \notin M \\ activation(C, X) }
  { \lcons((x \ltype M) \rightarrow C, X)}
  \rname{l-constr-arrow-2}
\end{mathpar}
For some type $C$ and type variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as follows.
\begin{mathpar}
  \inferrule
  { (\forall i = 1...n) \\ X \notin m_i }
  { activation(X\ m_i...m_n, X) }
  \rname{act-x}

  \inferrule
  { postive(P, X) \\ activation(C, X) }
  { activation((\_ \stype P) \multimap C, X)}
  \rname{act-pos}

  \inferrule
  { X \notin M \\ activation(C, X) }
  { activation((x \stype M) \multimap C, X)}
  \rname{act-lolli}
\end{mathpar}
Linear constructors are used for constructing linear objects that must be used once. These constructors can be applied to linear terms as restricted usage prevent duplication of linear variables.

An unapplied linear constructor is a non-linear entity as they can be created freely ``out of thin air". However, once a linear constructor has been partially applied to a linear term, its linearity is ``activated'' as seen in rules \rname{l-constr-pos-2} and \rname{l-constr-arrow-2}, essentially forcing the rest of its arguments to be fully applied. This is done to prevent partially applied constructors from duplicating linear variables.

\pagebreak

\subsection{Introduction Rules}
The formation of new inductive types is defined as follows.
\begin{mathpar}
  \inferrule
  { (\forall i = 1...n) \\ arity(A, s) \\ \scons(C_i, X) \\\\
    \pure{\Gamma} \\ 
    \Gamma \vdash A : U_{k+1} \\ 
    \Gamma, X \utype A \vdash C_i : U_k }
  { \Gamma \vdash Ind_s(X : A)\{C_1|...|C_n\} : A }
  \rname{ind-intro}
\end{mathpar}
The formation of constructors for defined inductive types is defined as follows.
\begin{mathpar}
  \inferrule
  { \pure{\Gamma} \\   
    I_s := Ind_s(X : A)\{C_1|...|C_n\} \\
    \Gamma \vdash I_s :A \\
    1 \leq i \leq n }
  { \Gamma \vdash Constr(i, I_s) : C_i[I_s/X] }
  \rname{constr-intro}
\end{mathpar}

\subsection{Non-Dependent Case}
\subsection{Dependent Case}
\subsection{Fixpoint}
\subsection{Examples}

\end{document}

### 2021-11-06 notes/scratch1.tex
scratch

@@
 \newcommand{\scons}{constructor_{s}}
+\newcommand{\indu}{\text{Ind}_{\scriptscriptstyle U}}
+\newcommand{\indl}{\text{Ind}_{\scriptscriptstyle L}}
+\newcommand{\inds}{\text{Ind}_s}
+\newcommand{\constr}{\text{Constr}}
+\newcommand{\case}{\text{Case}}
+\newcommand{\dcase}{\text{DCase}}
+\newcommand{\fix}{\text{Fix }}
 \newcommand{\pure}[1]{|#1|}
+\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
+\newcommand{\restr}[1]{\overline{#1}}
+\newcommand{\lrangle}[1]{\langle #1 \rangle}
 
@@
 \end{mathpar}
+For some arity type $A$ and sort $s'$, the term $A\lrangle{s'}$ is inductively defined over the structure of $A$ as follows.
+\begin{align*}
+  ((x \utype M) \rightarrow A)\lrangle{s'} &= (x \utype M) \rightarrow A\lrangle{s'} \\
+  (s)\lrangle{s'} &= s'
+\end{align*}
+For some arity type $A$, term $I$ and sort $s'$, the term $A\{I, s'\}$ is inductively defined over the structure of $A$ as follows.
+\begin{align*}
+  ((x \utype M) \rightarrow A)\lrangle{I, s'} &= (x \utype M) \rightarrow A\lrangle{(I\ x), s'} \\
+  (s)\lrangle{I, s'} &= (\_ \utype I) \rightarrow s'
+\end{align*}
 
@@
   { (\forall i = 1...n) \\ X \notin m_i }
-  { postive(X\ m_1...m_n, X) }
+  { postive((X\ m_1...m_n), X) }
   \rname{pos-x}
@@
   { (\forall i = 1...n) \\ X \notin m_i }
-  { \ucons(X\ m_i...m_n, X) }
+  { \ucons((X\ m_i...m_n), X) }
   \rname{u-constr-x}
@@
   { (\forall i = 1...n) \\ X \notin m_i }
-  { \lcons(X\ m_i...m_n, X) }
+  { \lcons((X\ m_i...m_n), X) }
   \rname{l-constr-x} \\
@@
   { (\forall i = 1...n) \\ X \notin m_i }
-  { activation(X\ m_i...m_n, X) }
+  { activation((X\ m_i...m_n), X) }
   \rname{act-x}
@@
 
-\pagebreak
-
 \subsection{Introduction Rules}
@@
     \Gamma, X \utype A \vdash C_i : U_k }
-  { \Gamma \vdash Ind_s(X : A)\{C_1|...|C_n\} : A }
+  { \Gamma \vdash \inds(X : A)\{C_1|...|C_n\} : A }
   \rname{ind-intro}
@@
   { \pure{\Gamma} \\   
-    I_s := Ind_s(X : A)\{C_1|...|C_n\} \\
-    \Gamma \vdash I_s :A \\
+    I := \inds(X : A)\{C_1|...|C_n\} \\
+    \Gamma \vdash I :A \\
     1 \leq i \leq n }
-  { \Gamma \vdash Constr(i, I_s) : C_i[I_s/X] }
+  { \Gamma \vdash \constr(i, I) : C_i[I/X] }
   \rname{constr-intro}
@@
 \subsection{Non-Dependent Case}
+For some constructor type $C$ and type variables $X$ and $Q$, the term $C\{X, Q\}$ is inductively defined over the structure of $C$ as follows.
+\begin{align*}
+  ((\_ \stype P) \rightarrow C)\{X, Q\} &= (\_ \stype P) \multimap C\{X, Q\} \\
+  ((\_ \stype P) \multimap C)\{X, Q\} &= (\_ \stype P) \multimap C\{X, Q\} \\
+  ((x \stype M) \rightarrow C)\{X, Q\} &= (x \stype P) \multimap C\{X, Q\} \\
+  ((x \stype M) \multimap C)\{X, Q\} &= (x \stype P) \multimap C\{X, Q\} \\
+  (X\ m_1...m_n)\{X, Q\} &= (Q\ m_1...m_n)
+\end{align*}
+We define the notation $C[I, P] := C\{X, Q\}[I/X, P/Q]$.
+
+Inductive objects can be eliminated by a case expression as follows.
+\begin{mathpar}
+  \inferrule
+  { (\forall i = 1...n) \\ 
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\ 
+    arity(A, s) \\
+    I := \inds(X : A)\{C_1|...|C_n\} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_n) \\
+    \restr{\Gamma_2} \vdash Q : A\lrangle{s'} \\
+    \Gamma_2 \vdash f_i : C_i[I, Q] }
+  { \Gamma \vdash \case(m, Q)\{f_1|...|f_n\} : (Q\ a_1...a_n) }
+  \rname{case}
+\end{mathpar}
+
 \subsection{Dependent Case}
+For some non-linear constructor type $C$ and type variables $X$, $Q$ and $c$, the term $C\{X, Q, c\}$ is inductively defined over the structure of $C$ as follows.
+\begin{align*}
+  ((\_ \utype P) \rightarrow C)\{X, Q, c\} &= (p \utype P) \multimap C\{X, Q, (c\ p)\} \\
+  ((x \utype M) \rightarrow C)\{X, Q, c\} &= (x \utype M) \multimap C\{X, Q, (c\ x)\} \\
+  (X\ m_1...m_n)\{X, Q, c\} &= (Q\ m_1...m_n\ c)
+\end{align*}
+We define the notation $C[I, P, t] := C\{X, Q, c\}[I/X, P/Q, t/c]$.
+
+Non-linear inductive objects can be eliminated by a dependent case expression as follows.
+\begin{mathpar}
+  \inferrule
+  { (\forall i = 1...n) \\ 
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\ 
+    arity(A, U_i) \\
+    I := \indu(X : A)\{C_1|...|C_n\} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_n) \\
+    \restr{\Gamma_2} \vdash Q : A\lrangle{I, s'} \\
+    \Gamma_2 \vdash f_i : C_i[I, Q, Constr(i, I)] }
+  { \Gamma \vdash \dcase(m, Q)\{f_1|...|f_n\} : (Q\ a_1...a_n) }
+  \rname{dcase}
+\end{mathpar}
+
 \subsection{Fixpoint}
-\subsection{Examples}
+\begin{mathpar}
+  \inferrule
+  { \pure{\Gamma} \\ 
+    \Gamma \vdash T : U_i \\ 
+    \Gamma, f \utype T \vdash m : T \\
+    \text{guard condition} }
+  { \Gamma \vdash \fix f.m : T }
+  \rname{fix}
+\end{mathpar}
 

### 2021-11-27 writing/draft4/lics22.tex
draft4

--- 2021-10-27 writing/lics22/lics22.tex
@@
 \usepackage{tikz-cd}
+\usepackage{enumitem}
 
@@
 \newtheorem{definition}{Definition}[section]
-
+\setlist[itemize]{leftmargin=1.6em}
 \renewcommand{\syntleft}{}
@@
 \newcommand{\pure}[1]{|#1|}
-\newcommand{\sort}{\textbf{sort}}
-\newcommand{\type}{\textbf{type}}
 \newcommand{\refl}{\text{refl}}
@@
 \newcommand{\unit}{\text{unit}}
-\newcommand{\new}{\textbf{new}}
-\newcommand{\free}{\textbf{free}}
-\newcommand{\get}{\textbf{get}}
-\newcommand{\set}{\textbf{set}}
+\newcommand{\new}{\text{new}}
+\newcommand{\free}{\text{free}}
+\newcommand{\get}{\text{get}}
+\newcommand{\set}{\text{set}}
+\newcommand{\session}{\text{session}}
+\newcommand{\channel}{\text{channel}}
+\newcommand{\open}{\text{open}}
+\newcommand{\close}{\text{close}}
+\newcommand{\send}{\text{send}}
+\newcommand{\recv}{\text{recv}}
+\newcommand{\SEND}{\texttt{SEND}}
+\newcommand{\RECV}{\texttt{RECV}}
+\newcommand{\END}{\texttt{END}}
 \newcommand{\utype}{:_{\scriptscriptstyle U}}
@@
 We collapse the sort annotations on arrows $(x \utype A) \rightarrow B$ and $(x \ltype A) \rightarrow B$ into $(x : A) \rightarrow B$ whenever $A$'s sort is inferable from context.
+We also omit writing universe levels for the sake of clarity.
 
@@
     \begin{itemize}
-      \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U_k) \rightarrow A \rightarrow L_k$ \\
+      \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U) \rightarrow A \rightarrow L$ \\
       The $\mapsto$ is a linear type constructor.
@@
 Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
-      \item $\new : (A : U_k) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
+      \item $\textbf{new} : (A : U) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
       $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
-      \item $\free : (A : U_k) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \mathbf{1}$ \\
+      \item $\textbf{free} : (A : U) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \unit$ \\
       $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use.
 This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
-      \item $\get : (A : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
+      \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
       $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
@@
 The proof $(l \mapsto_A m)$ is returned unchanged as well.
-      \item $\set : (A\ B : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
+      \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
       $\set$ has the most complicated type of all the axioms presented here.
@@
 
+  \subsection{Session Reasoning}
+  The inclusion of dependent types and linear types in CLC makes it extremely expressive when encoding communication protocols.
+By indexing each channel with a protocol, CLC can enforce communication on a channel to strictly adhere to its specified protocol.
+
+  In Figure \ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
+Intuitively, the \texttt{session}  forms a stack of possible operations executed in a protocol.
+The $\SEND$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after sending a message of type $A$, the protocol progresses to stage $ss$.
+Likewise, the $\RECV$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after receiving a message of type $A$, the protocol progresses to stage $ss$.
+Finally, the $\END$ constructor indicates that the protocol has finished and communication is over.
+
+  \begin{figure}[h]
+  \caption{Inductive Session Type}
+  \begin{lstlisting}
+  Inductive session : U :=
+  | SEND : U -> session -> session
+  | RECV : U -> session -> session
+  | END : session.
+  \end{lstlisting}
+  \label{session}
+  \Description{}
+  \end{figure}
+
+  In Definition \ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
+
+  \begin{definition}
+    Communication axioms.
+    \begin{itemize}
+      \item $\textbf{channel} : \session \rightarrow L$ \\
+      The $\channel$ axiom is a type constructor. It is used to form the types of communication channels that obey the protocol specified by its session.
+      \item $\textbf{open} : (ss : \session) \rightarrow \channel\ ss$ \\
+      When given a protocol, $\open$ creates a channel indexed by this protocol.
+      \item $\textbf{close} : \channel\ \END \rightarrow \unit$ \\
+      Once all communications specified on a given channel has finished, $\close$ will close and reclaim the channel.
+      \item $\textbf{send} : (A : U) \rightarrow A \rightarrow (ss : \session) \rightarrow$ \\
+      \phantom{send : (A :)}
+      $\channel\ (\SEND\ A\ ss) \rightarrow \channel\ ss$ \\
+      After a channel has reached the point in its protocol where a message of type $A$ should be sent, $\send$ may send a term of type $A$ on this channel, causing the channel to progress to the next stage of its protocol.
+      \item $\textbf{recv} : (A : U) \rightarrow (ss : \session) \rightarrow $ \\
+      \phantom{recv : (A :)}
+      $\channel\ (\RECV\ A\ ss) \rightarrow F \_:A.\channel\ ss$
+      After a channel has reached the point in its protocol where a message of type $A$ should be received, $\recv$ may receive such a term of type $A$ from the channel, causing the channel to progress to the next stage of its protocol.
+    \end{itemize}
+    \label{comm}
+  \end{definition}
+
+  The important detail to notice here is that each channel created by $\open$ is of linear type $(\channel\ ss)$ where $ss$ is some arbitrary protocol, hence must be used exactly once.
+However, the $\send$ and $\recv$ operations return back channels after consuming them, progressing their protocols one stage forward.
+Coupled with the fact that the $\close$ operation is only able to close channels whose protocols have ended, this forces channels to communicate strictly by their specified protocols.
+
+  Dependent types also allows for more precise specifications of protocols themselves.
+A simple example would be the specification of message sizes.
+Figure \ref{communication} shows such an example.
+The inductive type \texttt{ilist} is a length indexed list.
+Using \texttt{ilist}, we construct a protocol \texttt{ss} that first sends two natural numbers, then receives eight natural numbers and finally sends back a boolean acknowledgement.
+
+  \begin{figure}[h]
+  \caption{Sized Communication}
+  \begin{lstlisting}
+  Inductive ilist (A : U) : Nat -> U :=
+  | nil  : ilist A 0
+  | cons : A -> (n : Nat) -> 
+              ilist A n -> ilist A (S n).
+
+  let ss := 
+    SEND (ilist Nat 2) 
+      (RECV (ilist Nat 8) (SEND Bool END)) 
+  in
+  let send_msg : ilist Nat 2 := 
+    (cons 1 _ (cons 2 _ nil)) 
+  in
+  let ch := open ss in
+  let ch := send _ send_msg _ ch in
+  let [ recv_msg, ch ] := recv _ _ ch in
+  let ch := send _ true _ ch in
+  close ch            
+  \end{lstlisting}
+  \label{communication}
+  \Description{}
+  \end{figure}
+
   \subsection{Mutable Data}

### 2021-12-02 writing/draft5/draft5.tex
draft5

--- 2021-10-27 writing/lics22/lics22.tex
@@
-% document layout
-%\documentclass[sigplan,screen,review,authordraft]{acmart}
-\documentclass[sigplan,screen]{acmart}
+\documentclass[sigplan,screen,review,authordraft]{acmart}
+
 \usepackage[utf8]{inputenc}
@@
 \usepackage{listings}
-\usepackage{lstlangcoq}
 \usepackage{float}
 \usepackage{tikz-cd}
-
-\lstset{basicstyle=\small,language=Coq,showstringspaces=false}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
+\usepackage{enumitem}
+
+\setlist[itemize]{leftmargin=1.6em}
 \renewcommand{\syntleft}{}
@@
 \newcommand{\pure}[1]{|#1|}
-\newcommand{\sort}{\textbf{sort}}
-\newcommand{\type}{\textbf{type}}
 \newcommand{\refl}{\text{refl}}
 \newcommand{\letin}[3]{\ $\text{let }#1\text{ := }#2\text{ in }#3$\ }
-\newcommand{\ind}{\ $\text{ind}$\ }
-\newcommand{\iter}{\ $\text{iter}$\ }
+\newcommand{\ind}[1]{\text{Ind}_{#1}}
+\newcommand{\constr}{\text{Constr}}
+\newcommand{\case}{\text{Case}}
+\newcommand{\dcase}{\text{DCase}}
+\newcommand{\fix}{\text{Fix }}
+\newcommand{\sep}{\text{ | }}
 \newcommand{\unit}{\text{unit}}
-\newcommand{\new}{\textbf{new}}
-\newcommand{\free}{\textbf{free}}
-\newcommand{\get}{\textbf{get}}
-\newcommand{\set}{\textbf{set}}
+\newcommand{\new}{\text{new}}
+\newcommand{\free}{\text{free}}
+\newcommand{\get}{\text{get}}
+\newcommand{\set}{\text{set}}
+\newcommand{\session}{\text{session}}
+\newcommand{\channel}{\text{channel}}
+\newcommand{\open}{\text{open}}
+\newcommand{\close}{\text{close}}
+\newcommand{\send}{\text{send}}
+\newcommand{\recv}{\text{recv}}
+\newcommand{\SEND}{\texttt{SEND}}
+\newcommand{\RECV}{\texttt{RECV}}
+\newcommand{\END}{\texttt{END}}
 \newcommand{\utype}{:_{\scriptscriptstyle U}}
@@
 \newcommand{\lift}[1]{\llparenthesis #1 \rrparenthesis}
-
-% title and author
+\newcommand{\lrangle}[1]{\langle #1 \rangle}
+\newcommand{\ucons}{constructor_{\scriptscriptstyle U}}
+\newcommand{\lcons}{constructor_{\scriptscriptstyle L}}
+\newcommand{\scons}{constructor_{s}}
+\newcommand{\inl}{\text{inl}}
+\newcommand{\inr}{\text{inr}}
+
 \title{The Calculus of Linear Constructions}
@@
 
-% document body
+\keywords{Dependent types, linear types, inductive types, Calculus of Constructions}
+
 \begin{document}
@@
   The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends CC with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
+Specifically, CLC extends the predicative CC$\omega$ with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
 We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
-CLC is backwards compatible with CC, allowing CLC to enjoy the fruits of decades of CC research.
-We have formalized and proven correct all major results of the core calculus in the Coq Proof Assistant.
-We extend CLC with linear inductive types, showing that CLC facilitates correct by construction imperative programming.
+We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
+Through examples, we demonstrate that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
+We have formalized and proven correct all major results in the Coq Proof Assistant.
 \end{abstract}
@@
 In CC types can depend on terms, allowing one to write precise specifications as types.
-Today, CC and its variations CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
+Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
 These theorem provers have found great success in the fields of software verification, and constructive mathematics.
@@
 Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
-Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide additional reasoning principles for dealing with resource.
-The design and embedding of these logics are a difficult problem in their own right, requiring additional proofs to justify their soundness, and effectiveness.
+Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
 We propose an alternative solution: extend CC with linear types.
@@
 This paper presents a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
-CLC extends CC$\omega$ with linear types.
+CLC extends the predicative CC$\omega$ with linear types.
 CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
@@
 
-  Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach to extending CC$\omega$ with linear types, akin to Mazurak et al.'s work on System F \cite{mazurak}.
-Prior works have either utilized a literal !
-exponential or a pair of adjoint connectives to bridge the gap between linear and non-linear types, adding clutter that greatly hinders their practical usage.
-Our method requires neither exponential nor adjoint connectives, allowing for a straightforward modeling of CLC in CC$\omega$, and lifting of CC$\omega$ into CLC with minimal annotations, endowing CLC with the fruits of decades of CC research.
-
-  We further extend CLC with inductive types, showing that CLC facilitates correct by construction imperative programming.
-We have formalized all major results in Coq, and implemented a prototype in OCaml.
+Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach, akin to Mazurak et al.'s work on System F \cite{mazurak}.
+Prior works have either utilized an explicit !
+exponential or a pair of adjoint connectives to mediate between the worlds of linear and non-linear types.
+Our method requires neither exponential nor adjoint connectives explicitly in the syntax, allowing for a straightforward modeling of CLC in the predicative CC$\omega$, and lifting of predicative CC$\omega$ into CLC with minimal annotations needed.
+
+We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
+Linear connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
+Through examples, we show that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
+
+We have formalized all major results in Coq, and implemented a prototype in OCaml, both of which are available in the first author's Github repository.
 
@@
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-    \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, propagation and subject reduction.
+  \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, validity and subject reduction.
   \item Next, we prove the linearity theorem, showing the tangible impact linear types have on the structure of terms.
 We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
-    \item Additionally, we observe that CLC is highly backwards compatible with CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
-We also prove that all valid CC$\omega$ terms after adding minimal amounts of annotation are valid CLC terms.
+  \item Additionally, we observe that CLC is backwards compatible with the predicative CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
+We prove that all valid CC$\omega$ terms are valid CLC terms after adding minimal amounts of annotation.
+  \item Furthermore, we extend CLC with rules to define inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
+This is the first formalization of a generalized mechanism for defining inductive linear types that we are aware of.
   \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
 To the best of our knowledge, our development is the first machine checked formalization of a linear dependent type theory.
-    \item Furthermore, we extend CLC with linear inductive data, demonstrating CLC's capabilities for correct by construction imperative programming.
-    \item Finally, we give an implementation extended with user definable linear and non-linear inductive types.
-Algorithmic type checking employed by the implementation streamlines the process of writing CLC.
+  \item Finally, we give an implementation of CILC in OCaml. Algorithmic type checking employed by the typechecker streamlines the process of writing CILC programs.
 \end{itemize}
@@
 \begin{figure}[h]
-    \vspace{-1em}
-    \caption{Syntax}
-    \vspace{1em}
+  \caption{Syntax of CLC}
   \begin{grammar}
-      <$k$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} predicative levels
-
-      <$i$> := * | 0 | 1 | 2 ... \hspace*{3em} all levels
+    <$i$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} universe levels
 
@@
 
-      <$m, n, A, B, C$> ::= $U_i$ | $L_k$ | $x$ \hspace*{4em} expressions
+    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4.2em} expressions
     \indalt $(x :_s A) \rightarrow B$
     \indalt $(x :_s A) \multimap B$
-      \indalt $\lambda x. n$ | $m\ n$
+    \indalt $\lambda x :_s A. n$
+    \indalt $m\ n$
   \end{grammar}
-    \label{syntax}
-  \end{figure}
-  The syntax of the core type theory is presented in Figure \ref{syntax}.
+  \Description{}
+  \label{syntax1}
+\end{figure}
+
+The syntax of the core type theory is presented in Figure \ref{syntax1}.
 Our type theory contains two sorts of universes $U$ and $L$.
-We use the meta variable $k$ to specifically quantify over levels $0, 1, 2, ...$ that correspond to the predicative universes.
-We use the meta variable $i$ to quantify over all levels $*, 0, 1, 2, ...$.
-Here, $U_*$ is the impredicative universe of propositions, in the same spirit as CC$\omega$'s $Prop$ universe.
-$U_k$, and $L_k$ are the predicative universes of non-linear and linear types respectively.
-  \begin{figure}[h]
-    \vspace{-1em}
+We use the meta variable $i$ to quantify over universe levels $0, 1, 2, ...$.
+$U_i$, and $L_i$ are the universes at level $i$ of non-linear and linear types respectively.
+\begin{figure}[h]
   \caption{Correspondence of CLC types and MELL implications}
   \Description{}
-    \vspace{-1em}
   \begin{align}
-      (\_ \utype A) \rightarrow B \quad &\Leftrightarrow \quad !(!A \multimap B) \\
-      (\_ \ltype A) \rightarrow B \quad &\Leftrightarrow \quad !(A \multimap B) \\
-      (\_ \utype A) \multimap B \quad &\Leftrightarrow \quad !A \multimap B \\
-      (\_ \ltype A) \multimap B \quad &\Leftrightarrow \quad A \multimap B
+    (\_ \utype A) \rightarrow B \quad & \Leftrightarrow \quad !(!A \multimap B) \\
+    (\_ \ltype A) \rightarrow B \quad & \Leftrightarrow \quad !(A \multimap B)  \\
+    (\_ \utype A) \multimap B \quad   & \Leftrightarrow \quad !A \multimap B    \\
+    (\_ \ltype A) \multimap B \quad   & \Leftrightarrow \quad A \multimap B
   \end{align}
-    \vspace{-1em}
   \label{correspondence}
@@
 The reason for these function types is that we have built the !
-exponential of linear logic directly into universe sorts.
+exponential of linear logic implicitly into universe sorts.
 The sort annotation $s$ here records the universe sort of the function domain.
@@
 is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into universe sorts, we implicitly limit !
+By integrating the exponential directly into universe sorts, we have limited !
 to only be canonically usable.
@@
 
-  Figure \ref{correspondence} illustrates the correspondence between CLC function types and Multiplicative Exponential Linear Logic (MELL) implications.
+Figure \ref{correspondence} illustrates the correspondence between CLC functions and Multiplicative Exponential Linear Logic (MELL) implications.
 MELL lacks counterparts for the cases (1), (3) if the co-domain $B$ is dependent on arguments of domain $A$.
@@
 
-  In practice, algorithmic type checking techniques allow users to omit writing sort indices.
-Our implementation employs bi-directional type checking and users never interact with indices in the surface syntax.
+Algorithmic type checking techniques allow users to omit writing sort indices for many situations in practice.
+Our implementation employs bi-directional type checking and users rarely interact with sort indices in the surface syntax.
 
 \subsection{Universes and Cumulativity}
-  CLC features two sorts of universes $U$ and $L$ with level indices $*, 0, 1, 2, \cdots$.
-$U_*$ is the impredicative universe of propositions.
-$U_k$ and $L_k$ are the predicative universes of non-linear types and linear types respectively.
+CLC features two sorts of universes $U$ and $L$ with level indices $0, 1, 2, \cdots$.
+$U_i$ and $L_i$ are the predicative universes of non-linear types and linear types respectively.
 The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
 Basically, terms with types that occur within $U_i$ are unrestricted in their usage.
-Terms with types that occur within $L_k$ are restricted to being used exactly once.
+Terms with types that occur within $L_i$ are restricted to being used exactly once.
 
@@
           \end{enumerate}
-      \item $U_* \preceq U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
+    \item $U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
     \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
@@
 Figure \ref{universe} illustrates the structure of our universe hierarchy.
-Each linear universe $L_k$ has $U_{k+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
-However, $L_k$ cumulates to $L_{k+1}$.
+Each linear universe $L_i$ has $U_{i+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
+However, $L_i$ cumulates to $L_{i+1}$.
 These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes, and subsequently losing track of linearity.
 
-  \begin{figure}[H]
-    \vspace{-1em}
+\begin{figure}[h]
   \caption{The Universe Hierarchy}
-    \vspace{0.8em}
+  \vspace{1em}
   \centering
   \begin{tikzcd}
-      U_* \arrow[r, ":"', "\preceq", dash] 
-      & U_0 \arrow[r, ":"', "\preceq", dash] 
+    U_0 \arrow[r, ":"', "\preceq", dash]
     & U_1 \arrow[r, ":"', "\preceq", dash]
@@
     & \cdots \\
-      & L_0 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash] 
+    L_0 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
     & L_1 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
@@
   \end{tikzcd}
-    \vspace{-0.5em}
   \label{universe}
@@
 \begin{figure}[h]
-    \vspace{-1em}
   \caption{Structural Judgments}
-    \vspace{-0.5em}
   \begin{mathpar}
@@
     \rname{Wf-L}
-      \\
 
@@
     \rname{Pure-U}
-      \\
 
@@
   \end{mathpar}
-    \vspace{-1em}
   \label{structural}
@@
 
-  \begin{definition} We formally define the terms \textit{non-linear}, \textit{linear}, \textit{unrestricted} and \textit{restricted}.
+\begin{definition} We formally define the terminology \textit{non-linear}, \textit{linear}, \textit{unrestricted} and \textit{restricted}.
   \begin{enumerate}
@@
 
-  The axiom rules \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
 In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
@@
 To support linear types, we must restrict weakening to non-linear variables.
-This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{Prop-Axiom}, \rname{U-Axiom}, \rname{L-Axiom}.
-From \rname{L-Axiom} we can see that the universe of linear types $L_k$ is an inhabitant of $U_{k+1}$.
+This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{U-Axiom}, \rname{L-Axiom}.
+From \rname{L-Axiom} we can see that the universe of linear types $L_i$ is an inhabitant of $U_{i+1}$.
 This is inspired by Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
 
-  The \rname{Prop} rule is used for forming propositions.
-From the judgment $\Gamma \vdash A : U_i$ we can see that \rname{Prop} allows impredicative quantification over non-linear types of arbitrary level $i$.
-The judgment $\Gamma, x \utype A \vdash B : U_*$ asserts that the co-domain must be in the impredicative universe $U_*$.
-The final resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_*$ has sort $U$, indicating that terms with type $(x \utype A) \rightarrow B$ enjoy unrestricted usage.
-
 \begin{figure}[h]
@@
     { \pure{\Gamma} }
-      { \Gamma \vdash U_* : U_0 } 
-      \rname{Prop-Axiom}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash s_k : U_{k+1} } 
+    { \Gamma \vdash s_i : U_{i+1} }
     \rname{Sort-Axiom}
@@
       \Gamma \vdash A : U_i \\
-        \Gamma, x \utype A \vdash B : U_* }
-      { \Gamma \vdash (x \utype A) \rightarrow B : U_* } 
-      \rname{Prop}
-      \\
+      \Gamma, x \utype A \vdash B : s_i }
+    { \Gamma \vdash (x \utype A) \rightarrow B : U_i }
+    \rname{U$\rightarrow$}
 
@@
     { \pure{\Gamma} \\
-        \Gamma \vdash A : U_k \\ 
-        \Gamma, x \utype A \vdash B : s_k }
-      { \Gamma \vdash (x \utype A) \rightarrow B : U_k } 
-      \rname{U-Prod}
+      \Gamma \vdash A : L_i \\
+      \Gamma \vdash B : s_i \\
+      x \notin \Gamma }
+    { \Gamma \vdash (x \ltype A) \rightarrow B : U_i }
+    \rname{L$\rightarrow$}
 
@@
     { \pure{\Gamma} \\
-        \Gamma \vdash A : L_k \\ 
-        \Gamma \vdash B : s_k \\
+      \Gamma \vdash A : U_i \\
+      \Gamma, x \utype A \vdash B : s_i }
+    { \Gamma \vdash (x \utype A) \multimap B : L_i }
+    \rname{U$\multimap$}
+
+    \inferrule
+    { \pure{\Gamma} \\
+      \Gamma \vdash A : L_i \\
+      \Gamma \vdash B : s_i \\
       x \notin \Gamma }
-      { \Gamma \vdash (x \ltype A) \rightarrow B : U_k } 
-      \rname{L-Prod}
-      \\
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : U_k \\ 
-        \Gamma, x \utype A \vdash B : s_k }
-      { \Gamma \vdash (x \utype A) \multimap B : L_k } 
-      \rname{U-Lolli}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : L_k \\ 
-        \Gamma \vdash B : s_k \\
-        x \notin \Gamma }
-      { \Gamma \vdash (x \ltype A) \multimap B : L_k } 
-      \rname{L-Lolli}
+    { \Gamma \vdash (x \ltype A) \multimap B : L_i }
+    \rname{L$\multimap$}
   \end{mathpar}
@@
 
-  The \rname{U-Prod} rule is used for forming non-linear function types with non-linear domains.
-This is evident from the judgment $\Gamma \vdash A : U_k$.
-Due to the fact that $A$ is in the predicative non-linear universe $U_k$, terms of type $A$ have unrestricted usage.
-The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_k$.
-The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_k$, we see that the overall type is a non-linear type.
+The \rname{U$\rightarrow$} rule is used for forming non-linear function types with non-linear domains.
+This is evident from the judgment $\Gamma \vdash A : U_i$.
+Due to the fact that $A$ is in the non-linear universe $U_i$, terms of type $A$ have unrestricted usage.
+The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_i$.
+The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_i$ and $L_i$.
+From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_i$, we see that the overall type is a non-linear type.
 The term level $\lambda$-abstractions of these types can be used unrestrictedly.
 
-  The \rname{L-Prod} rule is used for forming non-linear function types with linear domains.
-From the judgment $\Gamma \vdash A : L_k$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
-Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B : s_k$.
-Like \rname{U-Prod}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_k$ and $L_k$.
-The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_k$.
+The \rname{L$\rightarrow$} rule is used for forming non-linear function types with linear domains.
+From the judgment $\Gamma \vdash A : L_i$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
+Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B : s_i$.
+Like \rname{U$\rightarrow$}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
+The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_i$.
 Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
@@
 
-  The \rname{U-Lolli}, \rname{L-Lolli} rules are similar to \rname{U-Prod}, \rname{L-Prod} in spirit.
+The \rname{U$\multimap$}, \rname{L$\multimap$} rules are similar to \rname{U$\rightarrow$}, \rname{L$\rightarrow$} in spirit.
 The dependency considerations for domain $A$ and co-domain $B$ are exactly the same.
-The main difference between \rname{U-Lolli}, \rname{L-Lolli} and \rname{U-Prod}, \rname{L-Prod} is the universe sort of the resulting type.
-The sorts of the function types formed by \rname{U-Lolli}, \rname{L-Lolli} are $L$, meaning they are linear function types.
+The main difference between \rname{U$\multimap$}, \rname{L$\multimap$} and \rname{U$\rightarrow$}, \rname{L$\rightarrow$} is the universe sort of the resulting type.
+The sorts of the function types formed by \rname{U$\multimap$}, \rname{L$\multimap$} are $L$, meaning they are linear function types.
 The term level $\lambda$-abstractions of these types must be used exactly once.
@@
     \rname{L-Var}
-      \\
 
@@
       \Gamma, x \stype{s} A \vdash n : B }
-      { \Gamma \vdash \lambda x . n : (x \stype{s} A) \rightarrow B }
-      \rname{Prod-$\lambda$}
-      \\
+    { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \rightarrow B }
+    \rname{$\lambda$$\rightarrow$}
 
@@
       \Gamma, x \stype{s} A \vdash n : B }
-      { \Gamma \vdash \lambda x . n : (x \stype{s} A) \multimap B }
-      \rname{Lolli-$\lambda$}
-      \\
+    { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \multimap B }
+    \rname{$\lambda$$\multimap$}
 
@@
     { \Gamma \vdash m\ n : B[n/x] }
-      \rname{U-Prod-App}
+    \rname{App-U$\rightarrow$}
 
@@
     { \Gamma \vdash m\ n : B[n/x] }
-      \rname{L-Prod-App} 
-      \\
+    \rname{App-L$\rightarrow$}
 
@@
     { \Gamma \vdash m\ n : B[n/x] }
-      \rname{U-Lolli-App}
+    \rname{App-U$\multimap$}
 
@@
     { \Gamma \vdash m\ n : B[n/x] }
-      \rname{L-Lolli-App} 
-      \\
+    \rname{App-L$\multimap$}
 
@@
     { \Gamma \vdash m : B }
-      \rname{Conv}
+    \rname{Conversion}
   \end{mathpar}
@@
 
-  In \rname{Prod-$\lambda$}, the function type being addressed has form $(x \stype{s} A) \rightarrow B$.
+In \rname{$\lambda$$\rightarrow$}, the function type being addressed has form $(x \stype{s} A) \rightarrow B$.
 $\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
@@
 
-  In contrast to \rname{Prod-$\lambda$}, the \rname{Lolli-$\lambda$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
+In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
 Due to the fact these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
@@
 
-  For \rname{U-Prod-App}, domain $A$ is a non-linear type, as seen by its annotation in $(x \utype A) \rightarrow B$.
+For \rname{App-U$\rightarrow$}, domain $A$ is a non-linear type, as seen by its annotation in $(x \utype A) \rightarrow B$.
 Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
@@
 
-  Now for \rname{L-Prod-App}, domain $A$ is a linear type, as seen by its annotation in $(x \ltype A) \rightarrow B$.
+Now for \rname{App-L$\rightarrow$}, domain $A$ is a linear type, as seen by its annotation in $(x \ltype A) \rightarrow B$.
 Intuitively, this tells us that $x$ must be used once within the body of $m$.
@@
 
-  The rules \rname{U-Lolli-App}, \rname{L-Lolli-App} follow the same considerations as \rname{U-Prod-App}, \rname{L-Prod-App}. Additional conditions are not required for these two rules to be sound.
-
-  Finally, the \rname{Conv} rule allows judgment $\Gamma \vdash m : A$ to convert to judgment $\Gamma \vdash m : B$ if $B$ is a valid type in context $\overline{\Gamma}$.
-Futhermore, $A$ must be a subtype of $B$ satisfying the cumulativity relation $A \preceq B$.
-This rule gives rise to large eliminations (compute types from terms), as computations embedded at the type level can convert to canonical types.
+The rules \rname{App-U$\multimap$}, \rname{App-L$\multimap$} follow the same considerations as \rname{App-U$\rightarrow$}, \rname{App-L$\rightarrow$}.
+Additional conditions are not required for these two rules to be sound.
+
+Finally, the \rname{Conversion} rule allows judgment $\Gamma \vdash m : A$ to convert to judgment $\Gamma \vdash m : B$ if $B$ is a valid type in context $\overline{\Gamma}$.
+Futhermore, $A$ must be a subtype of $B$ satisfying the relation $A \preceq B$.
+This rule gives rise to large eliminations (compute types from terms), as computations embedded at the type level can reduce to canonical types.
 
 \subsection{Equality and Reduction} \label{reduction}
-  The operational semantics and definitional $\beta$-equality of CLC are presented in Figure \ref{parallel}, all of which are entirely standard.
-  \begin{figure}[h]
-    \caption{Equality and Parallel Reduction}
+The operational semantics and $\beta$-equality of CLC terms are presented in Figure \ref{red}, all of which are entirely standard.
+
+As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
+
+\begin{figure}[h]
+  \caption{Equality and Reduction}
   \begin{mathpar}
@@
     {  }
-      { (\lambda x.m)\ n \pstep m[n/x] }
-      \rname{P-$\beta$}
+    { (\lambda x \stype{s}A.m)\ n \pstep m[n/x] }
+    \rname{Step-$\beta$}
+
+    \inferrule
+    { A \step A' }
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A' .m }
+    \rname{Step-$\lambda$L}
 
@@
     { m \step m' }
-      { \lambda x.m \step \lambda x.m' }
-      \rname{Step-$\lambda$}
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A.m' }
+    \rname{Step-$\lambda$R}
 
@@
     { (x \stype{s} A) \rightarrow B \step (x \stype{s} A') \rightarrow B }
-      \rname{Step-ProdL}
+    \rname{Step-L$\rightarrow$}
 
@@
     { (x \stype{s} A) \rightarrow B \step (x \stype{s} A) \rightarrow B' }
-      \rname{Step-ProdR}
+    \rname{Step-R$\rightarrow$}
 
@@
     { (x \stype{s} A) \multimap B \step (x \stype{s} A') \multimap B }
-      \rname{Step-LolliL}
+    \rname{Step-L$\multimap$}
 
@@
     { (x \stype{s} A) \multimap B \step (x \stype{s} A) \multimap B' }
-      \rname{Step-LolliR}
+    \rname{Step-R$\multimap$}
 
@@
   \end{mathpar}
-    \label{parallel}
-    \Description{}
-  \end{figure}
-  As we have discussed previously, the elimination of the explicit !
-exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
+  \label{red}
+  \Description{}
+\end{figure}
 
@@
 First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
-Next, we show that CLC is a valid linear type theory through the \textit{linearity} theorem.
+Next, we show that CLC is an effective linear type theory through the \textit{linearity} theorem.
 Furthermore, we show that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
@@
 
-  Our design of integrating ! into universe sorts removes the need for ! manipulating syntax and semantics. The following substitution lemmas are directly proved by induction on typing derivations.
+Our design of integrating ! into universe sorts removes the need for ! manipulating syntax and semantics. The following substitution lemmas are directly proved by induction on typing derivation.
 
@@
 \subsubsection{Type Soundness}
-  In order to prove subject reduction, we first prove the propagation theorem.
-The main purpose of propagation is to lower types down to the term level, enabling the application of various inversion lemmas.
+In order to prove subject reduction, we first prove the validity theorem. The main purpose of validity is to lower types down to the term level, enabling the application of various inversion lemmas.
 
 \begin{theorem}
-    Propagation.
+  Validity.
 For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exist some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
-    \label{propagation}
+  \label{validity}
 \end{theorem}
@@
 
-  Before we proceed, we first prove the seemingly obvious narity lemma.
-
-  \begin{lemma} 
-    Narity. If $\Gamma \vdash m : A$ is a valid judgment, for any $x \notin \Gamma$, there is $occurs(x, m) = 0$.
-  \end{lemma}
-
-  For cases with branched syntax in the linearity theorem, such as application $(m\ n)$, the $occurs$ function sums up the occurrences of variable $x$ in branches $m$ and $n$.
-The narity lemma is used to prove that either $occurs(x, m) = 1 \wedge occurs(x, n) = 0$ or $occurs(x, m) = 0 \wedge occurs(x, n) = 1$ is true.
-
 \begin{theorem}
@@
 \begin{theorem}
-    Promotion.
-If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
+  Promotion. If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
 \end{theorem}
@@
   \begin{align*}
-      \erase{x} &= x \\
-      \erase{U_*} &= Prop \\
-      \erase{U_k} &= Type_k \\
-      \erase{L_k} &= Type_k \\
-      \erase{(x \stype{s} A) \rightarrow B} &= (x : \erase{A}) \rightarrow \erase{B} \\
-      \erase{(x \stype{s} A) \multimap B} &= (x : \erase{A}) \rightarrow \erase{B} \\
-      \erase{\lambda x.n} &= \lambda x.\erase{n} \\
-      \erase{m\ n} &= \erase{m}\ \erase{n}
+    \erase{x}                             & = x                                     \\
+    \erase{U_i}                           & = Type_i                                \\
+    \erase{L_i}                           & = Type_i                                \\
+    \erase{(x \stype{s} A) \rightarrow B} & = (x : \erase{A}) \rightarrow \erase{B} \\
+    \erase{(x \stype{s} A) \multimap B}   & = (x : \erase{A}) \rightarrow \erase{B} \\
+    \erase{\lambda x\stype{s}A.n}         & = \lambda x : \erase{A}.\erase{n}       \\
+    \erase{m\ n}                          & = \erase{m}\ \erase{n}
   \end{align*}
@@
   \begin{align*}
-      \erase{\epsilon} &= \epsilon \\
-      \erase{\Gamma, x \stype{s} A} &= \erase{\Gamma}, x : \erase{A}
+    \erase{\epsilon}              & = \epsilon                      \\
+    \erase{\Gamma, x \stype{s} A} & = \erase{\Gamma}, x : \erase{A}
   \end{align*}
@@
 We prove the following lemma to commute erasure and substitution whenever needed.
+
 \begin{lemma}
@@
 For the following lemma, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
+
 \begin{theorem} \label{preserve}
-    Erasure preserves reduction.
-For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  Reduction Erasure. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
 \end{theorem}
@@
 \subsubsection{Backwards Compatibility}
-  To show that CLC is backwards compatible with CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
+To show that CLC is backwards compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
 
@@
   \begin{align*}
-      \lift{x} &= x \\
-      \lift{Prop} &= U_* \\
-      \lift{Type_k} &= U_k \\
-      \lift{(x : A) \rightarrow B} &= (x \utype \lift{A}) \rightarrow \lift{B} \\
-      \lift{\lambda x.n} &= \lambda x.\lift{n} \\
-      \lift{m\ n} &= \lift{m}\ \lift{n}
+    \lift{x}                     & = x                                        \\
+    \lift{Type_i}                & = U_i                                      \\
+    \lift{(x : A) \rightarrow B} & = (x \utype \lift{A}) \rightarrow \lift{B} \\
+    \lift{\lambda x : A.n}       & = \lambda x\utype\lift{A}.\lift{n}         \\
+    \lift{m\ n}                  & = \lift{m}\ \lift{n}
   \end{align*}
@@
   \begin{align*}
-      \lift{\epsilon} &= \epsilon \\
-      \lift{\Gamma, x : A} &= \lift{\Gamma}, x \utype \lift{A}
+    \lift{\epsilon}      & = \epsilon                         \\
+    \lift{\Gamma, x : A} & = \lift{\Gamma}, x \utype \lift{A}
   \end{align*}
@@
 \begin{theorem}
-    Lift preserves reduction.
-For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+  Lift preserves reduction. For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
 \end{theorem}
 
-  \begin{theorem}
+\begin{theorem} \label{lift}
   Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
@@
 The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domain of function types are annotated with sort $U$.
-In practice, even these annotations can be omitted if algorithmic type checking is able to infer them from context.
-
-  \section{Extensions}
-  Readers familiar with the literature of linear logic or linear type theory will have noticed at this point that CLC does not possess the iconic $\otimes$ connective.
-Though it is entirely possible to encode $\otimes$ in core CLC through a linear Church encoding, proofs and programs written in this style will be incredibly difficult to maintain.
-
-  \begin{figure}[h]
-  \vspace{-1em}
-  \caption{Inductive Type Definition and Elimination}
-  \begin{lstlisting}
-  (* Sigma Definition *)
-  Inductive Sigma (A: U) (F: A -> U): U :=
-  | pair : (x: A) -> F x -> Sigma A F.
-
-  (* Tensor Definition *)
-  Inductive Tensor (A: L) (B: L): L :=
-  | tpair: A -> B -> Tensor A B.
-
-  (* FTensor Definition *)
-  Inductive FTensor (A: U) (F: A -> L): L :=
-  | fpair: (x: A) -> F x -> FTensor A F.
-
-  (* Sigma Elimination *)
-  match m with
-  | pair x y => (* clause body *)
-  end
-
-  (* Tensor Elimination *)
-  match m with
-  | tpair x y => (* clause body *)
-  end
-
-  (* FTensor Elimination *)
-  match m with
-  | fpair x y => (* clause body *)
-  end
-  \end{lstlisting}
-  \vspace{-1em}
-  \label{inductive}
-  \Description{}
-  \end{figure}
-
-  A major design goal of CLC is to support an ergonomic interface for users to define inductive types \cite{inductive,cic} as an extension to the core theory.
-The presence of universe sorts $U$ and $L$ allow users to declare inductive types to be non-linear or linear.
-The lack of !
-exponential simplifies the construction and elimination of inductive terms through direct use of constructors and pattern-matching.
-The $\otimes$ connective becomes definable as an inductive type.
-
-  Figure \ref{inductive} is an excerpt taken from our implementation's prelude, with syntax heavily inspired by Coq.
-
-  \begin{itemize}
-    \item \texttt{Sigma} defines the standard dependent sum type.
-Since first component of the pair constructor is of non-linear type, the type of the second component is allowed to depend on the first component.
-    \item \texttt{Tensor} defines the $\otimes$ connective for CLC.
-Both components of the \texttt{tpair} constructor are of linear types.
-When a \texttt{Tensor} term is eliminated through pattern-matching, its first and second components must be used exactly once in the body of its pattern clause.
-    \item \texttt{FTensor} fulfills a role similar to Krishnaswami et al.'s $F$ adjoint connective \cite{neel15}.
-The type of \texttt{FTensor}'s first component is non-linear.
-The linear type of its second component is allowed to depend on the first component.
-Due to the fact that \texttt{FTensor}'s second component is linear, the entire \texttt{FTensor} must be linear as well, otherwise duplication of restricted variables may occur.
-  \end{itemize}
-
-  Due to the complexity of checking soundness for dependent linear type definitions (such as \texttt{FTensor}) and pattern-matching based elimination, we have not verified the algorithm employed by our implementation.
-
-  Figure \ref{formation}, Figure \ref{introduction} and Figure \ref{elimination} presents the formation, introduction and elimination rules for select inductive types respectively.
-We will use these inductive types during the construction of examples.
-
-  \begin{figure}[H]
-    \caption{Select Formation Rules}
-    \begin{mathpar}
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash \unit : U_0 }
-      \rname{Unit-Form}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash \mathbb{N} : U_0 }
-      \rname{$\mathbb{N}$-Form}
-
+In practice, these annotations can often be omitted if algorithmic type checking is able to infer them from context.
+
+\section{The Extension to CILC}
+Although the core CLC language is extremely expressive, formally shown by Theorem \ref{lift} to be at least as expressive as the predicative CC$\omega$, the complexity of Church-style encodings to do so will quickly become unmanageable.
+To address this, we extend CLC with rules for defining inductive linear types in the style of CIC.
+
+We have formalized CILC in our Coq development and proven its soundness.
+
+\subsection{Syntax of CILC}
+\begin{figure}[h]
+  \caption{Syntax}
+  \begin{grammar}
+    <$C,I,P,Q,f$> ::= ... \hspace*{10em} expressions
+    \indalt $\ind{s}(X : A)\{C_1 \sep ... \sep C_k\}$
+    \indalt $\constr(k, I)$
+    \indalt $\case(m, Q)\{f_1 \sep ... \sep f_k \}$
+    \indalt $\dcase(m, Q)\{f_1 \sep ... \sep f_k \}$
+    \indalt $\fix f : A := m$
+  \end{grammar}
+  \Description{}
+  \label{syntax2}
+\end{figure}
+
+As shown in Figure \ref{syntax2}, CILC extends CIC with for the introduction and elimination of inductive types.
+Ind is used during the formation of new inductive types, and Constr is used for introducing their constructors.
+Case and DCase are non-dependent and dependent case eliminators respectively.
+CIC-style primitive recursion has been factored out of case eliminators in favor of a Coq-style Fix construct.
+The flexibility of an independent Fix construct allows for a cleaner formalization of inductive linear type elimination.
+
+\subsection{Arity}
+For type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as depicted in Figure \ref{arity}.
+
+\begin{figure}[h]
+  \caption{Arity}
+  \begin{mathpar}
+    \inferrule
+    { }
+    { arity(s,s) }
+    \rname{A-Sort}
+
+    \inferrule
+    { arity(A, s) }
+    { arity((x \utype M) \rightarrow A, s)}
+    \rname{A$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{arity}
+\end{figure}
+
+\begin{definition}
+  For some arity type $A$ and sort $s'$, the term $A\lrangle{s'}$ is inductively defined over the structure of $A$ as follows.
+  \begin{align*}
+    ((x \utype M) \rightarrow A)\lrangle{s'} & = (x \utype M) \rightarrow A\lrangle{s'} \\
+    (s)\lrangle{s'}                          & = s'
+  \end{align*}
+\end{definition}
+
+\begin{definition}
+  For some arity type $A$, term $I$ and sort $s'$, the term $A\{I, s'\}$ is inductively defined over the structure of $A$ as follows.
+  \begin{align*}
+    ((x \utype M) \rightarrow A)\lrangle{I, s'} & = (x \utype M) \rightarrow A\lrangle{(I\ x), s'} \\
+    (s)\lrangle{I, s'}                          & = (\_ \utype I) \rightarrow s'
+  \end{align*}
+\end{definition}
+
+\subsection{Strict Positivity}
+For type $P$ and variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as depicted in Figure \ref{positive}.
+
+We say that $X$ occurs strictly positive in $P$ if the judgment $postive(P,X)$ is valid.
+
+\begin{figure}[h]
+  \caption{Strict Positivity}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { positive((X\ m_1...m_k), X) }
+    \rname{Pos-X}
+
+    \inferrule
+    { X \notin M \\ positive(P, X)}
+    { positive((x \stype{s} M) \rightarrow P, X) }
+    \rname{Pos$\rightarrow$}
+
+    \inferrule
+    { X \notin M \\ positive(P, X) }
+    { positive((x \stype{s} M) \multimap P, X) }
+    \rname{Pos$\multimap$}
+  \end{mathpar}
+  \Description{}
+  \label{positive}
+\end{figure}
+
+\subsection{Non-Linear Constructor}
+For type $C$ and variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{nlconstr}.
+
+Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not applied to restricted terms as this may cause duplication.
+
+\begin{figure}[h]
+  \caption{Non-linear Constructor}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { \ucons((X\ m_i...m_k), X) }
+    \rname{U-Constr-X}
+
+    \inferrule
+    { positive(P, X) \\ \ucons(C, X) }
+    { \ucons((\_ \utype P) \rightarrow C, X)}
+    \rname{U-Constr-pos}
+
+    \inferrule
+    { X \notin M \\ \ucons(C, X) }
+    { \ucons((x \utype M) \rightarrow C, X)}
+    \rname{U-Constr$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{nlconstr}
+\end{figure}
+
+\subsection{Linear Constructor}
+For type $C$ and variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{lconstr}.
+
+Linear constructors are used for constructing linear objects that must be used once. These constructors may be applied to linear terms as restricted usage prevent duplication and enforce single usage.
+
+An unapplied linear constructor is a non-linear entity as they can be created freely ``out of thin air".
+However, once a linear constructor has been partially applied to a linear term, its linearity is ``activated'' in rules \rname{L-Constr-Pos-2} and \rname{L-Constr-2$\rightarrow$}, essentially forcing the rest of its arguments to be fully applied.
+This is to prevent partially applied constructors from duplicating linear variables.
+
+\begin{figure}[h]
+  \caption{Linear Constructor}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { \lcons((X\ m_i...m_k), X) }
+    \rname{L-Constr-X}
+
+    \inferrule
+    { positive(P, X) \\ \lcons(C, X) }
+    { \lcons((\_ \utype P) \rightarrow C, X)}
+    \rname{L-Constr-Pos-1}
+
+    \inferrule
+    { X \notin M \\ \lcons(C, X) }
+    { \lcons((x \utype M) \rightarrow C, X)}
+    \rname{L-Constr-1$\rightarrow$}
+
+    \inferrule
+    { positive(P, X) \\ activation(C, X) }
+    { \lcons((\_ \ltype P) \rightarrow C, X)}
+    \rname{L-Constr-Pos-2}
+
+    \inferrule
+    { X \notin M \\ activation(C, X) }
+    { \lcons((x \ltype M) \rightarrow C, X)}
+    \rname{L-Constr-2$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{lconstr}
+\end{figure}
+
+For type $C$ and variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{activation}.
+Intuitively, $activation(C,X)$ ensures the linearity of $C$.
+
+\begin{figure}[h]
+  \caption{Activation}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { activation((X\ m_i...m_k), X) }
+    \rname{Act-X}
+
+    \inferrule
+    { positive(P, X) \\ activation(C, X) }
+    { activation((\_ \stype P) \multimap C, X)}
+    \rname{Act-Pos}
+
+    \inferrule
+    { X \notin M \\ activation(C, X) }
+    { activation((x \stype M) \multimap C, X)}
+    \rname{Act$\multimap$}
+  \end{mathpar}
+  \Description{}
+  \label{activation}
+\end{figure}
+
+\subsection{Introduction Rules}
+The formation of new inductive types and their constructors are presented in Figure \ref{ind}.
+
+\begin{figure}[h]
+  \caption{Inductive Type Formation}
+  \begin{mathpar}
+    \inferrule
+    { (\forall j = 1...n) \\ arity(A, s) \\ \scons(C_j, X) \\\\
+      \pure{\Gamma} \\
+      \Gamma \vdash A : U_i \\
+      \Gamma, X \utype A \vdash C_j : t_i }
+    { \Gamma \vdash \ind{s}(X : A)\{C_1|...|C_k\} : A }
+    \rname{Ind}
+
+    \inferrule
+    { 1 \leq i \leq k \\\\
+    \pure{\Gamma} \\
+    I := \ind{s}(X : A)\{C_1|...|C_k\} \\
+    \Gamma \vdash I :A }
+    { \Gamma \vdash \constr(i, I) : C_i[I/X] }
+    \rname{Constr}
+  \end{mathpar}
+  \Description{}
+  \label{ind}
+\end{figure}
+
+\subsection{Non-Dependent Case}
+
+\begin{definition}
+  For constructor type $C$ and variables $X$ and $Q$, the term $C\{X, Q\}$ is inductively defined over the structure of $C$ as follows.
+  \begin{align*}
+    ((\_ \stype{s} P) \rightarrow C)\{X, Q\} & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
+    ((\_ \stype{s} P) \multimap C)\{X, Q\}   & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
+    ((x \stype{s} M) \rightarrow C)\{X, Q\}  & = (x \stype{s} P) \multimap C\{X, Q\}  \\
+    ((x \stype{s} M) \multimap C)\{X, Q\}    & = (x \stype{s} P) \multimap C\{X, Q\}  \\
+    (X\ m_1...m_k)\{X, Q\}                   & = (Q\ m_1...m_k)
+  \end{align*}
+  We define the notation $C[I, P] := C\{X, Q\}[I/X, P/Q]$.
+\end{definition}
+
+The rule for non-dependent case elimination is presented in Figure \ref{case}. This elimination rule is non-dependent because the motive $Q$ may not depend on the discriminated term $m$ itself.
+
+\begin{figure}[h]
+  \caption{Non-dependent Case Elimination}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k_1) \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+    arity(A, s) \\\\
+    I := \ind{s}(X : A)\{C_1|...|C_{k_1}\} \\
+    \overline{\Gamma_2} \vdash Q : A\lrangle{s'} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_{k_2}) \\
+    \Gamma_2 \vdash f_i : C_i[I, Q] }
+    { \Gamma \vdash \case(m, Q)\{f_1|...|f_{k_1}\} : (Q\ a_1...a_{k_2}) }
+    \rname{Case}
+  \end{mathpar}
+  \Description{}
+  \label{case}
+\end{figure}
+
+\subsection{Dependent Case}
+
+\begin{definition}
+  For some non-linear constructor type $C$ and type variables $X$, $Q$ and $c$, the term $C\{X, Q, c\}$ is inductively defined over the structure of $C$ as follows.
+  \begin{align*}
+    ((\_ \utype P) \rightarrow C)\{X, Q, c\} & = (p \utype P) \multimap C\{X, Q, (c\ p)\} \\
+    ((x \utype M) \rightarrow C)\{X, Q, c\}  & = (x \utype M) \multimap C\{X, Q, (c\ x)\} \\
+    (X\ m_1...m_n)\{X, Q, c\}                & = (Q\ m_1...m_n\ c)
+  \end{align*}
+  We define the notation $C[I, P, t] := C\{X, Q, c\}[I/X, P/Q, t/c]$.
+\end{definition}
+
+Dependent elimination of non-linear inductive objects is presented in Figure \ref{dcase}.
+This elimination rule is dependent because the motive $Q$ may depend on the discriminated term $m$.
+For this very reason, $m$ is not allowed to have linear type as doing so may transport restricted variable into types.
+
+\begin{figure}[h]
+  \caption{Dependent Case Elimination}
+  \begin{mathpar}
+    \inferrule
+    { \pure{\Gamma_1} \\
+    (\forall i = 1...n) \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+    arity(A, U_i) \\\\
+    I := \ind{U}(X : A)\{C_1|...|C_n\} \\
+    \overline{\Gamma_2} \vdash Q : A\lrangle{I, s'} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_n) \\
+    \Gamma_2 \vdash f_i : C_i[I, Q, Constr(i, I)] }
+    { \Gamma \vdash \dcase(m, Q)\{f_1|...|f_n\} : (Q\ a_1...a_n\ m) }
+    \rname{DCase}
+  \end{mathpar}
+  \Description{}
+  \label{dcase}
+\end{figure}
+
+\subsection{Fixpoint}
+The typing rule of fix-points presented in Figure \ref{fix} is mostly standard.
+A syntactic guard condition is utilized to conservatively ensure termination.
+The fix-point term must have non-linear type in a pure context, otherwise recursive calls may cause duplicated usage of restricted variables.
+
+\begin{figure}[h]
+  \caption{Fixpoint}
+  \begin{mathpar}
     \inferrule
@@
       \Gamma \vdash A : U_i \\
-        \Gamma \vdash m : A \\
-        \Gamma \vdash n : A }
-      { \Gamma \vdash m =_A n : U_* }
-      \rname{=-Form}
-
-      \inferrule
-      { \pure{\Gamma} \\ 
-        \Gamma \vdash A : U_k \\ 
-        \Gamma x \utype A \vdash B : U_k  }
-      { \Gamma \vdash \Sigma x:A.B : U_k }
-      \rname{$\Sigma$-Form}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : L_k \\
-        \Gamma \vdash B : L_k }
-      { \Gamma \vdash A \otimes B : L_k }
-      \rname{$\otimes$-Form}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash A : U_k \\ 
-        \Gamma x \utype A \vdash B : L_k }
-      { \Gamma \vdash F x:A.B : L_k }
-      \rname{$F$-Form}
-    \end{mathpar}
-    \label{formation}
-    \Description{}
-  \end{figure}
-
-  \begin{figure}[H]
-    \vspace{-0.8em}
-    \caption{Select Introduction Rules}
-    \begin{mathpar}
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash () : \unit }
-      \rname{Unit-Intro}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash O : \mathbb{N} }
-      \rname{O-Intro}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash n : \mathbb{N} }
-      { \Gamma \vdash S\ n : \mathbb{N} }
-      \rname{S-Intro}
-
-      \inferrule
-      { \pure{\Gamma} \\ 
-        \Gamma \vdash A : U_i \\
-        \Gamma \vdash m : A }
-      { \Gamma \vdash \refl_A m : m =_A m }
-      \rname{=-Intro}
-
-      \inferrule
-      { \pure{\Gamma} \\
-        \Gamma \vdash m : A \\
-        \Gamma \vdash n : B[m/x] \\\\
-        \Gamma \vdash \Sigma x:A.B : U_k }
-      { \Gamma \vdash (m, n) : \Sigma x:A.B }
-      \rname{$\Sigma$-Intro}
-
-      \inferrule
-      { \Gamma_1 \vdash m : A \\
-        \Gamma_2 \vdash n : B \\ 
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\\\
-        \Gamma \vdash A \otimes B : L_k }
-      { \Gamma \vdash [m, n] : A \otimes B }
-      \rname{$\otimes$-Intro}
-
-      \inferrule
-      { \pure{\Gamma_1} \\
-        \Gamma_1 \vdash m : A \\
-        \Gamma_2 \vdash n : B[m/x] \\\\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-        \Gamma \vdash F x:A.B : L_k }
-      { \Gamma \vdash \{m, n\} : F x:A.B }
-      \rname{$F$-Intro}
-    \end{mathpar}
-    \vspace{-1em}
-    \label{introduction}
-    \Description{}
-  \end{figure}
-
-  \begin{figure}[h]
-    \vspace{-1em}
-    \caption{Select Elimination Rules}
-    \begin{mathpar}
-      \inferrule
-      { \Gamma_1 \vdash m : \unit \\ 
-        \Gamma_2 \vdash n : A \\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{()}{m}{n} : A }
-      \rname{Unit-Elim}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash \iter_\mathbb{N}^U :
-        (C \utype \mathbb{N} \rightarrow U_i) \rightarrow C\ O \rightarrow \\
-        ((n \utype \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)) \rightarrow
-        (n \utype \mathbb{N}) \rightarrow C\ n }
-      \rname{$\mathbb{N}$-U-Elim}
-
-      \inferrule
-      { \pure{\Gamma} }
-      { \Gamma \vdash \iter_\mathbb{N}^L :
-        (C \utype \mathbb{N} \rightarrow L_k) \rightarrow C\ O \multimap \\
-        ((n \utype \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)) \multimap
-        (n \utype \mathbb{N}) \multimap C\ n }
-      \rname{$\mathbb{N}$-L-Elim}
-
-      \inferrule
-      { \Gamma_1 \vdash m : \Sigma x:A.B \\ 
-        \Gamma_2, x \utype A, y \utype B \vdash n : C \\\\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{$(x, y)$}{$m$}{$n$} \vdash C }
-      \rname{$\Sigma$-Elim}
-
-      \inferrule
-      { \Gamma_1 \vdash m : A \otimes B \\ 
-        \Gamma_2, x \ltype A, y \ltype B \vdash n : C \\\\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{$[x, y]$}{$m$}{$n$} \vdash C }
-      \rname{$\otimes$-Elim}
-
-      \inferrule
-      { \Gamma_1 \vdash m : F x:A.B \\
-        \Gamma_2, x \utype A, y \ltype B \vdash n : C \\\\
-        \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-      { \Gamma \vdash \letin{$\{x, y\}$}{$m$}{$n$} \vdash C }
-      \rname{$F$-Elim}
-    \end{mathpar}
-    \vspace{-0.8em}
-    \label{elimination}
-    \Description{}
-  \end{figure}
-
-  Of all the rules presented here, perhaps the most surprising to readers familiar with linear types or dependent types is the \rname{$\mathbb{N}$-L-Elim} rule.
-This rule constructs a term with linear type through a primitive recursion principle.
-The $\multimap$ arrows protect the O-case element of linear type $C\ O$ from duplication by currying.
-The S-case function is of non-linear type $(n : \mathbb{N}) \rightarrow C\ n \rightarrow C\ (S\ n)$ because it must be applied repeatedly during iteration.
-
-  \section{Applications}
-  In this section, we demonstrate some of the possible applications of CLC through examples.
-We collapse the sort annotations on arrows $(x \utype A) \rightarrow B$ and $(x \ltype A) \rightarrow B$ into $(x : A) \rightarrow B$ whenever $A$'s sort is inferable from context.
-
-  \subsection{Stateful Reasoning}
-  We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
-  \begin{definition}
-    State axioms.
-    \begin{itemize}
-      \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U_k) \rightarrow A \rightarrow L_k$ \\
-      The $\mapsto$ is a linear type constructor.
-It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
-Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
-      \item $\new : (A : U_k) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
-      $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
-      \item $\free : (A : U_k) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \mathbf{1}$ \\
-      $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use.
-This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
-      \item $\get : (A : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
-      $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
-The type of $\get$'s output is the rather complicated $F \_: (\Sigma x.
-x =_A m).
-(l \mapsto_A m)$.
-This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored.
-The proof $(l \mapsto_A m)$ is returned unchanged as well.
-      \item $\set : (A\ B : U_k) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
-      $\set$ has the most complicated type of all the axioms presented here.
-Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term.
-This is the so called "strong update" operation.
-The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
-    \end{itemize}
-  \end{definition}
-  Due to the fact that $(l \mapsto_A m)$ is a linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
-This is the same non-sharing principle that empowers Separation Logic.
-Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CLC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
-
-  \subsection{Internalization of Effect Laws}
-  The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CLC framework.
-
-  \begin{enumerate}
-    \item Performing $\get$ twice on the same address will produce two terms that are provably equal.
-    \item After using $\set$ to update the store at an address $l$ to term $m$, performing $\get$ again on the same address will retrieve $m$.
-  \end{enumerate}
-
-  Figure \ref{law1} is an excerpt taken from our implementation.
-It demonstrates the internalization of Law 1.
-The usage of implicit parameters allows one to omit writing much of the redundant arguments that could be inferred from context.
-Here, $\new$ allocates a store for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-When $\get$ is applied to $c$ twice, the equality proofs \textit{pf1} $: (1 =_\mathbb{N} x)$ and \textit{pf2} $: (1 =_\mathbb{N} y)$ generated by $\get$ allow one to create proof \textit{pf} $: (x =_\mathbb{N} y)$ that the retrieved values $x$ and $y$ are equal.
-Finally, due to $c$ having linear type, it is required for it to be freed at the very end.
-
-  \begin{figure}[h]
-  \vspace{-0.6em}
-  \caption{Law 1}
-  \begin{lstlisting}
-  let [ _, c ] := new _ 1 in
-  let [ xeq, c ] := get _ _ _ c in
-  let [ yeq, c ] := get _ _ _ c in
-  let ( x, pf1 ) := xeq in
-  let ( y, pf2 ) := yeq in
-  let pf1 := Eq_sym _ _ _ pf1 in
-  let pf : Eq _ x y := 
-    Eq_trans _ _ _ _ pf1 pf2 
-  in free _ _ _ c.
-  \end{lstlisting}
-  \vspace{-0.6em}
-  \label{law1}
-  \Description{}
-  \end{figure}
-
-  Figure \ref{law2} demonstrates the internalization of Law 2.
-Again, implicit parameters fill in the annoying bureaucratic arguments when they can be inferred.
-Here, a store is allocated by $\new$ for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-After using $\set$ to update the store at $l$ to 2 and then performing $\get$ on $l$, one can create proof \textit{pf} $: (2 =_\mathbb{N} z)$ that the value $z$ retrieved by $\get$ is exactly equal to 2.
-
-  \begin{figure}[h]
-  \vspace{-0.6em}
-  \caption{Law 2}
-  \begin{lstlisting}
-  let [ _, c ] := new _ 1 in
-  let c := set _ _ _ _ c 2 in
-  let [ zeq, c ] := get _ _ _ c in
-  let ( z, pf1 ) := zeq in
-  let pf : Eq _ 2 z := pf1 in
-  free _ _ _ c.
-  \end{lstlisting}
-  \vspace{-0.6em}
-  \label{law2}
-  \Description{}
-  \end{figure}
-
-  Though we have not implemented program extraction for CLC yet, our preliminary experiments show the feasibility of compiling the state axioms to constant time imperative operations after proof erasure.
-This can greatly improve the runtime efficiency of CLC over pure functional languages when used for programming.
-More complex examples such as safe array indexing are available in the first author's Github repository.
-
-  \subsection{Mutable Data}
-  Immutable data in functional programming languages are easy to reason about due their values staying constant over their lifespan.
-However, this convenience comes at the price of efficiency when attempting to perform "updates" on existing data.
-In order to update immutable data, an entirely new copy must be created with new entries substituted in for the original ones.
-So after update, there are two data structures in memory, the original one and the updated one.
-A garbage collector is necessary to reclaim memory allocated for the original copy.
-
-  Due to the non-sharing property of linear types, it is possible to safely update linear data in-place without encountering synchronization issues.
-Garbage collection is no longer necessary as extra copies of data are not created.
-Languages such as ATS that implement mutable linear data have tiny runtimes when compared to immutable garbage collector based ones such as OCaml or Haskell, making them suitable for programming in computationally constrained settings such as embedded programming.
-
-  \begin{figure}[h]
-  \caption{Linear Lists}
-  \begin{lstlisting}
-  Inductive list (A : U) : L :=
-  | nil : list A
-  | cons : A -> list A -> list A.
-
-  Fixpoint append (A : U) : list A -o 
-    list A -o (list A -o list A) -o list A 
-  :=
-    fun ls1 ls2 k =>
-      match ls1 with
-      | nil => k ls2
-      | (cons h t) as _cons_ =>
-        append _ t ls2 
-          (fun res => k (_cons_ h res))
-      end.
-  \end{lstlisting}
-  \label{list}
-  \Description{}
-  \end{figure}
-
-  Figure \ref{list} presents a CPS append function for linear lists.
-In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
-Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
-Basically, each \texttt{_cons_} must be applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
-Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
-Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.
-
-  \section{Related Work}
-  Linear types are a class of type systems inspired by Girard' substructural Linear Logic \cite{girard}.
-Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
-Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
-A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
-Benton \cite{benton1994} investigates the ramifications of the !
-exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
-Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
-
-  Over the years, work has been done to enrich linear type theories with dependent types.
-Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
-Krishnaswami et al.
-present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
-Luo et al.
-\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
-The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
-
-  \section{Future Work}
-  The integration of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
-
-  On the theoretical side, we intend to investigate the semantic models of CLC, with the goal of explaining the natural cohesion between its rules.
-The impredicative nature of the \textit{Prop}-like universe $U_*$ hints at the possibility of extending dependency to terms of linear types in a philosophically satisfying way.
-Linear types also appear to augment parametricity, which can produce stronger free theorems.
-
-  On the application side, we intend to bolster our implementation with more features, making it a full-featured theorem prover/programming language.
-We aim to fully verify the typechecking algorithms and extensions employed by the current implementation.
-The unification algorithm used to resolve implicit parameters is extremely ad-hoc, as the problem of unification for linear types has not thoroughly studied, we hope to improve this situation.
+      \Gamma, f \utype A \vdash m : A \\
+      \text{guard condition} }
+    { \Gamma \vdash \fix f : A := m : A }
+    \rname{fix}
+  \end{mathpar}
+  \Description{}
+  \label{fix}
+\end{figure}
+
+\subsection{Extended Reduction}
+The reduction semantics for the extended language is standard, we only present the most interesting $\iota$-reduction cases.
+
+\begin{figure}[h]
+  \caption{$\iota$-Reductions}
+  \begin{mathpar}
+    \inferrule
+    { }
+    { \case((\constr(i, I)\ a_1...a_n),Q)\{f_1|...f_k\} \\\\
+      \leadsto (f_i\ a_1...a_k) }
+    \rname{Step-Case-$\iota$}
+
+    \inferrule
+    { }
+    { \dcase((\constr(i, I)\ a_1...a_k),Q)\{f_1|...f_k\} \\\\
+      \leadsto (f_i\ a_1...a_k) }
+    \rname{Step-DCase-$\iota$}
+
+    \inferrule
+    { }
+    { \fix f : A := m \leadsto m[(\fix f : A := m)/f] }
+    \rname{Step-Fix-$\iota$}
+  \end{mathpar}
+  \Description{}
+  \label{stepfix}
+\end{figure}
+
+\subsection{Linear Connectives}
+In Figure \ref{connectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
+For any term with inductive linear type, linearity will ultimately force the term to be eliminated by a Case expression.
+
+\begin{figure}[h]
+  \caption{Inductively Defined Linear Connectives}
+  \begin{align*}
+    \_\otimes\_ :=\
+     & \ind{L}(X : (x \utype L_i) \rightarrow (y \utype L_i) \rightarrow L)\{        \\
+     & \sep\ (A \utype L_i) \rightarrow (B \utype L_i)                               \\
+     & \qquad \rightarrow (\_ \ltype A)  \multimap (\_ \ltype B) \multimap (X\ A\ B) \\
+     & \}
+    \\
+    \lrangle{\_,\_} := \
+     & \constr(1, \otimes)
+    \\\\
+    \_\oplus\_ := \
+     & \ind{L}(X : (x \utype L_i) \rightarrow (y \utype L_i) \rightarrow L)\{        \\
+     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
+    \rightarrow (\_ \ltype A)  \multimap (X\ A\ B)                                   \\
+     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
+    \rightarrow (\_ \ltype B)  \multimap (X\ A\ B)                                   \\
+     & \}
+    \\
+    \inl(\_) := \
+     & \constr(1, \oplus)
+    \\
+    \inr(\_) := \
+     & \constr(2, \oplus)
+  \end{align*}
+  \Description{}
+  \label{connectives}
+\end{figure}
 
@@
 \bibliography{../ref}
-
 \end{document}

### 2021-12-02 writing/draft5/draft5.tex
writing

@@
 \usepackage{enumitem}
+\usepackage{minted}
 
@@
 
+\subsection{Non-Linear Connectives}
+Non-linear inductive types have been extensively studied since their inception.
+We will omit discussion of non-linear inductive types here as all inductive definitions in the predicative fragment of CIC are available in CILC.
+
 \subsection{Linear Connectives}
-In Figure \ref{connectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
+In Figure \ref{lconnectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
 For any term with inductive linear type, linearity will ultimately force the term to be eliminated by a Case expression.
@@
     \_\otimes\_ :=\
-     & \ind{L}(X : (x \utype L_i) \rightarrow (y \utype L_i) \rightarrow L)\{        \\
+     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
      & \sep\ (A \utype L_i) \rightarrow (B \utype L_i)                               \\
@@
     \_\oplus\_ := \
-     & \ind{L}(X : (x \utype L_i) \rightarrow (y \utype L_i) \rightarrow L)\{        \\
+     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
      & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
@@
   \Description{}
-  \label{connectives}
-\end{figure}
+  \label{lconnectives}
+\end{figure}
+
+\subsection{Mixed Connectives}
+In Figure \ref{mconnectives}, we demonstrate how the $F$ connective of LNL$_D$ \cite{neel15} can be expressed as an inductive linear type.
+The interesting property of $F$ is that the type of its linear component is dependent on the non-linear component.
+
+\begin{figure}[h]
+  \caption{Linear Non-Linear Connective}
+  \begin{align*}
+    F := \
+     & \ind{L}(X : (A \utype U_i) \rightarrow
+    (B \utype (\_ \utype A) \rightarrow L_i) \rightarrow L_i)\{ \\
+     & \sep \ (A \utype U_i) \rightarrow
+    (B \utype (\_ \utype A) \rightarrow L_i)                    \\
+     & \qquad \rightarrow (m \utype A) \rightarrow
+    (\_ \ltype B\ m) \multimap (X\ A\ B)                        \\
+     & \}
+    \\
+    [\_,\_] :=\
+     & \constr(1, F)
+  \end{align*}
+  \Description{}
+  \label{mconnectives}
+\end{figure}
+
+\section{Examples}
+We shall write examples in the concrete syntax of our implementation.
+We deliberately chose to model our concrete syntax after Coq such that it would be easier for readers already familiar with Coq to adapt to our syntax.
+
+\subsection{Stateful Reasoning}
+We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
+\begin{definition}
+  State axioms.
+  \begin{itemize}
+    \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U) \rightarrow A \rightarrow L$ \\
+          The $\mapsto$ is a linear type constructor.
+It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
+Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
+    \item $\textbf{new} : (A : U) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
+          $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
+    \item $\textbf{free} : (A : U) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \unit$ \\
+          $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use.
+This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
+    \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
+          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
+The type of $\get$'s output is the rather complicated $F \_: (\Sigma x.
+x =_A m).
+(l \mapsto_A m)$.
+This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored.
+The proof $(l \mapsto_A m)$ is returned unchanged as well.
+    \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
+          $\set$ has the most complicated type of all the axioms presented here.
+Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term.
+This is the so called "strong update" operation.
+The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
+  \end{itemize}
+\end{definition}
+Due to the fact that $(l \mapsto_A m)$ is a linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
+This is the same non-sharing principle that empowers Separation Logic.
+Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
+
+\subsection{Internalization of Effect Laws}
+The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CLC framework.
+
+\begin{enumerate}
+  \item Performing $\get$ twice on the same address will produce two terms that are provably equal.
+  \item After using $\set$ to update the store at an address $l$ to term $m$, performing $\get$ again on the same address will retrieve $m$.
+\end{enumerate}
+
+Figure \ref{law1} is an excerpt taken from our implementation.
+It demonstrates the internalization of Law 1.
+The usage of implicit parameters allows one to omit writing much of the redundant arguments that could be inferred from context.
+Here, $\new$ allocates a store for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
+When $\get$ is applied to $c$ twice, the equality proofs \textit{pf1} $: (1 =_\mathbb{N} x)$ and \textit{pf2} $: (1 =_\mathbb{N} y)$ generated by $\get$ allow one to create proof \textit{pf} $: (x =_\mathbb{N} y)$ that the retrieved values $x$ and $y$ are equal.
+Finally, due to $c$ having linear type, it is required for it to be freed at the very end.
+
+\begin{figure}[h]
+  \caption{Law 1}
+  \begin{minted}{coq}
+  let [ _, c ] := new _ 1 in
+  let [ xeq, c ] := get _ _ _ c in
+  let [ yeq, c ] := get _ _ _ c in
+  let ( x, pf1 ) := xeq in
+  let ( y, pf2 ) := yeq in
+  let pf1 := Eq_sym _ _ _ pf1 in
+  let pf : Eq _ x y := 
+    Eq_trans _ _ _ _ pf1 pf2 
+  in free _ _ _ c.
+  \end{minted}
+  \label{law1}
+  \Description{}
+\end{figure}
+
+Figure \ref{law2} demonstrates the internalization of Law 2.
+Again, implicit parameters fill in the annoying bureaucratic arguments whenever they can be inferred.
+Here, a store is allocated by $\new$ for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
+After using $\set$ to update the store at $l$ to 2 and then performing $\get$ on $l$, one can create proof \textit{pf} $: (2 =_\mathbb{N} z)$ that the value $z$ retrieved by $\get$ is exactly equal to 2.
+
+\begin{figure}[h]
+  \caption{Law 2}
+  \begin{minted}{coq}
+  let [ _, c ] := new _ 1 in
+  let c := set _ _ _ _ c 2 in
+  let [ zeq, c ] := get _ _ _ c in
+  let ( z, pf1 ) := zeq in
+  let pf : Eq _ 2 z := pf1 in
+  free _ _ _ c.
+  \end{minted}
+  \label{law2}
+  \Description{}
+\end{figure}
+
+\subsection{Protocol Enforcement}
+The inclusion of dependent types and linear types in CILC makes it extremely expressive when encoding communication protocols.
+By indexing each channel with a protocol, CLC can enforce communication on a channel to strictly adhere to its specified protocol.
+
+In Figure \ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
+Intuitively, the \texttt{session}  forms a stack of possible operations executed in a protocol.
+The $\SEND$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after sending a message of type $A$, the protocol progresses to stage $ss$.
+Likewise, the $\RECV$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after receiving a message of type $A$, the protocol progresses to stage $ss$.
+Finally, the $\END$ constructor indicates that the protocol has finished and communication is over.
+
+\begin{figure}[h]
+  \caption{Inductive Session Type}
+  \begin{minted}{coq}
+  Inductive session : U :=
+  | SEND : U -> session -> session
+  | RECV : U -> session -> session
+  | END : session.
+  \end{minted}
+  \label{session}
+  \Description{}
+\end{figure}
+
+In Definition \ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
+
+\begin{definition}
+  Communication axioms.
+  \begin{itemize}
+    \item $\textbf{channel} : \session \rightarrow L$ \\
+          The $\channel$ axiom is a type constructor. It is used to form the types of communication channels that obey the protocol specified by its session.
+    \item $\textbf{open} : (ss : \session) \rightarrow \channel\ ss$ \\
+          When given a protocol, $\open$ creates a channel indexed by this protocol.
+    \item $\textbf{close} : \channel\ \END \rightarrow \unit$ \\
+          Once all communications specified on a given channel has finished, $\close$ will close and reclaim the channel.
+    \item $\textbf{send} : (A : U) \rightarrow A \rightarrow (ss : \session) \rightarrow$ \\
+          \phantom{send : (A :)}
+          $\channel\ (\SEND\ A\ ss) \rightarrow \channel\ ss$ \\
+          After a channel has reached the point in its protocol where a message of type $A$ should be sent, $\send$ may send a term of type $A$ on this channel, causing the channel to progress to the next stage of its protocol.
+    \item $\textbf{recv} : (A : U) \rightarrow (ss : \session) \rightarrow $ \\
+          \phantom{recv : (A :)}
+          $\channel\ (\RECV\ A\ ss) \rightarrow F \_:A.\channel\ ss$
+          After a channel has reached the point in its protocol where a message of type $A$ should be received, $\recv$ may receive such a term of type $A$ from the channel, causing the channel to progress to the next stage of its protocol.
+  \end{itemize}
+  \label{comm}
+\end{definition}
+
+The important detail to notice here is that each channel created by $\open$ is of linear type $(\channel\ ss)$ where $ss$ is some arbitrary protocol, hence must be used exactly once.
+However, the $\send$ and $\recv$ operations return back channels after consuming them, progressing their protocols one stage forward.
+Coupled with the fact that the $\close$ operation is only able to close channels whose protocols have ended, this forces channels to communicate strictly by their specified protocols.
+
+Dependent types also allows for more precise specifications of protocols themselves.
+A simple example would be the specification of message sizes.
+Figure \ref{communication} shows such an example.
+The inductive type \texttt{ilist} is a length indexed list.
+Using \texttt{ilist}, we construct a protocol \texttt{ss} that first sends two natural numbers, then receives eight natural numbers and finally sends back a boolean acknowledgement.
+
+\begin{figure}[h]
+  \caption{Sized Communication}
+  \begin{minted}{coq}
+  Inductive ilist (A : U) : Nat -> U :=
+  | nil  : ilist A 0
+  | cons : A -> (n : Nat) -> 
+             ilist A n -> ilist A (S n).
+
+  let ss := 
+    SEND (ilist Nat 2) 
+      (RECV (ilist Nat 8) (SEND Bool END)) 
+  in
+  let send_msg : ilist Nat 2 := 
+    (cons 1 _ (cons 2 _ nil)) 
+  in
+  let ch := open ss in
+  let ch := send _ send_msg _ ch in
+  let [ recv_msg, ch ] := recv _ _ ch in
+  let ch := send _ true _ ch in
+  close ch.            
+  \end{minted}
+  \label{communication}
+  \Description{}
+\end{figure}
+
+\section{Related Work}
+Linear types are a class of type systems inspired by Girard' substructural Linear Logic \cite{girard}.
+Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
+Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
+A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
+Benton \cite{benton1994} investigates the ramifications of the !
+exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
+Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
+
+Over the years, work has been done to enrich linear type theories with dependent types.
+Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
+Krishnaswami et al.
+present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
+Luo et al.
+\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
+The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
+
+\section{Future Work}
+The integration of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
+
+On the theoretical side, we intend to investigate the semantic models of CLC.
+In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy.
+This has since been removed due to the additional difficulty of working with $Prop$ in the soundness proof of CILC.
+We believe that these difficulties are not intrinsic and $Prop$ could be added back with careful proof work.
+
+On the application side, we intend to bolster our implementation with more features such as compilation to executable binary, making it a full-featured theorem prover/programming language.
+We aim to fully verify the typechecking algorithm employed by the current implementation.
+The unification algorithm used for implicit argument resolution is extremely ad-hoc, as the problem of unification for linear types has been not thoroughly studied, we hope to improve this situation as well.
 

### 2021-12-02 writing/draft5/draft5.tex
writing

@@
 \end{definition}
-Due to the fact that $(l \mapsto_A m)$ is a linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
+Due to the fact that $(l \mapsto_A m)$ is an abstract linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
 This is the same non-sharing principle that empowers Separation Logic.

### 2021-12-02 writing/draft5/draft5.tex
writing
draft5

@@
 \begin{equation*}
-  (x \ltype Money\ 5) \rightarrow (Camels + Marlboro)
+  (x \ltype Money\ 5) \rightarrow (Camels \oplus Marlboro)
 \end{equation*}
@@
   \item Furthermore, we extend CLC with rules to define inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-This is the first formalization of a generalized mechanism for defining inductive linear types that we are aware of.
+This is the first presentation of a generalized mechanism for defining inductive linear types that we are aware of.
   \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
@@
 
-\subsection{Internalization of Effect Laws}
-The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CLC framework.
+The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CILC framework.
 
@@
   let ( y, pf2 ) := yeq in
-  let pf1 := Eq_sym _ _ _ pf1 in
-  let pf : Eq _ x y := 
-    Eq_trans _ _ _ _ pf1 pf2 
+  let pf1 := eq_sym _ _ _ pf1 in
+  let pf : eq _ x y := 
+    eq_trans _ _ _ _ pf1 pf2 
   in free _ _ _ c.
@@
   let ( z, pf1 ) := zeq in
-  let pf : Eq _ 2 z := pf1 in
+  let pf : eq _ 2 z := pf1 in
   free _ _ _ c.
@@
 The inclusion of dependent types and linear types in CILC makes it extremely expressive when encoding communication protocols.
-By indexing each channel with a protocol, CLC can enforce communication on a channel to strictly adhere to its specified protocol.
+By indexing each channel with a protocol, CILC can enforce communication on a channel to strictly adhere to its specified protocol.
 

### 2021-12-02 writing/draft5/draft5.tex
draft5

@@
 
+\subsection{Mutable Data}
+Figure \ref{list} presents a CPS append function for linear lists.
+In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
+Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
+Each \texttt{_cons_} is applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
+Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
+Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.
+
+\begin{figure}[h]
+  \caption{Linear Lists}
+  \begin{minted}{coq}
+  Inductive list (A : U) : L :=
+  | nil : list A
+  | cons : A -> list A -> list A.
+
+  Fixpoint append (A : U) : list A -o 
+    list A -o (list A -o list A) -o list A 
+  :=
+    fun ls1 ls2 k =>
+      match ls1 with
+      | nil => k ls2
+      | (cons h t) as _cons_ =>
+        append _ t ls2 
+          (fun res => k (_cons_ h res))
+      end.
+  \end{minted}
+  \label{list}
+  \Description{}
+\end{figure}
+
 \section{Related Work}
@@
 \section{Future Work}
-The integration of linear types and dependent types opens the door to many topics for research, both in theory and in its applications.
-
 On the theoretical side, we intend to investigate the semantic models of CLC.

### 2021-12-02 writing/draft5/draft5.tex
draft5

@@
 \subsection{Extended Reduction}
-The reduction semantics for the extended language is standard, we only present the most interesting $\iota$-reduction cases.
+The reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
 

### 2021-12-03 writing/draft5/draft5.tex
draft

@@
 Linear connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
-Through examples, we show that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
-
-We have formalized all major results in Coq, and implemented a prototype in OCaml, both of which are available in the first author's Github repository.
+Through examples, we show that CILC facilitates lightweight reasoning about stateful programs and protocol enforcement.
+
+We have formalized all major results in Coq and implemented a prototype in OCaml, both of which are available in the first author's Github repository.
 

### 2021-12-11 writing/appendix/draft.tex
coq

--- 2021-12-03 writing/draft5/draft5.tex
@@
-\documentclass[sigplan,screen,review,authordraft]{acmart}
+\documentclass{article}
 
 \usepackage[utf8]{inputenc}
+\usepackage[margin=1in]{geometry}
 
@@
 \usepackage{amsmath}
+\usepackage{amssymb}
 \usepackage{mdwtab}
@@
 \setlength{\grammarindent}{7em}
-\newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
+\newcommand{\indalt}[1][2]{\\\hspace*{-0.2em}\textbar\quad}
 \newcommand{\rname}[1]{\textsc{\footnotesize #1}}
@@
 \newcommand{\step}{\leadsto}
-\newcommand{\pstep}{\leadsto}
+\newcommand{\pstep}{\leadsto_p}
 \newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
@@
 
-\title{The Calculus of Linear Constructions}
+\title{The Calculus of Linear Constructions --- Technical Report}
 \author{Qiancheng Fu}
-\affiliation{
-  \institution{Boston University}
-  \city{Boston}
-  \state{MA}
-  \country{USA}
-}
-\email{qcfu@bu.edu}
-
-\keywords{Dependent types, linear types, inductive types, Calculus of Constructions}
 
 \begin{document}
-\begin{abstract}
-  The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends the predicative CC$\omega$ with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
-We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Through examples, we demonstrate that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
-We have formalized and proven correct all major results in the Coq Proof Assistant.
-\end{abstract}
 \maketitle
+\tableofcontents
 
 \section{Introduction}
-The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
-In CC types can depend on terms, allowing one to write precise specifications as types.
-Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
-These theorem provers have found great success in the fields of software verification, and constructive mathematics.
+This extended report is meant to accompany our paper of the same title.
+Here, we describe the meta-theory of CILC and their proofs in detail.
+All the results presented here have been formalized and proven correct in the Coq Proof Assistant.
 
-However, due to its origins as a logical framework for constructive mathematics, it is quite difficult for CC to encode and reason about resources.
-Intuitively, a mathematical theorem can be applied an unrestricted number of times.
-Comparatively, the usage of resources is more limited.
-For example, if we encode Girard's classical example \cite{girard95} of purchasing cigarettes literally into CC as a function of type:
-\begin{equation*}
-  Money \rightarrow Camels + Marlboro
-\end{equation*}
-If viewed as a propositional implication, the customer will still maintain full ownership of their money after paying the vendor, because implication does not diminish the validity of its antecedent.
-Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
-Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
-We propose an alternative solution: extend CC with linear types.
-
-This paper presents a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
-CLC extends the predicative CC$\omega$ with linear types.
-CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
-We add an extra universe sort $L$ of linear types with cumulativity parallel to the sort $U$ of non-linear types.
-This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
-
-The presence of both linear and dependent types enables CLC to write specifications that faithfully encode the usage of resources.
-The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$ as follows.
-\begin{equation*}
-  (x \ltype Money\ 5) \rightarrow (Camels \oplus Marlboro)
-\end{equation*}
-This new specification for the transaction demands a payment of 5 units of money, and the customer be relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
-
-Compared to preexisting approaches for integrating linear types and dependent types, CLC offers a ``lightweight'' approach, akin to Mazurak et al.'s work on System F \cite{mazurak}.
-Prior works have either utilized an explicit !
-exponential or a pair of adjoint connectives to mediate between the worlds of linear and non-linear types.
-Our method requires neither exponential nor adjoint connectives explicitly in the syntax, allowing for a straightforward modeling of CLC in the predicative CC$\omega$, and lifting of predicative CC$\omega$ into CLC with minimal annotations needed.
-
-We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Linear connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
-Through examples, we show that CILC facilitates lightweight reasoning about stateful programs and protocol enforcement.
-
-We have formalized all major results in Coq and implemented a prototype in OCaml, both of which are available in the first author's Github repository.
-
-\noindent \textbf{\textit{Contributions}}:
-Our contributions can be summarized as follows.
-\begin{itemize}
-  \item Fist, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types.
-The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
-  \item Second, we study the meta-theory of CLC directly, showing that it satisfies the standard properties of confluence, validity and subject reduction.
-  \item Next, we prove the linearity theorem, showing the tangible impact linear types have on the structure of terms.
-We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
-  \item Additionally, we observe that CLC is backwards compatible with the predicative CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
-We prove that all valid CC$\omega$ terms are valid CLC terms after adding minimal amounts of annotation.
-  \item Furthermore, we extend CLC with rules to define inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-This is the first presentation of a generalized mechanism for defining inductive linear types that we are aware of.
-  \item All major results have been formalized and proven correct in the Coq Proof Assistant with help from the Autosubst \cite{autosubst} library.
-To the best of our knowledge, our development is the first machine checked formalization of a linear dependent type theory.
-  \item Finally, we give an implementation of CILC in OCaml. Algorithmic type checking employed by the typechecker streamlines the process of writing CILC programs.
-\end{itemize}
-
-\section{The Language of CLC}
-\subsection{Syntax}
-\begin{figure}[h]
-  \caption{Syntax of CLC}
+\section{Syntax of CLC \texttt{(clc_ast.v)}}
+\begin{figure}[H]
   \begin{grammar}
-    <$i$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} universe levels
+    <$i$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{2.4em} universe levels
 
@@
 
-    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4.2em} expressions
+    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{3em} expressions
     \indalt $(x :_s A) \rightarrow B$
@@
   \end{grammar}
-  \Description{}
-  \label{syntax1}
 \end{figure}
 
-The syntax of the core type theory is presented in Figure \ref{syntax1}.
-Our type theory contains two sorts of universes $U$ and $L$.
-We use the meta variable $i$ to quantify over universe levels $0, 1, 2, ...$.
-$U_i$, and $L_i$ are the universes at level $i$ of non-linear and linear types respectively.
-\begin{figure}[h]
-  \caption{Correspondence of CLC types and MELL implications}
-  \Description{}
-  \begin{align}
-    (\_ \utype A) \rightarrow B \quad & \Leftrightarrow \quad !(!A \multimap B) \\
-    (\_ \ltype A) \rightarrow B \quad & \Leftrightarrow \quad !(A \multimap B)  \\
-    (\_ \utype A) \multimap B \quad   & \Leftrightarrow \quad !A \multimap B    \\
-    (\_ \ltype A) \multimap B \quad   & \Leftrightarrow \quad A \multimap B
-  \end{align}
-  \label{correspondence}
-\end{figure}
-
-A clear departure of our language from standard presentations of both linear type theory and dependent type theory is the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
-The reason for these function types is that we have built the !
-exponential of linear logic implicitly into universe sorts.
-The sort annotation $s$ here records the universe sort of the function domain.
-The behavior of !
-is difficult to account for even in simple linear type theories.
-Subtle issues arise if !!
-is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
-By integrating the exponential directly into universe sorts, we have limited !
-to only be canonically usable.
-This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
-
-Figure \ref{correspondence} illustrates the correspondence between CLC functions and Multiplicative Exponential Linear Logic (MELL) implications.
-MELL lacks counterparts for the cases (1), (3) if the co-domain $B$ is dependent on arguments of domain $A$.
-We will discuss function type formation in Section \ref{tyformation} in greater detail.
-
-Algorithmic type checking techniques allow users to omit writing sort indices for many situations in practice.
-Our implementation employs bi-directional type checking and users rarely interact with sort indices in the surface syntax.
-
-\subsection{Universes and Cumulativity}
-CLC features two sorts of universes $U$ and $L$ with level indices $0, 1, 2, \cdots$.
-$U_i$ and $L_i$ are the predicative universes of non-linear types and linear types respectively.
-The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
-Basically, terms with types that occur within $U_i$ are unrestricted in their usage.
-Terms with types that occur within $L_i$ are restricted to being used exactly once.
-
-In order to lift terms from lower universes to higher ones, there exists cumulativity between universe levels of the same sort. We define cumulativity as follows.
-
-\begin{definition}
-  The cumulativity relation ($\preceq$) is the smallest binary relation over terms such that
-  \begin{enumerate}
-    \item $\preceq$ is a partial order with respect to equality.
-          \begin{enumerate}
-            \item If $A \equiv B$, then $A \preceq B$.
-            \item If $A \preceq B$ and $B \preceq A$, then $A \equiv B$.
-            \item If $A \preceq B$ and $B \preceq C$, then $A \preceq B$.
-          \end{enumerate}
-    \item $U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
-    \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
-    \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
-          $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{s} A_2) \rightarrow B_2$
-    \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
-          $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{s} A_2) \multimap B_2$
-  \end{enumerate}
-\end{definition}
-
-Figure \ref{universe} illustrates the structure of our universe hierarchy.
-Each linear universe $L_i$ has $U_{i+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
-However, $L_i$ cumulates to $L_{i+1}$.
-These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes, and subsequently losing track of linearity.
-
-\begin{figure}[h]
-  \caption{The Universe Hierarchy}
-  \vspace{1em}
-  \centering
-  \begin{tikzcd}
-    U_0 \arrow[r, ":"', "\preceq", dash]
-    & U_1 \arrow[r, ":"', "\preceq", dash]
-    & U_2 \arrow[r, ":"', "\preceq", dash]
-    & \cdots \\
-    L_0 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
-    & L_1 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
-    & L_2 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
-    & \cdots
-  \end{tikzcd}
-  \label{universe}
-  \Description{}
-\end{figure}
-
-\subsection{Context and Structural Judgments}
-The context of our language employs a mixed linear/non-linear representation in the style of Luo\cite{luo}.
-Variables in the context are annotated to indicate whether they are linear or non-linear.
-A non-linear variable is annotated as $\Gamma, x \utype A$, whereas a linear variable is annotated as $\Gamma, x \ltype A$.
-
-Next, we define a $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ relation that merges two mixed contexts $\Gamma_1$, and $\Gamma_2$ into $\Gamma$, by performing contraction on shared non-linear variables.
-For linear variables, the $\mrg{\_}{\_}{\_}$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
-This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for unrestricted variables whilst forbidding it for restricted ones.
-
-An auxiliary judgment $\pure{\Gamma}$ is defined to assert that a context $\Gamma$ does not contain linear variables.
-In other words, all variables found in $\pure{\Gamma}$ are annotated of the form $x \utype A$.
-The full rules for structural judgments are presented in Figure \ref{structural}.
-
-\begin{figure}[h]
-  \caption{Structural Judgments}
-  \begin{mathpar}
-    \inferrule
-    { }
-    { \epsilon \vdash }
-    \rname{Wf-$\epsilon$}
-
-    \inferrule
-    { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A \utype U_i }
-    { \Gamma, x \utype A \vdash }
-    \rname{Wf-U}
-
-    \inferrule
-    { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A \utype L_k }
-    { \Gamma, x \ltype A\ \vdash }
-    \rname{Wf-L}
-
-    \inferrule
-    { }
-    { \pure{\epsilon} }
-    \rname{Pure-$\epsilon$}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A \utype U }
-    { \pure{\Gamma, x \utype A} }
-    \rname{Pure-U}
-
-    \inferrule
-    { }
-    { \mrg{\epsilon}{\epsilon}{\epsilon} }
-    \rname{Merge-$\epsilon$}
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \mrg{\Gamma_1, x \utype A}
-      {\Gamma_2, x \utype A}
-      {\Gamma, x \utype A} }
-    \rname{Merge-U}
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-      x \notin \Gamma_2 }
-    { \mrg{\Gamma_1, x \ltype A}
-      {\Gamma_2}
-      {\Gamma, x \ltype A} }
-    \rname{Merge-L1}
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-      x \notin \Gamma_1 }
-    { \mrg{\Gamma_1}
-      {\Gamma_2, x \ltype A}
-      {\Gamma, x \ltype A} }
-    \rname{Merge-L2}
-  \end{mathpar}
-  \label{structural}
-  \Description{}
-\end{figure}
-
-\begin{definition}
-  The context restriction function $\overline{\Gamma}$ is defined as a recursive filter over $\Gamma$ as follows.
-All linear variables are removed from context $\Gamma$.
-The result of context restriction is the non-linear subset of the original context.
-  \begin{align*}
-    \overline{\epsilon} = \epsilon
-    \hspace*{4em}
-    \overline{\Gamma, x \utype A} = \overline{\Gamma}, x \utype A
-    \hspace*{4em}
-    \overline{\Gamma, x \ltype A} = \overline{\Gamma}
-  \end{align*}
-\end{definition}
-
-\subsection{Typing Judgment}
-Typing judgments in CLC take on the form of $\Gamma \vdash m : A$. Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$, with free variables typed in $\Gamma$.
-
-\begin{definition} We formally define the terminology \textit{non-linear}, \textit{linear}, \textit{unrestricted} and \textit{restricted}.
-  \begin{enumerate}
-    \item A \textit{type} $A$ is \textit{non-linear} under context $\Gamma$ if $\Gamma \vdash A : U_i$.
-    \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A : L_k$.
-    \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$, such that $\Gamma \vdash m : A$.
-A unrestricted term may be used an arbitrary number of times.
-    \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$, such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
-  \end{enumerate}
-\end{definition}
-
-\subsection{Type Formation} \label{tyformation}
-The rules for forming types are presented in Figure \ref{type}.
-In CLC, we forbid types from depending on linear terms similar to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
-
-The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
-In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
-Variables not pertaining to actual proofs could be introduced this way, thus giving rise to the admissibility of weakening.
-To support linear types, we must restrict weakening to non-linear variables.
-This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{U-Axiom}, \rname{L-Axiom}.
-From \rname{L-Axiom} we can see that the universe of linear types $L_i$ is an inhabitant of $U_{i+1}$.
-This is inspired by Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
-
-\begin{figure}[h]
-  \caption{Type Formation}
-  \begin{mathpar}
-    \inferrule
-    { \pure{\Gamma} }
-    { \Gamma \vdash s_i : U_{i+1} }
-    \rname{Sort-Axiom}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : U_i \\
-      \Gamma, x \utype A \vdash B : s_i }
-    { \Gamma \vdash (x \utype A) \rightarrow B : U_i }
-    \rname{U$\rightarrow$}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : L_i \\
-      \Gamma \vdash B : s_i \\
-      x \notin \Gamma }
-    { \Gamma \vdash (x \ltype A) \rightarrow B : U_i }
-    \rname{L$\rightarrow$}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : U_i \\
-      \Gamma, x \utype A \vdash B : s_i }
-    { \Gamma \vdash (x \utype A) \multimap B : L_i }
-    \rname{U$\multimap$}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : L_i \\
-      \Gamma \vdash B : s_i \\
-      x \notin \Gamma }
-    { \Gamma \vdash (x \ltype A) \multimap B : L_i }
-    \rname{L$\multimap$}
-  \end{mathpar}
-  \label{type}
-  \Description{}
-\end{figure}
-
-The \rname{U$\rightarrow$} rule is used for forming non-linear function types with non-linear domains.
-This is evident from the judgment $\Gamma \vdash A : U_i$.
-Due to the fact that $A$ is in the non-linear universe $U_i$, terms of type $A$ have unrestricted usage.
-The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_i$.
-The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_i$ and $L_i$.
-From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_i$, we see that the overall type is a non-linear type.
-The term level $\lambda$-abstractions of these types can be used unrestrictedly.
-
-The \rname{L$\rightarrow$} rule is used for forming non-linear function types with linear domains.
-From the judgment $\Gamma \vdash A : L_i$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
-Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B : s_i$.
-Like \rname{U$\rightarrow$}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
-The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_i$.
-Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
-Again, the overall resulting type is non-linear, so the term level $\lambda$-abstractions of these types can be used unrestrictedly.
-
-The \rname{U$\multimap$}, \rname{L$\multimap$} rules are similar to \rname{U$\rightarrow$}, \rname{L$\rightarrow$} in spirit.
-The dependency considerations for domain $A$ and co-domain $B$ are exactly the same.
-The main difference between \rname{U$\multimap$}, \rname{L$\multimap$} and \rname{U$\rightarrow$}, \rname{L$\rightarrow$} is the universe sort of the resulting type.
-The sorts of the function types formed by \rname{U$\multimap$}, \rname{L$\multimap$} are $L$, meaning they are linear function types.
-The term level $\lambda$-abstractions of these types must be used exactly once.
-
-\subsection{Term Formation} \label{teformation}
-The rules for term formation are presented in Figure~\ref{term}.
-
-\begin{figure}[h]
-  \caption{Term Formation}
-  \begin{mathpar}
-    \inferrule
-    { \pure{\Gamma_1, \Gamma_2} }
-    { \Gamma_1, x \utype A, \Gamma_2 \vdash x : A }
-    \rname{U-Var}
-
-    \inferrule
-    { \pure{\Gamma_1, \Gamma_2} }
-    { \Gamma_1, x \ltype A, \Gamma_2 \vdash x : A }
-    \rname{L-Var}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash (x \stype{s} A) \rightarrow B : t_i \\
-      \Gamma, x \stype{s} A \vdash n : B }
-    { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \rightarrow B }
-    \rname{$\lambda$$\rightarrow$}
-
-    \inferrule
-    { \overline{\Gamma} \vdash (x \stype{s} A) \multimap B : t_i \\
-      \Gamma, x \stype{s} A \vdash n : B }
-    { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \multimap B }
-    \rname{$\lambda$$\multimap$}
-
-    \inferrule
-    { \Gamma_1 \vdash m : (x \utype A) \rightarrow B \\
-      \pure{\Gamma_2} \\
-      \Gamma_2 \vdash n : A \\
-      \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \Gamma \vdash m\ n : B[n/x] }
-    \rname{App-U$\rightarrow$}
-
-    \inferrule
-    { \Gamma_1 \vdash m : (x \ltype A) \rightarrow B \\
-      \Gamma_2 \vdash n : A \\
-      \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \Gamma \vdash m\ n : B[n/x] }
-    \rname{App-L$\rightarrow$}
-
-    \inferrule
-    { \Gamma_1 \vdash m : (x \utype A) \multimap B \\
-      \pure{\Gamma_2} \\
-      \Gamma_2 \vdash n : A \\
-      \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \Gamma \vdash m\ n : B[n/x] }
-    \rname{App-U$\multimap$}
-
-    \inferrule
-    { \Gamma_1 \vdash m : (x \ltype A) \multimap B \\
-      \Gamma_2 \vdash n : A \\
-      \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \Gamma \vdash m\ n : B[n/x] }
-    \rname{App-L$\multimap$}
-
-    \inferrule
-    { \Gamma \vdash m : A \\
-      \overline{\Gamma} \vdash B : s_i \\ A \preceq B }
-    { \Gamma \vdash m : B }
-    \rname{Conversion}
-  \end{mathpar}
-  \label{term}
-  \Description{}
-\end{figure}
-
-The rules \rname{U-Var}, and \rname{L-Var} are used for typing free variables.
-The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \utype A, \Gamma_2$ with a non-linear type $A$.
-The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \ltype A, \Gamma_2$ with linear type $A$.
-For both rules, the side condition $\pure{\Gamma_1, \Gamma_2}$ forbids irrelevant variables of linear types from occurring within the context.
-This prevents another vector of weakening variables with linear type.
-
-In \rname{$\lambda$$\rightarrow$}, the function type being addressed has form $(x \stype{s} A) \rightarrow B$.
-$\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
-This consideration is realized by the side condition $\pure{\Gamma}$, asserting all variables in context $\Gamma$ are unrestricted.
-Next, the body of the abstraction $n$ is typed as $\Gamma, x \stype{s} A \vdash n : B$, where $s$ is the sort of $A$.
-Finally, the resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \rightarrow B$ asserts that the $\lambda$-abstraction can be used unrestrictedly.
-
-In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
-Due to the fact these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
-However, in judgment $\overline{\Gamma} \vdash (x \stype{s} A) \multimap B : L_k$ the context must be filtered, because types are not allowed to depend on restricted variables.
-The final resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \multimap B$ asserts that the $\lambda$-abstraction must be used exactly once.
-
-For \rname{App-U$\rightarrow$}, domain $A$ is a non-linear type, as seen by its annotation in $(x \utype A) \rightarrow B$.
-Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
-Thus, the supplied argument $n$ must not depend on restricted variables in context $\Gamma_2$.
-Otherwise, substitution may put multiple copies of $n$ into $m$ during $\beta$-reduction, duplicating variables that should have been restricted.
-This justifies the side condition of $\pure{\Gamma_2}$.
-The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
-
-Now for \rname{App-L$\rightarrow$}, domain $A$ is a linear type, as seen by its annotation in $(x \ltype A) \rightarrow B$.
-Intuitively, this tells us that $x$ must be used once within the body of $m$.
-During $\beta$-reduction, substitution will only put a single copy of $n$ into the body of $m$, so $n$ can depend on restricted variables within $\Gamma_2$ without fear of duplicating them.
-This justifies the lack of side condition $\pure{\Gamma_2}$.
-The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
-
-The rules \rname{App-U$\multimap$}, \rname{App-L$\multimap$} follow the same considerations as \rname{App-U$\rightarrow$}, \rname{App-L$\rightarrow$}.
-Additional conditions are not required for these two rules to be sound.
-
-Finally, the \rname{Conversion} rule allows judgment $\Gamma \vdash m : A$ to convert to judgment $\Gamma \vdash m : B$ if $B$ is a valid type in context $\overline{\Gamma}$.
-Futhermore, $A$ must be a subtype of $B$ satisfying the relation $A \preceq B$.
-This rule gives rise to large eliminations (compute types from terms), as computations embedded at the type level can reduce to canonical types.
-
-\subsection{Equality and Reduction} \label{reduction}
-The operational semantics and $\beta$-equality of CLC terms are presented in Figure \ref{red}, all of which are entirely standard.
-
-As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
-
-\begin{figure}[h]
-  \caption{Equality and Reduction}
+\section{Reduction and Equality of CLC \texttt{(clc_ast.v)}}
+\begin{figure}[H]
   \begin{mathpar}
@@
     {  }
-    { (\lambda x \stype{s}A.m)\ n \pstep m[n/x] }
+    { (\lambda x \stype{s}A.m)\ n \step m[n/x] }
     \rname{Step-$\beta$}
@@
   \label{red}
-  \Description{}
 \end{figure}
 
-\subsection{Meta Theory} \label{meta}
-In this section, we focus our discussion on the properties of CLC.
-First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
-Next, we show that CLC is an effective linear type theory through the \textit{linearity} theorem.
-Furthermore, we show that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
-Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, showing that CLC is strongly normalizing.
+\section{Confluence of CLC \texttt{(clc_confluence.v)}}
 
-All proofs have been formalized in Coq with help from the Autosubst \cite{autosubst} library.
-The Coq development is publicly available on the first author's Github repository.
-To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
-We give a hand written version of the proof in the appendix as well.
+\subsection{Parallel Reduction}
+To prove the confluence property of CLC, we employ the standard technique utilizing parallel reductions.
 
-\subsubsection{Reduction and Confluence}
-
-The following lemmas and proofs are entirely standard using parallel step technique \cite{takahashi}. The presence of linear types do not pose any complications as reductions are untyped.
-
-\begin{lemma}
-  Parallel reduction satisfies the diamond property. If $m \pstep_p m_1$ and $m \pstep_p m_2$ then there exists $m'$ such that $m_1 \pstep_p m'$ and $m_2 \pstep_p m'$.
-\end{lemma}
-
-\begin{lemma}
-  Each $\pstep_p$ is reachable with $\step^*$. If $m \pstep_p m'$ then $m \step^* m'$.
-\end{lemma}
-
-\begin{theorem}
-  The transitive reflexive closure of reduction is confluent. If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \pstep^* m'$ and $m_2 \pstep^* m'$.
-\end{theorem}
-
-\begin{corollary}
-  The definitional equality relation $\equiv$ is an equivalence relation.
-\end{corollary}
-
-\subsubsection{Weakening} \label{weakening}
-CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.
-
-\begin{lemma}
-  Weakening. If $\Gamma \vdash m : A$ is a valid judgment, then for any $x \notin \Gamma$, the judgment $\Gamma, x \utype B \vdash m : A$ is derivable.
-\end{lemma}
-
-\subsubsection{Substitution} \label{subst}
-Though the substitution lemma is regarded as a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
-Much of the difficulty arises during the substitution of arguments containing !
-exponential.
-Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
-He defines additional syntax and semantics for the intricate unboxing of !
-terms, solving the lack of substitute in Abramsky's term calculus.
-
-Our design of integrating ! into universe sorts removes the need for ! manipulating syntax and semantics. The following substitution lemmas are directly proved by induction on typing derivation.
-
-\begin{lemma}
-  Non-linear Substitution.
-For $\Gamma_1, x \utype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ are valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
-\end{lemma}
-
-\begin{lemma}
-  Linear Substitution. For $\Gamma_1, x \ltype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
-\end{lemma}
-
-\subsubsection{Type Soundness}
-In order to prove subject reduction, we first prove the validity theorem. The main purpose of validity is to lower types down to the term level, enabling the application of various inversion lemmas.
-
-\begin{theorem}
-  Validity.
-For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exist some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
-  \label{validity}
-\end{theorem}
-
-With weakening, substitution, propagation and various inversion lemmas proven, subject reduction can now be proved by induction on typing derivation.
-
-\begin{theorem}
-  Subject Reduction. For $\Gamma \vdash m : A$, if $m \pstep n$ then $\Gamma \vdash n : A$.
-\end{theorem}
-
-\subsubsection{Linearity}
-At this point, we have proven that CLC is type sound.
-However, we still need to prove that the removal of weakening and contraction for restricted variables yields tangible impact on the structure of terms.
-For this purpose, we define a binding aware recursive function $occurs(x, m)$ that counts the number of times variable $x$ appears within term $m$.
-The linearity theorem asserts that restricted variables are used exactly once within a term, subsuming safe resource usage.
-
-\begin{theorem}
-  Linearity. If $\Gamma \vdash m : A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
-\end{theorem}
-
-\subsubsection{Promotion and Dereliction}
-Linear types of CLC are not obtained through packing and unpacking !, so there are no explicit rules for \textit{promotion} and \textit{dereliction} of Linear Logic.
-Arbitrary computations existing at the type level also muddle the association between which types can be promoted or derelicted to which.
-Nevertheless, \textit{promotion} and \textit{dereliction} for canonical types are derivable as theorems through $\eta$-expansion.
-
-\begin{theorem}
-  Promotion. If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
-\end{theorem}
-
-\begin{theorem}
-  Dereliction. If $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$ is a valid judgment, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$ is derivable.
-\end{theorem}
-
-\subsubsection{Strong Normalization}
-The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function to CC$\omega$.
-We assume familiarity with CC$\omega$ syntax here, and define the erasure function as follows.
-
-\begin{definition}
-  \begin{align*}
-    \erase{x}                             & = x                                     \\
-    \erase{U_i}                           & = Type_i                                \\
-    \erase{L_i}                           & = Type_i                                \\
-    \erase{(x \stype{s} A) \rightarrow B} & = (x : \erase{A}) \rightarrow \erase{B} \\
-    \erase{(x \stype{s} A) \multimap B}   & = (x : \erase{A}) \rightarrow \erase{B} \\
-    \erase{\lambda x\stype{s}A.n}         & = \lambda x : \erase{A}.\erase{n}       \\
-    \erase{m\ n}                          & = \erase{m}\ \erase{n}
-  \end{align*}
-\end{definition}
-
-With slight overloading of notation, we define erasure for CLC contexts recursively.
-
-\begin{definition}
-  \begin{align*}
-    \erase{\epsilon}              & = \epsilon                      \\
-    \erase{\Gamma, x \stype{s} A} & = \erase{\Gamma}, x : \erase{A}
-  \end{align*}
-\end{definition}
-
-We prove the following lemma to commute erasure and substitution whenever needed.
-
-\begin{lemma}
-  Erasure commutes with substitution. For any CLC terms $m$, $n$ and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$.
-\end{lemma}
-
-For the following lemma, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
-
-\begin{theorem} \label{preserve}
-  Reduction Erasure. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
-\end{theorem}
-
-\begin{theorem} \label{embedding}
-  Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
-\end{theorem}
-
-If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}.
-This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
-
-\begin{theorem}
-  Well-typed CLC terms are strongly normalizing.
-\end{theorem}
-
-\subsubsection{Backwards Compatibility}
-To show that CLC is backwards compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
-
-\begin{definition}
-  \begin{align*}
-    \lift{x}                     & = x                                        \\
-    \lift{Type_i}                & = U_i                                      \\
-    \lift{(x : A) \rightarrow B} & = (x \utype \lift{A}) \rightarrow \lift{B} \\
-    \lift{\lambda x : A.n}       & = \lambda x\utype\lift{A}.\lift{n}         \\
-    \lift{m\ n}                  & = \lift{m}\ \lift{n}
-  \end{align*}
-\end{definition}
-
-With slight overloading of notation, we define lifting for CC$\omega$ recursively.
-\begin{definition}
-  \begin{align*}
-    \lift{\epsilon}      & = \epsilon                         \\
-    \lift{\Gamma, x : A} & = \lift{\Gamma}, x \utype \lift{A}
-  \end{align*}
-\end{definition}
-
-The following lemmas and theorems are all proved following a similar procedure to proving the soundness of erasure.
-\begin{lemma}
-  Lift commutes with substitution. For any CC$\omega$ terms $m$, $n$ and some variable $x$, $\lift{m[n/x]} = \lift{m}[\lift{n}/x]$.
-\end{lemma}
-
-\begin{theorem}
-  Lift preserves reduction. For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
-\end{theorem}
-
-\begin{theorem} \label{lift}
-  Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
-\end{theorem}
-
-The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domain of function types are annotated with sort $U$.
-In practice, these annotations can often be omitted if algorithmic type checking is able to infer them from context.
-
-\section{The Extension to CILC}
-Although the core CLC language is extremely expressive, formally shown by Theorem \ref{lift} to be at least as expressive as the predicative CC$\omega$, the complexity of Church-style encodings to do so will quickly become unmanageable.
-To address this, we extend CLC with rules for defining inductive linear types in the style of CIC.
-
-We have formalized CILC in our Coq development and proven its soundness.
-
-\subsection{Syntax of CILC}
-\begin{figure}[h]
-  \caption{Syntax}
-  \begin{grammar}
-    <$C,I,P,Q,f$> ::= ... \hspace*{10em} expressions
-    \indalt $\ind{s}(X : A)\{C_1 \sep ... \sep C_k\}$
-    \indalt $\constr(k, I)$
-    \indalt $\case(m, Q)\{f_1 \sep ... \sep f_k \}$
-    \indalt $\dcase(m, Q)\{f_1 \sep ... \sep f_k \}$
-    \indalt $\fix f : A := m$
-  \end{grammar}
-  \Description{}
-  \label{syntax2}
-\end{figure}
-
-As shown in Figure \ref{syntax2}, CILC extends CIC with for the introduction and elimination of inductive types.
-Ind is used during the formation of new inductive types, and Constr is used for introducing their constructors.
-Case and DCase are non-dependent and dependent case eliminators respectively.
-CIC-style primitive recursion has been factored out of case eliminators in favor of a Coq-style Fix construct.
-The flexibility of an independent Fix construct allows for a cleaner formalization of inductive linear type elimination.
-
-\subsection{Arity}
-For type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as depicted in Figure \ref{arity}.
-
-\begin{figure}[h]
-  \caption{Arity}
+\begin{figure}[H]
   \begin{mathpar}
@@
     { }
-    { arity(s,s) }
-    \rname{A-Sort}
-
-    \inferrule
-    { arity(A, s) }
-    { arity((x \utype M) \rightarrow A, s)}
-    \rname{A$\rightarrow$}
-  \end{mathpar}
-  \Description{}
-  \label{arity}
-\end{figure}
-
-\begin{definition}
-  For some arity type $A$ and sort $s'$, the term $A\lrangle{s'}$ is inductively defined over the structure of $A$ as follows.
-  \begin{align*}
-    ((x \utype M) \rightarrow A)\lrangle{s'} & = (x \utype M) \rightarrow A\lrangle{s'} \\
-    (s)\lrangle{s'}                          & = s'
-  \end{align*}
-\end{definition}
-
-\begin{definition}
-  For some arity type $A$, term $I$ and sort $s'$, the term $A\{I, s'\}$ is inductively defined over the structure of $A$ as follows.
-  \begin{align*}
-    ((x \utype M) \rightarrow A)\lrangle{I, s'} & = (x \utype M) \rightarrow A\lrangle{(I\ x), s'} \\
-    (s)\lrangle{I, s'}                          & = (\_ \utype I) \rightarrow s'
-  \end{align*}
-\end{definition}
-
-\subsection{Strict Positivity}
-For type $P$ and variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as depicted in Figure \ref{positive}.
-
-We say that $X$ occurs strictly positive in $P$ if the judgment $postive(P,X)$ is valid.
-
-\begin{figure}[h]
-  \caption{Strict Positivity}
-  \begin{mathpar}
-    \inferrule
-    { (\forall i = 1...k) \\ X \notin m_i }
-    { positive((X\ m_1...m_k), X) }
-    \rname{Pos-X}
-
-    \inferrule
-    { X \notin M \\ positive(P, X)}
-    { positive((x \stype{s} M) \rightarrow P, X) }
-    \rname{Pos$\rightarrow$}
-
-    \inferrule
-    { X \notin M \\ positive(P, X) }
-    { positive((x \stype{s} M) \multimap P, X) }
-    \rname{Pos$\multimap$}
-  \end{mathpar}
-  \Description{}
-  \label{positive}
-\end{figure}
-
-\subsection{Non-Linear Constructor}
-For type $C$ and variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{nlconstr}.
-
-Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not applied to restricted terms as this may cause duplication.
-
-\begin{figure}[h]
-  \caption{Non-linear Constructor}
-  \begin{mathpar}
-    \inferrule
-    { (\forall i = 1...k) \\ X \notin m_i }
-    { \ucons((X\ m_i...m_k), X) }
-    \rname{U-Constr-X}
-
-    \inferrule
-    { positive(P, X) \\ \ucons(C, X) }
-    { \ucons((\_ \utype P) \rightarrow C, X)}
-    \rname{U-Constr-pos}
-
-    \inferrule
-    { X \notin M \\ \ucons(C, X) }
-    { \ucons((x \utype M) \rightarrow C, X)}
-    \rname{U-Constr$\rightarrow$}
-  \end{mathpar}
-  \Description{}
-  \label{nlconstr}
-\end{figure}
-
-\subsection{Linear Constructor}
-For type $C$ and variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{lconstr}.
-
-Linear constructors are used for constructing linear objects that must be used once. These constructors may be applied to linear terms as restricted usage prevent duplication and enforce single usage.
-
-An unapplied linear constructor is a non-linear entity as they can be created freely ``out of thin air".
-However, once a linear constructor has been partially applied to a linear term, its linearity is ``activated'' in rules \rname{L-Constr-Pos-2} and \rname{L-Constr-2$\rightarrow$}, essentially forcing the rest of its arguments to be fully applied.
-This is to prevent partially applied constructors from duplicating linear variables.
-
-\begin{figure}[h]
-  \caption{Linear Constructor}
-  \begin{mathpar}
-    \inferrule
-    { (\forall i = 1...k) \\ X \notin m_i }
-    { \lcons((X\ m_i...m_k), X) }
-    \rname{L-Constr-X}
-
-    \inferrule
-    { positive(P, X) \\ \lcons(C, X) }
-    { \lcons((\_ \utype P) \rightarrow C, X)}
-    \rname{L-Constr-Pos-1}
-
-    \inferrule
-    { X \notin M \\ \lcons(C, X) }
-    { \lcons((x \utype M) \rightarrow C, X)}
-    \rname{L-Constr-1$\rightarrow$}
-
-    \inferrule
-    { positive(P, X) \\ activation(C, X) }
-    { \lcons((\_ \ltype P) \rightarrow C, X)}
-    \rname{L-Constr-Pos-2}
-
-    \inferrule
-    { X \notin M \\ activation(C, X) }
-    { \lcons((x \ltype M) \rightarrow C, X)}
-    \rname{L-Constr-2$\rightarrow$}
-  \end{mathpar}
-  \Description{}
-  \label{lconstr}
-\end{figure}
-
-For type $C$ and variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{activation}.
-Intuitively, $activation(C,X)$ ensures the linearity of $C$.
-
-\begin{figure}[h]
-  \caption{Activation}
-  \begin{mathpar}
-    \inferrule
-    { (\forall i = 1...k) \\ X \notin m_i }
-    { activation((X\ m_i...m_k), X) }
-    \rname{Act-X}
-
-    \inferrule
-    { positive(P, X) \\ activation(C, X) }
-    { activation((\_ \stype P) \multimap C, X)}
-    \rname{Act-Pos}
-
-    \inferrule
-    { X \notin M \\ activation(C, X) }
-    { activation((x \stype M) \multimap C, X)}
-    \rname{Act$\multimap$}
-  \end{mathpar}
-  \Description{}
-  \label{activation}
-\end{figure}
-
-\subsection{Introduction Rules}
-The formation of new inductive types and their constructors are presented in Figure \ref{ind}.
-
-\begin{figure}[h]
-  \caption{Inductive Type Formation}
-  \begin{mathpar}
-    \inferrule
-    { (\forall j = 1...n) \\ arity(A, s) \\ \scons(C_j, X) \\\\
-      \pure{\Gamma} \\
-      \Gamma \vdash A : U_i \\
-      \Gamma, X \utype A \vdash C_j : t_i }
-    { \Gamma \vdash \ind{s}(X : A)\{C_1|...|C_k\} : A }
-    \rname{Ind}
-
-    \inferrule
-    { 1 \leq i \leq k \\\\
-    \pure{\Gamma} \\
-    I := \ind{s}(X : A)\{C_1|...|C_k\} \\
-    \Gamma \vdash I :A }
-    { \Gamma \vdash \constr(i, I) : C_i[I/X] }
-    \rname{Constr}
-  \end{mathpar}
-  \Description{}
-  \label{ind}
-\end{figure}
-
-\subsection{Non-Dependent Case}
-
-\begin{definition}
-  For constructor type $C$ and variables $X$ and $Q$, the term $C\{X, Q\}$ is inductively defined over the structure of $C$ as follows.
-  \begin{align*}
-    ((\_ \stype{s} P) \rightarrow C)\{X, Q\} & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
-    ((\_ \stype{s} P) \multimap C)\{X, Q\}   & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
-    ((x \stype{s} M) \rightarrow C)\{X, Q\}  & = (x \stype{s} P) \multimap C\{X, Q\}  \\
-    ((x \stype{s} M) \multimap C)\{X, Q\}    & = (x \stype{s} P) \multimap C\{X, Q\}  \\
-    (X\ m_1...m_k)\{X, Q\}                   & = (Q\ m_1...m_k)
-  \end{align*}
-  We define the notation $C[I, P] := C\{X, Q\}[I/X, P/Q]$.
-\end{definition}
-
-The rule for non-dependent case elimination is presented in Figure \ref{case}. This elimination rule is non-dependent because the motive $Q$ may not depend on the discriminated term $m$ itself.
-
-\begin{figure}[h]
-  \caption{Non-dependent Case Elimination}
-  \begin{mathpar}
-    \inferrule
-    { (\forall i = 1...k_1) \\
-    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-    arity(A, s) \\\\
-    I := \ind{s}(X : A)\{C_1|...|C_{k_1}\} \\
-    \overline{\Gamma_2} \vdash Q : A\lrangle{s'} \\\\
-    \Gamma_1 \vdash m : (I\ a_1...a_{k_2}) \\
-    \Gamma_2 \vdash f_i : C_i[I, Q] }
-    { \Gamma \vdash \case(m, Q)\{f_1|...|f_{k_1}\} : (Q\ a_1...a_{k_2}) }
-    \rname{Case}
-  \end{mathpar}
-  \Description{}
-  \label{case}
-\end{figure}
-
-\subsection{Dependent Case}
-
-\begin{definition}
-  For some non-linear constructor type $C$ and type variables $X$, $Q$ and $c$, the term $C\{X, Q, c\}$ is inductively defined over the structure of $C$ as follows.
-  \begin{align*}
-    ((\_ \utype P) \rightarrow C)\{X, Q, c\} & = (p \utype P) \multimap C\{X, Q, (c\ p)\} \\
-    ((x \utype M) \rightarrow C)\{X, Q, c\}  & = (x \utype M) \multimap C\{X, Q, (c\ x)\} \\
-    (X\ m_1...m_n)\{X, Q, c\}                & = (Q\ m_1...m_n\ c)
-  \end{align*}
-  We define the notation $C[I, P, t] := C\{X, Q, c\}[I/X, P/Q, t/c]$.
-\end{definition}
-
-Dependent elimination of non-linear inductive objects is presented in Figure \ref{dcase}.
-This elimination rule is dependent because the motive $Q$ may depend on the discriminated term $m$.
-For this very reason, $m$ is not allowed to have linear type as doing so may transport restricted variable into types.
-
-\begin{figure}[h]
-  \caption{Dependent Case Elimination}
-  \begin{mathpar}
-    \inferrule
-    { \pure{\Gamma_1} \\
-    (\forall i = 1...n) \\
-    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-    arity(A, U_i) \\\\
-    I := \ind{U}(X : A)\{C_1|...|C_n\} \\
-    \overline{\Gamma_2} \vdash Q : A\lrangle{I, s'} \\\\
-    \Gamma_1 \vdash m : (I\ a_1...a_n) \\
-    \Gamma_2 \vdash f_i : C_i[I, Q, Constr(i, I)] }
-    { \Gamma \vdash \dcase(m, Q)\{f_1|...|f_n\} : (Q\ a_1...a_n\ m) }
-    \rname{DCase}
-  \end{mathpar}
-  \Description{}
-  \label{dcase}
-\end{figure}
-
-\subsection{Fixpoint}
-The typing rule of fix-points presented in Figure \ref{fix} is mostly standard.
-A syntactic guard condition is utilized to conservatively ensure termination.
-The fix-point term must have non-linear type in a pure context, otherwise recursive calls may cause duplicated usage of restricted variables.
-
-\begin{figure}[h]
-  \caption{Fixpoint}
-  \begin{mathpar}
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : U_i \\
-      \Gamma, f \utype A \vdash m : A \\
-      \text{guard condition} }
-    { \Gamma \vdash \fix f : A := m : A }
-    \rname{fix}
-  \end{mathpar}
-  \Description{}
-  \label{fix}
-\end{figure}
-
-\subsection{Extended Reduction}
-The reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
-
-\begin{figure}[h]
-  \caption{$\iota$-Reductions}
-  \begin{mathpar}
-    \inferrule
-    { }
-    { \case((\constr(i, I)\ a_1...a_n),Q)\{f_1|...f_k\} \\\\
-      \leadsto (f_i\ a_1...a_k) }
-    \rname{Step-Case-$\iota$}
+    { x \pstep x }
+    \rname{Pstep-Var}
 
@@
     { }
-    { \dcase((\constr(i, I)\ a_1...a_k),Q)\{f_1|...f_k\} \\\\
-      \leadsto (f_i\ a_1...a_k) }
-    \rname{Step-DCase-$\iota$}
+    { s_i \pstep s_i }
+    \rname{Pstep-Sort}
 
     \inferrule
+    { m \pstep m' }
     { }
-    { \fix f : A := m \leadsto m[(\fix f : A := m)/f] }
-    \rname{Step-Fix-$\iota$}
   \end{mathpar}
-  \Description{}
-  \label{stepfix}
+  \label{pred}
 \end{figure}
 
-\subsection{Non-Linear Connectives}
-Non-linear inductive types have been extensively studied since their inception.
-We will omit discussion of non-linear inductive types here as all inductive definitions in the predicative fragment of CIC are available in CILC.
-
-\subsection{Linear Connectives}
-In Figure \ref{lconnectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
-For any term with inductive linear type, linearity will ultimately force the term to be eliminated by a Case expression.
-
-\begin{figure}[h]
-  \caption{Inductively Defined Linear Connectives}
-  \begin{align*}
-    \_\otimes\_ :=\
-     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
-     & \sep\ (A \utype L_i) \rightarrow (B \utype L_i)                               \\
-     & \qquad \rightarrow (\_ \ltype A)  \multimap (\_ \ltype B) \multimap (X\ A\ B) \\
-     & \}
-    \\
-    \lrangle{\_,\_} := \
-     & \constr(1, \otimes)
-    \\\\
-    \_\oplus\_ := \
-     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
-     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
-    \rightarrow (\_ \ltype A)  \multimap (X\ A\ B)                                   \\
-     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
-    \rightarrow (\_ \ltype B)  \multimap (X\ A\ B)                                   \\
-     & \}
-    \\
-    \inl(\_) := \
-     & \constr(1, \oplus)
-    \\
-    \inr(\_) := \
-     & \constr(2, \oplus)
-  \end{align*}
-  \Description{}
-  \label{lconnectives}
-\end{figure}
-
-\subsection{Mixed Connectives}
-In Figure \ref{mconnectives}, we demonstrate how the $F$ connective of LNL$_D$ \cite{neel15} can be expressed as an inductive linear type.
-The interesting property of $F$ is that the type of its linear component is dependent on the non-linear component.
-
-\begin{figure}[h]
-  \caption{Linear Non-Linear Connective}
-  \begin{align*}
-    F := \
-     & \ind{L}(X : (A \utype U_i) \rightarrow
-    (B \utype (\_ \utype A) \rightarrow L_i) \rightarrow L_i)\{ \\
-     & \sep \ (A \utype U_i) \rightarrow
-    (B \utype (\_ \utype A) \rightarrow L_i)                    \\
-     & \qquad \rightarrow (m \utype A) \rightarrow
-    (\_ \ltype B\ m) \multimap (X\ A\ B)                        \\
-     & \}
-    \\
-    [\_,\_] :=\
-     & \constr(1, F)
-  \end{align*}
-  \Description{}
-  \label{mconnectives}
-\end{figure}
-
-\section{Examples}
-We shall write examples in the concrete syntax of our implementation.
-We deliberately chose to model our concrete syntax after Coq such that it would be easier for readers already familiar with Coq to adapt to our syntax.
-
-\subsection{Stateful Reasoning}
-We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator.
-\begin{definition}
-  State axioms.
-  \begin{itemize}
-    \item $l \mapsto_A m : \mathbb{N} \rightarrow (A : U) \rightarrow A \rightarrow L$ \\
-          The $\mapsto$ is a linear type constructor.
-It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
-Intuitively, it is a proof that a term $m$ of non-linear type $A$ is currently stored at address $l$.
-    \item $\textbf{new} : (A : U) \rightarrow (m : A) \rightarrow F l:\mathbb{N}.(l \mapsto_A m)$ \\
-          $\new$ is a function that allocates a new memory cell for term $m$ of non-linear type $A$. It results in an address $l$ paired with an assertion $(l \mapsto_A m)$.
-    \item $\textbf{free} : (A : U) \rightarrow F l:\mathbb{N}.F m:A.(l \mapsto_A m) \rightarrow \unit$ \\
-          $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use.
-This can be seen in its input $F l:\mathbb{N}.F m:A.(l \mapsto_A m)$, where the assertion component $(l \mapsto_A m)$ is a proof that location $l$ is currently allocated.
-    \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
-          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
-The type of $\get$'s output is the rather complicated $F \_: (\Sigma x.
-x =_A m).
-(l \mapsto_A m)$.
-This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored.
-The proof $(l \mapsto_A m)$ is returned unchanged as well.
-    \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \multimap (n : B) \multimap (l \mapsto_B n)$ \\
-          $\set$ has the most complicated type of all the axioms presented here.
-Intuitively, if address $l$ is allocated and storing some term $m$ of type $A$, $\set$ may put term $n$ of type $B$ into the the store at address $l$, overwriting the original term.
-This is the so called "strong update" operation.
-The consumption of $(l \mapsto_A m)$ and return of $(l \mapsto_B n)$ accurately characterize this update process.
-  \end{itemize}
-\end{definition}
-Due to the fact that $(l \mapsto_A m)$ is an abstract linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
-This is the same non-sharing principle that empowers Separation Logic.
-Unlike Separation Logic, $(l \mapsto_A m)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
-
-The combination of linear types and dependent types within the state axioms enable the internalization of the so called "effect laws" for state, where these laws are now provable theorems within the CILC framework.
-
-\begin{enumerate}
-  \item Performing $\get$ twice on the same address will produce two terms that are provably equal.
-  \item After using $\set$ to update the store at an address $l$ to term $m$, performing $\get$ again on the same address will retrieve $m$.
-\end{enumerate}
-
-Figure \ref{law1} is an excerpt taken from our implementation.
-It demonstrates the internalization of Law 1.
-The usage of implicit parameters allows one to omit writing much of the redundant arguments that could be inferred from context.
-Here, $\new$ allocates a store for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-When $\get$ is applied to $c$ twice, the equality proofs \textit{pf1} $: (1 =_\mathbb{N} x)$ and \textit{pf2} $: (1 =_\mathbb{N} y)$ generated by $\get$ allow one to create proof \textit{pf} $: (x =_\mathbb{N} y)$ that the retrieved values $x$ and $y$ are equal.
-Finally, due to $c$ having linear type, it is required for it to be freed at the very end.
-
-\begin{figure}[h]
-  \caption{Law 1}
-  \begin{minted}{coq}
-  let [ _, c ] := new _ 1 in
-  let [ xeq, c ] := get _ _ _ c in
-  let [ yeq, c ] := get _ _ _ c in
-  let ( x, pf1 ) := xeq in
-  let ( y, pf2 ) := yeq in
-  let pf1 := eq_sym _ _ _ pf1 in
-  let pf : eq _ x y := 
-    eq_trans _ _ _ _ pf1 pf2 
-  in free _ _ _ c.
-  \end{minted}
-  \label{law1}
-  \Description{}
-\end{figure}
-
-Figure \ref{law2} demonstrates the internalization of Law 2.
-Again, implicit parameters fill in the annoying bureaucratic arguments whenever they can be inferred.
-Here, a store is allocated by $\new$ for natural number 1, creating a proof $c$ of the form $(l \mapsto_\mathbb{N} 1)$.
-After using $\set$ to update the store at $l$ to 2 and then performing $\get$ on $l$, one can create proof \textit{pf} $: (2 =_\mathbb{N} z)$ that the value $z$ retrieved by $\get$ is exactly equal to 2.
-
-\begin{figure}[h]
-  \caption{Law 2}
-  \begin{minted}{coq}
-  let [ _, c ] := new _ 1 in
-  let c := set _ _ _ _ c 2 in
-  let [ zeq, c ] := get _ _ _ c in
-  let ( z, pf1 ) := zeq in
-  let pf : eq _ 2 z := pf1 in
-  free _ _ _ c.
-  \end{minted}
-  \label{law2}
-  \Description{}
-\end{figure}
-
-\subsection{Protocol Enforcement}
-The inclusion of dependent types and linear types in CILC makes it extremely expressive when encoding communication protocols.
-By indexing each channel with a protocol, CILC can enforce communication on a channel to strictly adhere to its specified protocol.
-
-In Figure \ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
-Intuitively, the \texttt{session}  forms a stack of possible operations executed in a protocol.
-The $\SEND$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after sending a message of type $A$, the protocol progresses to stage $ss$.
-Likewise, the $\RECV$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after receiving a message of type $A$, the protocol progresses to stage $ss$.
-Finally, the $\END$ constructor indicates that the protocol has finished and communication is over.
-
-\begin{figure}[h]
-  \caption{Inductive Session Type}
-  \begin{minted}{coq}
-  Inductive session : U :=
-  | SEND : U -> session -> session
-  | RECV : U -> session -> session
-  | END : session.
-  \end{minted}
-  \label{session}
-  \Description{}
-\end{figure}
-
-In Definition \ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
-
-\begin{definition}
-  Communication axioms.
-  \begin{itemize}
-    \item $\textbf{channel} : \session \rightarrow L$ \\
-          The $\channel$ axiom is a type constructor. It is used to form the types of communication channels that obey the protocol specified by its session.
-    \item $\textbf{open} : (ss : \session) \rightarrow \channel\ ss$ \\
-          When given a protocol, $\open$ creates a channel indexed by this protocol.
-    \item $\textbf{close} : \channel\ \END \rightarrow \unit$ \\
-          Once all communications specified on a given channel has finished, $\close$ will close and reclaim the channel.
-    \item $\textbf{send} : (A : U) \rightarrow A \rightarrow (ss : \session) \rightarrow$ \\
-          \phantom{send : (A :)}
-          $\channel\ (\SEND\ A\ ss) \rightarrow \channel\ ss$ \\
-          After a channel has reached the point in its protocol where a message of type $A$ should be sent, $\send$ may send a term of type $A$ on this channel, causing the channel to progress to the next stage of its protocol.
-    \item $\textbf{recv} : (A : U) \rightarrow (ss : \session) \rightarrow $ \\
-          \phantom{recv : (A :)}
-          $\channel\ (\RECV\ A\ ss) \rightarrow F \_:A.\channel\ ss$
-          After a channel has reached the point in its protocol where a message of type $A$ should be received, $\recv$ may receive such a term of type $A$ from the channel, causing the channel to progress to the next stage of its protocol.
-  \end{itemize}
-  \label{comm}
-\end{definition}
-
-The important detail to notice here is that each channel created by $\open$ is of linear type $(\channel\ ss)$ where $ss$ is some arbitrary protocol, hence must be used exactly once.
-However, the $\send$ and $\recv$ operations return back channels after consuming them, progressing their protocols one stage forward.
-Coupled with the fact that the $\close$ operation is only able to close channels whose protocols have ended, this forces channels to communicate strictly by their specified protocols.
-
-Dependent types also allows for more precise specifications of protocols themselves.
-A simple example would be the specification of message sizes.
-Figure \ref{communication} shows such an example.
-The inductive type \texttt{ilist} is a length indexed list.
-Using \texttt{ilist}, we construct a protocol \texttt{ss} that first sends two natural numbers, then receives eight natural numbers and finally sends back a boolean acknowledgement.
-
-\begin{figure}[h]
-  \caption{Sized Communication}
-  \begin{minted}{coq}
-  Inductive ilist (A : U) : Nat -> U :=
-  | nil  : ilist A 0
-  | cons : A -> (n : Nat) -> 
-             ilist A n -> ilist A (S n).
-
-  let ss := 
-    SEND (ilist Nat 2) 
-      (RECV (ilist Nat 8) (SEND Bool END)) 
-  in
-  let send_msg : ilist Nat 2 := 
-    (cons 1 _ (cons 2 _ nil)) 
-  in
-  let ch := open ss in
-  let ch := send _ send_msg _ ch in
-  let [ recv_msg, ch ] := recv _ _ ch in
-  let ch := send _ true _ ch in
-  close ch.            
-  \end{minted}
-  \label{communication}
-  \Description{}
-\end{figure}
-
-\subsection{Mutable Data}
-Figure \ref{list} presents a CPS append function for linear lists.
-In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
-Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
-Each \texttt{_cons_} is applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
-Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
-Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.
-
-\begin{figure}[h]
-  \caption{Linear Lists}
-  \begin{minted}{coq}
-  Inductive list (A : U) : L :=
-  | nil : list A
-  | cons : A -> list A -> list A.
-
-  Fixpoint append (A : U) : list A -o 
-    list A -o (list A -o list A) -o list A 
-  :=
-    fun ls1 ls2 k =>
-      match ls1 with
-      | nil => k ls2
-      | (cons h t) as _cons_ =>
-        append _ t ls2 
-          (fun res => k (_cons_ h res))
-      end.
-  \end{minted}
-  \label{list}
-  \Description{}
-\end{figure}
-
-\section{Related Work}
-Linear types are a class of type systems inspired by Girard' substructural Linear Logic \cite{girard}.
-Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
-Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
-A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
-Benton \cite{benton1994} investigates the ramifications of the !
-exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
-Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
-
-Over the years, work has been done to enrich linear type theories with dependent types.
-Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
-Krishnaswami et al.
-present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
-Luo et al.
-\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
-The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
-
-\section{Future Work}
-On the theoretical side, we intend to investigate the semantic models of CLC.
-In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy.
-This has since been removed due to the additional difficulty of working with $Prop$ in the soundness proof of CILC.
-We believe that these difficulties are not intrinsic and $Prop$ could be added back with careful proof work.
-
-On the application side, we intend to bolster our implementation with more features such as compilation to executable binary, making it a full-featured theorem prover/programming language.
-We aim to fully verify the typechecking algorithm employed by the current implementation.
-The unification algorithm used for implicit argument resolution is extremely ad-hoc, as the problem of unification for linear types has been not thoroughly studied, we hope to improve this situation as well.
-
-\bibliographystyle{acm}
-\bibliography{../ref}
 \end{document}

### 2021-12-13 writing/appendix/draft.tex
appendix

@@
 
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+\theoremstyle{definition}
+\newtheorem{definition}{Definition}[section]
+
 \setlist[itemize]{leftmargin=1.6em}
@@
 \renewcommand{\syntright}{}
-\setlength{\grammarparsep}{7pt}
-\setlength{\grammarindent}{7em}
-\newcommand{\indalt}[1][2]{\\\hspace*{-0.2em}\textbar\quad}
+\setlength{\grammarparsep}{15pt}
+\setlength{\grammarindent}{10em}
+\newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
 \newcommand{\rname}[1]{\textsc{\footnotesize #1}}
@@
 \newcommand{\ltype}{:_{\scriptscriptstyle L}}
-\newcommand{\stype}[1]{:_#1}
+\newcommand{\stype}[1]{:_{#1}}
 \newcommand{\step}{\leadsto}
+\newcommand{\red}{\leadsto^*}
 \newcommand{\pstep}{\leadsto_p}
@@
   \begin{grammar}
-    <$i$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{2.4em} universe levels
+    <$i$> ::= 0 | 1 | 2 ... \phantom{* |} \hspace*{2.4em} universe levels
 
@@
 
-    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{3em} expressions
+    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4em} expressions
     \indalt $(x :_s A) \rightarrow B$
@@
     \inferrule
-    { m_1 \step^* n \\ m_2 \step^* n }
+    { m_1 \red n \\ m_2 \red n }
     { m_1 \equiv m_2 : A }
@@
     { x \pstep x }
-    \rname{Pstep-Var}
+    \rname{PStep-Var}
 
@@
     { s_i \pstep s_i }
-    \rname{Pstep-Sort}
+    \rname{PStep-Sort}
 
     \inferrule
-    { m \pstep m' }
-    { }
+    { A \pstep A' \\ m \pstep m' }
+    { \lambda x \stype{s} A.m \pstep \lambda x \stype{s} A'.m' }
+    \rname{PStep-$\lambda$}
+
+    \inferrule
+    { m \pstep m' \\ n \pstep n' }
+    { m\ n \pstep m'\ n' }
+    \rname{PStep-App}
+
+    \inferrule
+    { m \pstep m' \\ n \pstep n'}
+    { (\lambda x \stype{s} A.m)\ n \pstep m'[n'/x] }
+    \rname{PStep-$\beta$}
+
+    \inferrule
+    { A \pstep A' \\ B \pstep B' }
+    { (x \stype{s} A) \rightarrow B \pstep (x \stype{s} A') \rightarrow B' }
+    \rname{PStep$\rightarrow$}
+
+    \inferrule
+    { A \pstep A' \\ B \pstep B' }
+    { (x \stype{s} A) \multimap B \pstep (x \stype{s} A') \multimap B' }
+    \rname{PStep$\multimap$}
   \end{mathpar}
@@
 
+\subsection{Reduction Lemmas}
+Here, we prove some simple lemmas concerning $\step$, $\red$ and substitution.
+
+\begin{definition}
+  For a term $m$ and a map $\sigma$ from variables to terms, let $m[\sigma]$ be the term obtained by applying $\sigma$ uniformly to all free variables in $m$.
+\end{definition}
+
+\begin{definition}
+  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ reduces to $\tau$ if for any variable $x$ there exists a reduction $(\sigma\ x) \red (\tau\ x)$.
+We write $\sigma \red \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{stepsubst}
+  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a step $m \step n$, then there exists a step $m[\sigma] \step n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step n$.
+\end{proof}
+
+\begin{lemma}\label{redapp}
+  For terms $m_1, m_2, n_1, n_2$, if these exists reductions $m_1 \red m_2$ and $n_1 \red n_2$, then there exists reduction $(m_1\ n_1) \red (m_2\ n_2)$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-AppL}, \rname{Step-AppR}.
+\end{proof}
+
+\begin{lemma}\label{redlam}
+  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $m_1 \red m_2$, then there exists reduction $\lambda x \stype{s}A_1.m_1 \red \lambda x \stype{s}A_2.m_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
+\end{proof}
+
+\begin{lemma}\label{redarrow}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \rightarrow B_1 \red (x \stype{s} A_2) \rightarrow B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
+\end{proof}
+
+\begin{lemma}\label{redlolli}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \multimap B_1 \red (x \stype{s} A_2) \multimap B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
+\end{proof}
+
+\begin{lemma}\label{redsubst}
+  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a reduction $m \red n$, then there exists a reduction $m[\sigma] \red n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\red$, the transitivity of $\red$ and Lemma \ref{stepsubst}.
+\end{proof}
+
+\begin{lemma}\label{redcompat}
+  For maps $\sigma, \tau$ from variables to terms, if there is a map reduction $\sigma \red \tau$, then for any term $m$ these is a reduction $m[\sigma] \red m[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$, applying Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}.
+\end{proof}
+
+\subsection{Equality Lemmas}
+Here, we prove some simple lemmas concerning $\red$, $\equiv$ and substitution.
+
+\begin{definition}
+  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ is equal to $\tau$ if for any variable $x$ there exists an equality $(\sigma\ x) \equiv (\tau\ x)$.
+We write $\sigma \equiv \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{convhom}
+  For any map $f$ from terms to terms, if for any terms $m, n$ such that $m \step n$ implies $f\ m \equiv f\ n$, then for any terms $m, n$ equality $m \equiv n$ implies $f\ m \equiv f\ n$.
+\end{lemma}
+\begin{proof}
+  By the properties of the transitive reflexive closure $\red$ and that $\equiv$ is an equivalence relation.
+\end{proof}
+
+\begin{lemma}\label{convapp}
+  For terms $m_1, m_2, n_1, n_2$, if there exists equalities $m_1 \equiv m_2$ and $n_1 \equiv n_2$, then there exists equality $(m_1\ n_1) \equiv (m_2\ n_2)$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-AppL}, \rname{Step-AppR}.
+\end{proof}
+
+\begin{lemma}\label{convlam}
+  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $m_1 \equiv m_2$, then there exists equality $\lambda x \stype{s} A_1.m_1 \equiv \lambda x \stype{s} A_2.m_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
+\end{proof}
+
+\begin{lemma}\label{convarrow}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{s} A_2) \rightarrow B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
+\end{proof}
+
+\begin{lemma}\label{convlolli}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{s} A_2) \multimap B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
+\end{proof}
+
+\begin{lemma}
+  For terms $m, n$ and map $\sigma$ from variables to terms, if there is equality $m \equiv n$, then there is equality $m[\sigma] \equiv n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By Lemmas \ref{convhom} and \ref{stepsubst}.
+\end{proof}
+
+\begin{lemma}\label{convcompat}
+  For maps $\sigma, \tau$ from variables to terms and term $m$, if these is map equality $\sigma \equiv \tau$, then there is equality $m[\sigma] \equiv m[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$, applying Lemmas \ref{convapp}, \ref{convlam}, \ref{convarrow}, \ref{convlolli}.
+\end{proof}
+
+\begin{lemma}
+  For terms $m_1, m_2, n$, if there is equality $m_1 \equiv m_2$, then there is equality $n[m_1/x] \equiv n[m_2/x]$ for any variable $x \in FV(n)$.
+\end{lemma}
+\begin{proof}
+  This is a special case of Lemma \ref{convcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+\end{proof}
+
+\subsection{Parallel Reduction Lemmas}
+
+\begin{definition}\label{psstep}
+  For maps $\sigma, \tau$ from variables to terms, we say $\sigma$ parallel reduces to $\tau$ if for any variable $x$ there exists a parallel reduction $(\sigma\ x) \pstep (\tau\ x)$.
+We write $\sigma \pstep \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{psteprefl}
+  For any term $m$, there exists a reflexive parallel reduction $m \pstep m$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$.
+\end{proof}
+
+\begin{lemma}
+  For any map $\sigma$ from variables to terms, there exists a reflexive parallel map reduction $\sigma \pstep \sigma$.
+\end{lemma}
+\begin{proof}
+  By Definition \ref{psstep} and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{steppstep}
+  For any terms $m, n$, if there exists step $m \step n$, then there exists a parallel reduction $m \pstep n$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step n$ and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{pstepred}
+  For terms $m, n$, if there exists parallel reduction $m \pstep n$, then there exists a reduction $m \red n$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$, utilizing the transitive property of $\red$ and Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}, \ref{redsubst}, \ref{redcompat}.
+\end{proof}
+
+\begin{lemma}
+  For terms $m, n$ and map $\sigma$ from variables to terms, if there exists parallel reduction $m \pstep n$, there exists parallel reduction $m[\sigma] \pstep n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$ and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{pstepcompat}
+  For terms $m, n$ and maps $\sigma, \tau$ from variables to terms, if there exists parallel reduction $m \pstep n$ and parallel map reduction $\sigma \pstep \tau$, there exists parallel reduction $m[\sigma] \pstep n[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$.
+\end{proof}
+
+\begin{lemma}
+  For terms $m_1, m_2, n$, if there is parallel reduction $m_1 \pstep m_2$, then there is parallel reduction $n[m_1/x] \pstep n[m_2/x]$ for any variable $x \in FV(n)$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{psteprefl}, this is a special case of Lemma \ref{pstepcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+\end{proof}
+
+\subsection{Confluence Theorem}
+We first show that $\pstep$ satisfies the diamond property. Using the diamond property, we ultimately prove the confluence theorem.
+
+\begin{lemma}\label{diamond}
+  CLC term reduction has the diamond property.
+For terms $m, m_1, m_2$, if there are parallel reductions $m \pstep m_1$ and $m \pstep m_2$, then there exists term $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep m_1$.
+Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and apply the induction hypothesis.
+The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma \ref{pstepcompat} is used to push these cases through.
+\end{proof}
+
+\begin{lemma}\label{strip}
+  Strip lemma. For terms $m, m_1, m_2$, if there is parallel reduction $m \pstep m_1$ and reduction $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \pstep m'$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{diamond}.
+\end{proof}
+
+\begin{theorem}
+  CLC term reduction is confluent. For terms $m, m_1, m_2$, if there are reductions $m \red m_1$ and $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \red m'$.
+\end{theorem}
+\begin{proof}
+  By induction on the derivation of $m \red m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{strip}.
+\end{proof}
+
+\subsection{Corollaries of Confluence}
+The following results are all corollaries of confluence, proven using a combination of induction, transitivity and confluence.
+These corollaries allow us to refute false reductions and equalities in future proofs.
+
+\begin{corollary}
+  For a universe $s_i$ and term $m$, if there is reduction $s_i \red m$, then $m = s_i$.
+\end{corollary}
+
+\begin{corollary}
+  For variable $x$ and term $m$, if there is reduction $x \red m$, then $m = x$.
+\end{corollary}
+
+\begin{corollary}
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
+\end{corollary}
+
+\begin{corollary}
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
+\end{corollary}
+
+\begin{corollary}
+  For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exists $A',m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
+\end{corollary}
+
+\begin{corollary}
+  For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there is $s = t$ and $i = j$.
+\end{corollary}
+
+\begin{corollary}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{t} A_2) \rightarrow B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
+\end{corollary}
+
+\begin{corollary}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{t} A_2) \multimap B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
+\end{corollary}
+
+\section{Context of CLC \texttt{(clc_context.v)}}
+Contexts of CLC are of the form $x_1 \stype{s_1} A_1, x_2 \stype{s_2} A_2, ...
+x_k \stype{s_k} A_k$ where each free variable $x_i$ is assigned a type $A_i$ and sort $s_i$.
+Contexts will be referred to by meta variables $\Gamma$ and $\Delta$.
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \inferrule
+    { }
+    { \epsilon \vdash }
+    \rname{Wf-$\epsilon$}
+
+    \inferrule
+    { \Gamma\ \vdash \\
+      \overline{\Gamma} \vdash A : U_i }
+    { \Gamma, x \utype A \vdash }
+    \rname{Wf-U}
+
+    \inferrule
+    { \Gamma\ \vdash \\
+      \overline{\Gamma} \vdash A : L_i }
+    { \Gamma, x \ltype A\ \vdash }
+    \rname{Wf-L}
+    \\
+
+    \inferrule
+    { }
+    { \pure{\epsilon} }
+    \rname{Pure-$\epsilon$}
+
+    \inferrule
+    { \pure{\Gamma} \\
+      \Gamma \vdash A : U_i }
+    { \pure{\Gamma, x \utype A} }
+    \rname{Pure-U}
+    \\
+
+    \inferrule
+    { }
+    { \mrg{\epsilon}{\epsilon}{\epsilon} }
+    \rname{Merge-$\epsilon$}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+    { \mrg{\Gamma_1, x \utype A}
+      {\Gamma_2, x \utype A}
+      {\Gamma, x \utype A} }
+    \rname{Merge-U}
+    \\
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_2 }
+    { \mrg{\Gamma_1, x \ltype A}
+      {\Gamma_2}
+      {\Gamma, x \ltype A} }
+    \rname{Merge-L1}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_1 }
+    { \mrg{\Gamma_1}
+      {\Gamma_2, x \ltype A}
+      {\Gamma, x \ltype A} }
+    \rname{Merge-L2}
+  \end{mathpar}
+  \label{structural}
+\end{figure}
+
+\subsection{Merge Lemmas}
+Since weakening and contraction rules will not be allowed on restricted variables, it is necessary to have lemmas that enable the manipulation of contexts.
+
+\begin{lemma}\label{mergesym}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\mrg{\Gamma_2}{\Gamma_1}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure}
+  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\mrg{\Gamma}{\Gamma}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\pure{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergere1}
+  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\Gamma}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergere2}
+  For any context $\Gamma$, there is $\mrg{\Gamma}{\overline{\Gamma}}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergepureinv}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma}$, then there is $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure1}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_1}$, then there is $\Gamma = \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure2}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$, then there is $\Gamma = \Gamma_1$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepurepure}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\pure{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepureeq}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\Gamma_1 = \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergerere}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergererere}
+  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\overline{\Gamma}}{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergesplit1}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\subsection{Restriction and Purity Lemmas}
+
+\begin{lemma}\label{rere}
+  For any context $\Gamma$, there is $\overline{\Gamma} = \overline{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{purere}
+  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\Gamma = \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{repure}
+  For any context $\Gamma$ , there is $\pure{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
 \end{document}

### 2021-12-30 writing/appendix/draft.tex
writing

@@
 This extended report is meant to accompany our paper of the same title.
-Here, we describe the meta-theory of CILC and their proofs in detail.
+Here, we describe the meta-theory of CLC and CILC and their proofs in detail.
 All the results presented here have been formalized and proven correct in the Coq Proof Assistant.
@@
 
-\begin{lemma}
+\begin{lemma}\label{convsubst}
   For terms $m, n$ and map $\sigma$ from variables to terms, if there is equality $m \equiv n$, then there is equality $m[\sigma] \equiv n[\sigma]$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{redsortinv}
   For a universe $s_i$ and term $m$, if there is reduction $s_i \red m$, then $m = s_i$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{redvarinv}
   For variable $x$ and term $m$, if there is reduction $x \red m$, then $m = x$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{redarrowinv}
   For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{redlolliinv}
   For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{redlaminv}
   For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exists $A',m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{sortinj}
   For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there is $s = t$ and $i = j$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{arrowinj}
   For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{t} A_2) \rightarrow B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
@@
 
-\begin{corollary}
+\begin{corollary}\label{lolliinj}
   For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{t} A_2) \multimap B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
@@
 
-\subsection{Merge Lemmas}
+\subsection{Merge Lemmas}\label{mergeprop}
 Since weakening and contraction rules will not be allowed on restricted variables, it is necessary to have lemmas that enable the manipulation of contexts.
@@
 
-\begin{lemma}\label{mergesplit1}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
 \subsection{Restriction and Purity Lemmas}
@@
 
+\begin{lemma}\label{hasure}
+  For any context $\Gamma$, variable $x$ and type $A$, if there is $x \utype A \in \Gamma$, then there is $x \utype A \in \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $x \utype A \in \Gamma$.
+\end{proof}
+
+\begin{lemma}\label{haslre}
+  For any context $\Gamma$, variable $x$ and type $A$, there is $x \ltype A \notin \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergesplit1}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_2}{\Gamma_2}{\Delta}$ and $\mrg{\Delta_1}{\Delta}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\section{Subtyping of CLC \texttt{(clc_subtype.v)}}
+
+The cumulativity relation ($\preceq$) is the smallest binary relation over terms such that
+\begin{enumerate}
+  \item $\preceq$ is a partial order with respect to equality.
+        \begin{enumerate}
+          \item If $A \equiv B$, then $A \preceq B$.
+          \item If $A \preceq B$ and $B \preceq A$, then $A \equiv B$.
+          \item If $A \preceq B$ and $B \preceq C$, then $A \preceq B$.
+        \end{enumerate}
+  \item $U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
+  \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
+  \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
+        $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{s} A_2) \rightarrow B_2$
+  \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
+        $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{s} A_2) \multimap B_2$
+\end{enumerate}
+
+\noindent
+Here, we give an inductive definition of the cumulativity relation ($\preceq$) that is suitable for writing proofs.
+
+\begin{mathpar}
+  \inferrule
+  { }
+  { A \prec A }
+  \rname{$\prec$-Refl}
+
+  \inferrule
+  { i_1 \leq i_2 }
+  { s_{i_1} \prec s_{i_2} }
+  \rname{$\prec$-Sort}
+
+  \inferrule
+  { B_1 \prec B_2 }
+  { (x \stype{s} A) \rightarrow B_1 \prec (x \stype{s} A) \rightarrow B_2 }
+  \rname{$\prec$$\rightarrow$}
+
+  \inferrule
+  { B_1 \prec B_2 }
+  { (x \stype{s} A) \multimap B_1 \prec (x \stype{s} A) \multimap B_2 }
+  \rname{$\prec$$\multimap$}
+
+  \inferrule
+  { A' \prec B' \\ A \equiv A' \\ B \equiv B' }
+  { A \preceq B }
+  \rname{$\prec$-$\preceq$}
+\end{mathpar}
+
+\subsection{Subtyping Lemmas}
+
+\begin{lemma}\label{sub1sub}
+  For terms $A, B$, if there is $A \prec B$, then there is $A \preceq B$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the reflexivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{sub1conv}
+  For terms $A, B, C$, if there is $A \prec B$ and $B \equiv C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{convsub1}
+  For terms $A, B, C$, if there is $A \equiv B$ and $B \prec C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{convsub}
+  For terms $A, B$, if there is $A \equiv B$, then there is $A \preceq B$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{convsub1} and \rname{$\prec$-Refl}.
+\end{proof}
+
+\begin{lemma}\label{subrefl}
+  For term $A$, there is $A \preceq A$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{sub1sub} and \rname{$\prec$-Refl}.
+\end{proof}
+
+\begin{lemma}\label{subprop}
+  For natural numbers $i, j$ and sort $s$ such that $i \leq j$, there is $s_i \preceq s_j$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{sub1sub} and \rname{$\prec$-Sort}.
+\end{proof}
+
+\begin{lemma}\label{sub1trans}
+  For terms $A, B, C, D$, if there is $A \prec B$, $B \equiv C$ and $C \prec D$, then there is $A \preceq D$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $A \prec B$, definition of $\prec$ and Lemmas \ref{sub1sub}, \ref{sub1conv}, \ref{convsub1}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A, B, C$, if there is $A \preceq B$ and $B \preceq C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$, rule \rname{$\prec$-$\preceq$} and Lemma \ref{sub1trans}.
+\end{proof}
+
+\begin{lemma}
+  For sorts $s, t$ and natural numbers $i, j$, if there is $s_i \preceq t_j$, then there is $s = t$ and $i \leq j$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{sortinj}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{t} A_2) \rightarrow B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{arrowinj}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{t} A_2) \multimap B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{lolliinj}.
+\end{proof}
+
+\begin{lemma}\label{sub1subst}
+  For terms $A, B$  and map $\sigma$ from variables to terms, if there is $A \prec B$, then there is $A[\sigma] \prec B[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $A \prec B$ and the definition of $\prec$.
+\end{proof}
+
+\begin{lemma}
+  For terms $A, B$ and map $\sigma$ from variables to terms, if there is $A \preceq B$, then there is $A[\sigma] \preceq B[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By rule \rname{$\prec$-$\preceq$} and Lemmas \ref{convsubst}, \ref{sub1subst}.
+\end{proof}
+
+\section{Typing of CLC \texttt{(clc_typing.v)}}
+
+The following rules define well-formed contexts.
+\begin{mathpar}
+  \inferrule
+  { }
+  { \epsilon \vdash }
+  \rname{$\epsilon$-Ok}
+
+  \inferrule
+  { \Gamma \vdash \\
+    \overline{\Gamma} \vdash A : U_i }
+  { \Gamma, x \utype A \vdash }
+  \rname{U-Ok}
+
+  \inferrule
+  { \Gamma \vdash \\
+    \overline{\Gamma} \vdash A : L_i }
+  { \Gamma, x \ltype A \vdash }
+  \rname{L-Ok}
+\end{mathpar}
+
+\noindent
+The typing rules of CLC are presented below.
+\begin{mathpar}
+  \inferrule
+  { \pure{\Gamma} }
+  { \Gamma \vdash s_i : U_{i+1} }
+  \rname{Sort-Axiom}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : U_i \\
+    \Gamma, x \utype A \vdash B : s_i }
+  { \Gamma \vdash (x \utype A) \rightarrow B : U_i }
+  \rname{U$\rightarrow$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : L_i \\
+    \Gamma \vdash B : s_i \\
+    x \notin \Gamma }
+  { \Gamma \vdash (x \ltype A) \rightarrow B : U_i }
+  \rname{L$\rightarrow$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : U_i \\
+    \Gamma, x \utype A \vdash B : s_i }
+  { \Gamma \vdash (x \utype A) \multimap B : L_i }
+  \rname{U$\multimap$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : L_i \\
+    \Gamma \vdash B : s_i \\
+    x \notin \Gamma }
+  { \Gamma \vdash (x \ltype A) \multimap B : L_i }
+  \rname{L$\multimap$}
+
+  \inferrule
+  { \pure{\Gamma_1, \Gamma_2} }
+  { \Gamma_1, x \utype A, \Gamma_2 \vdash x : A }
+  \rname{U-Var}
+
+  \inferrule
+  { \pure{\Gamma_1, \Gamma_2} }
+  { \Gamma_1, x \ltype A, \Gamma_2 \vdash x : A }
+  \rname{L-Var}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash (x \stype{s} A) \rightarrow B : t_i \\
+    \Gamma, x \stype{s} A \vdash n : B }
+  { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \rightarrow B }
+  \rname{$\lambda$$\rightarrow$}
+
+  \inferrule
+  { \overline{\Gamma} \vdash (x \stype{s} A) \multimap B : t_i \\
+    \Gamma, x \stype{s} A \vdash n : B }
+  { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \multimap B }
+  \rname{$\lambda$$\multimap$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \utype A) \rightarrow B \\
+    \pure{\Gamma_2} \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-U$\rightarrow$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \ltype A) \rightarrow B \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-L$\rightarrow$}
+\end{mathpar}
+
+\begin{mathpar}
+  \inferrule
+  { \Gamma_1 \vdash m : (x \utype A) \multimap B \\
+    \pure{\Gamma_2} \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-U$\multimap$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \ltype A) \multimap B \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-L$\multimap$}
+
+  \inferrule
+  { \Gamma \vdash m : A \\
+    \overline{\Gamma} \vdash B : s_i \\ A \preceq B }
+  { \Gamma \vdash m : B }
+  \rname{Conversion}
+\end{mathpar}
+
+\section{Inversion Lemmas of CLC \texttt{(clc_inversion.v)}}
+
+\begin{lemma}\label{uarrowinv}
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \rightarrow B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash (x \utype A) \rightarrow B : s$.
+\end{proof}
+
+\begin{lemma}\label{larrowinv}
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\rightarrow B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash (x \ltype A) \rightarrow B : s$.
+\end{proof}
+
+\begin{lemma}\label{ulolliinv}
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \multimap B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash (x \utype A) \multimap B : s$.
+\end{proof}
+
+\begin{lemma}\label{llolliinv}
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\multimap B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash (x \ltype A) \multimap B : s$.
+\end{proof}
+
+\begin{lemma}\label{arrowlaminvx}
+  For any context $\Gamma$, terms $A, n, C$ and sort $s$, if there is $\Gamma \vdash \lambda x \stype{s} A .
+n : C$, then for all terms $A', B$, sorts $s', t$ and natural number $i$ such that $C \preceq (x \stype{s'} A') \rightarrow B$ and $\overline{\Gamma, x \stype{s'} A'} \vdash B : t_i$, there is $\Gamma, x \stype{s'} A' \vdash n : B$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas \ref{uarrowinv}, \ref{larrowinv}.
+\end{proof}
+
+\begin{lemma}\label{lollilaminvx}
+  For any context $\Gamma$, terms $A, n, C$ and sort $s$, if there is $\Gamma \vdash \lambda x \stype{s} A .
+n : C$, then for all terms $A', B$, sorts $s', t$ and natural number $i$ such that $C \preceq (x \stype{s'} A') \multimap B$ and $\overline{\Gamma, x \stype{s'} A'} \vdash B : t_i$, there is $\Gamma, x \stype{s'} A' \vdash n : B$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas \ref{ulolliinv}, \ref{llolliinv}.
+\end{proof}
+
+\begin{lemma}\label{arrowlaminv}
+  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there is $\overline{\Gamma} \vdash (x \stype{s'} A') \rightarrow B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
+n : (x \stype{s'} A') \rightarrow B$, then there is $\Gamma, x \stype{s'} A' \vdash n : B$.
+\end{lemma}
+\begin{proof}
+  Direct consequence of Lemmas \ref{uarrowinv}, \ref{larrowinv} and \ref{arrowlaminvx}.
+\end{proof}
+
+\begin{lemma}\label{lollilaminv}
+  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there is $\overline{\Gamma} \vdash (x \stype{s'} A') \multimap B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
+n : (x \stype{s'} A') \multimap B$, then there is $\Gamma, x \stype{s'} A' \vdash n : B$.
+\end{lemma}
+\begin{proof}
+  Direct consequence of Lemmas \ref{ulolliinv}, \ref{llolliinv} and \ref{lollilaminvx}.
+\end{proof}
+
+\section{Weakening Lemmas of CLC \texttt{(clc_weakening.v)}}
+
+Weakening for non-linear types is admissible in CLC. To prove this, we first define an $agreeR$ relation between two contexts $\Gamma, \Gamma'$ and a mapping $\xi$ from variables to variables.
+
+\begin{mathpar}
+  \inferrule
+  { }
+  { agreeR\ \xi\ \epsilon\ \epsilon }
+  \rname{agreeR-$\epsilon$}
+
+  \inferrule
+  { agreeR\ \xi\ \Gamma\ \Gamma' \\
+    x \notin FV(\Gamma) \cup FV(\Gamma') }
+  { agreeR\ (\xi \cup (x, x))\ (\Gamma, x \utype A) (\Gamma', x \utype A[\xi]) }
+  \rname{agreeR-U}
+
+  \inferrule
+  { agreeR\ \xi\ \Gamma\ \Gamma' \\
+    x \notin FV(\Gamma) \cup FV(\Gamma') }
+  { agreeR\ (\xi \cup (x, x))\ (\Gamma, x \ltype A) (\Gamma', x \ltype A[\xi]) }
+  \rname{agreeR-L}
+
+  \inferrule
+  { agreeR\ \xi\ \Gamma\ \Gamma' \\
+    x \notin FV(\Gamma) \cup FV(\Gamma') }
+  { agreeR\ \xi\ \Gamma\ (\Gamma', x \utype A) }
+  \rname{agreeR-wk}
+\end{mathpar}
+
+\subsection{Properties of $agreeR$}\label{agreeRprop}
+
+\begin{lemma}\label{agreerenrefl}
+  For any context $\Gamma$ and the identity map $id$ from variables to variables, $agreeR\ id\ \Gamma\ \Gamma$ is always true.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$ and the definition of $agreeR$.
+\end{proof}
+
+\begin{lemma}\label{agreerenpure}
+  For contexts $\Gamma, \Gamma'$ and mapping $\xi$, if there is $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\pure{\Gamma}$, then there is $\pure{\Gamma'}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $agreeR\ \xi\ \Gamma\ \Gamma'$.
+\end{proof}
+
+\begin{lemma}\label{agreerenrere}
+  For contexts $\Gamma, \Gamma'$ and mapping $\xi$, if there is $agreeR\ \xi\ \Gamma\ \Gamma'$, then there is $agreeR\ \xi\ \pure{\Gamma}\ \pure{\Gamma'}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $agreeR\ \xi\ \Gamma\ \Gamma'$.
+\end{proof}
+
+\subsection{Weakening Theorem}
+
+\begin{lemma}\label{mergeagreereninv}
+  For contexts $\Gamma, \Gamma', \Gamma_1, \Gamma_2$ and mapping $\xi$, if there is $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exists $\Gamma_1', \Gamma_2'$ such that $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $agreeR\ \xi\ \Gamma\ \Gamma'$ and lemmas in Section \ref{agreeRprop}.
+\end{proof}
+
+\begin{lemma}\label{renaming}
+  For context $\Gamma, \Gamma'$, terms $m, A$ and mapping $\xi$, if there is $\Gamma \vdash m : A$ and $agreeR\ \xi\ \Gamma\ \Gamma'$, then there is $\Gamma' \vdash m[\xi] : A[\xi]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash m : A$.
+We shall only discuss the application case in detail, as the other cases are proven by application of the induction hypothesis and the lemmas in Section \ref{agreeRprop}.
+  \begin{itemize}
+    \item For the \rname{App-U$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
+At this point, the induction hypothesis allows us to apply \rname{App-U$\rightarrow$} to prove the goal.
+    \item For the \rname{App-L$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+At this point, the induction hypothesis allows us to apply \rname{App-L$\rightarrow$} to prove the goal.
+    \item For the \rname{App-U$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
+At this point, the induction hypothesis allows us to apply \rname{App-U$\multimap$} to prove the goal.
+    \item For the \rname{App-L$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+At this point, the induction hypothesis allows us to apply \rname{App-L$\multimap$} to prove the goal.
+  \end{itemize}
+\end{proof}
+
+\begin{theorem}
+  Weakening is admissible for CLC variables of non-linear type. For context $\Gamma$ and terms $m, A, B$, if there is $\Gamma \vdash m : A$, then there is $\Gamma, x \utype B \vdash m : A$.
+\end{theorem}
+\begin{proof}
+  Using \rname{agreeR-wk} and Lemma \ref{agreerenrefl} a proof of $agreeR\ id\ \Gamma\ (\Gamma, x \utype B)$ can be constructed. Then by Lemma \ref{renaming}, the theorem can be proven.
+\end{proof}
+
+\section{Substitution Lemmas of CLC \texttt{(clc_substitution.v)}}
+
+Similar to the proof of weakening, we first define an $agreeS$ relation between two contexts $\Gamma, \Delta$ and a mapping $\sigma$ from variables to terms.
+
+\begin{mathpar}
+  \inferrule
+  { }
+  { agreeS\ \sigma\ \epsilon\ \epsilon }
+  \rname{agreeS-$\epsilon$}
+
+  \inferrule
+  { agreeS\ \sigma\ \Delta\ \Gamma \\
+    x \notin FV(\Delta) \cup FV(\Gamma) }
+  { agreeS\ (\sigma \cup (x, x))\ (\Delta, x \utype A[\sigma])\ (\Gamma, x \utype A) }
+  \rname{agreeS-U}
+
+  \inferrule
+  { agreeS\ \sigma\ \Delta\ \Gamma \\
+    x \notin FV(\Delta) \cup FV(\Gamma) }
+  { agreeS\ (\sigma \cup (x, x))\ (\Delta, x \ltype A[\sigma])\ (\Gamma, x \ltype A) }
+  \rname{agreeS-L}
+
+  \inferrule
+  { agreeS\ \sigma\ \Delta\ \Gamma \\
+    \overline{\Delta} \vdash n : A[\sigma] \\
+    x \notin FV(\Delta) \cup FV(\Gamma) }
+  { agreeS\ (\sigma \cup (x, n))\ \Delta\ (\Gamma, x \utype A) }
+  \rname{agreeS-wkU}
+
+  \inferrule
+  { \mrg{\Delta_1}{\Delta_2}{\Delta} \\
+    agreeS\ \sigma\ \Delta_1\ \Gamma \\
+    \Delta_2 \vdash n : A[\sigma] \\
+    x \notin FV(\Delta) \cup FV(\Gamma) }
+  { agreeS\ (\sigma \cup (x, n))\ \Delta\ (\Gamma, x \ltype A) }
+  \rname{agreeS-wkL}
+\end{mathpar}
+
+\begin{mathpar}
+  \inferrule
+  { A \preceq B \\
+    \overline{\Delta} \vdash B[\sigma] : U_i \\
+    agreeS\ \sigma\ \Delta\ (\Gamma, x \utype A) }
+  { agreeS\ \sigma\ \Delta\ (\Gamma, x \utype B) }
+  \rname{agreeS-convU}
+
+  \inferrule
+  { A \preceq B \\
+    \overline{\Delta} \vdash B[\sigma] : L_i \\
+    \overline{\Gamma} \vdash B : L_i \\
+    agreeS\ \sigma\ \Delta\ (\Gamma, x \ltype A) }
+  { agreeS\ \sigma\ \Delta\ (\Gamma, x \ltype B) }
+  \rname{agreeS-convL}
+\end{mathpar}
+
+\subsection{Properties of $agreeS$}\label{agreeSprop}
+
+\begin{lemma}\label{agreesubstrefl}
+  For any context $\Gamma$ and identity mapping id, there is $agreeS\ id\ \Gamma\ \Gamma$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{agreesubstrere}
+  For contexts $\Delta, \Gamma$ and mapping $\sigma$, if there is $agreeS\ \sigma\ \Delta\ \Gamma$, then there is $agreeS\ \sigma\ \overline{\Delta}\ \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $agreeS\ \sigma\ \Delta\ \Gamma$.
+\end{proof}
+
+\subsection{Substitution Lemma}
+
+\begin{lemma}\label{mergeagreesubstinv}
+  For contexts $\Delta, \Gamma, \Gamma_1, \Gamma_2$ and mapping $\sigma$, if there is $agreeS\ \sigma\ \Delta\ \Gamma$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exists contexts $\Delta_1, \Delta_2$ such that $\mrg{\Delta_1}{\Delta_2}{\Delta}$ and $agreeS\ \sigma\ \Delta_1\ \Gamma_1$ and $agreeS\ \sigma\ \Delta_2\ \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $agreeS\ \sigma\ \Delta\ \Gamma$ and lemmas in Section \ref{agreeSprop}.
+\end{proof}
+
+\begin{lemma}
+  Generalized Substitution Lemma.
+For context $\Gamma, \Delta$, terms $m, A$ and mapping $\sigma$, if there is $\Gamma \vdash m : A$ and $agreeS\ \sigma\ \Delta\ \Gamma$, then there is $\Delta \vdash m[\sigma] : A[\sigma]$.
+\end{lemma}
+\begin{proof}
+  The proof proceeds by induction on the derivation of $\Gamma \vdash m : A$.
+Similar to the proof of Lemma \ref{renaming}, the interesting cases are the application cases where Lemma \ref{mergeagreesubstinv} must be utilized to split the $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ judgments for use in the induction hypothesis.
+\end{proof}
+
+\subsection{Corollaries of Substitution}
+
+\begin{corollary}\label{substitutionu}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there is $\Gamma_1, x \utype A \vdash m : B$ and $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
+\end{corollary}
+
+\begin{corollary}\label{substitutionl}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there is $\Gamma_1, x \ltype A \vdash m : B$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
+\end{corollary}
+
+\begin{corollary}\label{contextconvu}
+  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there is $B \equiv A$ and $\overline{\Gamma} \vdash A : U_i$ and $\Gamma, x \utype A \vdash m : C$, then there is $\Gamma, x \utype B \vdash m : C$.
+\end{corollary}
+
+\begin{corollary}\label{contextconvl}
+  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there is $B \equiv A$ and $\overline{\Gamma} \vdash A : L_i$ and $\Gamma, x \ltype A \vdash m : C$, then there is $\Gamma, x \ltype B \vdash m : C$.
+\end{corollary}
+
+\section{Typing Validity of CLC \texttt{(clc_validity.v)}}
+
+In this section, we prove that the types of all CLC terms are themselves well-sorted.
+
+\begin{lemma}\label{mergecontextokinv}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$, then there is $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and the properties of $\mrg{\_}{\_}{\_}$ discussed in Section \ref{mergeprop}.
+\end{proof}
+
+\begin{theorem}\label{validity}
+  The validity of typing theorem.
+For any context $\Gamma$ and terms $m, A$, if there is $\Gamma \vdash$ and $\Gamma \vdash m : A$, then there exists sort $s$ and natural number $i$ such that $\overline{\Gamma} \vdash A : s_i$.
+\end{theorem}
+
+\section{Subject Reduction of CLC \texttt{(clc_soundness.v)}}
+
+\begin{theorem}
+  For any context $\Gamma$ and terms $m, n, A$, if $\Gamma \vdash$ and $\Gamma \vdash m : A$ and $m \step n$, then there is $\Gamma \vdash n : A$.
+\end{theorem}
+\begin{proof}
+  The proof proceeds by induction on the derivation of $\Gamma \vdash m : A$. The interesting cases are the application cases which we shall discuss in detail.
+  \begin{itemize}
+    \item For the \rname{App-U$\rightarrow$} case,
+  \end{itemize}
+\end{proof}
+
 \end{document}

### 2022-01-03 writing/appendix/draft.tex
writing

@@
 \title{The Calculus of Linear Constructions --- Technical Report}
-\author{Qiancheng Fu}
+\author{Anonymous Submission}
 
@@
 This extended report is meant to accompany our paper of the same title.
-Here, we describe the meta-theory of CLC and CILC and their proofs in detail.
+Here, we describe the meta-theory of CLC and its proofs in detail.
 All the results presented here have been formalized and proven correct in the Coq Proof Assistant.
@@
 
-\section{Subtyping of CLC \texttt{(clc_subtype.v)}}
+\section{Subtyping of CLC \texttt{(clc_subtype.v)}}\label{subtyping}
 
@@
   \begin{itemize}
-    \item For the \rname{App-U$\rightarrow$} case,
+    \item For the \rname{App-U$\rightarrow$} case, from assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$ and Lemmas \ref{mergepure2}, \ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+Applying Lemma \ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+Now by the induction hypothesis, we can conclude there exists sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
+Applying Lemma \ref{uarrowinv} to assumption $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : U_{i'}$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$ where $s'$ is a sort and $i', j'$ are natural numbers.
+The goal can finally be proven by applying the substitution Lemma \ref{substitutionu} on assumptions $\Gamma_2 \vdash n : A$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$.
+    \item For the \rname{App-L$\rightarrow$} case, from assumption $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and Lemmas \ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+Applying Lemma \ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+Now by the induction hypothesis, we can conclude that there exists sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \ltype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
+Applying Lemma \ref{larrowinv} to assumption $\overline{\Gamma_1} \vdash (x \ltype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : L_{i'}$ and $\overline{\Gamma_1} \vdash B : s'_{j'}$.
+Due to the fact that variable $x$ is not a free variable in $B$, the substitution occurring in goal $\exists s \in sort, i \in \mathbb{N}, \overline{\Gamma} \vdash B[n/x] : s_i$ is trivial, thus the judgment $\overline{\Gamma_1} \vdash B : s'_{j'}$ that we have proven shows the existence of the goal.
+    \item For the \rname{App-U$\multimap$} case, the proof is similar to the \rname{App-U$\rightarrow$} case, the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
+    \item For the \rname{App-L$\multimap$} case, the proof is similar to the \rname{App-L$\rightarrow$} case, the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
   \end{itemize}
@@
 
+\section{Linearity Theorems of CLC \texttt{(clc_linearity.v)}}
+
+\subsection{Linearity}
+
+We introduce a meta-function $occurs$ that counts the number of times a given variable occurs in a term.
+
+\begin{align*}
+  occurs\ x\ y                               & =
+  \begin{dcases}
+    1 & x =_\alpha y    \\
+    0 & x \neq_\alpha y
+  \end{dcases}                                                \\
+  occurs\ x\ s_i                             & = 0                           \\
+  occurs\ x\ ((y \stype{s} A) \rightarrow B) & = occurs\ x\ A + occurs\ x\ B \\
+  occurs\ x\ ((y \stype{s} A) \multimap B)   & = occurs\ x\ A + occurs\ x\ B \\
+  occurs\ x\ (\lambda x \stype{s} A.n)       & = occurs\ x\ A + occurs\ x\ n \\
+  occurs\ x\ (m\ n)                          & = occurs\ x\ m + occurs\ x\ n
+\end{align*}
+
+\begin{lemma}\label{islmergeinv}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable with linear type $x \in \Gamma$ there is $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable $x \notin \Gamma$ there is $x \notin \Gamma_1$ and $x \notin \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{narity}
+  For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable $x \notin \Gamma$ there is $occurs\ x\ m = 0$
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash m : A$.
+\end{proof}
+
+\begin{theorem}\label{linearity}
+  Linearity. For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable with linear type $x \in \Gamma$ there is $occurs\ x\ m = 1$.
+\end{theorem}
+\begin{proof}
+  The proof proceeds by induction on the derivation of $\Gamma \vdash m : A$, we will discuss the application cases in detail.
+  \begin{itemize}
+    \item For case \rname{App-U$\rightarrow$}, by assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $(x \ltype A) \in \Gamma$ and Lemma \ref{islmergeinv} we can conclude that $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
+In both cases, applying the induction hypothesis and Lemma \ref{narity} proves the goal.
+    \item For case \rname{App-L$\rightarrow$}, the proof is the same as \rname{App-U$\rightarrow$}.
+    \item For case \rname{App-U$\multimap$}, the proof is the same as \rname{App-U$\rightarrow$}.
+    \item For case \rname{App-L$\multimap$}, the proof is the same as \rname{App-U$\rightarrow$}.
+  \end{itemize}
+\end{proof}
+
+\subsection{Promotion}
+\begin{theorem}\label{promotion}
+  Promotion.
+For context $\Gamma$, terms $m, A, B$ and sort $s$, if there is $\pure{\Gamma}$ and $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \multimap B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$.
+\end{theorem}
+\begin{proof}
+  Set $n = \lambda x \stype{s} A. (m\ x)$. The proof proceeds by case analysis on the sort $s$.
+  \begin{itemize}
+    \item If $s = U$, then we may apply Theorem \ref{validity} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
+Now applying Lemma \ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+Now by \rname{U$\rightarrow$} and Lemma \ref{purere} the goal is proven.
+    \item If $s = L$, then we may apply Theorem \ref{validity} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
+Now applying Lemma \ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+Now by \rname{L$\rightarrow$} and Lemma \ref{purere} the goal is proven.
+  \end{itemize}
+\end{proof}
+
+\subsection{Dereliction}
+
+\begin{theorem}
+  Dereliction.
+For context $\Gamma$, terms $m, A, B$ and sort $s$, if there is $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$.
+\end{theorem}
+\begin{proof}
+  Set $n = \lambda x \stype{s} A . (m\ x)$. The proof proceeds by case analysis on the sort $s$.
+  \begin{itemize}
+    \item If $s = U$, then we may apply Theorem \ref{validity} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
+Now applying Lemma \ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+By \rname{U$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \utype A) \multimap B : L_{i'}$.
+By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
+    \item If $s = L$, then we may apply Theorem \ref{validity} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
+Now applying Lemma \ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+By \rname{L$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \ltype A) \multimap B : L_{i'}$.
+By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
+  \end{itemize}
+\end{proof}
+
+\section{Logical Consistency of CLC \texttt{clc_consistent.v}}
+
+\subsection{Strong Normalization}
+
+The proof of the logical consistency of CLC proceeds by construction of a reduction preserving erasure from CLC to CC$\omega$. As CC$\omega$ is consistent, CLC must be consistent as well.
+
+The erasure procedure is recursively defined as follows.
+\begin{align*}
+  \erase{x}                             & = x                                     \\
+  \erase{U_i}                           & = Type_i                                \\
+  \erase{L_i}                           & = Type_i                                \\
+  \erase{(x \stype{s} A) \rightarrow B} & = (x : \erase{A}) \rightarrow \erase{B} \\
+  \erase{(x \stype{s} A) \multimap B}   & = (x : \erase{A}) \rightarrow \erase{B} \\
+  \erase{\lambda x\stype{s}A.n}         & = \lambda x : \erase{A}.\erase{n}       \\
+  \erase{m\ n}                          & = \erase{m}\ \erase{n}
+\end{align*}
+
+With slight overloading of notation, we define erasure for CLC contexts recursively.
+
+\begin{align*}
+  \erase{\epsilon}              & = \epsilon                      \\
+  \erase{\Gamma, x \stype{s} A} & = \erase{\Gamma}, x : \erase{A}
+\end{align*}
+
+\begin{lemma}
+  For CLC term $m$, map $\sigma$ from variables to CLC terms, map $\tau$ from variables to CC$\omega$ terms, if for all variables $x$ there is $\erase{\sigma\ x} = \tau\ x$, then $\erase{m[\sigma]} = \erase{m}[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of term $m$.
+\end{proof}
+
+For the following lemmas, we will index relations and judgments with subscript CLC or CC$\omega$ to emphasize the language it is defined over.
+
+\begin{lemma}\label{erasestep}
+  For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step_{\scriptscriptstyle \text{CLC}} n$.
+\end{proof}
+
+\begin{lemma}
+  For any CLC terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \equiv_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{CLC}} n$ and Lemma \ref{erasestep}.
+\end{proof}
+
+\begin{lemma}\label{erasesub1}
+  For any CLC terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \prec_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \prec_{\scriptscriptstyle \text{CLC}} n$.
+\end{proof}
+
+\begin{lemma}
+  For any CLC terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \preceq_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{CLC}} n$ and the properties of subtyping proven in Section \ref{subtyping}.
+\end{proof}
+
+\begin{theorem}\label{embedding}
+  Embedding.
+For any CLC context $\Gamma$ and CLC terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}} m : A$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{CC$\omega$}} \erase{m} : \erase{A}$.
+\end{theorem}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash_{\scriptscriptstyle \text{CLC}} m : A$.
+\end{proof}
+
+\begin{corollary}\label{embeddingcontext}
+  For any CLC context $\Gamma$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}}$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{CC$\omega$}}$.
+\end{corollary}
+\begin{proof}
+  Direct consequence of applying Theorem \ref{embedding} to all types in context $\Gamma$.
+\end{proof}
+
+\begin{theorem}
+  Strong normalization of CLC.
+\end{theorem}
+\begin{proof}
+  Suppose there exists a well-typed CLC term $m$ with an infinite sequence of reductions.
+Theorem \ref{embedding} shows that there must exist some term $\erase{m}$ that is well-typed in CC$\omega$.
+Additionally, this infinite sequence of reductions on $m$ can be translated in CC$\omega$ step-wise by Lemma \ref{erasestep}.
+This shows that we have constructed a non-normalizing CC$\omega$ term $\erase{m}$, which a contradiction to the strong normalization property of CC$\omega$, thus CLC must be strongly normalizing as well.
+\end{proof}
+
+\subsection{Embedding of CC$\omega$}
+To show that CLC is compatible with the predicative fragment of CC$\omega$, we construct a lifting procedure that lifts CC$\omega$ terms into CLC in a straightforward way.
+\begin{align*}
+  \lift{x}                     & = x                                        \\
+  \lift{Type_i}                & = U_i                                      \\
+  \lift{(x : A) \rightarrow B} & = (x \utype \lift{A}) \rightarrow \lift{B} \\
+  \lift{\lambda x : A.n}       & = \lambda x\utype\lift{A}.\lift{n}         \\
+  \lift{m\ n}                  & = \lift{m}\ \lift{n}
+\end{align*}
+
+With slight overloading of notation, we define lifting for CC$\omega$ recursively.
+\begin{align*}
+  \lift{\epsilon}      & = \epsilon                         \\
+  \lift{\Gamma, x : A} & = \lift{\Gamma}, x \utype \lift{A}
+\end{align*}
+
+\begin{lemma}
+  For CC$\omega$ context $\Gamma$, there is $\pure{\lift{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}
+  For CC$\omega$ term $m$, map $\sigma$ from variables to CC$\omega$ terms, map $\tau$ from variable to CLC terms, if for all variables $x$ there is $\lift{\sigma\ x} = \tau\ x$, then $\lift{m[\sigma]} = \lift{m}[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of term $m$.
+\end{proof}
+For the following lemmas, we will index relations and judgments with subscript CLC of CC$\omega$ to emphasize the language it is defined over.
+
+\begin{lemma}\label{liftstep}
+  For any CC$\omega$ terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step_{\scriptscriptstyle \text{CC$\omega$}} n$.
+\end{proof}
+
+\begin{lemma}
+  For any CC$\omega$ terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \equiv_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{CC$\omega$}} n$ and Lemma \ref{liftstep}.
+\end{proof}
+
+\begin{lemma}\label{liftsub1}
+  For any CC$\omega$ terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \prec_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \prec_{\scriptscriptstyle \text{CC$\omega$}} n$.
+\end{proof}
+
+\begin{lemma}
+  For any CC$\omega$ terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \preceq_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{CC$\omega$}} n$ and the properties of subtyping proven in Section \ref{subtyping}.
+\end{proof}
+
+\begin{theorem}\label{lifting}
+  Lifting.
+For any CC$\omega$ context $\Gamma$ and CC$\omega$ terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CC$\omega$}} m : A$, then there is $\lift{\Gamma} \vdash_{\scriptscriptstyle \text{CLC}} \lift{m} : \lift{A}$.
+\end{theorem}
+\begin{proof}
+  By induction on the derivation of $\Gamma \vdash_{\scriptscriptstyle \text{CC$\omega$}} m : A$.
+\end{proof}
+
 \end{document}

### 2022-01-03 writing/appendix/draft.tex
writing

@@
 \newcommand{\pstep}{\leadsto_p}
-\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
+\newcommand{\mrg}[3]{merge\ {#1}\ {#2}\ {#3}}
 \newcommand{\erase}[1]{\llbracket #1 \rrbracket}
@@
 \title{The Calculus of Linear Constructions --- Technical Report}
-\author{Anonymous Submission}
+\author{Anonymous Author(s)}
 
@@
     { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \mrg{\Gamma_1, x \utype A}
-      {\Gamma_2, x \utype A}
-      {\Gamma, x \utype A} }
+    { \mrg{(\Gamma_1, x \utype A)}
+      {(\Gamma_2, x \utype A)}
+      {(\Gamma, x \utype A)} }
     \rname{Merge-U}
@@
       x \notin \Gamma_2 }
-    { \mrg{\Gamma_1, x \ltype A}
+    { \mrg{(\Gamma_1, x \ltype A)}
       {\Gamma_2}
-      {\Gamma, x \ltype A} }
+      {(\Gamma, x \ltype A)} }
     \rname{Merge-L1}
@@
     { \mrg{\Gamma_1}
-      {\Gamma_2, x \ltype A}
-      {\Gamma, x \ltype A} }
+      {(\Gamma_2, x \ltype A)}
+      {(\Gamma, x \ltype A)} }
     \rname{Merge-L2}
@@
 \begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and the properties of $\mrg{\_}{\_}{\_}$ discussed in Section \ref{mergeprop}.
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and the properties of $merge$ discussed in Section \ref{mergeprop}.
 \end{proof}

### 2022-01-03 writing/draft6/draft6.tex
writing

--- 2021-12-03 writing/draft5/draft5.tex
@@
-\documentclass[sigplan,screen,review,authordraft]{acmart}
+\documentclass[sigplan,screen,review,anonymous]{acmart}
 
@@
 \newcommand{\pstep}{\leadsto}
-\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
+\newcommand{\mrg}[3]{merge\ {#1}\ {#2}\ {#3}}
 \newcommand{\erase}[1]{\llbracket #1 \rrbracket}
@@
 The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
-In CC types can depend on terms, allowing one to write precise specifications as types.
+In CC, types can depend on terms, allowing one to write precise specifications as types.
 Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
@@
 However, due to its origins as a logical framework for constructive mathematics, it is quite difficult for CC to encode and reason about resources.
-Intuitively, a mathematical theorem can be applied an unrestricted number of times.
-Comparatively, the usage of resources is more limited.
-For example, if we encode Girard's classical example \cite{girard95} of purchasing cigarettes literally into CC as a function of type:
-\begin{equation*}
-  Money \rightarrow Camels + Marlboro
-\end{equation*}
-If viewed as a propositional implication, the customer will still maintain full ownership of their money after paying the vendor, because implication does not diminish the validity of its antecedent.
-Unless the vendor is exceedingly generous, we are faced with the crime of counterfeiting.
 Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
@@
 This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
-
-The presence of both linear and dependent types enables CLC to write specifications that faithfully encode the usage of resources.
-The previous example of monetary transaction can be refined using an indexed linear type family $Money : \mathbb{N} \rightarrow L$ as follows.
-\begin{equation*}
-  (x \ltype Money\ 5) \rightarrow (Camels \oplus Marlboro)
-\end{equation*}
-This new specification for the transaction demands a payment of 5 units of money, and the customer be relieved of their ownership after the transaction finishes, effectively preventing the contradiction of having your cake and eating it too.
 
@@
 We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
-  \item Additionally, we observe that CLC is backwards compatible with the predicative CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
+  \item Additionally, we observe that CLC is compatible with the predicative CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
 We prove that all valid CC$\omega$ terms are valid CLC terms after adding minimal amounts of annotation.
@@
 Next, we define a $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ relation that merges two mixed contexts $\Gamma_1$, and $\Gamma_2$ into $\Gamma$, by performing contraction on shared non-linear variables.
-For linear variables, the $\mrg{\_}{\_}{\_}$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
-This definition of $\mrg{\_}{\_}{\_}$ is what allows contraction for unrestricted variables whilst forbidding it for restricted ones.
+For linear variables, the $merge$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
+This definition of $merge$ is what allows contraction for unrestricted variables whilst forbidding it for restricted ones.
 
@@
     { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A \utype U_i }
+      \overline{\Gamma} \vdash A : U_i }
     { \Gamma, x \utype A \vdash }
@@
     { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A \utype L_k }
+      \overline{\Gamma} \vdash A : L_i }
     { \Gamma, x \ltype A\ \vdash }
@@
     { \pure{\Gamma} \\
-      \Gamma \vdash A \utype U }
+      \Gamma \vdash A : U_i }
     { \pure{\Gamma, x \utype A} }
@@
 
-\subsubsection{Backwards Compatibility}
-To show that CLC is backwards compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
+\subsubsection{Compatibility}
+To show that CLC is compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
 
@@
     { }
-    { arity(s,s) }
+    { arity(s_i,s) }
     \rname{A-Sort}
@@
 \section{Future Work}
-On the theoretical side, we intend to investigate the semantic models of CLC.
 In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy.
@@
 
-On the application side, we intend to bolster our implementation with more features such as compilation to executable binary, making it a full-featured theorem prover/programming language.
 We aim to fully verify the typechecking algorithm employed by the current implementation.

### 2022-01-08 writing/appendix/draft.tex
draft

@@
 \begin{align*}
-  occurs\ x\ y                               & =
+  occurs(x, y)                               & =
   \begin{dcases}
@@
   \end{dcases}                                                \\
-  occurs\ x\ s_i                             & = 0                           \\
-  occurs\ x\ ((y \stype{s} A) \rightarrow B) & = occurs\ x\ A + occurs\ x\ B \\
-  occurs\ x\ ((y \stype{s} A) \multimap B)   & = occurs\ x\ A + occurs\ x\ B \\
-  occurs\ x\ (\lambda x \stype{s} A.n)       & = occurs\ x\ A + occurs\ x\ n \\
-  occurs\ x\ (m\ n)                          & = occurs\ x\ m + occurs\ x\ n
+  occurs(x, s_i)                             & = 0                           \\
+  occurs(x, ((y \stype{s} A) \rightarrow B)) & = occurs(x, A) + occurs(x, B) \\
+  occurs(x, ((y \stype{s} A) \multimap B))   & = occurs(x, A) + occurs(x, B) \\
+  occurs(x, (\lambda x \stype{s} A.n))       & = occurs(x, A) + occurs(x, n) \\
+  occurs(x, (m\ n))                          & = occurs(x, m) + occurs(x, n)
 \end{align*}
@@
 \begin{lemma}\label{narity}
-  For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable $x \notin \Gamma$ there is $occurs\ x\ m = 0$
+  For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable $x \notin \Gamma$ there is $occurs(x, m) = 0$
 \end{lemma}
@@
 \begin{theorem}\label{linearity}
-  Linearity. For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable with linear type $x \in \Gamma$ there is $occurs\ x\ m = 1$.
+  Linearity. For context $\Gamma$, terms $m, A$, if there is $\Gamma \vdash m : A$, then for any variable with linear type $x \in \Gamma$ there is $occurs(x, m) = 1$.
 \end{theorem}

### 2022-01-08 writing/draft6/draft6.tex
draft

@@
 
-However, due to its origins as a logical framework for constructive mathematics, it is quite difficult for CC to encode and reason about resources.
+However, due to its origins as a logical framework for constructive mathematics, it can be inconvenient for CC to encode and reason about resources.
 Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
@@
 We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Linear connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
+Standard Linear Logic connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
 Through examples, we show that CILC facilitates lightweight reasoning about stateful programs and protocol enforcement.
@@
   \item Next, we prove the linearity theorem, showing the tangible impact linear types have on the structure of terms.
-We also show that promotion and dereliction of linear logic are derivable as theorems for canonical types.
+We also show that promotion and dereliction of linear logic are derivable as theorems for function types.
   \item Additionally, we observe that CLC is compatible with the predicative CC$\omega$, allowing us to construct a reduction preserving model of CLC in CC$\omega$, showing that CLC is consistent.
@@
 
-In \rname{$\lambda$$\rightarrow$}, the function type being addressed has form $(x \stype{s} A) \rightarrow B$.
-$\lambda$-abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
+In \rname{$\lambda$$\rightarrow$}, the function type has form $(x \stype{s} A) \rightarrow B$.
+Abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
 This consideration is realized by the side condition $\pure{\Gamma}$, asserting all variables in context $\Gamma$ are unrestricted.
@@
 
-In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming $\lambda$-abstractions that must be used exactly once.
+In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming abstractions that must be used exactly once.
 Due to the fact these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
@@
 To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
-We give a hand written version of the proof in the appendix as well.
+We give a hand written version of the proof in an accompanying technical report.
 
@@
 Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
-He defines additional syntax and semantics for the intricate unboxing of !
+He defines intricate syntax and semantics for the unboxing of !
 terms, solving the lack of substitute in Abramsky's term calculus.

### 2022-01-09 writing/draft6/draft6.tex
draft

@@
 \newcommand{\sep}{\text{ | }}
-\newcommand{\unit}{\text{unit}}
-\newcommand{\new}{\text{new}}
-\newcommand{\free}{\text{free}}
-\newcommand{\get}{\text{get}}
-\newcommand{\set}{\text{set}}
-\newcommand{\session}{\text{session}}
-\newcommand{\channel}{\text{channel}}
-\newcommand{\open}{\text{open}}
-\newcommand{\close}{\text{close}}
-\newcommand{\send}{\text{send}}
-\newcommand{\recv}{\text{recv}}
+\newcommand{\unit}{\texttt{unit}}
+\newcommand{\new}{\texttt{new}}
+\newcommand{\free}{\texttt{free}}
+\newcommand{\get}{\texttt{get}}
+\newcommand{\set}{\texttt{set}}
+\newcommand{\session}{\texttt{session}}
+\newcommand{\channel}{\texttt{channel}}
+\newcommand{\open}{\texttt{open}}
+\newcommand{\close}{\texttt{close}}
+\newcommand{\send}{\texttt{send}}
+\newcommand{\recv}{\texttt{recv}}
 \newcommand{\SEND}{\texttt{SEND}}
@@
 
-The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are almost standard, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are similar to standard CC$\omega$, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
 In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
-Variables not pertaining to actual proofs could be introduced this way, thus giving rise to the admissibility of weakening.
+Variables not pertaining to actual proofs could be introduced this way, thus giving rise to the weakening rule.
 To support linear types, we must restrict weakening to non-linear variables.
@@
 However, once a linear constructor has been partially applied to a linear term, its linearity is ``activated'' in rules \rname{L-Constr-Pos-2} and \rname{L-Constr-2$\rightarrow$}, essentially forcing the rest of its arguments to be fully applied.
-This is to prevent partially applied constructors from duplicating linear variables.
+This is to prevent partially applied constructors from duplicating its linear arguments.
 
@@
 
-Dependent types also allows for more precise specifications of protocols themselves.
+Dependent types also allow for more precise specifications of protocols themselves.
 A simple example would be the specification of message sizes.
@@
 Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
-Using linear types for continuations is surprisingly accurate as control-flow itself is a linear concept.
+Using linear types for continuations is surprisingly accurate as control-flow itself cannot be freely duplicated nor discarded.
 
@@
 \section{Related Work}
-Linear types are a class of type systems inspired by Girard' substructural Linear Logic \cite{girard}.
+Linear types are a class of type systems inspired by Girard's substructural Linear Logic \cite{girard}.
 Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.

### 2022-01-09 writing/draft6/draft6.tex
draft

@@
 Standard Linear Logic connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
-Through examples, we show that CILC facilitates lightweight reasoning about stateful programs and protocol enforcement.
-
-We have formalized all major results in Coq and implemented a prototype in OCaml, both of which are available in the first author's Github repository.
+Through examples in Section \ref{examples}, we show that CILC facilitates lightweight reasoning about stateful programs, protocol enforcement and in-place update of data structures.
+
+We have formalized all major results in Coq and implemented a prototype in OCaml.
+Additional examples of applications such as safe array access and quantitative reasoning are included with our implementation.
 
@@
 
-\section{Examples}
+\section{Examples}\label{examples}
 We shall write examples in the concrete syntax of our implementation.

### 2022-01-12 writing/draft6/draft6.tex
draft

@@
 
-This paper presents a new linear dependent type system --- The Calculus of Linear Constructions (CLC).
+This paper presents a new linear dependent type system --- the Calculus of Linear Constructions (CLC).
 CLC extends the predicative CC$\omega$ with linear types.
@@
 We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Standard Linear Logic connectives such a $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
+Standard Linear Logic connectives such as $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
 Through examples in Section \ref{examples}, we show that CILC facilitates lightweight reasoning about stateful programs, protocol enforcement and in-place update of data structures.
@@
 \begin{itemize}
-  \item Fist, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types.
+  \item First, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types.
 The integration of linear types and dependent types allows CLC to directly and precisely reason about resource.
@@
   \begin{grammar}
-    <$i$> := 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} universe levels
+    <$i$> ::= 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} universe levels
 
@@
 to only be canonically usable.
-This allows us to derive the substitution lemma, and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
+This allows us to derive the substitution lemma and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
 

### 2022-01-14 writing/draft6/draft6.tex
draft

@@
 Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
-These theorem provers have found great success in the fields of software verification, and constructive mathematics.
+These theorem provers have found great success in the fields of software verification and constructive mathematics.
 
@@
 We use the meta variable $i$ to quantify over universe levels $0, 1, 2, ...$.
-$U_i$, and $L_i$ are the universes at level $i$ of non-linear and linear types respectively.
+$U_i$ and $L_i$ are the universes at level $i$ of non-linear and linear types respectively.
 \begin{figure}[h]
@@
 However, $L_i$ cumulates to $L_{i+1}$.
-These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes, and subsequently losing track of linearity.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes and subsequently losing track of linearity.
 
@@
     \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$, such that $\Gamma \vdash m : A$.
-A unrestricted term may be used an arbitrary number of times.
+An unrestricted term may be used an arbitrary number of times.
     \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$, such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
@@
 The rules for forming types are presented in Figure \ref{type}.
-In CLC, we forbid types from depending on linear terms similar to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
+In CLC, we forbid types from depending on linear terms similarly to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
 
@@
 From the judgment $\Gamma \vdash A : L_i$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
-Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evident in the judgment $\Gamma \vdash B : s_i$.
+Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evidently in the judgment $\Gamma \vdash B : s_i$.
 Like \rname{U$\rightarrow$}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
@@
 
-In \rname{$\lambda$$\rightarrow$}, the function type has form $(x \stype{s} A) \rightarrow B$.
+In \rname{$\lambda$$\rightarrow$}, the function type has the form $(x \stype{s} A) \rightarrow B$.
 Abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
@@
 In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming abstractions that must be used exactly once.
-Due to the fact these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evident in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
+Due to the fact that these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evidently in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
 However, in judgment $\overline{\Gamma} \vdash (x \stype{s} A) \multimap B : L_k$ the context must be filtered, because types are not allowed to depend on restricted variables.
@@
 This justifies the side condition of $\pure{\Gamma_2}$.
-The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
+The contexts $\Gamma_1$ and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$ and $\Gamma_2$.
 
@@
 This justifies the lack of side condition $\pure{\Gamma_2}$.
-The contexts $\Gamma_1$, and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$, and $\Gamma_2$.
+The contexts $\Gamma_1$ and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$ and $\Gamma_2$.
 
@@
 \subsubsection{Substitution} \label{subst}
-Though the substitution lemma is regarded as a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
+Although the substitution lemma is regarded as a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
 Much of the difficulty arises during the substitution of arguments containing !
@@
 \subsubsection{Strong Normalization}
-The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function to CC$\omega$.
-We assume familiarity with CC$\omega$ syntax here, and define the erasure function as follows.
+The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function from CLC to CC$\omega$.
+We assume familiarity with CC$\omega$ syntax here and define the erasure function as follows.
 
@@
 
-With slight overloading of notation, we define erasure for CLC contexts recursively.
+With slight overloading of notation we define erasure for CLC contexts recursively.
 
@@
 
-If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}.
+If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve} and \ref{embedding}.
 This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
@@
 
-As shown in Figure \ref{syntax2}, CILC extends CIC with for the introduction and elimination of inductive types.
+As shown in Figure \ref{syntax2}, CILC extends CIC with syntax for the introduction and elimination of inductive types.
 Ind is used during the formation of new inductive types, and Constr is used for introducing their constructors.
@@
 
-Linear constructors are used for constructing linear objects that must be used once. These constructors may be applied to linear terms as restricted usage prevent duplication and enforce single usage.
+Linear constructors are used for constructing linear objects that must be used once.
+These constructors may be applied to linear terms as restricted usage prevents duplication and enforces single usage.
 
@@
 A syntactic guard condition is utilized to conservatively ensure termination.
-The fix-point term must have non-linear type in a pure context, otherwise recursive calls may cause duplicated usage of restricted variables.
+The fix-point term must have non-linear type in a pure context, because recursive calls may cause duplicated usage of restricted variables.
 
@@
 \subsection{Extended Reduction}
-The reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
+As the reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
 
@@
 We shall write examples in the concrete syntax of our implementation.
-We deliberately chose to model our concrete syntax after Coq such that it would be easier for readers already familiar with Coq to adapt to our syntax.
+We deliberately choose to model our concrete syntax after Coq such that it would be easier for readers already familiar with Coq to adapt to our syntax.
 
@@
     \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (m : A) \rightarrow (l \mapsto_A m) \rightarrow F \_: (\Sigma x. x =_A m). (l \mapsto_A m)$ \\
-          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ that there is indeed a term $m$ currently stored there.
+          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto_A m)$ there is indeed a term $m$ currently stored there.
 The type of $\get$'s output is the rather complicated $F \_: (\Sigma x.
@@
 (l \mapsto_A m)$.
-This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claimed to have stored.
+This not only returns the term stored at $l$, but also a proof that the returned term is exactly the same as what $(l \mapsto_A m)$ claims to have stored.
 The proof $(l \mapsto_A m)$ is returned unchanged as well.
@@
 Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
-A term calculus for linear type theory was later realized by Abramsky \cite{abramsky1993}.
+A term calculus for linear type theory is realized by Abramsky \cite{abramsky1993}.
 Benton \cite{benton1994} investigates the ramifications of the !
@@
 \cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance, linear and affine types within a unified framework.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.
 The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
@@
 We aim to fully verify the typechecking algorithm employed by the current implementation.
-The unification algorithm used for implicit argument resolution is extremely ad-hoc, as the problem of unification for linear types has been not thoroughly studied, we hope to improve this situation as well.
+The unification algorithm used for implicit argument resolution is extremely ad-hoc.
+As the problem of unification for linear types has not been thoroughly studied, we hope to improve this situation as well.
 

### 2022-01-17 writing/draft7/draft7.tex
draft

--- 2022-01-08 writing/appendix/draft.tex
@@
-\documentclass{article}
+\documentclass[sigplan,screen,review,anonymous]{acmart}
 
 \usepackage[utf8]{inputenc}
-\usepackage[margin=1in]{geometry}
 
@@
 
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{corollary}{Corollary}[theorem]
-\newtheorem{lemma}[theorem]{Lemma}
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-
 \setlist[itemize]{leftmargin=1.6em}
@@
 \renewcommand{\syntright}{}
-\setlength{\grammarparsep}{15pt}
-\setlength{\grammarindent}{10em}
+\setlength{\grammarparsep}{7pt}
+\setlength{\grammarindent}{7em}
 \newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
@@
 \newcommand{\ltype}{:_{\scriptscriptstyle L}}
-\newcommand{\stype}[1]{:_{#1}}
+\newcommand{\stype}[1]{:_#1}
 \newcommand{\step}{\leadsto}
@@
 \newcommand{\lrangle}[1]{\langle #1 \rangle}
+\newcommand{\SigmaL}{\Sigma_{\scriptscriptstyle L}}
 \newcommand{\ucons}{constructor_{\scriptscriptstyle U}}
@@
 \newcommand{\inr}{\text{inr}}
-
-\title{The Calculus of Linear Constructions --- Technical Report}
-\author{Anonymous Author(s)}
+\newcommand{\arrv}{\texttt{ArrVec}}
+
+\title{The Calculus of Linear Constructions}
+\author{Qiancheng Fu}
+\affiliation{
+  \institution{Boston University}
+  \city{Boston}
+  \state{MA}
+  \country{USA}
+}
+\email{qcfu@bu.edu}
+
+\keywords{Dependent types, linear types, inductive types, Calculus of Constructions}
 
 \begin{document}
+\begin{abstract}
+  The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
+Specifically, CLC extends the predicative CC$\omega$ with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
+We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
+We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
+Through examples, we demonstrate that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
+We have formalized and proven correct all major results in the Coq Proof Assistant.
+\end{abstract}
 \maketitle
-\tableofcontents
 
 \section{Introduction}
-This extended report is meant to accompany our paper of the same title.
-Here, we describe the meta-theory of CLC and its proofs in detail.
-All the results presented here have been formalized and proven correct in the Coq Proof Assistant.
-
-\section{Syntax of CLC \texttt{(clc_ast.v)}}
-\begin{figure}[H]
+The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
+In CC, types can depend on terms, allowing one to write precise specifications as types.
+Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
+These theorem provers have found great success in the fields of software verification and constructive mathematics.
+
+However, due to its origins as a logical framework for constructive mathematics, it can be inconvenient for CC to encode and reason about resource.
+Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
+We propose an alternative solution: extend CC with linear types.
+
+This paper presents a new linear dependent type system --- the Calculus of Linear Constructions (CLC).
+CLC extends the predicative CC$\omega$ with linear types.
+CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
+We add an extra universe sort $L$ of linear types with cumulativity parallel to the sort $U$ of non-linear types.
+This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
+
+Preexisting approaches to integrating linear types and dependent types often relied on the !
+exponential or separate typing judgments for linear and non-linear types.
+The universe sorts of CLC is enough to completely distinguish between linear and non-linear types within a single typing judgment.
+As a result, CLC falls much closer to its CC roots than prior works in both syntax and semantics.
+
+We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
+Standard Linear Logic connectives such as $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
+Connectives that mix linear and non-linear types can be constructed just as easily using CILC.
+
+Through examples in Section \ref{examples}, we show some applications of CILC such as construction of safe random access array from the first principles, protocol enforcement through inductively defined session types and in-place update of data structures.
+
+We have formalized all major results in Coq and implemented a prototype in OCaml.
+Additional examples of applications such as provable effect laws and quantitative reasoning are included with our implementation.
+
+\vspace{4pt}
+\noindent \textbf{\textit{Contributions}}:
+Our contributions can be summarized as follows.
+\begin{itemize}
+  \item First, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types, enabling direct and precise reasoning about resource.
+  \item Second, we study the meta-theory of CLC, showing that it exhibits many desirable properties such as confluence, subject reduction and logical consistency.
+  \item Furthermore, we extend CLC with rules to define inductive linear types, forming the Calculus of Inductive Linear Constructions (CILC).
+This is the first complete presentation of a generalized mechanism for defining inductive linear types.
+  \item All major results have been proven correct in the Coq.
+Our development is the first machine checked formalization of a linear dependent type theory.
+We have also implemented a prototype of CILC in OCaml.
+\end{itemize}
+
+\section{The Language of CLC}
+\subsection{Syntax}
+\begin{figure}[h]
+  \caption{Syntax of CLC}
   \begin{grammar}
-    <$i$> ::= 0 | 1 | 2 ... \phantom{* |} \hspace*{2.4em} universe levels
+    <$i$> ::= 0 | 1 | 2 ... \phantom{* |} \hspace*{3em} universe levels
 
@@
 
-    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4em} expressions
+    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4.2em} expressions
     \indalt $(x :_s A) \rightarrow B$
@@
   \end{grammar}
-\end{figure}
-
-\section{Reduction and Equality of CLC \texttt{(clc_ast.v)}}
-\begin{figure}[H]
-  \begin{mathpar}
-    \inferrule
-    { m_1 \red n \\ m_2 \red n }
-    { m_1 \equiv m_2 : A }
-    \rname{Join}
-
-    \inferrule
-    {  }
-    { (\lambda x \stype{s}A.m)\ n \step m[n/x] }
-    \rname{Step-$\beta$}
-
-    \inferrule
-    { A \step A' }
-    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A' .m }
-    \rname{Step-$\lambda$L}
-
-    \inferrule
-    { m \step m' }
-    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A.m' }
-    \rname{Step-$\lambda$R}
-
-    \inferrule
-    { A \pstep A' }
-    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A') \rightarrow B }
-    \rname{Step-L$\rightarrow$}
-
-    \inferrule
-    { B \pstep B' }
-    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A) \rightarrow B' }
-    \rname{Step-R$\rightarrow$}
-
-    \inferrule
-    { A \pstep A' }
-    { (x \stype{s} A) \multimap B \step (x \stype{s} A') \multimap B }
-    \rname{Step-L$\multimap$}
-
-    \inferrule
-    { B \pstep B' }
-    { (x \stype{s} A) \multimap B \step (x \stype{s} A) \multimap B' }
-    \rname{Step-R$\multimap$}
-
-    \inferrule
-    { m \step m' }
-    { m\ n \step m'\ n }
-    \rname{Step-AppL}
-
-    \inferrule
-    { n \step n' }
-    { m\ n \step m\ n' }
-    \rname{Step-AppR}
-  \end{mathpar}
-  \label{red}
-\end{figure}
-
-\section{Confluence of CLC \texttt{(clc_confluence.v)}}
-
-\subsection{Parallel Reduction}
-To prove the confluence property of CLC, we employ the standard technique utilizing parallel reductions.
-
-\begin{figure}[H]
-  \begin{mathpar}
-    \inferrule
-    { }
-    { x \pstep x }
-    \rname{PStep-Var}
-
-    \inferrule
-    { }
-    { s_i \pstep s_i }
-    \rname{PStep-Sort}
-
-    \inferrule
-    { A \pstep A' \\ m \pstep m' }
-    { \lambda x \stype{s} A.m \pstep \lambda x \stype{s} A'.m' }
-    \rname{PStep-$\lambda$}
-
-    \inferrule
-    { m \pstep m' \\ n \pstep n' }
-    { m\ n \pstep m'\ n' }
-    \rname{PStep-App}
-
-    \inferrule
-    { m \pstep m' \\ n \pstep n'}
-    { (\lambda x \stype{s} A.m)\ n \pstep m'[n'/x] }
-    \rname{PStep-$\beta$}
-
-    \inferrule
-    { A \pstep A' \\ B \pstep B' }
-    { (x \stype{s} A) \rightarrow B \pstep (x \stype{s} A') \rightarrow B' }
-    \rname{PStep$\rightarrow$}
-
-    \inferrule
-    { A \pstep A' \\ B \pstep B' }
-    { (x \stype{s} A) \multimap B \pstep (x \stype{s} A') \multimap B' }
-    \rname{PStep$\multimap$}
-  \end{mathpar}
-  \label{pred}
-\end{figure}
-
-\subsection{Reduction Lemmas}
-Here, we prove some simple lemmas concerning $\step$, $\red$ and substitution.
+  \Description{}
+  \label{syntax1}
+\end{figure}
+
+The syntax of the core type theory is presented in Figure \ref{syntax1}.
+Our type theory contains two sorts of universes $U$ and $L$.
+We use the meta variable $i$ to quantify over universe levels $0, 1, 2, ...$.
+$U_i$ and $L_i$ are the universes at level $i$ of non-linear and linear types respectively.
+\begin{figure}[h]
+  \caption{Correspondence of CLC types and MELL implications}
+  \Description{}
+  \begin{align}
+    (\_ \utype A) \rightarrow B \quad & \Leftrightarrow \quad !(!A \multimap B) \\
+    (\_ \ltype A) \rightarrow B \quad & \Leftrightarrow \quad !(A \multimap B)  \\
+    (\_ \utype A) \multimap B \quad   & \Leftrightarrow \quad !A \multimap B    \\
+    (\_ \ltype A) \multimap B \quad   & \Leftrightarrow \quad A \multimap B
+  \end{align}
+  \label{correspondence}
+\end{figure}
+
+A clear departure of our language from standard presentations of both linear type theory and dependent type theory is the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
+The reason for these function types is that we have built the !
+exponential of linear logic implicitly into universe sorts.
+The sort annotation $s$ here records the universe sort of the function domain.
+The behavior of !
+is difficult to account for even in simple linear type theories.
+Subtle issues arise if !!
+is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
+By integrating the exponential directly into universe sorts, we have limited !
+to only be canonically usable.
+This allows us to derive the substitution lemma and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
+
+Figure \ref{correspondence} illustrates the correspondence between CLC functions and Multiplicative Exponential Linear Logic (MELL) implications.
+MELL lacks counterparts for the cases (1), (3) if the co-domain $B$ is dependent on arguments of domain $A$.
+We will discuss function type formation in Section \ref{tyformation} in greater detail.
+
+Algorithmic type checking techniques allow users to omit writing sort indices for many situations in practice.
+Our implementation employs bi-directional type checking and users rarely interact with sort indices in the surface syntax.
+
+\subsection{Universes and Cumulativity}
+CLC features two sorts of universes $U$ and $L$ with level indices $0, 1, 2, \cdots$.
+$U_i$ and $L_i$ are the predicative universes of non-linear types and linear types respectively.
+The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
+Basically, terms with types that occur within $U_i$ are unrestricted in their usage.
+Terms with types that occur within $L_i$ are restricted to being used exactly once.
+
+In order to lift terms from lower universes to higher ones, there exists cumulativity between universe levels of the same sort. We define cumulativity as follows.
 
 \begin{definition}
-  For a term $m$ and a map $\sigma$ from variables to terms, let $m[\sigma]$ be the term obtained by applying $\sigma$ uniformly to all free variables in $m$.
-\end{definition}
-
-\begin{definition}
-  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ reduces to $\tau$ if for any variable $x$ there exists a reduction $(\sigma\ x) \red (\tau\ x)$.
-We write $\sigma \red \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
-\end{definition}
-
-\begin{lemma}\label{stepsubst}
-  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a step $m \step n$, then there exists a step $m[\sigma] \step n[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \step n$.
-\end{proof}
-
-\begin{lemma}\label{redapp}
-  For terms $m_1, m_2, n_1, n_2$, if these exists reductions $m_1 \red m_2$ and $n_1 \red n_2$, then there exists reduction $(m_1\ n_1) \red (m_2\ n_2)$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\red$ and applying rules \rname{Step-AppL}, \rname{Step-AppR}.
-\end{proof}
-
-\begin{lemma}\label{redlam}
-  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $m_1 \red m_2$, then there exists reduction $\lambda x \stype{s}A_1.m_1 \red \lambda x \stype{s}A_2.m_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\red$ and applying rules \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
-\end{proof}
-
-\begin{lemma}\label{redarrow}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \rightarrow B_1 \red (x \stype{s} A_2) \rightarrow B_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\red$ and applying rules \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
-\end{proof}
-
-\begin{lemma}\label{redlolli}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \multimap B_1 \red (x \stype{s} A_2) \multimap B_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\red$ and applying rules \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
-\end{proof}
-
-\begin{lemma}\label{redsubst}
-  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a reduction $m \red n$, then there exists a reduction $m[\sigma] \red n[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\red$, the transitivity of $\red$ and Lemma \ref{stepsubst}.
-\end{proof}
-
-\begin{lemma}\label{redcompat}
-  For maps $\sigma, \tau$ from variables to terms, if there is a map reduction $\sigma \red \tau$, then for any term $m$ these is a reduction $m[\sigma] \red m[\tau]$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $m$, applying Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}.
-\end{proof}
-
-\subsection{Equality Lemmas}
-Here, we prove some simple lemmas concerning $\red$, $\equiv$ and substitution.
-
-\begin{definition}
-  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ is equal to $\tau$ if for any variable $x$ there exists an equality $(\sigma\ x) \equiv (\tau\ x)$.
-We write $\sigma \equiv \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
-\end{definition}
-
-\begin{lemma}\label{convhom}
-  For any map $f$ from terms to terms, if for any terms $m, n$ such that $m \step n$ implies $f\ m \equiv f\ n$, then for any terms $m, n$ equality $m \equiv n$ implies $f\ m \equiv f\ n$.
-\end{lemma}
-\begin{proof}
-  By the properties of the transitive reflexive closure $\red$ and that $\equiv$ is an equivalence relation.
-\end{proof}
-
-\begin{lemma}\label{convapp}
-  For terms $m_1, m_2, n_1, n_2$, if there exists equalities $m_1 \equiv m_2$ and $n_1 \equiv n_2$, then there exists equality $(m_1\ n_1) \equiv (m_2\ n_2)$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-AppL}, \rname{Step-AppR}.
-\end{proof}
-
-\begin{lemma}\label{convlam}
-  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $m_1 \equiv m_2$, then there exists equality $\lambda x \stype{s} A_1.m_1 \equiv \lambda x \stype{s} A_2.m_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
-\end{proof}
-
-\begin{lemma}\label{convarrow}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{s} A_2) \rightarrow B_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
-\end{proof}
-
-\begin{lemma}\label{convlolli}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{s} A_2) \multimap B_2$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
-\end{proof}
-
-\begin{lemma}\label{convsubst}
-  For terms $m, n$ and map $\sigma$ from variables to terms, if there is equality $m \equiv n$, then there is equality $m[\sigma] \equiv n[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By Lemmas \ref{convhom} and \ref{stepsubst}.
-\end{proof}
-
-\begin{lemma}\label{convcompat}
-  For maps $\sigma, \tau$ from variables to terms and term $m$, if these is map equality $\sigma \equiv \tau$, then there is equality $m[\sigma] \equiv m[\tau]$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $m$, applying Lemmas \ref{convapp}, \ref{convlam}, \ref{convarrow}, \ref{convlolli}.
-\end{proof}
-
-\begin{lemma}
-  For terms $m_1, m_2, n$, if there is equality $m_1 \equiv m_2$, then there is equality $n[m_1/x] \equiv n[m_2/x]$ for any variable $x \in FV(n)$.
-\end{lemma}
-\begin{proof}
-  This is a special case of Lemma \ref{convcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
-\end{proof}
-
-\subsection{Parallel Reduction Lemmas}
-
-\begin{definition}\label{psstep}
-  For maps $\sigma, \tau$ from variables to terms, we say $\sigma$ parallel reduces to $\tau$ if for any variable $x$ there exists a parallel reduction $(\sigma\ x) \pstep (\tau\ x)$.
-We write $\sigma \pstep \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
-\end{definition}
-
-\begin{lemma}\label{psteprefl}
-  For any term $m$, there exists a reflexive parallel reduction $m \pstep m$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $m$.
-\end{proof}
-
-\begin{lemma}
-  For any map $\sigma$ from variables to terms, there exists a reflexive parallel map reduction $\sigma \pstep \sigma$.
-\end{lemma}
-\begin{proof}
-  By Definition \ref{psstep} and Lemma \ref{psteprefl}.
-\end{proof}
-
-\begin{lemma}\label{steppstep}
-  For any terms $m, n$, if there exists step $m \step n$, then there exists a parallel reduction $m \pstep n$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \step n$ and Lemma \ref{psteprefl}.
-\end{proof}
-
-\begin{lemma}\label{pstepred}
-  For terms $m, n$, if there exists parallel reduction $m \pstep n$, then there exists a reduction $m \red n$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \pstep n$, utilizing the transitive property of $\red$ and Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}, \ref{redsubst}, \ref{redcompat}.
-\end{proof}
-
-\begin{lemma}
-  For terms $m, n$ and map $\sigma$ from variables to terms, if there exists parallel reduction $m \pstep n$, there exists parallel reduction $m[\sigma] \pstep n[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \pstep n$ and Lemma \ref{psteprefl}.
-\end{proof}
-
-\begin{lemma}\label{pstepcompat}
-  For terms $m, n$ and maps $\sigma, \tau$ from variables to terms, if there exists parallel reduction $m \pstep n$ and parallel map reduction $\sigma \pstep \tau$, there exists parallel reduction $m[\sigma] \pstep n[\tau]$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \pstep n$.
-\end{proof}
-
-\begin{lemma}
-  For terms $m_1, m_2, n$, if there is parallel reduction $m_1 \pstep m_2$, then there is parallel reduction $n[m_1/x] \pstep n[m_2/x]$ for any variable $x \in FV(n)$.
-\end{lemma}
-\begin{proof}
-  By Lemma \ref{psteprefl}, this is a special case of Lemma \ref{pstepcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
-\end{proof}
-
-\subsection{Confluence Theorem}
-We first show that $\pstep$ satisfies the diamond property. Using the diamond property, we ultimately prove the confluence theorem.
-
-\begin{lemma}\label{diamond}
-  CLC term reduction has the diamond property.
-For terms $m, m_1, m_2$, if there are parallel reductions $m \pstep m_1$ and $m \pstep m_2$, then there exists term $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \pstep m_1$.
-Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and apply the induction hypothesis.
-The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma \ref{pstepcompat} is used to push these cases through.
-\end{proof}
-
-\begin{lemma}\label{strip}
-  Strip lemma. For terms $m, m_1, m_2$, if there is parallel reduction $m \pstep m_1$ and reduction $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \pstep m'$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \pstep m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{diamond}.
-\end{proof}
-
-\begin{theorem}
-  CLC term reduction is confluent. For terms $m, m_1, m_2$, if there are reductions $m \red m_1$ and $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \red m'$.
-\end{theorem}
-\begin{proof}
-  By induction on the derivation of $m \red m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{strip}.
-\end{proof}
-
-\subsection{Corollaries of Confluence}
-The following results are all corollaries of confluence, proven using a combination of induction, transitivity and confluence.
-These corollaries allow us to refute false reductions and equalities in future proofs.
-
-\begin{corollary}\label{redsortinv}
-  For a universe $s_i$ and term $m$, if there is reduction $s_i \red m$, then $m = s_i$.
-\end{corollary}
-
-\begin{corollary}\label{redvarinv}
-  For variable $x$ and term $m$, if there is reduction $x \red m$, then $m = x$.
-\end{corollary}
-
-\begin{corollary}\label{redarrowinv}
-  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
-\end{corollary}
-
-\begin{corollary}\label{redlolliinv}
-  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
-\end{corollary}
-
-\begin{corollary}\label{redlaminv}
-  For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exists $A',m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
-\end{corollary}
-
-\begin{corollary}\label{sortinj}
-  For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there is $s = t$ and $i = j$.
-\end{corollary}
-
-\begin{corollary}\label{arrowinj}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{t} A_2) \rightarrow B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
-\end{corollary}
-
-\begin{corollary}\label{lolliinj}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{t} A_2) \multimap B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
-\end{corollary}
-
-\section{Context of CLC \texttt{(clc_context.v)}}
-Contexts of CLC are of the form $x_1 \stype{s_1} A_1, x_2 \stype{s_2} A_2, ...
-x_k \stype{s_k} A_k$ where each free variable $x_i$ is assigned a type $A_i$ and sort $s_i$.
-Contexts will be referred to by meta variables $\Gamma$ and $\Delta$.
-
-\begin{figure}[h]
-  \begin{mathpar}
-    \inferrule
-    { }
-    { \epsilon \vdash }
-    \rname{Wf-$\epsilon$}
-
-    \inferrule
-    { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A : U_i }
-    { \Gamma, x \utype A \vdash }
-    \rname{Wf-U}
-
-    \inferrule
-    { \Gamma\ \vdash \\
-      \overline{\Gamma} \vdash A : L_i }
-    { \Gamma, x \ltype A\ \vdash }
-    \rname{Wf-L}
-    \\
-
-    \inferrule
-    { }
-    { \pure{\epsilon} }
-    \rname{Pure-$\epsilon$}
-
-    \inferrule
-    { \pure{\Gamma} \\
-      \Gamma \vdash A : U_i }
-    { \pure{\Gamma, x \utype A} }
-    \rname{Pure-U}
-    \\
-
-    \inferrule
-    { }
-    { \mrg{\epsilon}{\epsilon}{\epsilon} }
-    \rname{Merge-$\epsilon$}
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
-    { \mrg{(\Gamma_1, x \utype A)}
-      {(\Gamma_2, x \utype A)}
-      {(\Gamma, x \utype A)} }
-    \rname{Merge-U}
-    \\
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-      x \notin \Gamma_2 }
-    { \mrg{(\Gamma_1, x \ltype A)}
-      {\Gamma_2}
-      {(\Gamma, x \ltype A)} }
-    \rname{Merge-L1}
-
-    \inferrule
-    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
-      x \notin \Gamma_1 }
-    { \mrg{\Gamma_1}
-      {(\Gamma_2, x \ltype A)}
-      {(\Gamma, x \ltype A)} }
-    \rname{Merge-L2}
-  \end{mathpar}
-  \label{structural}
-\end{figure}
-
-\subsection{Merge Lemmas}\label{mergeprop}
-Since weakening and contraction rules will not be allowed on restricted variables, it is necessary to have lemmas that enable the manipulation of contexts.
-
-\begin{lemma}\label{mergesym}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\mrg{\Gamma_2}{\Gamma_1}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergepure}
-  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\mrg{\Gamma}{\Gamma}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\pure{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergere1}
-  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\Gamma}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{mergere2}
-  For any context $\Gamma$, there is $\mrg{\Gamma}{\overline{\Gamma}}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{mergepureinv}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma}$, then there is $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergepure1}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_1}$, then there is $\Gamma = \Gamma_2$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergepure2}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$, then there is $\Gamma = \Gamma_1$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergepurepure}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\pure{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergepureeq}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\Gamma_1 = \Gamma_2$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergerere}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}\label{mergererere}
-  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\overline{\Gamma}}{\overline{\Gamma}}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\subsection{Restriction and Purity Lemmas}
-
-\begin{lemma}\label{rere}
-  For any context $\Gamma$, there is $\overline{\Gamma} = \overline{\overline{\Gamma}}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{purere}
-  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\Gamma = \overline{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{repure}
-  For any context $\Gamma$ , there is $\pure{\overline{\Gamma}}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{hasure}
-  For any context $\Gamma$, variable $x$ and type $A$, if there is $x \utype A \in \Gamma$, then there is $x \utype A \in \overline{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $x \utype A \in \Gamma$.
-\end{proof}
-
-\begin{lemma}\label{haslre}
-  For any context $\Gamma$, variable $x$ and type $A$, there is $x \ltype A \notin \overline{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the structure of $\Gamma$.
-\end{proof}
-
-\begin{lemma}\label{mergesplit1}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\begin{lemma}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_2}{\Gamma_2}{\Delta}$ and $\mrg{\Delta_1}{\Delta}{\Gamma}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
-\end{proof}
-
-\section{Subtyping of CLC \texttt{(clc_subtype.v)}}\label{subtyping}
-
   The cumulativity relation ($\preceq$) is the smallest binary relation over terms such that
@@
   \end{enumerate}
-
-\noindent
-Here, we give an inductive definition of the cumulativity relation ($\preceq$) that is suitable for writing proofs.
-
-\begin{mathpar}
-  \inferrule
-  { }
-  { A \prec A }
-  \rname{$\prec$-Refl}
-
-  \inferrule
-  { i_1 \leq i_2 }
-  { s_{i_1} \prec s_{i_2} }
-  \rname{$\prec$-Sort}
-
-  \inferrule
-  { B_1 \prec B_2 }
-  { (x \stype{s} A) \rightarrow B_1 \prec (x \stype{s} A) \rightarrow B_2 }
-  \rname{$\prec$$\rightarrow$}
-
-  \inferrule
-  { B_1 \prec B_2 }
-  { (x \stype{s} A) \multimap B_1 \prec (x \stype{s} A) \multimap B_2 }
-  \rname{$\prec$$\multimap$}
-
-  \inferrule
-  { A' \prec B' \\ A \equiv A' \\ B \equiv B' }
-  { A \preceq B }
-  \rname{$\prec$-$\preceq$}
-\end{mathpar}
-
-\subsection{Subtyping Lemmas}
-
-\begin{lemma}\label{sub1sub}
-  For terms $A, B$, if there is $A \prec B$, then there is $A \preceq B$.
-\end{lemma}
-\begin{proof}
-  By \rname{$\prec$-$\preceq$} and the reflexivity of equality $\equiv$.
-\end{proof}
-
-\begin{lemma}\label{sub1conv}
-  For terms $A, B, C$, if there is $A \prec B$ and $B \equiv C$, then there is $A \preceq C$.
-\end{lemma}
-\begin{proof}
-  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
-\end{proof}
-
-\begin{lemma}\label{convsub1}
-  For terms $A, B, C$, if there is $A \equiv B$ and $B \prec C$, then there is $A \preceq C$.
-\end{lemma}
-\begin{proof}
-  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
-\end{proof}
-
-\begin{lemma}\label{convsub}
-  For terms $A, B$, if there is $A \equiv B$, then there is $A \preceq B$.
-\end{lemma}
-\begin{proof}
-  By Lemma \ref{convsub1} and \rname{$\prec$-Refl}.
-\end{proof}
-
-\begin{lemma}\label{subrefl}
-  For term $A$, there is $A \preceq A$.
-\end{lemma}
-\begin{proof}
-  By Lemma \ref{sub1sub} and \rname{$\prec$-Refl}.
-\end{proof}
-
-\begin{lemma}\label{subprop}
-  For natural numbers $i, j$ and sort $s$ such that $i \leq j$, there is $s_i \preceq s_j$.
-\end{lemma}
-\begin{proof}
-  By Lemma \ref{sub1sub} and \rname{$\prec$-Sort}.
-\end{proof}
-
-\begin{lemma}\label{sub1trans}
-  For terms $A, B, C, D$, if there is $A \prec B$, $B \equiv C$ and $C \prec D$, then there is $A \preceq D$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $A \prec B$, definition of $\prec$ and Lemmas \ref{sub1sub}, \ref{sub1conv}, \ref{convsub1}.
-\end{proof}
-
-\begin{lemma}
-  For terms $A, B, C$, if there is $A \preceq B$ and $B \preceq C$, then there is $A \preceq C$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$, rule \rname{$\prec$-$\preceq$} and Lemma \ref{sub1trans}.
-\end{proof}
-
-\begin{lemma}
-  For sorts $s, t$ and natural numbers $i, j$, if there is $s_i \preceq t_j$, then there is $s = t$ and $i \leq j$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{sortinj}.
-\end{proof}
-
-\begin{lemma}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{t} A_2) \rightarrow B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{arrowinj}.
-\end{proof}
-
-\begin{lemma}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{t} A_2) \multimap B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
-\end{lemma}
-\begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{lolliinj}.
-\end{proof}
-
-\begin{lemma}\label{sub1subst}
-  For terms $A, B$  and map $\sigma$ from variables to terms, if there is $A \prec B$, then there is $A[\sigma] \prec B[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $A \prec B$ and the definition of $\prec$.
-\end{proof}
-
-\begin{lemma}
-  For terms $A, B$ and map $\sigma$ from variables to terms, if there is $A \preceq B$, then there is $A[\sigma] \preceq B[\sigma]$.
-\end{lemma}
-\begin{proof}
-  By rule \rname{$\prec$-$\preceq$} and Lemmas \ref{convsubst}, \ref{sub1subst}.
-\end{proof}
-
-\section{Typing of CLC \texttt{(clc_typing.v)}}
-
-The following rules define well-formed contexts.
+\end{definition}
+
+Figure \ref{universe} illustrates the structure of our universe hierarchy.
+Each linear universe $L_i$ has $U_{i+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
+However, $L_i$ cumulates to $L_{i+1}$.
+These two parallel threads of cumulativity prevent linear types from being transported to the non-linear universes and subsequently losing track of linearity.
+
+\begin{figure}[h]
+  \caption{The Universe Hierarchy}
+  \vspace{1em}
+  \centering
+  \begin{tikzcd}
+    U_0 \arrow[r, ":"', "\preceq", dash]
+    & U_1 \arrow[r, ":"', "\preceq", dash]
+    & U_2 \arrow[r, ":"', "\preceq", dash]
+    & \cdots \\
+    L_0 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
+    & L_1 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
+    & L_2 \arrow[r, "\preceq" description, dash] \arrow[ru, ":" description, dash]
+    & \cdots
+  \end{tikzcd}
+  \label{universe}
+  \Description{}
+\end{figure}
+
+\subsection{Context and Structural Judgments}
+The context of our language employs a mixed linear/non-linear representation in the style of Luo\cite{luo}.
+Variables in the context are annotated to indicate whether they are linear or non-linear.
+A non-linear variable is annotated as $\Gamma, x \utype A$, whereas a linear variable is annotated as $\Gamma, x \ltype A$.
+
+Next, we define a $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ relation that merges two mixed contexts $\Gamma_1$, and $\Gamma_2$ into $\Gamma$, by performing contraction on shared non-linear variables.
+For linear variables, the $merge$ relation is defined if and only if each variable occurs uniquely in one context and not the other.
+This definition of $merge$ is what allows contraction for unrestricted variables whilst forbidding it for restricted ones.
+
+An auxiliary judgment $\pure{\Gamma}$ is defined to assert that a context $\Gamma$ does not contain linear variables.
+In other words, all variables found in $\pure{\Gamma}$ are annotated of the form $x \utype A$.
+The full rules for structural judgments are presented in Figure \ref{structural}.
+
+\begin{figure}[h]
+  \caption{Structural Judgments}
   \begin{mathpar}
@@
     { \epsilon \vdash }
-  \rname{$\epsilon$-Ok}
-
-  \inferrule
-  { \Gamma \vdash \\
+    \rname{Wf-$\epsilon$}
+
+    \inferrule
+    { \Gamma\ \vdash \\
       \overline{\Gamma} \vdash A : U_i }
     { \Gamma, x \utype A \vdash }
-  \rname{U-Ok}
-
-  \inferrule
-  { \Gamma \vdash \\
+    \rname{Wf-U}
+
+    \inferrule
+    { \Gamma\ \vdash \\
       \overline{\Gamma} \vdash A : L_i }
-  { \Gamma, x \ltype A \vdash }
-  \rname{L-Ok}
+    { \Gamma, x \ltype A\ \vdash }
+    \rname{Wf-L}
+
+    \inferrule
+    { }
+    { \pure{\epsilon} }
+    \rname{Pure-$\epsilon$}
+
+    \inferrule
+    { \pure{\Gamma} \\
+      \Gamma \vdash A : U_i }
+    { \pure{\Gamma, x \utype A} }
+    \rname{Pure-U}
+
+    \inferrule
+    { }
+    { \mrg{\epsilon}{\epsilon}{\epsilon} }
+    \rname{Merge-$\epsilon$}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+    { \mrg{(\Gamma_1, x \utype A)}
+      {(\Gamma_2, x \utype A)}
+      {(\Gamma, x \utype A)} }
+    \rname{Merge-U}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_2 }
+    { \mrg{(\Gamma_1, x \ltype A)}
+      {\Gamma_2}
+      {(\Gamma, x \ltype A)} }
+    \rname{Merge-L1}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_1 }
+    { \mrg{\Gamma_1}
+      {(\Gamma_2, x \ltype A)}
+      {(\Gamma, x \ltype A)} }
+    \rname{Merge-L2}
   \end{mathpar}
-
-\noindent
-The typing rules of CLC are presented below.
+  \label{structural}
+  \Description{}
+\end{figure}
+
+\begin{definition}
+  The context restriction function $\overline{\Gamma}$ is defined as a recursive filter over $\Gamma$ as follows.
+All linear variables are removed from context $\Gamma$.
+The result of context restriction is the non-linear subset of the original context.
+  \begin{align*}
+    \overline{\epsilon} = \epsilon
+    \hspace*{4em}
+    \overline{\Gamma, x \utype A} = \overline{\Gamma}, x \utype A
+    \hspace*{4em}
+    \overline{\Gamma, x \ltype A} = \overline{\Gamma}
+  \end{align*}
+\end{definition}
+
+\subsection{Typing Judgment}
+Typing judgments in CLC take on the form of $\Gamma \vdash m : A$. Intuitively, this judgment states that the term $m$ is an inhabitant of type $A$, with free variables typed in $\Gamma$.
+
+\begin{definition} We formally define the terminology \textit{non-linear}, \textit{linear}, \textit{unrestricted} and \textit{restricted}.
+  \begin{enumerate}
+    \item A \textit{type} $A$ is \textit{non-linear} under context $\Gamma$ if $\Gamma \vdash A : U_i$.
+    \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A : L_k$.
+    \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$, such that $\Gamma \vdash m : A$.
+An unrestricted term may be used any number of times.
+    \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$, such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
+  \end{enumerate}
+\end{definition}
+
+\subsection{Type Formation} \label{tyformation}
+The rules for forming types are presented in Figure \ref{type}.
+In CLC, we forbid types from depending on linear terms similarly to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
+
+The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are similar to standard CC$\omega$, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
+Variables not pertaining to actual proofs could be introduced this way, thus giving rise to the weakening rule.
+To support linear types, we must restrict weakening to non-linear variables.
+This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{U-Axiom}, \rname{L-Axiom}.
+From \rname{L-Axiom} we can see that the universe of linear types $L_i$ is an inhabitant of $U_{i+1}$.
+This is inspired by Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
+
+\begin{figure}[h]
+  \caption{Type Formation}
   \begin{mathpar}
@@
     \rname{L$\multimap$}
-
+  \end{mathpar}
+  \label{type}
+  \Description{}
+\end{figure}
+
+The \rname{U$\rightarrow$} rule is used for forming non-linear function types with non-linear domains.
+This is evident from the judgment $\Gamma \vdash A : U_i$.
+Due to the fact that $A$ is in the non-linear universe $U_i$, terms of type $A$ have unrestricted usage.
+The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_i$.
+The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_i$ and $L_i$.
+From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_i$, we see that the overall type is a non-linear type.
+The term level $\lambda$-abstractions of these types can be used unrestrictedly.
+
+The \rname{L$\rightarrow$} rule is used for forming non-linear function types with linear domains.
+From the judgment $\Gamma \vdash A : L_i$, we see that $A$ is a linear type, and terms of type $A$ have restricted usage.
+Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evidently in the judgment $\Gamma \vdash B : s_i$.
+Like \rname{U$\rightarrow$}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
+The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_i$.
+Here, $x$ is a hypocritical unbinding variable whose only purpose is to preserve syntax uniformity.
+Again, the overall resulting type is non-linear, so the term level $\lambda$-abstractions of these types can be used unrestrictedly.
+
+The \rname{U$\multimap$}, \rname{L$\multimap$} rules are similar to \rname{U$\rightarrow$}, \rname{L$\rightarrow$} in spirit.
+The dependency considerations for domain $A$ and co-domain $B$ are exactly the same.
+The main difference between \rname{U$\multimap$}, \rname{L$\multimap$} and \rname{U$\rightarrow$}, \rname{L$\rightarrow$} is the universe sort of the resulting type.
+The sorts of the function types formed by \rname{U$\multimap$}, \rname{L$\multimap$} are $L$, meaning they are linear function types.
+The term level $\lambda$-abstractions of these types must be used exactly once.
+
+\subsection{Term Formation} \label{teformation}
+The rules for term formation are presented in Figure~\ref{term}.
+
+\begin{figure}[h]
+  \caption{Term Formation}
+  \begin{mathpar}
     \inferrule
@@
     \rname{App-L$\rightarrow$}
-\end{mathpar}
-
-\begin{mathpar}
+
     \inferrule
@@
   \end{mathpar}
-
-\section{Inversion Lemmas of CLC \texttt{(clc_inversion.v)}}
+  \label{term}
+  \Description{}
+\end{figure}
+
+The rules \rname{U-Var}, and \rname{L-Var} are used for typing free variables.
+The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \utype A, \Gamma_2$ with a non-linear type $A$.
+The \rname{L-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \ltype A, \Gamma_2$ with linear type $A$.
+For both rules, the side condition $\pure{\Gamma_1, \Gamma_2}$ forbids irrelevant variables of linear types from occurring within the context.
+This prevents another vector of weakening variables with linear type.
+
+In \rname{$\lambda$$\rightarrow$}, the function type has the form $(x \stype{s} A) \rightarrow B$.
+Abstractions of this type can be applied an unrestricted number of times, hence cannot depend on free variables with restricted usage without possibly duplicating them.
+This consideration is realized by the side condition $\pure{\Gamma}$, asserting all variables in context $\Gamma$ are unrestricted.
+Next, the body of the abstraction $n$ is typed as $\Gamma, x \stype{s} A \vdash n : B$, where $s$ is the sort of $A$.
+Finally, the resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \rightarrow B$ asserts that the $\lambda$-abstraction can be used unrestrictedly.
+
+In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming abstractions that must be used exactly once.
+Due to the fact that these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evidently in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
+However, in judgment $\overline{\Gamma} \vdash (x \stype{s} A) \multimap B : L_k$ the context must be filtered, because types are not allowed to depend on restricted variables.
+The final resulting judgment $\Gamma \vdash \lambda x.n : (x \stype{s} A) \multimap B$ asserts that the $\lambda$-abstraction must be used exactly once.
+
+For \rname{App-U$\rightarrow$}, domain $A$ is a non-linear type, as seen by its annotation in $(x \utype A) \rightarrow B$.
+Intuitively, this tells us that $x$ may be used an arbitrary number of times within the body of $m$.
+Thus, the supplied argument $n$ must not depend on restricted variables in context $\Gamma_2$.
+Otherwise, substitution may put multiple copies of $n$ into $m$ during $\beta$-reduction, duplicating variables that should have been restricted.
+This justifies the side condition of $\pure{\Gamma_2}$.
+The contexts $\Gamma_1$ and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$ and $\Gamma_2$.
+
+Now for \rname{App-L$\rightarrow$}, domain $A$ is a linear type, as seen by its annotation in $(x \ltype A) \rightarrow B$.
+Intuitively, this tells us that $x$ must be used once within the body of $m$.
+During $\beta$-reduction, substitution will only put a single copy of $n$ into the body of $m$, so $n$ can depend on restricted variables within $\Gamma_2$ without fear of duplicating them.
+This justifies the lack of side condition $\pure{\Gamma_2}$.
+The contexts $\Gamma_1$ and $\Gamma_2$ are finally merged together into $\Gamma$ by the relation $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, contracting all unrestricted variables shared between $\Gamma_1$ and $\Gamma_2$.
+
+The rules \rname{App-U$\multimap$}, \rname{App-L$\multimap$} follow the same considerations as \rname{App-U$\rightarrow$}, \rname{App-L$\rightarrow$}.
+Additional conditions are not required for these two rules to be sound.
+
+Finally, the \rname{Conversion} rule allows judgment $\Gamma \vdash m : A$ to convert to judgment $\Gamma \vdash m : B$ if $B$ is a valid type in context $\overline{\Gamma}$.
+Futhermore, $A$ must be a subtype of $B$ satisfying the relation $A \preceq B$.
+This rule gives rise to large eliminations (compute types from terms), as computations embedded at the type level can reduce to canonical types.
+
+\subsection{Equality and Reduction} \label{reduction}
+The operational semantics and $\beta$-equality of CLC terms are presented in Figure \ref{red}, all of which are entirely standard.
+
+As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
+
+\begin{figure}[h]
+  \caption{Equality and Reduction}
+  \begin{mathpar}
+    \inferrule
+    { m_1 \step^* n \\ m_2 \step^* n }
+    { m_1 \equiv m_2 : A }
+    \rname{Join}
+
+    \inferrule
+    {  }
+    { (\lambda x \stype{s}A.m)\ n \step m[n/x] }
+    \rname{Step-$\beta$}
+
+    \inferrule
+    { A \step A' }
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A' .m }
+    \rname{Step-$\lambda$L}
+
+    \inferrule
+    { m \step m' }
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A.m' }
+    \rname{Step-$\lambda$R}
+
+    \inferrule
+    { A \step A' }
+    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A') \rightarrow B }
+    \rname{Step-L$\rightarrow$}
+
+    \inferrule
+    { B \step B' }
+    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A) \rightarrow B' }
+    \rname{Step-R$\rightarrow$}
+
+    \inferrule
+    { A \step A' }
+    { (x \stype{s} A) \multimap B \step (x \stype{s} A') \multimap B }
+    \rname{Step-L$\multimap$}
+
+    \inferrule
+    { B \step B' }
+    { (x \stype{s} A) \multimap B \step (x \stype{s} A) \multimap B' }
+    \rname{Step-R$\multimap$}
+
+    \inferrule
+    { m \step m' }
+    { m\ n \step m'\ n }
+    \rname{Step-AppL}
+
+    \inferrule
+    { n \step n' }
+    { m\ n \step m\ n' }
+    \rname{Step-AppR}
+  \end{mathpar}
+  \label{red}
+  \Description{}
+\end{figure}
+
+\subsection{Meta Theory} \label{meta}
+In this section, we focus our discussion on the properties of CLC.
+First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
+Next, we show that CLC is an effective linear type theory through the \textit{linearity} theorem.
+Furthermore, we show that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
+Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, showing that CLC is strongly normalizing.
+
+All proofs have been formalized in Coq with help from the Autosubst \cite{autosubst} library.
+The Coq development is publicly available on the first author's Github repository.
+To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
+We give a hand written version of the proof in an accompanying technical report.
+
+\subsubsection{Reduction and Confluence}
+
+The proof of confluence is entirely standard using parallel step technique \cite{takahashi}. The presence of linear types do not pose any complications as reductions are untyped.
+
+\begin{theorem}
+  The transitive reflexive closure of reduction is confluent. If $m \step^* m_1$ and $m \step^* m_2$ then there exists $m'$ such that $m_1 \step^* m'$ and $m_2 \step^* m'$.
+\end{theorem}
+
+\subsubsection{Weakening} \label{weakening}
+CLC restricts the weakening rule for variables of linear types. However, weakening variables of non-linear types remain admissible.
+
+\begin{theorem}
+  Weakening. If $\Gamma \vdash m : A$ is a valid judgment, then for any $x \notin \Gamma$, the judgment $\Gamma, x \utype B \vdash m : A$ is derivable.
+\end{theorem}
+
+\subsubsection{Substitution} \label{subst}
+Although the substitution lemma is regarded as a boring and bureaucratic result, it is surprisingly hard to design linear typed languages where the substitution lemma is admissible.
+Much of the difficulty arises during the substitution of arguments containing !
+exponential.
+Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
+He defines intricate syntax and semantics for the unboxing of !
+terms, solving the lack of substitute in Abramsky's term calculus.
+
+Our design of integrating ! into universe sorts removes the need for ! manipulating syntax and semantics. The following substitution theorems are directly proved by induction on typing derivation.
+
+\begin{theorem}\label{nlsubst}
+  Non-linear Substitution.
+For $\Gamma_1, x \utype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ are valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
+\end{theorem}
+
+\begin{theorem}\label{lsubst}
+  Linear Substitution. For $\Gamma_1, x \ltype A \vdash m : B$ and $\Gamma_2 \vdash n : A$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ is valid for some $\Gamma$, then $\Gamma \vdash m[n/x] : B[n/x]$.
+\end{theorem}
+
+Notice in Theorem \ref{nlsubst} of non-linear substitutions, there exists an extra condition $\pure{\Gamma_2}$ which Theorem \ref{lsubst} of linear substitutions lack.
+This condition ensures that linear variables do not occur within the substituted term, thus cannot be duplicated by substitution.
+
+\subsubsection{Type Soundness}
+In order to prove subject reduction, we first prove the validity theorem. The main purpose of validity is to lower types down to the term level, enabling the application of various inversion lemmas.
+
+\begin{theorem}
+  Validity.
+For any context $\Gamma$, term $m$, type $A$, and sort $s$, if $\Gamma \vdash m : A$ is a valid judgment, then there exist some sort $s$ and level $i$ such that $\overline{\Gamma} \vdash A : s_i$.
+  \label{validity}
+\end{theorem}
+
+With weakening, substitution, validity and various inversion lemmas proven, subject reduction can now be proved by induction on typing derivation.
+
+\begin{theorem}
+  Subject Reduction. For $\Gamma \vdash m : A$, if $m \step n$ then $\Gamma \vdash n : A$.
+\end{theorem}
+
+\subsubsection{Linearity}
+At this point, we have proven that CLC is type sound.
+However, we still need to prove that the removal of weakening and contraction for restricted variables yields tangible impact on the structure of terms.
+For this purpose, we define a binding aware recursive function $occurs(x, m)$ that counts the number of times variable $x$ appears within term $m$.
+The linearity theorem asserts that restricted variables are used exactly once within a term, subsuming safe resource usage.
+
+\begin{theorem}
+  Linearity. If $\Gamma \vdash m : A$ is a valid judgment, for any $(x \ltype B) \in \Gamma$, there is $occurs(x, m) = 1$.
+\end{theorem}
+
+\subsubsection{Promotion and Dereliction}
+Linear types of CLC are not obtained through packing and unpacking !, so there are no explicit rules for \textit{promotion} and \textit{dereliction} of Linear Logic.
+Arbitrary computations existing at the type level also muddle the association between which types can be promoted or derelicted to which.
+Nevertheless, \textit{promotion} and \textit{dereliction} for canonical types are derivable as theorems through $\eta$-expansion.
+
+\begin{theorem}
+  Promotion. If $\Gamma \vdash m : (x \stype{s} A) \multimap B$ and $\pure{\Gamma}$ are valid judgments, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$ is derivable.
+\end{theorem}
+
+\begin{theorem}
+  Dereliction. If $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$ is a valid judgment, then there exists $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$ is derivable.
+\end{theorem}
+
+\subsubsection{Strong Normalization}
+The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function from CLC to CC$\omega$.
+We assume familiarity with CC$\omega$ syntax here and define the erasure function as follows.
+
+\begin{definition}
+  \begin{align*}
+    \erase{x}                             & = x                                     \\
+    \erase{U_i}                           & = Type_i                                \\
+    \erase{L_i}                           & = Type_i                                \\
+    \erase{(x \stype{s} A) \rightarrow B} & = (x : \erase{A}) \rightarrow \erase{B} \\
+    \erase{(x \stype{s} A) \multimap B}   & = (x : \erase{A}) \rightarrow \erase{B} \\
+    \erase{\lambda x\stype{s}A.n}         & = \lambda x : \erase{A}.\erase{n}       \\
+    \erase{m\ n}                          & = \erase{m}\ \erase{n}
+  \end{align*}
+\end{definition}
+
+With slight overloading of notation we define erasure for CLC contexts recursively.
+
+\begin{definition}
+  \begin{align*}
+    \erase{\epsilon}              & = \epsilon                      \\
+    \erase{\Gamma, x \stype{s} A} & = \erase{\Gamma}, x : \erase{A}
+  \end{align*}
+\end{definition}
+
+For the following theorems, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
+
+\begin{theorem} \label{preserve}
+  Reduction Erasure. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+\end{theorem}
+
+\begin{theorem} \label{embedding}
+  Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
+\end{theorem}
+
+If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve} and \ref{embedding}.
+This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
+
+\begin{theorem}
+  Well-typed CLC terms are strongly normalizing.
+\end{theorem}
+
+\subsubsection{Compatibility}
+To show that CLC is compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
+
+\begin{definition}
+  \begin{align*}
+    \lift{x}                     & = x                                        \\
+    \lift{Type_i}                & = U_i                                      \\
+    \lift{(x : A) \rightarrow B} & = (x \utype \lift{A}) \rightarrow \lift{B} \\
+    \lift{\lambda x : A.n}       & = \lambda x\utype\lift{A}.\lift{n}         \\
+    \lift{m\ n}                  & = \lift{m}\ \lift{n}
+  \end{align*}
+\end{definition}
+
+With slight overloading of notation, we define lifting for CC$\omega$ recursively.
+\begin{definition}
+  \begin{align*}
+    \lift{\epsilon}      & = \epsilon                         \\
+    \lift{\Gamma, x : A} & = \lift{\Gamma}, x \utype \lift{A}
+  \end{align*}
+\end{definition}
+
+The following theorems are proved following a similar procedure to proving the soundness of erasure.
+\begin{theorem}
+  Lift preserves reduction. For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{theorem}
+
+\begin{theorem} \label{lift}
+  Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
+\end{theorem}
+
+The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domain of function types are annotated with sort $U$.
+In practice, these annotations can often be omitted if algorithmic type checking is able to infer them from context.
+
+\section{The Extension to CILC}
+Although the core CLC language is extremely expressive, formally shown by Theorem \ref{lift} to be at least as expressive as the predicative CC$\omega$, the complexity of Church-style encodings to do so will quickly become unmanageable.
+To address this, we extend CLC with rules for defining inductive linear types in the style of CIC.
+
+We have formalized CILC in our Coq development and proven its soundness.
+
+\subsection{Syntax of CILC}
+\begin{figure}[h]
+  \caption{Syntax}
+  \begin{grammar}
+    <$C,I,P,Q,f$> ::= ... \hspace*{10em} expressions
+    \indalt $\ind{s}(X : A)\{C_1 \sep ... \sep C_k\}$
+    \indalt $\constr(k, I)$
+    \indalt $\case(m, Q)\{f_1 \sep ... \sep f_k \}$
+    \indalt $\dcase(m, Q)\{f_1 \sep ... \sep f_k \}$
+    \indalt $\fix f : A := m$
+  \end{grammar}
+  \Description{}
+  \label{syntax2}
+\end{figure}
+
+As shown in Figure \ref{syntax2}, CILC extends CIC with syntax for the introduction and elimination of inductive types.
+Ind is used during the formation of new inductive types, and Constr is used for introducing their constructors.
+Case and DCase are non-dependent and dependent case eliminators respectively.
+CIC-style primitive recursion has been factored out of case eliminators in favor of a Coq-style Fix construct.
+The flexibility of an independent Fix construct allows for a cleaner formalization of inductive linear type elimination.
+
+\subsection{Arity}
+For type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as depicted in Figure \ref{arity}.
+
+\begin{figure}[h]
+  \caption{Arity}
+  \begin{mathpar}
+    \inferrule
+    { }
+    { arity(s_i,s) }
+    \rname{A-Sort}
+
+    \inferrule
+    { arity(A, s) }
+    { arity((x \utype M) \rightarrow A, s)}
+    \rname{A$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{arity}
+\end{figure}
+
+\begin{definition}
+  For some arity type $A$ and sort $s'$, the term $A\lrangle{s'}$ is inductively defined over the structure of $A$ as follows.
+  \begin{align*}
+    ((x \utype M) \rightarrow A)\lrangle{s'} & = (x \utype M) \rightarrow A\lrangle{s'} \\
+    (s)\lrangle{s'}                          & = s'
+  \end{align*}
+\end{definition}
+
+\begin{definition}
+  For some arity type $A$, term $I$ and sort $s'$, the term $A\{I, s'\}$ is inductively defined over the structure of $A$ as follows.
+  \begin{align*}
+    ((x \utype M) \rightarrow A)\lrangle{I, s'} & = (x \utype M) \rightarrow A\lrangle{(I\ x), s'} \\
+    (s)\lrangle{I, s'}                          & = (\_ \utype I) \rightarrow s'
+  \end{align*}
+\end{definition}
+
+\subsection{Strict Positivity}
+For type $P$ and variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as depicted in Figure \ref{positive}.
+
+We say that $X$ occurs strictly positive in $P$ if the judgment $postive(P,X)$ is valid.
+
+\begin{figure}[h]
+  \caption{Strict Positivity}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { positive((X\ m_1...m_k), X) }
+    \rname{Pos-X}
+
+    \inferrule
+    { X \notin M \\ positive(P, X)}
+    { positive((x \stype{s} M) \rightarrow P, X) }
+    \rname{Pos$\rightarrow$}
+
+    \inferrule
+    { X \notin M \\ positive(P, X) }
+    { positive((x \stype{s} M) \multimap P, X) }
+    \rname{Pos$\multimap$}
+  \end{mathpar}
+  \Description{}
+  \label{positive}
+\end{figure}
+
+\subsection{Non-Linear Constructor}
+For type $C$ and variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{nlconstr}.
+
+Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not applied to restricted terms as this may cause duplication.
+
+\begin{figure}[h]
+  \caption{Non-linear Constructor}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { \ucons((X\ m_i...m_k), X) }
+    \rname{U-Constr-X}
+
+    \inferrule
+    { positive(P, X) \\ \ucons(C, X) }
+    { \ucons((\_ \utype P) \rightarrow C, X)}
+    \rname{U-Constr-pos}
+
+    \inferrule
+    { X \notin M \\ \ucons(C, X) }
+    { \ucons((x \utype M) \rightarrow C, X)}
+    \rname{U-Constr$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{nlconstr}
+\end{figure}
+
+\subsection{Linear Constructor}
+For type $C$ and variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{lconstr}.
+
+Linear constructors are used for constructing linear objects that must be used once.
+These constructors may be applied to linear terms as restricted usage prevents duplication and enforces single usage.
+
+An unapplied linear constructor is a non-linear entity as they can be created freely ``out of thin air".
+However, once a linear constructor has been partially applied to a linear term, its linearity is ``activated'' in rules \rname{L-Constr-Pos-2} and \rname{L-Constr-2$\rightarrow$}, essentially forcing the rest of its arguments to be fully applied.
+This is to prevent partially applied constructors from duplicating its linear arguments.
+
+\begin{figure}[h]
+  \caption{Linear Constructor}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { \lcons((X\ m_i...m_k), X) }
+    \rname{L-Constr-X}
+
+    \inferrule
+    { positive(P, X) \\ \lcons(C, X) }
+    { \lcons((\_ \utype P) \rightarrow C, X)}
+    \rname{L-Constr-Pos-1}
+
+    \inferrule
+    { X \notin M \\ \lcons(C, X) }
+    { \lcons((x \utype M) \rightarrow C, X)}
+    \rname{L-Constr-1$\rightarrow$}
+
+    \inferrule
+    { positive(P, X) \\ activation(C, X) }
+    { \lcons((\_ \ltype P) \rightarrow C, X)}
+    \rname{L-Constr-Pos-2}
+
+    \inferrule
+    { X \notin M \\ activation(C, X) }
+    { \lcons((x \ltype M) \rightarrow C, X)}
+    \rname{L-Constr-2$\rightarrow$}
+  \end{mathpar}
+  \Description{}
+  \label{lconstr}
+\end{figure}
+
+For type $C$ and variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{activation}.
+Intuitively, $activation(C,X)$ ensures the linearity of $C$.
+
+\begin{figure}[h]
+  \caption{Activation}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k) \\ X \notin m_i }
+    { activation((X\ m_i...m_k), X) }
+    \rname{Act-X}
+
+    \inferrule
+    { positive(P, X) \\ activation(C, X) }
+    { activation((\_ \stype P) \multimap C, X)}
+    \rname{Act-Pos}
+
+    \inferrule
+    { X \notin M \\ activation(C, X) }
+    { activation((x \stype M) \multimap C, X)}
+    \rname{Act$\multimap$}
+  \end{mathpar}
+  \Description{}
+  \label{activation}
+\end{figure}
+
+\subsection{Introduction Rules}
+The formation of new inductive types and their constructors are presented in Figure \ref{ind}.
+
+\begin{figure}[h]
+  \caption{Inductive Type Formation}
+  \begin{mathpar}
+    \inferrule
+    { (\forall j = 1...n) \\ arity(A, s) \\ \scons(C_j, X) \\\\
+      \pure{\Gamma} \\
+      \Gamma \vdash A : U_i \\
+      \Gamma, X \utype A \vdash C_j : t_i }
+    { \Gamma \vdash \ind{s}(X : A)\{C_1|...|C_k\} : A }
+    \rname{Ind}
+
+    \inferrule
+    { 1 \leq i \leq k \\\\
+    \pure{\Gamma} \\
+    I := \ind{s}(X : A)\{C_1|...|C_k\} \\
+    \Gamma \vdash I :A }
+    { \Gamma \vdash \constr(i, I) : C_i[I/X] }
+    \rname{Constr}
+  \end{mathpar}
+  \Description{}
+  \label{ind}
+\end{figure}
+
+\subsection{Non-Dependent Case}
+
+\begin{definition}
+  For constructor type $C$ and variables $X$ and $Q$, the term $C\{X, Q\}$ is inductively defined over the structure of $C$ as follows.
+  \begin{align*}
+    ((\_ \stype{s} P) \rightarrow C)\{X, Q\} & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
+    ((\_ \stype{s} P) \multimap C)\{X, Q\}   & = (\_ \stype{s} P) \multimap C\{X, Q\} \\
+    ((x \stype{s} M) \rightarrow C)\{X, Q\}  & = (x \stype{s} P) \multimap C\{X, Q\}  \\
+    ((x \stype{s} M) \multimap C)\{X, Q\}    & = (x \stype{s} P) \multimap C\{X, Q\}  \\
+    (X\ m_1...m_k)\{X, Q\}                   & = (Q\ m_1...m_k)
+  \end{align*}
+  We define the notation $C[I, P] := C\{X, Q\}[I/X, P/Q]$.
+\end{definition}
+
+The rule for non-dependent case elimination is presented in Figure \ref{case}. This elimination rule is non-dependent because the motive $Q$ may not depend on the discriminated term $m$ itself.
+
+\begin{figure}[h]
+  \caption{Non-dependent Case Elimination}
+  \begin{mathpar}
+    \inferrule
+    { (\forall i = 1...k_1) \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+    arity(A, s) \\\\
+    I := \ind{s}(X : A)\{C_1|...|C_{k_1}\} \\
+    \overline{\Gamma_2} \vdash Q : A\lrangle{s'} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_{k_2}) \\
+    \Gamma_2 \vdash f_i : C_i[I, Q] }
+    { \Gamma \vdash \case(m, Q)\{f_1|...|f_{k_1}\} : (Q\ a_1...a_{k_2}) }
+    \rname{Case}
+  \end{mathpar}
+  \Description{}
+  \label{case}
+\end{figure}
+
+\subsection{Dependent Case}
+
+\begin{definition}
+  For some non-linear constructor type $C$ and type variables $X$, $Q$ and $c$, the term $C\{X, Q, c\}$ is inductively defined over the structure of $C$ as follows.
+  \begin{align*}
+    ((\_ \utype P) \rightarrow C)\{X, Q, c\} & = (p \utype P) \multimap C\{X, Q, (c\ p)\} \\
+    ((x \utype M) \rightarrow C)\{X, Q, c\}  & = (x \utype M) \multimap C\{X, Q, (c\ x)\} \\
+    (X\ m_1...m_n)\{X, Q, c\}                & = (Q\ m_1...m_n\ c)
+  \end{align*}
+  We define the notation $C[I, P, t] := C\{X, Q, c\}[I/X, P/Q, t/c]$.
+\end{definition}
+
+Dependent elimination of non-linear inductive objects is presented in Figure \ref{dcase}.
+This elimination rule is dependent because the motive $Q$ may depend on the discriminated term $m$.
+For this very reason, $m$ is not allowed to have linear type as doing so may transport restricted variable into types.
+
+\begin{figure}[h]
+  \caption{Dependent Case Elimination}
+  \begin{mathpar}
+    \inferrule
+    { \pure{\Gamma_1} \\
+    (\forall i = 1...n) \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+    arity(A, U_i) \\\\
+    I := \ind{U}(X : A)\{C_1|...|C_n\} \\
+    \overline{\Gamma_2} \vdash Q : A\lrangle{I, s'} \\\\
+    \Gamma_1 \vdash m : (I\ a_1...a_n) \\
+    \Gamma_2 \vdash f_i : C_i[I, Q, Constr(i, I)] }
+    { \Gamma \vdash \dcase(m, Q)\{f_1|...|f_n\} : (Q\ a_1...a_n\ m) }
+    \rname{DCase}
+  \end{mathpar}
+  \Description{}
+  \label{dcase}
+\end{figure}
+
+\subsection{Fixpoint}
+The typing rule of fix-points presented in Figure \ref{fix} is mostly standard.
+A syntactic guard condition is utilized to conservatively ensure termination.
+The fix-point term must have non-linear type in a pure context, because recursive calls may cause duplicated usage of restricted variables.
+
+\begin{figure}[h]
+  \caption{Fixpoint}
+  \begin{mathpar}
+    \inferrule
+    { \pure{\Gamma} \\
+      \Gamma \vdash A : U_i \\
+      \Gamma, f \utype A \vdash m : A \\
+      \text{guard condition} }
+    { \Gamma \vdash \fix f : A := m : A }
+    \rname{fix}
+  \end{mathpar}
+  \Description{}
+  \label{fix}
+\end{figure}
+
+\subsection{Extended Reduction}
+As the reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
+
+\begin{figure}[h]
+  \caption{$\iota$-Reductions}
+  \begin{mathpar}
+    \inferrule
+    { }
+    { \case((\constr(i, I)\ a_1...a_n),Q)\{f_1|...f_k\} \\\\
+      \leadsto (f_i\ a_1...a_k) }
+    \rname{Step-Case-$\iota$}
+
+    \inferrule
+    { }
+    { \dcase((\constr(i, I)\ a_1...a_k),Q)\{f_1|...f_k\} \\\\
+      \leadsto (f_i\ a_1...a_k) }
+    \rname{Step-DCase-$\iota$}
+
+    \inferrule
+    { }
+    { \fix f : A := m \leadsto m[(\fix f : A := m)/f] }
+    \rname{Step-Fix-$\iota$}
+  \end{mathpar}
+  \Description{}
+  \label{stepfix}
+\end{figure}
+
+\subsection{Non-Linear Connectives}
+Non-linear inductive types have been extensively studied since their inception.
+We will omit discussion of non-linear inductive types here as all inductive definitions in the predicative fragment of CIC are available in CILC.
+
+\subsection{Linear Connectives}
+In Figure \ref{lconnectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
+For any term with inductive linear type, linearity will ultimately force the term to be eliminated by a Case expression.
+
+\begin{figure}[h]
+  \caption{Inductively Defined Linear Connectives}
+  \begin{align*}
+    \_\otimes\_ :=\
+     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
+     & \sep\ (A \utype L_i) \rightarrow (B \utype L_i)                               \\
+     & \qquad \rightarrow (\_ \ltype A)  \multimap (\_ \ltype B) \multimap (X\ A\ B) \\
+     & \}
+    \\
+    \lrangle{\_,\_} := \
+     & \constr(1, \otimes)
+    \\\\
+    \_\oplus\_ := \
+     & \ind{L}(X : (A \utype L_i) \rightarrow (B \utype L_i) \rightarrow L_i)\{      \\
+     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
+    \rightarrow (\_ \ltype A)  \multimap (X\ A\ B)                                   \\
+     & \sep \ (A \utype L_i) \rightarrow (B \utype L_i)
+    \rightarrow (\_ \ltype B)  \multimap (X\ A\ B)                                   \\
+     & \}
+    \\
+    \inl(\_) := \
+     & \constr(1, \oplus)
+    \\
+    \inr(\_) := \
+     & \constr(2, \oplus)
+  \end{align*}
+  \Description{}
+  \label{lconnectives}
+\end{figure}
+
+\subsection{Mixed Connectives}
+In Figure \ref{mconnectives}, we introduce a $\SigmaL$ connective.
+$\SigmaL$ is similar to the standard $\Sigma$ type where the type of the second component is dependent on the value of the first component.
+The key difference between these two type constructors is that $\Sigma$'s second component is non-linear, whereas $\SigmaL$'s second component is linear.
+
+\begin{figure}[h]
+  \caption{Mixed Connective}
+  \begin{align*}
+    \SigmaL := \
+     & \ind{L}(X : (A \utype U_i) \rightarrow
+    (B \utype (\_ \utype A) \rightarrow L_i) \rightarrow L_i)\{ \\
+     & \sep \ (A \utype U_i) \rightarrow
+    (B \utype (\_ \utype A) \rightarrow L_i)                    \\
+     & \qquad \rightarrow (m \utype A) \rightarrow
+    (\_ \ltype B\ m) \multimap (X\ A\ B)                        \\
+     & \}
+    \\
+    [\_,\_] :=\
+     & \constr(1, \SigmaL)
+  \end{align*}
+  \Description{}
+  \label{mconnectives}
+\end{figure}
+
+\section{Examples}\label{examples}
+In this section we will demonstrate some applications of CILC in practice.
+When it is clear from context whether a type is linear or non-linear, we will omit writing sort annotations on function types.
+We also use Coq style concrete syntax from our implementation for the sake of readability.
+The full code for each example is available with our implementation.
+
+\subsection{Stateful Reasoning}
+We first postulate five state axioms. These axioms could be viewed as an interface to a trusted memory allocator. Using these axioms, we will build safe random access arrays from the ground up.
+\begin{definition}
+  State axioms.
+  \begin{itemize}
+    \item $\mapsto\ : \mathbb{N} \rightarrow (A : U) \rightarrow L$ \\
+          The $\mapsto$ is an infix linear type constructor.
+It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
+Intuitively, a capability $(l \mapsto A)$ is a proof that a term of non-linear type $A$ is currently stored at address $l$.
+    \item $\textbf{new} : (A : U) \rightarrow A \rightarrow \SigmaL l:\mathbb{N}.(l \mapsto A)$ \\
+          $\new$ is a function that allocates a new memory cell for a term non-linear type $A$ and initializes the cell with its argument.
+This returns an address $l$ paired with a capability $(l \mapsto A)$.
+    \item $\textbf{free} : (A : U) \rightarrow (l:\mathbb{N}) \rightarrow (l \mapsto A) \rightarrow \unit$ \\
+          $\free$ is a function that de-allocates the memory cell at location $l$ which is currently in use. The capability $(l \mapsto A)$ is consumed so the memory cell can no longer be accessed.
+    \item $\textbf{get} : (A : U) \rightarrow (l : \mathbb{N}) \rightarrow (l \mapsto A) \rightarrow \SigmaL \_: A. (l \mapsto A)$ \\
+          $\get$ is a function that retrieves the term stored at address $l$ when given a proof $(l \mapsto A)$ there is indeed a term of type $A$ currently stored there.
+The type of $\get$'s output is the rather complicated $\SigmaL \_: A.
+(l \mapsto A)$.
+This not only returns the term stored at $l$, but also a proof $(l \mapsto A)$ that the memory state has not changed and may be accessed again.
+    \item $\textbf{set} : (A\ B : U) \rightarrow (l : \mathbb{N}) \rightarrow B \rightarrow (l \mapsto A) \rightarrow (l \mapsto B)$ \\
+          $\set$ has the most complicated type of all the axioms presented here.
+Intuitively, if address $l$ is allocated and storing some term of type $A$, $\set$ may put a term type $B$ into the the store at address $l$, overwriting the original term.
+This is the so called "strong update" operation.
+The consumption of $(l \mapsto A)$ and return of $(l \mapsto B)$ accurately characterize this update process.
+  \end{itemize}
+\end{definition}
+Due to the fact that $(l \mapsto A)$ is an abstract linear type, proofs of this type can not be duplicated and may only be consumed by other state axioms.
+This is the same non-sharing principle that empowers Separation Logic.
+Unlike Separation Logic, $(l \mapsto A)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
+
+A random access array is a contiguous region of memory that can be accessed by an address pointing to the start of this region and an offset.
+In Figure \ref{arrv}, we declare a linear inductive $\arrv$ to track the memory used by the array.
+$\arrv$ is parameterized by the type $A$ of elements stored in the array and the address $l$ of the initial element.
+$\arrv$ also takes an additional argument of type $\mathbb{N}$ indicating the length of the array.
+
+\begin{figure}[h]
+  \caption{Proof of Arrays}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
+Inductive ArrVec (A : U) (l : ?$\mathbb{N}$?) : ?$\mathbb{N}$? ?$\rightarrow$? L :=
+| Nil : ArrVec A l 0
+| Cons : (n : ?$\mathbb{N}$?) ?$\rightarrow$? ((l + n) ?$\mapsto$? A) ?$\rightarrow$? 
+    ArrVec A l n ?$\rightarrow$? ArrVec A l (S n).
+  \end{minted}
+  \label{arrv}
+  \Description{}
+\end{figure}
+
+Notice that for the $\texttt{Cons}$ constructor, each time a proof of $(l + n \mapsto A)$ is used to extend the current $\arrv$, the next application of $\texttt{Cons}$ expects a proof of $(l + S\ n \mapsto A)$.
+So for some type $A$ and natural numbers $l, n$, a term of type $\arrv\ A\ l\ n$ is a proof that a contiguous region of memory starting from address $l$ can be accessed by any offset less than $n$.
+Arrays can now be characterized as an address $l$ paired with an array proof $\arrv\ A\ l\ n$ as shown in Figure \ref{array}.
+
+\begin{figure}[h]
+  \caption{Arrays}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
+Definition Array (A : U) (n : ?$\mathbb{N}$?) : L := 
+  ?$\SigmaL$? l : ?$\mathbb{N}$?.ArrVec A l n.
+  \end{minted}
+  \label{array}
+  \Description{}
+\end{figure}
+
+To manipulate arrays, we construct a recursive helper function whose signature is shown in Figure \ref{trav}.
+When given an array proof $\arrv\ A\ l\ n$, an offset $m$ and proof $pf$ such that $m < n$, $\texttt{nth}$ will return a capability $(l + m \mapsto A)$ along with a function $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$.
+The capability $(l + m \mapsto A)$ grants the ability to manipulate the store at offset $m$ using the state axioms.
+The function $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$ is constructed using the rest of $\arrv\ A\ l\ n$ not at offset $m$.
+Intuitively, $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$ is the original array proof with a missing capability at offset $m$ which can be restored by applying capability $(l + m \mapsto A)$.
+
+\begin{figure}[H]
+  \caption{ArrVec Traversal}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
+Fixpoint nth 
+  (A : U) 
+  (l m n : ?$\mathbb{N}$?) 
+  (pf : m < n) 
+  (v : ArrVec A l n) 
+: (l + m ?$\mapsto$? A) ?$\otimes$? ((l + m ?$\mapsto$? A) ?$\multimap$? ArrVec A l n).
+  \end{minted}
+  \label{trav}
+  \Description{}
+\end{figure}
+
+An indexing function can now be constructed as a wrapper for $\texttt{nth}$ as shown in Figure \ref{index}.
+Given an array $\texttt{Array}\ A\ n$, an offset $m$ and proof that $m < n$, $\texttt{nth}$ can be used to extract out the capability at offset $m$, allowing $\get$ to retrieve the stored value.
+Finally, the array proof is restored and returned alongside the indexing result.
+This definition of $\texttt{index}$ is safe because the proof $m < n$ ensures that accessing offsets are always within bounds.
+
+\begin{figure}[h]
+  \caption{Array Indexing}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
+Definition index
+  (A : U)
+  (m n : ?$\mathbb{N}$?)
+  (pf : m < n)
+  (a : Array A n) 
+: ?$\SigmaL$? _ : A. Array A n
+  \end{minted}
+  \label{index}
+  \Description{}
+\end{figure}
+
+From our experiments with proof erasure, we observe that erasing the proof arguments of $\texttt{index}$ in a reasonable way can result in a single call to $\get$.
+This indicates that the arrays described here can be compiled for constant time access.
+
+\subsection{Protocol Enforcement}
+The inclusion of dependent types and linear types in CILC makes it extremely expressive when encoding communication protocols.
+By indexing each channel with a protocol, CILC can enforce communication on a channel to strictly adhere to its specified protocol.
+
+In Figure \ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
+Intuitively, the \texttt{session}  forms a stack of possible operations executed in a protocol.
+The $\SEND$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after sending a message of type $A$, the protocol progresses to stage $ss$.
+Likewise, the $\RECV$ constructor takes in a non-linear type $A$ and a session $ss$ as its argument, indicating that after receiving a message of type $A$, the protocol progresses to stage $ss$.
+Finally, the $\END$ constructor indicates that the protocol has finished and communication is over.
+
+\begin{figure}[h]
+  \caption{Inductive Session Type}
+  \begin{minted}{coq}
+Inductive session : U :=
+| SEND : U -> session -> session
+| RECV : U -> session -> session
+| END : session.
+  \end{minted}
+  \label{session}
+  \Description{}
+\end{figure}
+
+In Definition \ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
+
+\begin{definition}
+  Communication axioms.
+  \begin{itemize}
+    \item $\textbf{channel} : \session \rightarrow L$ \\
+          The $\channel$ axiom is a type constructor. It is used to form the types of communication channels that obey the protocol specified by its session.
+    \item $\textbf{open} : (ss : \session) \rightarrow \channel\ ss$ \\
+          When given a protocol, $\open$ creates a channel indexed by this protocol.
+    \item $\textbf{close} : \channel\ \END \rightarrow \unit$ \\
+          Once all communications specified on a given channel has finished, $\close$ will close and reclaim the channel.
+    \item $\textbf{send} : (A : U) \rightarrow A \rightarrow (ss : \session) \rightarrow$ \\
+          \phantom{send : (A :)}
+          $\channel\ (\SEND\ A\ ss) \rightarrow \channel\ ss$ \\
+          After a channel has reached the point in its protocol where a message of type $A$ should be sent, $\send$ may send a term of type $A$ on this channel, causing the channel to progress to the next stage of its protocol.
+    \item $\textbf{recv} : (A : U) \rightarrow (ss : \session) \rightarrow $ \\
+          \phantom{recv : (A :)}
+          $\channel\ (\RECV\ A\ ss) \rightarrow F \_:A.\channel\ ss$
+          After a channel has reached the point in its protocol where a message of type $A$ should be received, $\recv$ may receive such a term of type $A$ from the channel, causing the channel to progress to the next stage of its protocol.
+  \end{itemize}
+  \label{comm}
+\end{definition}
+
+The important detail to notice here is that each channel created by $\open$ is of linear type $(\channel\ ss)$ where $ss$ is some arbitrary protocol, hence must be used exactly once.
+However, the $\send$ and $\recv$ operations return back channels after consuming them, progressing their protocols one stage forward.
+Coupled with the fact that the $\close$ operation is only able to close channels whose protocols have ended, this forces channels to communicate strictly by their specified protocols.
+
+\subsection{Mutable Data}
+Figure \ref{list} presents a CPS append function for linear lists.
+In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
+Instead of de-allocating the memory of \texttt{cons}, it is bound to the variable \texttt{_cons_} of linear type $A \multimap list\ A \multimap list\ A$.
+Each \texttt{_cons_} is applied exactly once to reconstruct a list using the memory of the original, thus mutating the original list in-place.
+Also notice that the continuation $k$ is of linear type $(list\ A \multimap list\ A)$, indicating that $k$ must be applied exactly once.
+Using linear types for continuations is surprisingly accurate as control-flow itself cannot be freely duplicated nor discarded.
+
+\begin{figure}[h]
+  \caption{Linear Lists}
+  \begin{minted}{coq}
+Inductive list (A : U) : L :=
+| nil : list A
+| cons : A -> list A -> list A.
+
+Fixpoint append (A : U) : list A -o 
+  list A -o (list A -o list A) -o list A 
+:=
+  fun ls1 ls2 k =>
+    match ls1 with
+    | nil => k ls2
+    | (cons h t) as _cons_ =>
+      append _ t ls2 
+        (fun res => k (_cons_ h res))
+    end.
+  \end{minted}
+  \label{list}
+  \Description{}
+\end{figure}
+
+\section{Related Work}
+Linear types are a class of type systems inspired by Girard's substructural Linear Logic \cite{girard}.
+Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
+Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resource.
+A term calculus for linear type theory is realized by Abramsky \cite{abramsky1993}.
+Benton \cite{benton1994} investigates the ramifications of the !
+exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
+Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
+
+Over the years, work has been done to enrich linear type theories with dependent types.
+Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
+Krishnaswami et al.
+present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
+Luo et al.
+\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.
+The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
+
+\section{Future Work}
+In early versions of CLC, an impredicative $Prop$ universe sat at the bottom of the universe hierarchy.
+This has since been removed due to the additional difficulty of working with $Prop$ in the soundness proof of CILC.
+We believe that these difficulties are not intrinsic and $Prop$ could be added back with careful proof work.
+
+We aim to fully verify the typechecking algorithm employed by the current implementation.
+The unification algorithm used for implicit argument resolution is extremely ad-hoc.
+As the problem of unification for linear types has not been thoroughly studied, we hope to improve this situation as well.
+
+\bibliographystyle{acm}
+\bibliography{../ref}
+
+\appendix
+
+\section{Syntax of CLC}
+\begin{figure}[H]
+  \begin{grammar}
+    <$i$> ::= 0 | 1 | 2 ... \phantom{* |} \hspace*{2.4em} universe levels
+
+    <$s, t$> ::= $U$ | $L$ \phantom{| $x$} \hspace*{4.6em} sorts
+
+    <$m, n, A, B, M$> ::= $U_i$ | $L_i$ | $x$ \hspace*{4em} expressions
+    \indalt $(x :_s A) \rightarrow B$
+    \indalt $(x :_s A) \multimap B$
+    \indalt $\lambda x :_s A. n$
+    \indalt $m\ n$
+  \end{grammar}
+  \Description{}
+\end{figure}
+
+\section{Reduction and Equality of CLC}
+\begin{figure}[H]
+  \begin{mathpar}
+    \inferrule
+    { m_1 \red n \\ m_2 \red n }
+    { m_1 \equiv m_2 : A }
+    \rname{Join}
+
+    \inferrule
+    {  }
+    { (\lambda x \stype{s}A.m)\ n \step m[n/x] }
+    \rname{Step-$\beta$}
+
+    \inferrule
+    { A \step A' }
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A' .m }
+    \rname{Step-$\lambda$L}
+
+    \inferrule
+    { m \step m' }
+    { \lambda x \stype{s}A.m \step \lambda x \stype{s}A.m' }
+    \rname{Step-$\lambda$R}
+
+    \inferrule
+    { A \step A' }
+    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A') \rightarrow B }
+    \rname{Step-L$\rightarrow$}
+
+    \inferrule
+    { B \step B' }
+    { (x \stype{s} A) \rightarrow B \step (x \stype{s} A) \rightarrow B' }
+    \rname{Step-R$\rightarrow$}
+
+    \inferrule
+    { A \step A' }
+    { (x \stype{s} A) \multimap B \step (x \stype{s} A') \multimap B }
+    \rname{Step-L$\multimap$}
+
+    \inferrule
+    { B \step B' }
+    { (x \stype{s} A) \multimap B \step (x \stype{s} A) \multimap B' }
+    \rname{Step-R$\multimap$}
+
+    \inferrule
+    { m \step m' }
+    { m\ n \step m'\ n }
+    \rname{Step-AppL}
+
+    \inferrule
+    { n \step n' }
+    { m\ n \step m\ n' }
+    \rname{Step-AppR}
+  \end{mathpar}
+  \Description{}
+\end{figure}
+
+\section{Confluence of CLC}
+
+\subsection{Parallel Reduction}
+To prove the confluence property of CLC, we employ the standard technique utilizing parallel reductions.
+
+\begin{figure}[H]
+  \begin{mathpar}
+    \inferrule
+    { }
+    { x \pstep x }
+    \rname{PStep-Var}
+
+    \inferrule
+    { }
+    { s_i \pstep s_i }
+    \rname{PStep-Sort}
+
+    \inferrule
+    { A \pstep A' \\ m \pstep m' }
+    { \lambda x \stype{s} A.m \pstep \lambda x \stype{s} A'.m' }
+    \rname{PStep-$\lambda$}
+
+    \inferrule
+    { m \pstep m' \\ n \pstep n' }
+    { m\ n \pstep m'\ n' }
+    \rname{PStep-App}
+
+    \inferrule
+    { m \pstep m' \\ n \pstep n'}
+    { (\lambda x \stype{s} A.m)\ n \pstep m'[n'/x] }
+    \rname{PStep-$\beta$}
+
+    \inferrule
+    { A \pstep A' \\ B \pstep B' }
+    { (x \stype{s} A) \rightarrow B \pstep (x \stype{s} A') \rightarrow B' }
+    \rname{PStep$\rightarrow$}
+
+    \inferrule
+    { A \pstep A' \\ B \pstep B' }
+    { (x \stype{s} A) \multimap B \pstep (x \stype{s} A') \multimap B' }
+    \rname{PStep$\multimap$}
+  \end{mathpar}
+  \label{pred}
+  \Description{}
+\end{figure}
+
+\subsection{Reduction Lemmas}
+Here, we prove some simple lemmas concerning $\step$, $\red$ and substitution.
+
+\begin{definition}
+  For a term $m$ and a map $\sigma$ from variables to terms, let $m[\sigma]$ be the term obtained by applying $\sigma$ uniformly to all free variables in $m$.
+\end{definition}
+
+\begin{definition}
+  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ reduces to $\tau$ if for any variable $x$ there exists a reduction $(\sigma\ x) \red (\tau\ x)$.
+We write $\sigma \red \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{stepsubst}
+  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a step $m \step n$, then there exists a step $m[\sigma] \step n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step n$.
+\end{proof}
+
+\begin{lemma}\label{redapp}
+  For terms $m_1, m_2, n_1, n_2$, if these exists reductions $m_1 \red m_2$ and $n_1 \red n_2$, then there exists reduction $(m_1\ n_1) \red (m_2\ n_2)$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-AppL}, \rname{Step-AppR}.
+\end{proof}
+
+\begin{lemma}\label{redlam}
+  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $m_1 \red m_2$, then there exists reduction $\lambda x \stype{s}A_1.m_1 \red \lambda x \stype{s}A_2.m_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
+\end{proof}
+
+\begin{lemma}\label{redarrow}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \rightarrow B_1 \red (x \stype{s} A_2) \rightarrow B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
+\end{proof}
+
+\begin{lemma}\label{redlolli}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists reductions $A_1 \red A_2$ and $B_1 \red B_2$, then there exists reduction $(x \stype{s}A_1) \multimap B_1 \red (x \stype{s} A_2) \multimap B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\red$ and applying rules \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
+\end{proof}
+
+\begin{lemma}\label{redsubst}
+  For terms $m, n$ and a map $\sigma$ from variables to terms, if there exist a reduction $m \red n$, then there exists a reduction $m[\sigma] \red n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\red$, the transitivity of $\red$ and Lemma \ref{stepsubst}.
+\end{proof}
+
+\begin{lemma}\label{redcompat}
+  For maps $\sigma, \tau$ from variables to terms, if there is a map reduction $\sigma \red \tau$, then for any term $m$ these is a reduction $m[\sigma] \red m[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$, applying Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}.
+\end{proof}
+
+\subsection{Equality Lemmas}
+Here, we prove some simple lemmas concerning $\red$, $\equiv$ and substitution.
+
+\begin{definition}
+  For maps $\sigma, \tau$ from variables to terms, we say that $\sigma$ is equal to $\tau$ if for any variable $x$ there exists an equality $(\sigma\ x) \equiv (\tau\ x)$.
+We write $\sigma \equiv \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{convhom}
+  For any map $f$ from terms to terms, if for any terms $m, n$ such that $m \step n$ implies $f\ m \equiv f\ n$, then for any terms $m, n$ equality $m \equiv n$ implies $f\ m \equiv f\ n$.
+\end{lemma}
+\begin{proof}
+  By the properties of the transitive reflexive closure $\red$ and that $\equiv$ is an equivalence relation.
+\end{proof}
+
+\begin{lemma}\label{convapp}
+  For terms $m_1, m_2, n_1, n_2$, if there exists equalities $m_1 \equiv m_2$ and $n_1 \equiv n_2$, then there exists equality $(m_1\ n_1) \equiv (m_2\ n_2)$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-AppL}, \rname{Step-AppR}.
+\end{proof}
+
+\begin{lemma}\label{convlam}
+  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $m_1 \equiv m_2$, then there exists equality $\lambda x \stype{s} A_1.m_1 \equiv \lambda x \stype{s} A_2.m_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-$\lambda$L}, \rname{Step-$\lambda$R}.
+\end{proof}
+
+\begin{lemma}\label{convarrow}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{s} A_2) \rightarrow B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\rightarrow$}, \rname{Step-R$\rightarrow$}.
+\end{proof}
+
+\begin{lemma}\label{convlolli}
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{s} A_2) \multimap B_2$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and applying rules \rname{Join}, \rname{Step-L$\multimap$}, \rname{Step-R$\multimap$}.
+\end{proof}
+
+\begin{lemma}\label{convsubst}
+  For terms $m, n$ and map $\sigma$ from variables to terms, if there is equality $m \equiv n$, then there is equality $m[\sigma] \equiv n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By Lemmas \ref{convhom} and \ref{stepsubst}.
+\end{proof}
+
+\begin{lemma}\label{convcompat}
+  For maps $\sigma, \tau$ from variables to terms and term $m$, if these is map equality $\sigma \equiv \tau$, then there is equality $m[\sigma] \equiv m[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$, applying Lemmas \ref{convapp}, \ref{convlam}, \ref{convarrow}, \ref{convlolli}.
+\end{proof}
+
+\begin{lemma}
+  For terms $m_1, m_2, n$, if there is equality $m_1 \equiv m_2$, then there is equality $n[m_1/x] \equiv n[m_2/x]$ for any variable $x \in FV(n)$.
+\end{lemma}
+\begin{proof}
+  This is a special case of Lemma \ref{convcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+\end{proof}
+
+\subsection{Parallel Reduction Lemmas}
+
+\begin{definition}\label{psstep}
+  For maps $\sigma, \tau$ from variables to terms, we say $\sigma$ parallel reduces to $\tau$ if for any variable $x$ there exists a parallel reduction $(\sigma\ x) \pstep (\tau\ x)$.
+We write $\sigma \pstep \tau$ when it is clear from context that $\sigma, \tau$ are maps and not terms.
+\end{definition}
+
+\begin{lemma}\label{psteprefl}
+  For any term $m$, there exists a reflexive parallel reduction $m \pstep m$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $m$.
+\end{proof}
+
+\begin{lemma}
+  For any map $\sigma$ from variables to terms, there exists a reflexive parallel map reduction $\sigma \pstep \sigma$.
+\end{lemma}
+\begin{proof}
+  By Definition \ref{psstep} and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{steppstep}
+  For any terms $m, n$, if there exists step $m \step n$, then there exists a parallel reduction $m \pstep n$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step n$ and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{pstepred}
+  For terms $m, n$, if there exists parallel reduction $m \pstep n$, then there exists a reduction $m \red n$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$, utilizing the transitive property of $\red$ and Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}, \ref{redsubst}, \ref{redcompat}.
+\end{proof}
+
+\begin{lemma}
+  For terms $m, n$ and map $\sigma$ from variables to terms, if there exists parallel reduction $m \pstep n$, there exists parallel reduction $m[\sigma] \pstep n[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$ and Lemma \ref{psteprefl}.
+\end{proof}
+
+\begin{lemma}\label{pstepcompat}
+  For terms $m, n$ and maps $\sigma, \tau$ from variables to terms, if there exists parallel reduction $m \pstep n$ and parallel map reduction $\sigma \pstep \tau$, there exists parallel reduction $m[\sigma] \pstep n[\tau]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep n$.
+\end{proof}
+
+\begin{lemma}
+  For terms $m_1, m_2, n$, if there is parallel reduction $m_1 \pstep m_2$, then there is parallel reduction $n[m_1/x] \pstep n[m_2/x]$ for any variable $x \in FV(n)$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{psteprefl}, this is a special case of Lemma \ref{pstepcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+\end{proof}
+
+\subsection{Confluence Theorem}
+We first show that $\pstep$ satisfies the diamond property. Using the diamond property, we ultimately prove the confluence theorem.
+
+\begin{lemma}\label{diamond}
+  CLC term reduction has the diamond property.
+For terms $m, m_1, m_2$, if there are parallel reductions $m \pstep m_1$ and $m \pstep m_2$, then there exists term $m'$ such that $m_1 \pstep m'$ and $m_2 \pstep m'$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep m_1$.
+Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and apply the induction hypothesis.
+The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma \ref{pstepcompat} is used to push these cases through.
+\end{proof}
+
+\begin{lemma}\label{strip}
+  Strip lemma. For terms $m, m_1, m_2$, if there is parallel reduction $m \pstep m_1$ and reduction $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \pstep m'$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \pstep m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{diamond}.
+\end{proof}
+
+\begin{theorem}
+  CLC term reduction is confluent. For terms $m, m_1, m_2$, if there are reductions $m \red m_1$ and $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \red m'$.
+\end{theorem}
+\begin{proof}
+  By induction on the derivation of $m \red m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{strip}.
+\end{proof}
+
+\subsection{Corollaries of Confluence}
+The following results are all corollaries of confluence, proven using a combination of induction, transitivity and confluence.
+These corollaries allow us to refute false reductions and equalities in future proofs.
+
+\begin{corollary}\label{redsortinv}
+  For a universe $s_i$ and term $m$, if there is reduction $s_i \red m$, then $m = s_i$.
+\end{corollary}
+
+\begin{corollary}\label{redvarinv}
+  For variable $x$ and term $m$, if there is reduction $x \red m$, then $m = x$.
+\end{corollary}
+
+\begin{corollary}\label{redarrowinv}
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
+\end{corollary}
+
+\begin{corollary}\label{redlolliinv}
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
+\end{corollary}
+
+\begin{corollary}\label{redlaminv}
+  For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exists $A',m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
+\end{corollary}
+
+\begin{corollary}\label{sortinj}
+  For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there is $s = t$ and $i = j$.
+\end{corollary}
+
+\begin{corollary}\label{arrowinj}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{t} A_2) \rightarrow B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
+\end{corollary}
+
+\begin{corollary}\label{lolliinj}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{t} A_2) \multimap B_2$, then there are equalities $A_1 \equiv A_2$, $B_1 \equiv B_2$ and $s = t$.
+\end{corollary}
+
+\section{Context of CLC}
+Contexts of CLC are of the form $x_1 \stype{s1} A_1, x_2 \stype{s2} A_2, ...
+x_k \stype{sk} A_k$ where each free variable $x_i$ is assigned a type $A_i$ and sort $s_i$.
+Contexts will be referred to by meta variables $\Gamma$ and $\Delta$.
+
+\begin{figure}[h]
+  \begin{mathpar}
+    \inferrule
+    { }
+    { \epsilon \vdash }
+    \rname{Wf-$\epsilon$}
+
+    \inferrule
+    { \Gamma\ \vdash \\
+      \overline{\Gamma} \vdash A : U_i }
+    { \Gamma, x \utype A \vdash }
+    \rname{Wf-U}
+
+    \inferrule
+    { \Gamma\ \vdash \\
+      \overline{\Gamma} \vdash A : L_i }
+    { \Gamma, x \ltype A\ \vdash }
+    \rname{Wf-L}
+    \\
+
+    \inferrule
+    { }
+    { \pure{\epsilon} }
+    \rname{Pure-$\epsilon$}
+
+    \inferrule
+    { \pure{\Gamma} \\
+      \Gamma \vdash A : U_i }
+    { \pure{\Gamma, x \utype A} }
+    \rname{Pure-U}
+    \\
+
+    \inferrule
+    { }
+    { \mrg{\epsilon}{\epsilon}{\epsilon} }
+    \rname{Merge-$\epsilon$}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+    { \mrg{(\Gamma_1, x \utype A)}
+      {(\Gamma_2, x \utype A)}
+      {(\Gamma, x \utype A)} }
+    \rname{Merge-U}
+    \\
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_2 }
+    { \mrg{(\Gamma_1, x \ltype A)}
+      {\Gamma_2}
+      {(\Gamma, x \ltype A)} }
+    \rname{Merge-L1}
+
+    \inferrule
+    { \mrg{\Gamma_1}{\Gamma_2}{\Gamma} \\
+      x \notin \Gamma_1 }
+    { \mrg{\Gamma_1}
+      {(\Gamma_2, x \ltype A)}
+      {(\Gamma, x \ltype A)} }
+    \rname{Merge-L2}
+  \end{mathpar}
+  \label{struct}
+  \Description{}
+\end{figure}
+
+\subsection{Merge Lemmas}\label{mergeprop}
+Since weakening and contraction rules will not be allowed on restricted variables, it is necessary to have lemmas that enable the manipulation of contexts.
+
+\begin{lemma}\label{mergesym}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\mrg{\Gamma_2}{\Gamma_1}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure}
+  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\mrg{\Gamma}{\Gamma}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\pure{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergere1}
+  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\Gamma}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergere2}
+  For any context $\Gamma$, there is $\mrg{\Gamma}{\overline{\Gamma}}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergepureinv}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma}$, then there is $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure1}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_1}$, then there is $\Gamma = \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepure2}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$, then there is $\Gamma = \Gamma_1$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepurepure}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\pure{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergepureeq}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\Gamma_1 = \Gamma_2$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergerere}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}\label{mergererere}
+  For any context $\Gamma$, there is $\mrg{\overline{\Gamma}}{\overline{\Gamma}}{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\subsection{Restriction and Purity Lemmas}
+
+\begin{lemma}\label{rere}
+  For any context $\Gamma$, there is $\overline{\Gamma} = \overline{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{purere}
+  For any context $\Gamma$, if there is $\pure{\Gamma}$, then there is $\Gamma = \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{repure}
+  For any context $\Gamma$ , there is $\pure{\overline{\Gamma}}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{hasure}
+  For any context $\Gamma$, variable $x$ and type $A$, if there is $x \utype A \in \Gamma$, then there is $x \utype A \in \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $x \utype A \in \Gamma$.
+\end{proof}
+
+\begin{lemma}\label{haslre}
+  For any context $\Gamma$, variable $x$ and type $A$, there is $x \ltype A \notin \overline{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the structure of $\Gamma$.
+\end{proof}
+
+\begin{lemma}\label{mergesplit1}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\begin{lemma}
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_2}{\Gamma_2}{\Delta}$ and $\mrg{\Delta_1}{\Delta}{\Gamma}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$.
+\end{proof}
+
+\section{Subtyping of CLC}\label{subtyping}
+
+The cumulativity relation ($\preceq$) is the smallest binary relation over terms such that
+\begin{enumerate}
+  \item $\preceq$ is a partial order with respect to equality.
+        \begin{enumerate}
+          \item If $A \equiv B$, then $A \preceq B$.
+          \item If $A \preceq B$ and $B \preceq A$, then $A \equiv B$.
+          \item If $A \preceq B$ and $B \preceq C$, then $A \preceq B$.
+        \end{enumerate}
+  \item $U_0 \preceq U_1 \preceq U_2 \preceq \cdots$
+  \item $L_0 \preceq L_1 \preceq L_2 \preceq \cdots$
+  \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
+        $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{s} A_2) \rightarrow B_2$
+  \item If $A_1 \equiv A_2$ and $B_1 \preceq B_2$, \\ then
+        $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{s} A_2) \multimap B_2$
+\end{enumerate}
+
+\noindent
+Here, we give an inductive definition of the cumulativity relation ($\preceq$) that is suitable for writing proofs.
+
+\begin{mathpar}
+  \inferrule
+  { }
+  { A \prec A }
+  \rname{$\prec$-Refl}
+
+  \inferrule
+  { i_1 \leq i_2 }
+  { s_{i_1} \prec s_{i_2} }
+  \rname{$\prec$-Sort}
+
+  \inferrule
+  { B_1 \prec B_2 }
+  { (x \stype{s} A) \rightarrow B_1 \prec (x \stype{s} A) \rightarrow B_2 }
+  \rname{$\prec$$\rightarrow$}
+
+  \inferrule
+  { B_1 \prec B_2 }
+  { (x \stype{s} A) \multimap B_1 \prec (x \stype{s} A) \multimap B_2 }
+  \rname{$\prec$$\multimap$}
+
+  \inferrule
+  { A' \prec B' \\ A \equiv A' \\ B \equiv B' }
+  { A \preceq B }
+  \rname{$\prec$-$\preceq$}
+\end{mathpar}
+
+\subsection{Subtyping Lemmas}
+
+\begin{lemma}\label{sub1sub}
+  For terms $A, B$, if there is $A \prec B$, then there is $A \preceq B$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the reflexivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{sub1conv}
+  For terms $A, B, C$, if there is $A \prec B$ and $B \equiv C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{convsub1}
+  For terms $A, B, C$, if there is $A \equiv B$ and $B \prec C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By \rname{$\prec$-$\preceq$} and the transitivity of equality $\equiv$.
+\end{proof}
+
+\begin{lemma}\label{convsub}
+  For terms $A, B$, if there is $A \equiv B$, then there is $A \preceq B$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{convsub1} and \rname{$\prec$-Refl}.
+\end{proof}
+
+\begin{lemma}\label{subrefl}
+  For term $A$, there is $A \preceq A$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{sub1sub} and \rname{$\prec$-Refl}.
+\end{proof}
+
+\begin{lemma}\label{subprop}
+  For natural numbers $i, j$ and sort $s$ such that $i \leq j$, there is $s_i \preceq s_j$.
+\end{lemma}
+\begin{proof}
+  By Lemma \ref{sub1sub} and \rname{$\prec$-Sort}.
+\end{proof}
+
+\begin{lemma}\label{sub1trans}
+  For terms $A, B, C, D$, if there is $A \prec B$, $B \equiv C$ and $C \prec D$, then there is $A \preceq D$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $A \prec B$, definition of $\prec$ and Lemmas \ref{sub1sub}, \ref{sub1conv}, \ref{convsub1}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A, B, C$, if there is $A \preceq B$ and $B \preceq C$, then there is $A \preceq C$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$, rule \rname{$\prec$-$\preceq$} and Lemma \ref{sub1trans}.
+\end{proof}
+
+\begin{lemma}
+  For sorts $s, t$ and natural numbers $i, j$, if there is $s_i \preceq t_j$, then there is $s = t$ and $i \leq j$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{sortinj}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{t} A_2) \rightarrow B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{arrowinj}.
+\end{proof}
+
+\begin{lemma}
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{t} A_2) \multimap B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+\end{lemma}
+\begin{proof}
+  By transitivity of $\equiv$ and Corollary \ref{lolliinj}.
+\end{proof}
+
+\begin{lemma}\label{sub1subst}
+  For terms $A, B$  and map $\sigma$ from variables to terms, if there is $A \prec B$, then there is $A[\sigma] \prec B[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $A \prec B$ and the definition of $\prec$.
+\end{proof}
+
+\begin{lemma}
+  For terms $A, B$ and map $\sigma$ from variables to terms, if there is $A \preceq B$, then there is $A[\sigma] \preceq B[\sigma]$.
+\end{lemma}
+\begin{proof}
+  By rule \rname{$\prec$-$\preceq$} and Lemmas \ref{convsubst}, \ref{sub1subst}.
+\end{proof}
+
+\section{Typing of CLC}
+
+The following rules define well-formed contexts.
+\begin{mathpar}
+  \inferrule
+  { }
+  { \epsilon \vdash }
+  \rname{$\epsilon$-Ok}
+
+  \inferrule
+  { \Gamma \vdash \\
+    \overline{\Gamma} \vdash A : U_i }
+  { \Gamma, x \utype A \vdash }
+  \rname{U-Ok}
+
+  \inferrule
+  { \Gamma \vdash \\
+    \overline{\Gamma} \vdash A : L_i }
+  { \Gamma, x \ltype A \vdash }
+  \rname{L-Ok}
+\end{mathpar}
+
+\noindent
+The typing rules of CLC are presented below.
+\begin{mathpar}
+  \inferrule
+  { \pure{\Gamma} }
+  { \Gamma \vdash s_i : U_{i+1} }
+  \rname{Sort-Axiom}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : U_i \\
+    \Gamma, x \utype A \vdash B : s_i }
+  { \Gamma \vdash (x \utype A) \rightarrow B : U_i }
+  \rname{U$\rightarrow$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : L_i \\
+    \Gamma \vdash B : s_i \\
+    x \notin \Gamma }
+  { \Gamma \vdash (x \ltype A) \rightarrow B : U_i }
+  \rname{L$\rightarrow$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : U_i \\
+    \Gamma, x \utype A \vdash B : s_i }
+  { \Gamma \vdash (x \utype A) \multimap B : L_i }
+  \rname{U$\multimap$}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash A : L_i \\
+    \Gamma \vdash B : s_i \\
+    x \notin \Gamma }
+  { \Gamma \vdash (x \ltype A) \multimap B : L_i }
+  \rname{L$\multimap$}
+
+  \inferrule
+  { \pure{\Gamma_1, \Gamma_2} }
+  { \Gamma_1, x \utype A, \Gamma_2 \vdash x : A }
+  \rname{U-Var}
+
+  \inferrule
+  { \pure{\Gamma_1, \Gamma_2} }
+  { \Gamma_1, x \ltype A, \Gamma_2 \vdash x : A }
+  \rname{L-Var}
+
+  \inferrule
+  { \pure{\Gamma} \\
+    \Gamma \vdash (x \stype{s} A) \rightarrow B : t_i \\
+    \Gamma, x \stype{s} A \vdash n : B }
+  { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \rightarrow B }
+  \rname{$\lambda$$\rightarrow$}
+
+  \inferrule
+  { \overline{\Gamma} \vdash (x \stype{s} A) \multimap B : t_i \\
+    \Gamma, x \stype{s} A \vdash n : B }
+  { \Gamma \vdash \lambda x \stype{s} A . n : (x \stype{s} A) \multimap B }
+  \rname{$\lambda$$\multimap$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \utype A) \rightarrow B \\
+    \pure{\Gamma_2} \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-U$\rightarrow$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \ltype A) \rightarrow B \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-L$\rightarrow$}
+\end{mathpar}
+
+\begin{mathpar}
+  \inferrule
+  { \Gamma_1 \vdash m : (x \utype A) \multimap B \\
+    \pure{\Gamma_2} \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-U$\multimap$}
+
+  \inferrule
+  { \Gamma_1 \vdash m : (x \ltype A) \multimap B \\
+    \Gamma_2 \vdash n : A \\
+    \mrg{\Gamma_1}{\Gamma_2}{\Gamma} }
+  { \Gamma \vdash m\ n : B[n/x] }
+  \rname{App-L$\multimap$}
+
+  \inferrule
+  { \Gamma \vdash m : A \\
+    \overline{\Gamma} \vdash B : s_i \\ A \preceq B }
+  { \Gamma \vdash m : B }
+  \rname{Conversion}
+\end{mathpar}
+
+\section{Inversion Lemmas of CLC}
 
@@
 
-\section{Weakening Lemmas of CLC \texttt{(clc_weakening.v)}}
+\section{Weakening Lemmas of CLC}
 
@@
 
-\subsection{Properties of $agreeR$}\label{agreeRprop}
+\subsection{Properties of agreeR}\label{agreeRprop}
 
@@
 
-\section{Substitution Lemmas of CLC \texttt{(clc_substitution.v)}}
+\section{Substitution Lemmas of CLC}
 
@@
 
-\subsection{Properties of $agreeS$}\label{agreeSprop}
+\subsection{Properties of agreeS}\label{agreeSprop}
 
@@
 
-\section{Typing Validity of CLC \texttt{(clc_validity.v)}}
+\section{Typing Validity of CLC}
 
@@
 
-\begin{theorem}\label{validity}
+\begin{theorem}\label{valid}
   The validity of typing theorem.
@@
 
-\section{Subject Reduction of CLC \texttt{(clc_soundness.v)}}
+\section{Subject Reduction of CLC}
 
@@
 
-\section{Linearity Theorems of CLC \texttt{(clc_linearity.v)}}
+\section{Linearity Theorems of CLC}
 
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{validity} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
+    \item If $s = U$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
 Now applying Lemma \ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
 Now by \rname{U$\rightarrow$} and Lemma \ref{purere} the goal is proven.
-    \item If $s = L$, then we may apply Theorem \ref{validity} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
+    \item If $s = L$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
 Now applying Lemma \ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{validity} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
+    \item If $s = U$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
 Now applying Lemma \ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
@@
 By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
-    \item If $s = L$, then we may apply Theorem \ref{validity} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
+    \item If $s = L$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
 Now applying Lemma \ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
@@
 
-\section{Logical Consistency of CLC \texttt{clc_consistent.v}}
+\section{Logical Consistency of CLC}
 

### 2022-01-17 writing/draft7/draft7.tex
draft

@@
 \appendix
+\appendixname
 
@@
 
-\begin{theorem}\label{embedding}
+\begin{theorem}\label{embed}
   Embedding.
@@
 \begin{proof}
-  Direct consequence of applying Theorem \ref{embedding} to all types in context $\Gamma$.
+  Direct consequence of applying Theorem \ref{embed} to all types in context $\Gamma$.
 \end{proof}
@@
   Suppose there exists a well-typed CLC term $m$ with an infinite sequence of reductions.
-Theorem \ref{embedding} shows that there must exist some term $\erase{m}$ that is well-typed in CC$\omega$.
+Theorem \ref{embed} shows that there must exist some term $\erase{m}$ that is well-typed in CC$\omega$.
 Additionally, this infinite sequence of reductions on $m$ can be translated in CC$\omega$ step-wise by Lemma \ref{erasestep}.
@@
 
-\subsection{Embedding of CC$\omega$}
+\subsection{Embedding}
 To show that CLC is compatible with the predicative fragment of CC$\omega$, we construct a lifting procedure that lifts CC$\omega$ terms into CLC in a straightforward way.

### 2022-01-17 writing/draft7/draft7.tex
draft

@@
 We add an extra universe sort $L$ of linear types with cumulativity parallel to the sort $U$ of non-linear types.
-This ultimately cumulates in the \textit{linearity} theorem, stating that all resources are used exactly once.
+This ultimately cumulates in the \textit{linearity} theorem, stating that every resource is used exactly once.
 
@@
 exponential or separate typing judgments for linear and non-linear types.
-The universe sorts of CLC is enough to completely distinguish between linear and non-linear types within a single typing judgment.
+The universe sorts of CLC are enough to completely distinguish between linear and non-linear types within a single typing judgment.
 As a result, CLC falls much closer to its CC roots than prior works in both syntax and semantics.
@@
 
-Through examples in Section \ref{examples}, we show some applications of CILC such as construction of safe random access array from the first principles, protocol enforcement through inductively defined session types and in-place update of data structures.
+Through examples in Section \ref{examples}, we show some applications of CILC such as construction of safe random access array from the first principles, protocol enforcement through inductively defined session types and in-place update of data structures using continuations.
 
@@
 
-A clear departure of our language from standard presentations of both linear type theory and dependent type theory is the presence of two function types: $(x :_s A) \rightarrow B$, $(x :_s A) \multimap B$.
+A clear departure of our language from standard presentations of both linear type theory and dependent type theory is the presence of two function types: $(x :_s A) \rightarrow B$ and $(x :_s A) \multimap B$.
 The reason for these function types is that we have built the !
@@
 $U_i$ and $L_i$ are the predicative universes of non-linear types and linear types respectively.
-The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universe to which they belong.
+The main mechanism that CLC uses to distinguish between linear and non-linear types is by checking the universes to which they belong.
 Basically, terms with types that occur within $U_i$ are unrestricted in their usage.
@@
     \item A \textit{type} $A$ is \textit{linear} under context $\Gamma$ if $\Gamma \vdash A : L_k$.
-    \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$, such that $\Gamma \vdash m : A$.
-An unrestricted term may be used any number of times.
-    \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$, such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
+    \item A \textit{term} $m$ is \textit{unrestricted} under context $\Gamma$ if there exists non-linear type $A$ such that $\Gamma \vdash m : A$. An unrestricted term may be used any number of times.
+    \item A \textit{term} $m$ is \textit{restricted} under context $\Gamma$ if there exists linear type $A$ such that $\Gamma \vdash m : A$. A restricted term must be used exactly once.
   \end{enumerate}
@@
 
-The axiom rules \rname{U-Axiom}, \rname{L-Axiom} are similar to standard CC$\omega$, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+The axiom rules \rname{U-Axiom} and \rname{L-Axiom} are similar to standard CC$\omega$, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
 In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
@@
 To support linear types, we must restrict weakening to non-linear variables.
-This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{U-Axiom}, \rname{L-Axiom}.
+This justifies the restriction of $\Gamma$ to contain only non-linear variables for \rname{U-Axiom} and \rname{L-Axiom}.
 From \rname{L-Axiom} we can see that the universe of linear types $L_i$ is an inhabitant of $U_{i+1}$.
@@
 The non-linearity of domain $A$ allows $B$ to depend on terms of type $A$, as seen in judgment $\Gamma, x \utype A \vdash B : s_i$.
-The domain $B$ itself may be non-linear or linear, since $B$'s universe $s_k$ can vary between $U_i$ and $L_i$.
+The domain $B$ itself may be linear of non-linear, since $B$'s universe $s_k$ can vary between $U_i$ and $L_i$.
 From the resulting judgment $\Gamma \vdash (x \utype A) \rightarrow B : U_i$, we see that the overall type is a non-linear type.
@@
 Because of this, we forbid co-domain $B$ from depending on terms of type $A$, evidently in the judgment $\Gamma \vdash B : s_i$.
-Like \rname{U$\rightarrow$}, co-domain $B$ itself may be non-linear or linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
+Like \rname{U$\rightarrow$}, co-domain $B$ itself may be linear or non-linear, since $B$'s universe $s_i$ can vary between $U_i$ and $L_i$.
 The final resulting judgment is $\Gamma \vdash (x \ltype A) \rightarrow B : U_i$.
@@
 
-The rules \rname{U-Var}, and \rname{L-Var} are used for typing free variables.
+The rules \rname{U-Var} and \rname{L-Var} are used for typing free variables.
 The \rname{U-Var} rule asserts that free variable $x$ occurs within the context $\Gamma_1, x \utype A, \Gamma_2$ with a non-linear type $A$.
@@
 In contrast to \rname{$\lambda$$\rightarrow$}, the \rname{$\lambda$$\multimap$} rule is used for forming abstractions that must be used exactly once.
-Due to the fact that these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evidently in the judgment $\Gamma, x \stype{s} A \vdash n : B$, and lack of side condition $\pure{\Gamma}$.
+Due to the fact that these abstractions must be used once, they are allowed access to restricted variables within context $\Gamma$, evidently in the judgment $\Gamma, x \stype{s} A \vdash n : B$ and lack of side condition $\pure{\Gamma}$.
 However, in judgment $\overline{\Gamma} \vdash (x \stype{s} A) \multimap B : L_k$ the context must be filtered, because types are not allowed to depend on restricted variables.
@@
 
-The proof of confluence is entirely standard using parallel step technique \cite{takahashi}. The presence of linear types do not pose any complications as reductions are untyped.
+The proof of confluence is entirely standard using parallel step technique \cite{takahashi}. The presence of linear types does not pose any complications as reductions are untyped.
 
@@
 
-Notice in Theorem \ref{nlsubst} of non-linear substitutions, there exists an extra condition $\pure{\Gamma_2}$ which Theorem \ref{lsubst} of linear substitutions lack.
+Notice in Theorem \ref{nlsubst} of non-linear substitutions, there exists an extra condition $\pure{\Gamma_2}$ which Theorem \ref{lsubst} of linear substitutions lacks.
 This condition ensures that linear variables do not occur within the substituted term, thus cannot be duplicated by substitution.
@@
 
-If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve} and \ref{embedding}.
+If there exists some well-typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve} and \ref{embedding}.
 This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
@@
 
-The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domain of function types are annotated with sort $U$.
+The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domains of function types are annotated with sort $U$.
 In practice, these annotations can often be omitted if algorithmic type checking is able to infer them from context.
@@
 
-Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not applied to restricted terms as this may cause duplication.
+Non-linear constructors are used for constructing non-linear objects that may be freely used. These constructors may not be applied to restricted terms as this may cause duplication.
 
@@
 This is the same non-sharing principle that empowers Separation Logic.
-Unlike Separation Logic, $(l \mapsto A)$ itself is a first-class object in CILC, meaning that it can be the result of computation (large eliminations), input to computations, stored in data structures, etc.
+Unlike Separation Logic, $(l \mapsto A)$ itself is a first-class object in CILC, meaning that it can be the result of computations (large eliminations), input to computations and storage in data structures, etc.
 
@@
 The capability $(l + m \mapsto A)$ grants the ability to manipulate the store at offset $m$ using the state axioms.
-The function $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$ is constructed using the rest of $\arrv\ A\ l\ n$ not at offset $m$.
+The function $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$ is constructed using the rest of the capabilities in $\arrv\ A\ l\ n$ that are not at offset $m$.
 Intuitively, $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$ is the original array proof with a missing capability at offset $m$ which can be restored by applying capability $(l + m \mapsto A)$.
@@
 An indexing function can now be constructed as a wrapper for $\texttt{nth}$ as shown in Figure \ref{index}.
-Given an array $\texttt{Array}\ A\ n$, an offset $m$ and proof that $m < n$, $\texttt{nth}$ can be used to extract out the capability at offset $m$, allowing $\get$ to retrieve the stored value.
+Given an array $\texttt{Array}\ A\ n$, an offset $m$ and proof that $m < n$, $\texttt{nth}$ can be used to extract the capability at offset $m$, allowing $\get$ to retrieve the stored value.
 Finally, the array proof is restored and returned alongside the indexing result.
-This definition of $\texttt{index}$ is safe because the proof $m < n$ ensures that accessing offsets are always within bounds.
+This definition of $\texttt{index}$ is safe because the proof $m < n$ ensures that offsets used for indexing are always within bounds.
 
@@
     \item $\textbf{close} : \channel\ \END \rightarrow \unit$ \\
-          Once all communications specified on a given channel has finished, $\close$ will close and reclaim the channel.
+          Once all communications specified on a given channel have finished, $\close$ will close and reclaim the channel.
     \item $\textbf{send} : (A : U) \rightarrow A \rightarrow (ss : \session) \rightarrow$ \\
@@
 The important detail to notice here is that each channel created by $\open$ is of linear type $(\channel\ ss)$ where $ss$ is some arbitrary protocol, hence must be used exactly once.
-However, the $\send$ and $\recv$ operations return back channels after consuming them, progressing their protocols one stage forward.
+However, the $\send$ and $\recv$ operations return channels after consuming them, progressing their protocols one stage forward.
 Coupled with the fact that the $\close$ operation is only able to close channels whose protocols have ended, this forces channels to communicate strictly by their specified protocols.
@@
 
+\clearpage
 \appendix
@@
   By induction on the derivation of $m \pstep m_1$.
-Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and apply the induction hypothesis.
+Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and then to apply the induction hypothesis.
 The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma \ref{pstepcompat} is used to push these cases through.

### 2022-01-17 writing/draft7/draft7.tex
draft

@@
 First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
-Next, we show that CLC is an effective linear type theory through the \textit{linearity} theorem.
-Furthermore, we show that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
-Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, showing that CLC is strongly normalizing.
+Next, through the \textit{linearity} theorem we illustrate how CLC can effectively control usage of resource.
+Furthermore, we demonstrate that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
+Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, proving that CLC is strongly normalizing.
 
 All proofs have been formalized in Coq with help from the Autosubst \cite{autosubst} library.
-The Coq development is publicly available on the first author's Github repository.
-To the best of our knowledge, this is the first machined checked formalization of a linear dependently type theory.
+The Coq development is publicly available on the author's Github repository.
+To the best of our knowledge, this is the first machine checked formalization of a linear dependently type theory.
 We give a hand written version of the proof in an accompanying technical report.
@@
 
-The proof of confluence is entirely standard using parallel step technique \cite{takahashi}. The presence of linear types does not pose any complications as reductions are untyped.
+The proof of confluence is entirely standard using parallel reduction technique \cite{takahashi}. The presence of linear types does not pose any complications as reductions are untyped.
 
@@
 
-An indexing function can now be constructed as a wrapper for $\texttt{nth}$ as shown in Figure \ref{index}.
+An indexing function can now be constructed as a wrapper for $\texttt{nth}$ with its signature shown in Figure \ref{index}.
 Given an array $\texttt{Array}\ A\ n$, an offset $m$ and proof that $m < n$, $\texttt{nth}$ can be used to extract the capability at offset $m$, allowing $\get$ to retrieve the stored value.
@@
   (a : Array A n) 
-: ?$\SigmaL$? _ : A. Array A n
+: ?$\SigmaL$? _ : A. Array A n.
   \end{minted}
@@
   \caption{Inductive Session Type}
-  \begin{minted}{coq}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
 Inductive session : U :=
-| SEND : U -> session -> session
-| RECV : U -> session -> session
+| SEND : U ?$\rightarrow$? session ?$\rightarrow$? session
+| RECV : U ?$\rightarrow$? session ?$\rightarrow$? session
 | END : session.
@@
   \caption{Linear Lists}
-  \begin{minted}{coq}
+  \begin{minted}[escapeinside=??,mathescape=true]{coq}
 Inductive list (A : U) : L :=
 | nil : list A
-| cons : A -> list A -> list A.
-
-Fixpoint append (A : U) : list A -o 
-  list A -o (list A -o list A) -o list A 
+| cons : A ?$\rightarrow$? list A ?$\rightarrow$? list A.
+
+Fixpoint append (A : U) : list A ?$\multimap$? 
+  list A ?$\multimap$? (list A ?$\multimap$? list A) ?$\multimap$? list A 
 :=
-  fun ls1 ls2 k =>
+  fun ls1 ls2 k ?$\Rightarrow$?
     match ls1 with
-    | nil => k ls2
-    | (cons h t) as _cons_ =>
+    | nil ?$\Rightarrow$? k ls2
+    | (cons h t) as _cons_ ?$\Rightarrow$?
       append _ t ls2 
-        (fun res => k (_cons_ h res))
+        (fun res ?$\Rightarrow$? k (_cons_ h res))
     end.
@@
 \begin{lemma}\label{convapp}
-  For terms $m_1, m_2, n_1, n_2$, if there exists equalities $m_1 \equiv m_2$ and $n_1 \equiv n_2$, then there exists equality $(m_1\ n_1) \equiv (m_2\ n_2)$.
+  For terms $m_1, m_2, n_1, n_2$, if there exist equalities $m_1 \equiv m_2$ and $n_1 \equiv n_2$, then there exists equality $(m_1\ n_1) \equiv (m_2\ n_2)$.
 \end{lemma}
@@
 \begin{lemma}\label{convlam}
-  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $m_1 \equiv m_2$, then there exists equality $\lambda x \stype{s} A_1.m_1 \equiv \lambda x \stype{s} A_2.m_2$.
+  For terms $A_1, A_2, m_1, m_2$ and sort $s$, if there exist equalities $A_1 \equiv A_2$ and $m_1 \equiv m_2$, then there exists equality $\lambda x \stype{s} A_1.m_1 \equiv \lambda x \stype{s} A_2.m_2$.
 \end{lemma}
@@
 \begin{lemma}\label{convarrow}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{s} A_2) \rightarrow B_2$.
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exist equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \rightarrow B_1 \equiv (x \stype{s} A_2) \rightarrow B_2$.
 \end{lemma}
@@
 \begin{lemma}\label{convlolli}
-  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exists equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{s} A_2) \multimap B_2$.
+  For terms $A_1, A_2, B_1, B_2$ and sort $s$, if there exist equalities $A_1 \equiv A_2$ and $B_1 \equiv B_2$, then there exists equality $(x \stype{s} A_1) \multimap B_1 \equiv (x \stype{s} A_2) \multimap B_2$.
 \end{lemma}
@@
 \begin{lemma}\label{pstepcompat}
-  For terms $m, n$ and maps $\sigma, \tau$ from variables to terms, if there exists parallel reduction $m \pstep n$ and parallel map reduction $\sigma \pstep \tau$, there exists parallel reduction $m[\sigma] \pstep n[\tau]$.
+  For terms $m, n$ and maps $\sigma, \tau$ from variables to terms, if there exist parallel reduction $m \pstep n$ and parallel map reduction $\sigma \pstep \tau$, there exists parallel reduction $m[\sigma] \pstep n[\tau]$.
 \end{lemma}
@@
 \begin{lemma}\label{strip}
-  Strip lemma. For terms $m, m_1, m_2$, if there is parallel reduction $m \pstep m_1$ and reduction $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \pstep m'$.
+  Strip lemma. For terms $m, m_1, m_2$, if there are parallel reductions $m \pstep m_1$ and $m \red m_2$, then there exists term $m'$ such that $m_1 \red m'$ and $m_2 \pstep m'$.
 \end{lemma}
@@
 \begin{corollary}\label{redarrowinv}
-  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \rightarrow B \red m$, then there exist $A'$ and $B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \rightarrow B'$.
 \end{corollary}
@@
 \begin{corollary}\label{redlolliinv}
-  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exists $A', B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
+  For terms $A, B, m$ and sort $s$, if there is reduction $(x \stype{s} A) \multimap B \red m$, then there exist $A'$ and $B'$ such that there are reductions $A \red A'$, $B \red B'$ and $m = (x \stype{s} A') \multimap B'$.
 \end{corollary}
@@
 \begin{corollary}\label{redlaminv}
-  For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exists $A',m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
+  For terms $A, m, n$ and sort $s$, if there is reduction $\lambda x \stype{s} A.m \red n$, then there exist $A'$ and $m'$ such that there are reductions $A \red A'$, $m \red m'$ and $n = \lambda x \stype{s} A'.m'$.
 \end{corollary}
@@
 \begin{corollary}\label{sortinj}
-  For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there is $s = t$ and $i = j$.
+  For sorts $s, t$ and levels $i, j$, if there is equality $s_i \equiv t_j$, then there are $s = t$ and $i = j$.
 \end{corollary}
@@
 \begin{lemma}\label{mergepureinv}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma}$, then there is $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma}$, then there are $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$.
 \end{lemma}
@@
 \begin{lemma}\label{mergepure1}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_1}$, then there is $\Gamma = \Gamma_2$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_1}$, then there is $\Gamma = \Gamma_2$.
 \end{lemma}
@@
 \begin{lemma}\label{mergepure2}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$, then there is $\Gamma = \Gamma_1$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ , if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$, then there is $\Gamma = \Gamma_1$.
 \end{lemma}
@@
 \begin{lemma}\label{mergepurepure}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\pure{\Gamma}$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$, then there is $\pure{\Gamma}$.
 \end{lemma}
@@
 \begin{lemma}\label{mergepureeq}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, and also $\pure{\Gamma_1}$, $\pure{\Gamma_2}$, then there is $\Gamma_1 = \Gamma_2$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, $\pure{\Gamma_1}$ and $\pure{\Gamma_2}$, then there is $\Gamma_1 = \Gamma_2$.
 \end{lemma}
@@
 \begin{lemma}\label{mergerere}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there is $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there are $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
 \end{lemma}
@@
 \begin{lemma}\label{mergesplit1}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$ are valid, then there exists $\Delta$ such that $\mrg{\Delta_1}{\Gamma_2}{\Delta}$ and $\mrg{\Delta}{\Delta_2}{\Gamma}$ are valid.
 \end{lemma}
@@
 \begin{lemma}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$, then there exists $\Delta$ such that $\mrg{\Delta_2}{\Gamma_2}{\Delta}$ and $\mrg{\Delta_1}{\Delta}{\Gamma}$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma, \Delta_1, \Delta_2$, if $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\mrg{\Delta_1}{\Delta_2}{\Gamma_1}$ are valid, then there exists $\Delta$ such that $\mrg{\Delta_2}{\Gamma_2}{\Delta}$ and $\mrg{\Delta_1}{\Delta}{\Gamma}$ are valid.
 \end{lemma}
@@
 \begin{lemma}\label{sub1conv}
-  For terms $A, B, C$, if there is $A \prec B$ and $B \equiv C$, then there is $A \preceq C$.
+  For terms $A, B, C$, if there are $A \prec B$ and $B \equiv C$, then there is $A \preceq C$.
 \end{lemma}
@@
 \begin{lemma}\label{convsub1}
-  For terms $A, B, C$, if there is $A \equiv B$ and $B \prec C$, then there is $A \preceq C$.
+  For terms $A, B, C$, if there are $A \equiv B$ and $B \prec C$, then there is $A \preceq C$.
 \end{lemma}
@@
 \begin{lemma}\label{sub1trans}
-  For terms $A, B, C, D$, if there is $A \prec B$, $B \equiv C$ and $C \prec D$, then there is $A \preceq D$.
+  For terms $A, B, C, D$, if there are $A \prec B$, $B \equiv C$ and $C \prec D$, then there is $A \preceq D$.
 \end{lemma}
@@
 \begin{lemma}
-  For terms $A, B, C$, if there is $A \preceq B$ and $B \preceq C$, then there is $A \preceq C$.
+  For terms $A, B, C$, if there are $A \preceq B$ and $B \preceq C$, then there is $A \preceq C$.
 \end{lemma}
@@
 \begin{lemma}
-  For sorts $s, t$ and natural numbers $i, j$, if there is $s_i \preceq t_j$, then there is $s = t$ and $i \leq j$.
+  For sorts $s, t$ and natural numbers $i, j$, if there is $s_i \preceq t_j$, then there are $s = t$ and $i \leq j$.
 \end{lemma}
@@
 \begin{lemma}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{t} A_2) \rightarrow B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \rightarrow B_1 \preceq (x \stype{t} A_2) \rightarrow B_2$, then there are $A_1 \equiv A_2$, $B_1 \preceq B_2$ and $s = t$.
 \end{lemma}
@@
 \begin{lemma}
-  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{t} A_2) \multimap B_2$, then there is $A_1 \equiv A_2$ and $B_1 \preceq B_2$ and $s = t$.
+  For terms $A_1, A_2, B_1, B_2$ and sorts $s, t$, if there is $(x \stype{s} A_1) \multimap B_1 \preceq (x \stype{t} A_2) \multimap B_2$, then there are $A_1 \equiv A_2$, $B_1 \preceq B_2$ and $s = t$.
 \end{lemma}
@@
 \begin{lemma}\label{uarrowinv}
-  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \rightarrow B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \rightarrow B : s$, then there exist sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
 \end{lemma}
@@
 \begin{lemma}\label{larrowinv}
-  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\rightarrow B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\rightarrow B : s$, then there exist sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
 \end{lemma}
@@
 \begin{lemma}\label{ulolliinv}
-  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \multimap B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \utype A) \multimap B : s$, then there exist sort $t$ and natural number $i$ such that $\Gamma \vdash A : U_i$ and $\Gamma, x \utype A \vdash B : t_i$.
 \end{lemma}
@@
 \begin{lemma}\label{llolliinv}
-  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\multimap B : s$, then there exists sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
+  For any context $\Gamma$ and terms $A, B, s$, if there is $\Gamma \vdash (x \ltype A)\multimap B : s$, then there exist sort $t$ and natural number $i$ such that $\Gamma \vdash A : L_i$ and $\Gamma \vdash B : t_i$.
 \end{lemma}
@@
 \begin{lemma}\label{arrowlaminv}
-  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there is $\overline{\Gamma} \vdash (x \stype{s'} A') \rightarrow B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
+  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there are $\overline{\Gamma} \vdash (x \stype{s'} A') \rightarrow B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
 n : (x \stype{s'} A') \rightarrow B$, then there is $\Gamma, x \stype{s'} A' \vdash n : B$.
@@
 \begin{lemma}\label{lollilaminv}
-  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there is $\overline{\Gamma} \vdash (x \stype{s'} A') \multimap B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
+  For any context $\Gamma$, terms $A, A', B, n$, sorts $s, s', t$ and natural number $i$, if there are $\overline{\Gamma} \vdash (x \stype{s'} A') \multimap B : t_i$ and $\Gamma \vdash \lambda x \stype{s} A .
 n : (x \stype{s'} A') \multimap B$, then there is $\Gamma, x \stype{s'} A' \vdash n : B$.
@@
 \begin{lemma}\label{agreerenpure}
-  For contexts $\Gamma, \Gamma'$ and mapping $\xi$, if there is $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\pure{\Gamma}$, then there is $\pure{\Gamma'}$.
+  For contexts $\Gamma, \Gamma'$ and mapping $\xi$, if there are $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\pure{\Gamma}$, then there is $\pure{\Gamma'}$.
 \end{lemma}
@@
 \begin{lemma}\label{mergeagreereninv}
-  For contexts $\Gamma, \Gamma', \Gamma_1, \Gamma_2$ and mapping $\xi$, if there is $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exists $\Gamma_1', \Gamma_2'$ such that $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+  For contexts $\Gamma, \Gamma', \Gamma_1, \Gamma_2$ and mapping $\xi$, if there are $agreeR\ \xi\ \Gamma\ \Gamma'$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exist $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 \end{lemma}
@@
 \begin{lemma}\label{renaming}
-  For context $\Gamma, \Gamma'$, terms $m, A$ and mapping $\xi$, if there is $\Gamma \vdash m : A$ and $agreeR\ \xi\ \Gamma\ \Gamma'$, then there is $\Gamma' \vdash m[\xi] : A[\xi]$.
+  For context $\Gamma, \Gamma'$, terms $m, A$ and mapping $\xi$, if there are $\Gamma \vdash m : A$ and $agreeR\ \xi\ \Gamma\ \Gamma'$, then there is $\Gamma' \vdash m[\xi] : A[\xi]$.
 \end{lemma}
@@
   \begin{itemize}
-    \item For the \rname{App-U$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-U$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
 At this point, the induction hypothesis allows us to apply \rname{App-U$\rightarrow$} to prove the goal.
-    \item For the \rname{App-L$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-L$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 At this point, the induction hypothesis allows us to apply \rname{App-L$\rightarrow$} to prove the goal.
-    \item For the \rname{App-U$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-U$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
 At this point, the induction hypothesis allows us to apply \rname{App-U$\multimap$} to prove the goal.
-    \item For the \rname{App-L$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there is $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$ and $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-L$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 At this point, the induction hypothesis allows us to apply \rname{App-L$\multimap$} to prove the goal.
@@
 \begin{proof}
-  Using \rname{agreeR-wk} and Lemma \ref{agreerenrefl} a proof of $agreeR\ id\ \Gamma\ (\Gamma, x \utype B)$ can be constructed. Then by Lemma \ref{renaming}, the theorem can be proven.
+  Using \rname{agreeR-wk} and Lemma \ref{agreerenrefl} a proof of the relation $agreeR\ id\ \Gamma\ (\Gamma, x \utype B)$ can be constructed. Then by Lemma \ref{renaming}, the theorem can be proven.
 \end{proof}
@@
 
-Similar to the proof of weakening, we first define an $agreeS$ relation between two contexts $\Gamma, \Delta$ and a mapping $\sigma$ from variables to terms.
+Similarly to the proof of weakening, we first define an $agreeS$ relation between two contexts $\Gamma, \Delta$ and a mapping $\sigma$ from variables to terms.
 
@@
   \inferrule
-  { agreeS\ \sigma\ \Delta\ \Gamma \\
-    \overline{\Delta} \vdash n : A[\sigma] \\
-    x \notin FV(\Delta) \cup FV(\Gamma) }
+  { x \notin FV(\Delta) \cup FV(\Gamma) \\\\
+    agreeS\ \sigma\ \Delta\ \Gamma \\
+    \overline{\Delta} \vdash n : A[\sigma] }
   { agreeS\ (\sigma \cup (x, n))\ \Delta\ (\Gamma, x \utype A) }
@@
   \inferrule
-  { \mrg{\Delta_1}{\Delta_2}{\Delta} \\
+  { x \notin FV(\Delta) \cup FV(\Gamma) \\
+    \mrg{\Delta_1}{\Delta_2}{\Delta} \\\\
     agreeS\ \sigma\ \Delta_1\ \Gamma \\
-    \Delta_2 \vdash n : A[\sigma] \\
-    x \notin FV(\Delta) \cup FV(\Gamma) }
+    \Delta_2 \vdash n : A[\sigma] }
   { agreeS\ (\sigma \cup (x, n))\ \Delta\ (\Gamma, x \ltype A) }
@@
   \inferrule
-  { A \preceq B \\
-    \overline{\Delta} \vdash B[\sigma] : U_i \\
-    agreeS\ \sigma\ \Delta\ (\Gamma, x \utype A) }
+  { agreeS\ \sigma\ \Delta\ (\Gamma, x \utype A) \\\\
+    A \preceq B \\
+    \overline{\Delta} \vdash B[\sigma] : U_i }
   { agreeS\ \sigma\ \Delta\ (\Gamma, x \utype B) }
@@
   { A \preceq B \\
+    agreeS\ \sigma\ \Delta\ (\Gamma, x \ltype A) \\\\
     \overline{\Delta} \vdash B[\sigma] : L_i \\
-    \overline{\Gamma} \vdash B : L_i \\
-    agreeS\ \sigma\ \Delta\ (\Gamma, x \ltype A) }
+    \overline{\Gamma} \vdash B : L_i }
   { agreeS\ \sigma\ \Delta\ (\Gamma, x \ltype B) }
@@
 \begin{lemma}\label{mergeagreesubstinv}
-  For contexts $\Delta, \Gamma, \Gamma_1, \Gamma_2$ and mapping $\sigma$, if there is $agreeS\ \sigma\ \Delta\ \Gamma$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exists contexts $\Delta_1, \Delta_2$ such that $\mrg{\Delta_1}{\Delta_2}{\Delta}$ and $agreeS\ \sigma\ \Delta_1\ \Gamma_1$ and $agreeS\ \sigma\ \Delta_2\ \Gamma_2$.
+  For contexts $\Delta, \Gamma, \Gamma_1, \Gamma_2$ and mapping $\sigma$, if there are $agreeS\ \sigma\ \Delta\ \Gamma$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then there exist contexts $\Delta_1, \Delta_2$ such that there are $\mrg{\Delta_1}{\Delta_2}{\Delta}$, $agreeS\ \sigma\ \Delta_1\ \Gamma_1$ and $agreeS\ \sigma\ \Delta_2\ \Gamma_2$.
 \end{lemma}
@@
   Generalized Substitution Lemma.
-For context $\Gamma, \Delta$, terms $m, A$ and mapping $\sigma$, if there is $\Gamma \vdash m : A$ and $agreeS\ \sigma\ \Delta\ \Gamma$, then there is $\Delta \vdash m[\sigma] : A[\sigma]$.
+For context $\Gamma, \Delta$, terms $m, A$ and mapping $\sigma$, if there are $\Gamma \vdash m : A$ and $agreeS\ \sigma\ \Delta\ \Gamma$, then there is $\Delta \vdash m[\sigma] : A[\sigma]$.
 \end{lemma}
@@
   The proof proceeds by induction on the derivation of $\Gamma \vdash m : A$.
-Similar to the proof of Lemma \ref{renaming}, the interesting cases are the application cases where Lemma \ref{mergeagreesubstinv} must be utilized to split the $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ judgments for use in the induction hypothesis.
+Similarly to the proof of Lemma \ref{renaming}, the interesting cases are the application cases where Lemma \ref{mergeagreesubstinv} must be utilized to split the $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ judgments for use in the induction hypothesis.
 \end{proof}
@@
 \begin{corollary}\label{substitutionu}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there is $\Gamma_1, x \utype A \vdash m : B$ and $\pure{\Gamma_2}$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there are $\Gamma_1, x \utype A \vdash m : B$, $\pure{\Gamma_2}$, $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
 \end{corollary}
@@
 \begin{corollary}\label{substitutionl}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there is $\Gamma_1, x \ltype A \vdash m : B$ and $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$ and terms $A, B, m, n$, if there are $\Gamma_1, x \ltype A \vdash m : B$, $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma_2 \vdash n : A$, then there is $\Gamma \vdash m[n/x]: B[n/x]$.
 \end{corollary}
@@
 \begin{corollary}\label{contextconvu}
-  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there is $B \equiv A$ and $\overline{\Gamma} \vdash A : U_i$ and $\Gamma, x \utype A \vdash m : C$, then there is $\Gamma, x \utype B \vdash m : C$.
+  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there are $B \equiv A$, $\overline{\Gamma} \vdash A : U_i$ and $\Gamma, x \utype A \vdash m : C$, then there is $\Gamma, x \utype B \vdash m : C$.
 \end{corollary}
@@
 \begin{corollary}\label{contextconvl}
-  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there is $B \equiv A$ and $\overline{\Gamma} \vdash A : L_i$ and $\Gamma, x \ltype A \vdash m : C$, then there is $\Gamma, x \ltype B \vdash m : C$.
+  For context $\Gamma$, terms $m, A, B, C$ and natural number $i$, if there are $B \equiv A$, $\overline{\Gamma} \vdash A : L_i$ and $\Gamma, x \ltype A \vdash m : C$, then there is $\Gamma, x \ltype B \vdash m : C$.
 \end{corollary}
@@
 \begin{lemma}\label{mergecontextokinv}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$, then there is $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there are $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$, then there are $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
 \end{lemma}
@@
   The validity of typing theorem.
-For any context $\Gamma$ and terms $m, A$, if there is $\Gamma \vdash$ and $\Gamma \vdash m : A$, then there exists sort $s$ and natural number $i$ such that $\overline{\Gamma} \vdash A : s_i$.
+For any context $\Gamma$ and terms $m, A$, if there are $\Gamma \vdash$ and $\Gamma \vdash m : A$, then there exist sort $s$ and natural number $i$ such that $\overline{\Gamma} \vdash A : s_i$.
 \end{theorem}
@@
 Applying Lemma \ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
-Now by the induction hypothesis, we can conclude there exists sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
+Now by the induction hypothesis, we can conclude there exist sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
 Applying Lemma \ref{uarrowinv} to assumption $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : U_{i'}$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$ where $s'$ is a sort and $i', j'$ are natural numbers.
@@
 Due to the fact that variable $x$ is not a free variable in $B$, the substitution occurring in goal $\exists s \in sort, i \in \mathbb{N}, \overline{\Gamma} \vdash B[n/x] : s_i$ is trivial, thus the judgment $\overline{\Gamma_1} \vdash B : s'_{j'}$ that we have proven shows the existence of the goal.
-    \item For the \rname{App-U$\multimap$} case, the proof is similar to the \rname{App-U$\rightarrow$} case, the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
-    \item For the \rname{App-L$\multimap$} case, the proof is similar to the \rname{App-L$\rightarrow$} case, the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
+    \item For the \rname{App-U$\multimap$} case, the proof is similar to the \rname{App-U$\rightarrow$} case; the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
+    \item For the \rname{App-L$\multimap$} case, the proof is similar to the \rname{App-L$\rightarrow$} case; the only difference is that the inversion lemmas used correspond to $\multimap$ instead of $\rightarrow$.
   \end{itemize}
@@
 \begin{lemma}\label{islmergeinv}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable with linear type $x \in \Gamma$ there is $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable with linear type $x \in \Gamma$ there are $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
 \end{lemma}
@@
 \begin{lemma}
-  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable $x \notin \Gamma$ there is $x \notin \Gamma_1$ and $x \notin \Gamma_2$.
+  For contexts $\Gamma_1, \Gamma_2, \Gamma$, if there is $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$, then for any variable $x \notin \Gamma$ there are $x \notin \Gamma_1$ and $x \notin \Gamma_2$.
 \end{lemma}
@@
   Promotion.
-For context $\Gamma$, terms $m, A, B$ and sort $s$, if there is $\pure{\Gamma}$ and $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \multimap B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$.
+For context $\Gamma$, terms $m, A, B$ and sort $s$, if there are $\pure{\Gamma}$, $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \multimap B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \rightarrow B$.
 \end{theorem}
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
-Now applying Lemma \ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+    \item If $s = U$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
+Now applying Lemma \ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
 Now by \rname{U$\rightarrow$} and Lemma \ref{purere} the goal is proven.
-    \item If $s = L$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exists sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
-Now applying Lemma \ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+    \item If $s = L$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
+Now applying Lemma \ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
 Now by \rname{L$\rightarrow$} and Lemma \ref{purere} the goal is proven.
@@
   Dereliction.
-For context $\Gamma$, terms $m, A, B$ and sort $s$, if there is $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$.
+For context $\Gamma$, terms $m, A, B$ and sort $s$, if there are $\Gamma \vdash$ and $\Gamma \vdash m : (x \stype{s} A) \rightarrow B$, then there exists term $n$ such that $\Gamma \vdash n : (x \stype{s} A) \multimap B$.
 \end{theorem}
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
-Now applying Lemma \ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+    \item If $s = U$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
+Now applying Lemma \ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
 By \rname{U$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \utype A) \multimap B : L_{i'}$.
 By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
-    \item If $s = L$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exists sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
-Now applying Lemma \ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exists sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+    \item If $s = L$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
+Now applying Lemma \ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
 By \rname{L$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \ltype A) \multimap B : L_{i'}$.
@@
 \begin{lemma}
-  For CC$\omega$ term $m$, map $\sigma$ from variables to CC$\omega$ terms, map $\tau$ from variable to CLC terms, if for all variables $x$ there is $\lift{\sigma\ x} = \tau\ x$, then $\lift{m[\sigma]} = \lift{m}[\tau]$.
+  For CC$\omega$ term $m$, map $\sigma$ from variables to CC$\omega$ terms, map $\tau$ from variables to CLC terms, if for all variables $x$ there is $\lift{\sigma\ x} = \tau\ x$, then $\lift{m[\sigma]} = \lift{m}[\tau]$.
 \end{lemma}
@@
 \end{proof}
-For the following lemmas, we will index relations and judgments with subscript CLC of CC$\omega$ to emphasize the language it is defined over.
+For the following lemmas, we will index relations and judgments with subscript CLC or CC$\omega$ to emphasize the language it is defined over.
 

### 2022-01-20 writing/draft7/draft7.tex
draft

@@
 The universe sorts of CLC are enough to completely distinguish between linear and non-linear types within a single typing judgment.
-As a result, CLC falls much closer to its CC roots than prior works in both syntax and semantics.
+As a result, CLC falls closer to its CC roots than prior works in both syntax and semantics.
 
@@
 Krishnaswami et al.
-present a dependent linear type theory \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus, demonstrating the ability to internalize imperative programming in the style of Hoare Type Theory \cite{htt}.
-Luo et al.
-\cite{luo} introduce the property of essential linearity and a mixed linear/non-linear context, describing the first type theory that allows types to depend on linear terms.
+present a dependent linear type theory LNL$_D$ \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus.
+The inclusion of universe sorts in CLC draws inspiration from LNL$_D$, they however, differ in their usage.
+LNL$_D$ uses two typing judgments for linear and non-linear types and two adjoint connectives to mediate between these two judgments, CLC uses a single typing judgment without any mediating connectives.
 Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.
-The Idris 2 programming language \cite{idris2} implements QTT as its core type system.
 

### 2022-01-20 writing/draft7/draft7.tex
draft

@@
 present a dependent linear type theory LNL$_D$ \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus.
-The inclusion of universe sorts in CLC draws inspiration from LNL$_D$, they however, differ in their usage.
-LNL$_D$ uses two typing judgments for linear and non-linear types and two adjoint connectives to mediate between these two judgments, CLC uses a single typing judgment without any mediating connectives.
+The inclusion of universe sorts in CLC takes inspiration from LNL$_D$ but the two theories differ in overall design.
+LNL$_D$ uses two typing judgments for linear and non-linear types and two adjoint connectives to mediate between these two judgments whereas CLC uses a single typing judgment without any mediating connectives.
 Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.

### 2022-01-21 writing/draft7/draft7.tex
draft

--- 2022-01-20 writing/draft7/draft7.tex
@@
 
+\def\CComg{{CC}$\omega$}
+
 \begin{document}
@@
   The Calculus of Linear Constructions (CLC) is an extension of the Calculus of Constructions (CC) with linear types.
-Specifically, CLC extends the predicative CC$\omega$ with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
-We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resource.
+Specifically, CLC extends the predicative \CComg with a hierarchy of linear universes that precisely controls the weakening and contraction of its term level inhabitants.
+We study the meta-theory of CLC, showing that it is a sound logical framework for reasoning about resources.
 We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Through examples, we demonstrate that CILC facilitates correct by construction imperative programming and lightweight protocol enforcement.
-We have formalized and proven correct all major results in the Coq Proof Assistant.
+Through examples, we demonstrate that CILC facilitates correct-by-construction imperative programming and lightweight protocol enforcement.
+We have formalized and proven correct all the reported major results in the Coq Proof Assistant.
 \end{abstract}
@@
 \section{Introduction}
-The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work \cite{cc}.
+The Calculus of Constructions (CC) is a dependent type theory introduced by Coquand and Huet in their landmark work~\cite{cc}.
 In CC, types can depend on terms, allowing one to write precise specifications as types.
-Today, CC and its extensions CIC \cite{cic} and ECC \cite{ecc} lie at the core of popular proof assistants such as Coq \cite{coq}, Agda \cite{agda}, Lean \cite{lean} and others.
+Today, CC and its extensions CIC~\cite{cic} and ECC~\cite{ecc} lie at the core of popular proof assistants such as Coq~\cite{coq}, Agda~\cite{agda}, Lean~\cite{lean} and others.
 These theorem provers have found great success in the fields of software verification and constructive mathematics.
 
-However, due to its origins as a logical framework for constructive mathematics, it can be inconvenient for CC to encode and reason about resource.
-Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resource.
+However, due to its origins as a logical framework for constructive mathematics, it can be inconvenient for CC to encode and reason about resources.
+Users of proof assistants based on CC often need to embed external logics such as Separation Logic to provide domain-specific reasoning principles for dealing with resources.
 We propose an alternative solution: extend CC with linear types.
 
-This paper presents a new linear dependent type system --- the Calculus of Linear Constructions (CLC).
-CLC extends the predicative CC$\omega$ with linear types.
-CC$\omega$ itself is an extension of CC with a cumulative hierarchy of type universes.
+This paper presents a new linear dependent type system --- the Calculus of Linear Constructions (CLC), which extends the predicative \CComg with linear types.
+\CComg itself is an extension of CC with a cumulative hierarchy of type universes.
 We add an extra universe sort $L$ of linear types with cumulativity parallel to the sort $U$ of non-linear types.
@@
 The universe sorts of CLC are enough to completely distinguish between linear and non-linear types within a single typing judgment.
-As a result, CLC falls closer to its CC roots than prior works in both syntax and semantics.
+As a result, CLC falls closer to its CC roots than prior works in terms of both syntax and semantics.
 
 We further extend CLC with rules for defining inductive linear types in the style of CIC, forming the Calculus of Inductive Linear Constructions (CILC).
-Standard Linear Logic connectives such as $\otimes$ and $\oplus$ are definable using this generalized mechanism in a straightforward way.
-Connectives that mix linear and non-linear types can be constructed just as easily using CILC.
-
-Through examples in Section \ref{examples}, we show some applications of CILC such as construction of safe random access array from the first principles, protocol enforcement through inductively defined session types and in-place update of data structures using continuations.
-
-We have formalized all major results in Coq and implemented a prototype in OCaml.
+Standard Linear Logic connectives such as $\otimes$ and $\oplus$ can then be defined straightforwardly.
+And connectives that mix linear and non-linear types can be constructed just as easily in CILC.
+
+Through examples in Section~\ref{examples}, we show some applications of CILC such as construction of safe random access array from the first principles, protocol enforcement through inductively defined session types and in-place update of data structures using continuations.
+
+We have formalized in Coq all the major results reported here and implemented in OCaml a prototype.
 Additional examples of applications such as provable effect laws and quantitative reasoning are included with our implementation.
@@
 \begin{itemize}
-  \item First, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types, enabling direct and precise reasoning about resource.
+  \item First, we describe the Calculus of Linear Constructions, an extension to the Calculus of Constructions with linear types, enabling direct and precise reasoning about resources.
   \item Second, we study the meta-theory of CLC, showing that it exhibits many desirable properties such as confluence, subject reduction and logical consistency.
@@
 This is the first complete presentation of a generalized mechanism for defining inductive linear types.
-  \item All major results have been proven correct in the Coq.
+  \item All major results have been proven correct in Coq.
 Our development is the first machine checked formalization of a linear dependent type theory.
@@
 
-The syntax of the core type theory is presented in Figure \ref{syntax1}.
+The syntax of the core type theory is presented in Figure~\ref{syntax1}.
 Our type theory contains two sorts of universes $U$ and $L$.
@@
 Subtle issues arise if !!
-is not canonically isomorphic to !, which may invalidate the substitution lemma \cite{substitute}.
+is not canonically isomorphic to !, which may invalidate the substitution lemma~\cite{substitute}.
 By integrating the exponential directly into universe sorts, we have limited !
 to only be canonically usable.
-This allows us to derive the substitution lemma and construct a direct modeling of CLC in CC$\omega$ without requiring any additional machinery for manipulating exponential.
-
-Figure \ref{correspondence} illustrates the correspondence between CLC functions and Multiplicative Exponential Linear Logic (MELL) implications.
+This allows us to derive the substitution lemma and construct a direct modeling of CLC in \CComg without requiring any additional machinery for manipulating exponential.
+
+Figure~\ref{correspondence} illustrates the correspondence between CLC functions and Multiplicative Exponential Linear Logic (MELL) implications.
 MELL lacks counterparts for the cases (1), (3) if the co-domain $B$ is dependent on arguments of domain $A$.
-We will discuss function type formation in Section \ref{tyformation} in greater detail.
+We will discuss function type formation in Section~\ref{tyformation} in greater detail.
 
@@
 
-Figure \ref{universe} illustrates the structure of our universe hierarchy.
+Figure~\ref{universe} illustrates the structure of our universe hierarchy.
 Each linear universe $L_i$ has $U_{i+1}$ as its type, allowing functions to freely quantify over linear \textit{types}.
@@
 In other words, all variables found in $\pure{\Gamma}$ are annotated of the form $x \utype A$.
-The full rules for structural judgments are presented in Figure \ref{structural}.
+The full rules for structural judgments are presented in Figure~\ref{structural}.
 
@@
 \subsection{Type Formation} \label{tyformation}
-The rules for forming types are presented in Figure \ref{type}.
-In CLC, we forbid types from depending on linear terms similarly to \cite{llf,neel15} for the same reason of avoiding philosophical troubles.
-
-The axiom rules \rname{U-Axiom} and \rname{L-Axiom} are similar to standard CC$\omega$, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
+The rules for forming types are presented in Figure~\ref{type}.
+In CLC, we forbid types from depending on linear terms similarly to~\cite{llf,neel15} for the same reason of avoiding philosophical troubles.
+
+The axiom rules \rname{U-Axiom} and \rname{L-Axiom} are similar to standard \CComg, the main difference being the extra side-condition of judgment $\pure{\Gamma}$.
 In most presentations of dependent type theories without linear types, the universe axioms are derivable under any well-formed context $\Gamma$.
@@
 From \rname{L-Axiom} we can see that the universe of linear types $L_i$ is an inhabitant of $U_{i+1}$.
-This is inspired by Krishnaswami et al.'s treatment of linear universes \cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
+This is inspired by Krishnaswami et al.'s treatment of linear universes~\cite{neel15}, where linear \textit{types} themselves can be used unrestrictedly.
 
@@
 \subsection{Equality and Reduction} \label{reduction}
-The operational semantics and $\beta$-equality of CLC terms are presented in Figure \ref{red}, all of which are entirely standard.
-
-As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain the standard operational semantics of CC$\omega$, whose $\beta$-reductions are very well behaved.
+The operational semantics and $\beta$-equality of CLC terms are presented in Figure~\ref{red}, all of which are entirely standard.
+
+As we have discussed previously, the elimination of the explicit ! exponential allows CLC to maintain the standard operational semantics of \CComg, whose $\beta$-reductions are very well behaved.
 
@@
 First, we show the type soundness of CLC through the \textit{subject reduction} theorem.
-Next, through the \textit{linearity} theorem we illustrate how CLC can effectively control usage of resource.
+Next, through the \textit{linearity} theorem we illustrate how CLC can effectively control usage of resources.
 Furthermore, we demonstrate that the \textit{promotion} and \textit{dereliction} rules can be encoded as $\eta$-expansions.
-Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed CC$\omega$ terms, proving that CLC is strongly normalizing.
-
-All proofs have been formalized in Coq with help from the Autosubst \cite{autosubst} library.
+Finally, we construct a reduction preserving erasure function that maps well-typed CLC terms to well-typed \CComg terms, proving that CLC is strongly normalizing.
+
+All proofs have been formalized in Coq with help from the Autosubst~\cite{autosubst} library.
 The Coq development is publicly available on the author's Github repository.
@@
 
-The proof of confluence is entirely standard using parallel reduction technique \cite{takahashi}. The presence of linear types does not pose any complications as reductions are untyped.
+The proof of confluence is entirely standard using parallel reduction technique~\cite{takahashi}. The presence of linear types does not pose any complications as reductions are untyped.
 
@@
 exponential.
-Perhaps the most famous work detailing the issues of substitution is due to Wadler \cite{substitute}.
+Perhaps the most famous work detailing the issues of substitution is due to Wadler~\cite{substitute}.
 He defines intricate syntax and semantics for the unboxing of !
@@
 
-Notice in Theorem \ref{nlsubst} of non-linear substitutions, there exists an extra condition $\pure{\Gamma_2}$ which Theorem \ref{lsubst} of linear substitutions lacks.
+Notice in Theorem~\ref{nlsubst} of non-linear substitutions, there exists an extra condition $\pure{\Gamma_2}$ which Theorem~\ref{lsubst} of linear substitutions lacks.
 This condition ensures that linear variables do not occur within the substituted term, thus cannot be duplicated by substitution.
@@
 \subsubsection{Strong Normalization}
-The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function from CLC to CC$\omega$.
-We assume familiarity with CC$\omega$ syntax here and define the erasure function as follows.
+The strong normalization theorem of CLC is proven by construction of a typing and reduction preserving erasure function from CLC to \CComg.
+We assume familiarity with \CComg syntax here and define the erasure function as follows.
 
@@
 
-For the following theorems, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
+For the following theorems, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in \CComg as $\step_{\scriptscriptstyle \text{\CComg}}$.
 
 \begin{theorem} \label{preserve}
-  Reduction Erasure. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  Reduction Erasure. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{\CComg}} \erase{n}$.
 \end{theorem}
@@
 \begin{theorem} \label{embedding}
-  Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
+  Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in \CComg.
 \end{theorem}
 
-If there exists some well-typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve} and \ref{embedding}.
-This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.
+If there exists some well-typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed \CComg term along with its infinite sequence of reductions by virtue of theorems~\ref{preserve} and~\ref{embedding}.
+This is contradictory to the strong normalization property of \CComg proven through the Girard-Tait method~\cite{ecc}, so this hypothetical term does not exist in CLC.
 
@@
 \subsubsection{Compatibility}
-To show that CLC is compatible with the predicative CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.
+To show that CLC is compatible with the predicative \CComg, we construct a function that annotates \CComg terms with sort $U$, lifting them into the non-linear fragment of CLC.
 
@@
 
-With slight overloading of notation, we define lifting for CC$\omega$ recursively.
+With slight overloading of notation, we define lifting for \CComg recursively.
 \begin{definition}
@@
 \begin{theorem}
-  Lift preserves reduction. For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+  Lift preserves reduction. For any \CComg term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
 \end{theorem}
@@
 \begin{theorem} \label{lift}
-  Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
+  Lifting. If $\Gamma \vdash m : A$ is a valid judgment in \CComg, then $\lift{\Gamma} \vdash \lift{m} : \lift{A}$ is a valid judgment in CLC.
 \end{theorem}
 
-The lifting theorem shows that all valid CC$\omega$ terms are valid CLC terms if the domains of function types are annotated with sort $U$.
+The lifting theorem shows that all valid \CComg terms are valid CLC terms if the domains of function types are annotated with sort $U$.
 In practice, these annotations can often be omitted if algorithmic type checking is able to infer them from context.
@@
 \section{The Extension to CILC}
-Although the core CLC language is extremely expressive, formally shown by Theorem \ref{lift} to be at least as expressive as the predicative CC$\omega$, the complexity of Church-style encodings to do so will quickly become unmanageable.
+Although the core CLC language is extremely expressive, formally shown by Theorem~\ref{lift} to be at least as expressive as the predicative \CComg, the complexity of Church-style encodings to do so will quickly become unmanageable.
 To address this, we extend CLC with rules for defining inductive linear types in the style of CIC.
@@
 
-As shown in Figure \ref{syntax2}, CILC extends CIC with syntax for the introduction and elimination of inductive types.
+As shown in Figure~\ref{syntax2}, CILC extends CIC with syntax for the introduction and elimination of inductive types.
 Ind is used during the formation of new inductive types, and Constr is used for introducing their constructors.
@@
 \subsection{Arity}
-For type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as depicted in Figure \ref{arity}.
+For type $A$ and sort $s$, the judgment $arity(A, s)$ is inductively defined over the structure of $A$ as depicted in Figure~\ref{arity}.
 
@@
 \subsection{Strict Positivity}
-For type $P$ and variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as depicted in Figure \ref{positive}.
+For type $P$ and variable $X$, the judgment $postive(P, X)$ is inductively defined over the structure of $P$ as depicted in Figure~\ref{positive}.
 
@@
 \subsection{Non-Linear Constructor}
-For type $C$ and variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{nlconstr}.
+For type $C$ and variable $X$, the judgment $\ucons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure~\ref{nlconstr}.
 
@@
 \subsection{Linear Constructor}
-For type $C$ and variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{lconstr}.
+For type $C$ and variable $X$, the judgment $\lcons(C, X)$ is inductively defined over the structure of $C$ as shown in Figure~\ref{lconstr}.
 
@@
 
-For type $C$ and variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as shown in Figure \ref{activation}.
+For type $C$ and variable $X$, the judgment $activation(C, X)$ is inductively defined over the structure of $C$ as shown in Figure~\ref{activation}.
 Intuitively, $activation(C,X)$ ensures the linearity of $C$.
@@
 \subsection{Introduction Rules}
-The formation of new inductive types and their constructors are presented in Figure \ref{ind}.
+The formation of new inductive types and their constructors are presented in Figure~\ref{ind}.
 
@@
 
-The rule for non-dependent case elimination is presented in Figure \ref{case}. This elimination rule is non-dependent because the motive $Q$ may not depend on the discriminated term $m$ itself.
+The rule for non-dependent case elimination is presented in Figure~\ref{case}. This elimination rule is non-dependent because the motive $Q$ may not depend on the discriminated term $m$ itself.
 
@@
 
-Dependent elimination of non-linear inductive objects is presented in Figure \ref{dcase}.
+Dependent elimination of non-linear inductive objects is presented in Figure~\ref{dcase}.
 This elimination rule is dependent because the motive $Q$ may depend on the discriminated term $m$.
@@
 \subsection{Fixpoint}
-The typing rule of fix-points presented in Figure \ref{fix} is mostly standard.
+The typing rule of fix-points presented in Figure~\ref{fix} is mostly standard.
 A syntactic guard condition is utilized to conservatively ensure termination.
@@
 \subsection{Extended Reduction}
-As the reduction semantics for the extended language is standard \cite{cic}, we only present the most interesting $\iota$-reduction cases.
+As the reduction semantics for the extended language is standard~\cite{cic}, we only present the most interesting $\iota$-reduction cases.
 
@@
 \subsection{Linear Connectives}
-In Figure \ref{lconnectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
+In Figure~\ref{lconnectives}, we demonstrate how the $\otimes$ and $\oplus$ connectives of linear logic could be defined in CILC as inductive linear types.
 For any term with inductive linear type, linearity will ultimately force the term to be eliminated by a Case expression.
@@
 \subsection{Mixed Connectives}
-In Figure \ref{mconnectives}, we introduce a $\SigmaL$ connective.
+In Figure~\ref{mconnectives}, we introduce a $\SigmaL$ connective.
 $\SigmaL$ is similar to the standard $\Sigma$ type where the type of the second component is dependent on the value of the first component.
@@
           The $\mapsto$ is an infix linear type constructor.
-It fulfills a role similar to L3's \cite{l3} memory access capability, or Separation Logic's \cite{reynolds02} points-to assertion.
+It fulfills a role similar to L3's~\cite{l3} memory access capability, or Separation Logic's~\cite{reynolds02} points-to assertion.
 Intuitively, a capability $(l \mapsto A)$ is a proof that a term of non-linear type $A$ is currently stored at address $l$.
@@
 A random access array is a contiguous region of memory that can be accessed by an address pointing to the start of this region and an offset.
-In Figure \ref{arrv}, we declare a linear inductive $\arrv$ to track the memory used by the array.
+In Figure~\ref{arrv}, we declare a linear inductive $\arrv$ to track the memory used by the array.
 $\arrv$ is parameterized by the type $A$ of elements stored in the array and the address $l$ of the initial element.
@@
 So for some type $A$ and natural numbers $l, n$, a term of type $\arrv\ A\ l\ n$ is a proof that a contiguous region of memory starting from address $l$ can be accessed by any offset less than $n$.
-Arrays can now be characterized as an address $l$ paired with an array proof $\arrv\ A\ l\ n$ as shown in Figure \ref{array}.
+Arrays can now be characterized as an address $l$ paired with an array proof $\arrv\ A\ l\ n$ as shown in Figure~\ref{array}.
 
@@
 
-To manipulate arrays, we construct a recursive helper function whose signature is shown in Figure \ref{trav}.
+To manipulate arrays, we construct a recursive helper function whose signature is shown in Figure~\ref{trav}.
 When given an array proof $\arrv\ A\ l\ n$, an offset $m$ and proof $pf$ such that $m < n$, $\texttt{nth}$ will return a capability $(l + m \mapsto A)$ along with a function $(l + m \mapsto A) \multimap \arrv\ A\ l\ n$.
@@
 
-An indexing function can now be constructed as a wrapper for $\texttt{nth}$ with its signature shown in Figure \ref{index}.
+An indexing function can now be constructed as a wrapper for $\texttt{nth}$ with its signature shown in Figure~\ref{index}.
 Given an array $\texttt{Array}\ A\ n$, an offset $m$ and proof that $m < n$, $\texttt{nth}$ can be used to extract the capability at offset $m$, allowing $\get$ to retrieve the stored value.
@@
 
-In Figure \ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
+In Figure~\ref{session}, we declare an inductive \texttt{session} type for specifying protocols.
 Intuitively, the \texttt{session}  forms a stack of possible operations executed in a protocol.
@@
 
-In Definition \ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
+In Definition~\ref{comm}, we postulate five axioms for communication using the $\session$ type. These axioms could be viewed as an interface to trusted communication libraries.
 
@@
 \subsection{Mutable Data}
-Figure \ref{list} presents a CPS append function for linear lists.
+Figure~\ref{list} presents a CPS append function for linear lists.
 In the \texttt{cons} case, the constructor is opened, exposing its data element \texttt{h} and tail list \texttt{t}.
@@
 \section{Related Work}
-Linear types are a class of type systems inspired by Girard's substructural Linear Logic \cite{girard}.
-Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resource.
-Wadler \cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resource.
-A term calculus for linear type theory is realized by Abramsky \cite{abramsky1993}.
-Benton \cite{benton1994} investigates the ramifications of the !
+Linear types are a class of type systems inspired by Girard's substructural Linear Logic~\cite{girard}.
+Girard notices that the weakening and contraction rules of Classical Logic when restricted carefully, give rise to a new logical foundation for reasoning about resources.
+Wadler~\cite{wadler1990,wadler1991} then applies an analogous restriction to variable usage in simple type theory, leading to the development of linear type theory where terms respect resources.
+A term calculus for linear type theory is realized by Abramsky~\cite{abramsky1993}.
+Benton~\cite{benton1994} investigates the ramifications of the !
 exponential in linear term calculi, decomposing it to adjoint connectives $F$ and $G$ that map between linear and non-linear judgments.
-Programming languages \cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource safe software in practical applications.
+Programming languages~\cite{l3,ats,linear-haskell} featuring linear types have also been implemented, allowing programmers to write resource-safe software in practical applications.
 
 Over the years, work has been done to enrich linear type theories with dependent types.
-Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types \cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
-V\'{a}k\'{a}r \cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL \cite{dill}.
+Cervesato and Pfenning extend the Edinburgh Logical Framework with linear types~\cite{lf,llf}, being the first to demonstrate that dependent types and linear types can coexist within a type theory.
+V\'{a}k\'{a}r~\cite{vakar14} presents a linear dependent type theory, with syntax and semantics drawing inspiration from DILL~\cite{dill}.
 Krishnaswami et al.
-present a dependent linear type theory LNL$_D$ \cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus.
-The inclusion of universe sorts in CLC draws inspiration from LNL$_D$, they however, differ in their usage.
-LNL$_D$ uses two typing judgments for linear and non-linear types and two adjoint connectives to mediate between these two judgments, CLC uses a single typing judgment without any mediating connectives.
-Based on initial ideas of McBride \cite{nothing}, Atkey's Quantitative Type Theory (QTT) \cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.
+present a dependent linear type theory LNL$_D$~\cite{neel15} based on Benton's earlier work on mixed linear and non-linear calculus~\cite{benton1994}.
+While the inclusion of universe sorts in CLC draws inspiration from LNL$_D$, they differ in their usage.
+In particular, LNL$_D$ uses two typing judgments for linear and non-linear types and two adjoint connectives to mediate between these two judgments, but CLC uses a single typing judgment without any mediating connectives.
+Based on initial ideas of McBride~\cite{nothing}, Atkey's Quantitative Type Theory (QTT)~\cite{qtt} uses semi-ring annotations to track variable occurrence, simulating irrelevance and linear and affine types within a unified framework.
 
@@
 
-We aim to fully verify the typechecking algorithm employed by the current implementation.
-The unification algorithm used for implicit argument resolution is extremely ad-hoc.
-As the problem of unification for linear types has not been thoroughly studied, we hope to improve this situation as well.
+Currently, the linear types of CILC allow certain effects such as the ones described in Section~\ref{examples} to coexist with dependent types.
+However, these effects appeal to axioms and are not intrinsic to the core type theory.
+We plan to investigate how and to what extent can effects be internalized in future extensions to CILC.
+Our goal is to strive towards a unifying language whose effects could be built using linearity as its building blocks, possessing both static provable safety and runtime efficiency.
 
@@
 \begin{proof}
-  By induction on the derivation of $\red$, the transitivity of $\red$ and Lemma \ref{stepsubst}.
+  By induction on the derivation of $\red$, the transitivity of $\red$ and Lemma~\ref{stepsubst}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the structure of $m$, applying Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}.
+  By induction on the structure of $m$, applying Lemmas~\ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}.
 \end{proof}
@@
 \begin{proof}
-  By Lemmas \ref{convhom} and \ref{stepsubst}.
+  By Lemmas~\ref{convhom} and~\ref{stepsubst}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the structure of $m$, applying Lemmas \ref{convapp}, \ref{convlam}, \ref{convarrow}, \ref{convlolli}.
+  By induction on the structure of $m$, applying Lemmas~\ref{convapp}, \ref{convlam}, \ref{convarrow}, \ref{convlolli}.
 \end{proof}
@@
 \begin{proof}
-  This is a special case of Lemma \ref{convcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+  This is a special case of Lemma~\ref{convcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
 \end{proof}
@@
 \begin{proof}
-  By Definition \ref{psstep} and Lemma \ref{psteprefl}.
+  By Definition~\ref{psstep} and Lemma~\ref{psteprefl}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $m \step n$ and Lemma \ref{psteprefl}.
+  By induction on the derivation of $m \step n$ and Lemma~\ref{psteprefl}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $m \pstep n$, utilizing the transitive property of $\red$ and Lemmas \ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}, \ref{redsubst}, \ref{redcompat}.
+  By induction on the derivation of $m \pstep n$, utilizing the transitive property of $\red$ and Lemmas~\ref{redapp}, \ref{redlam}, \ref{redarrow}, \ref{redlolli}, \ref{redsubst}, \ref{redcompat}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $m \pstep n$ and Lemma \ref{psteprefl}.
+  By induction on the derivation of $m \pstep n$ and Lemma~\ref{psteprefl}.
 \end{proof}
@@
 \begin{proof}
-  By Lemma \ref{psteprefl}, this is a special case of Lemma \ref{pstepcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
+  By Lemma~\ref{psteprefl}, this is a special case of Lemma~\ref{pstepcompat} where $\sigma$ maps $x$ to $m_1$ and $\tau$ maps $x$ to $m_2$.
 \end{proof}
@@
 Each case in the induction specializes $m$ appearing in $m \pstep m_2$, allowing one to invert its derivation in a syntax directed way and then to apply the induction hypothesis.
-The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma \ref{pstepcompat} is used to push these cases through.
+The difficult cases are due to \rname{PStep-$\beta$} as it concerns substitution, so Lemma~\ref{pstepcompat} is used to push these cases through.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $m \pstep m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{diamond}.
+  By induction on the derivation of $m \pstep m_1$, utilizing transitivity of $\red$ and Lemmas~\ref{steppstep}, \ref{pstepred}, \ref{diamond}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $m \red m_1$, utilizing transitivity of $\red$ and Lemmas \ref{steppstep}, \ref{pstepred}, \ref{strip}.
+  By induction on the derivation of $m \red m_1$, utilizing transitivity of $\red$ and Lemmas~\ref{steppstep}, \ref{pstepred}, \ref{strip}.
 \end{proof}
@@
 \begin{proof}
-  By Lemma \ref{convsub1} and \rname{$\prec$-Refl}.
+  By Lemma~\ref{convsub1} and \rname{$\prec$-Refl}.
 \end{proof}
@@
 \begin{proof}
-  By Lemma \ref{sub1sub} and \rname{$\prec$-Refl}.
+  By Lemma~\ref{sub1sub} and \rname{$\prec$-Refl}.
 \end{proof}
@@
 \begin{proof}
-  By Lemma \ref{sub1sub} and \rname{$\prec$-Sort}.
+  By Lemma~\ref{sub1sub} and \rname{$\prec$-Sort}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $A \prec B$, definition of $\prec$ and Lemmas \ref{sub1sub}, \ref{sub1conv}, \ref{convsub1}.
+  By induction on the derivation of $A \prec B$, definition of $\prec$ and Lemmas~\ref{sub1sub}, \ref{sub1conv}, \ref{convsub1}.
 \end{proof}
@@
 \begin{proof}
-  By transitivity of $\equiv$, rule \rname{$\prec$-$\preceq$} and Lemma \ref{sub1trans}.
+  By transitivity of $\equiv$, rule \rname{$\prec$-$\preceq$} and Lemma~\ref{sub1trans}.
 \end{proof}
@@
 \begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{sortinj}.
+  By transitivity of $\equiv$ and Corollary~\ref{sortinj}.
 \end{proof}
@@
 \begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{arrowinj}.
+  By transitivity of $\equiv$ and Corollary~\ref{arrowinj}.
 \end{proof}
@@
 \begin{proof}
-  By transitivity of $\equiv$ and Corollary \ref{lolliinj}.
+  By transitivity of $\equiv$ and Corollary~\ref{lolliinj}.
 \end{proof}
@@
 \begin{proof}
-  By rule \rname{$\prec$-$\preceq$} and Lemmas \ref{convsubst}, \ref{sub1subst}.
+  By rule \rname{$\prec$-$\preceq$} and Lemmas~\ref{convsubst}, \ref{sub1subst}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas \ref{uarrowinv}, \ref{larrowinv}.
+  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas~\ref{uarrowinv}, \ref{larrowinv}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas \ref{ulolliinv}, \ref{llolliinv}.
+  By induction on the derivation of $\Gamma \vdash \lambda x \stype{s} A . n : C$ and Lemmas~\ref{ulolliinv}, \ref{llolliinv}.
 \end{proof}
@@
 \begin{proof}
-  Direct consequence of Lemmas \ref{uarrowinv}, \ref{larrowinv} and \ref{arrowlaminvx}.
+  Direct consequence of Lemmas~\ref{uarrowinv}, \ref{larrowinv} and~\ref{arrowlaminvx}.
 \end{proof}
@@
 \begin{proof}
-  Direct consequence of Lemmas \ref{ulolliinv}, \ref{llolliinv} and \ref{lollilaminvx}.
+  Direct consequence of Lemmas~\ref{ulolliinv}, \ref{llolliinv} and~\ref{lollilaminvx}.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $agreeR\ \xi\ \Gamma\ \Gamma'$ and lemmas in Section \ref{agreeRprop}.
+  By induction on the derivation of $agreeR\ \xi\ \Gamma\ \Gamma'$ and lemmas in Section~\ref{agreeRprop}.
 \end{proof}
@@
   By induction on the derivation of $\Gamma \vdash m : A$.
-We shall only discuss the application case in detail, as the other cases are proven by application of the induction hypothesis and the lemmas in Section \ref{agreeRprop}.
+We shall only discuss the application case in detail, as the other cases are proven by application of the induction hypothesis and the lemmas in Section~\ref{agreeRprop}.
   \begin{itemize}
-    \item For the \rname{App-U$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
-From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
+    \item For the \rname{App-U$\rightarrow$} case, Lemma~\ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+From $\pure{\Gamma_2}$ and Lemma~\ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
 At this point, the induction hypothesis allows us to apply \rname{App-U$\rightarrow$} to prove the goal.
-    \item For the \rname{App-L$\rightarrow$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-L$\rightarrow$} case, Lemma~\ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 At this point, the induction hypothesis allows us to apply \rname{App-L$\rightarrow$} to prove the goal.
-    \item For the \rname{App-U$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
-From $\pure{\Gamma_2}$ and Lemma \ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
+    \item For the \rname{App-U$\multimap$} case, Lemma~\ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+From $\pure{\Gamma_2}$ and Lemma~\ref{agreerenpure} we know that there is $\pure{\Gamma_2'}$.
 At this point, the induction hypothesis allows us to apply \rname{App-U$\multimap$} to prove the goal.
-    \item For the \rname{App-L$\multimap$} case, Lemma \ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
+    \item For the \rname{App-L$\multimap$} case, Lemma~\ref{mergeagreereninv} is applied to split the context $\Gamma$ into two contexts $\Gamma_1'$ and $\Gamma_2'$ such that there are $\mrg{\Gamma_1'}{\Gamma_2'}{\Gamma'}$, $agreeR\ \xi\ \Gamma_1\ \Gamma_1'$ and $agreeR\ \xi\ \Gamma_2\ \Gamma_2'$.
 At this point, the induction hypothesis allows us to apply \rname{App-L$\multimap$} to prove the goal.
@@
 \begin{proof}
-  Using \rname{agreeR-wk} and Lemma \ref{agreerenrefl} a proof of the relation $agreeR\ id\ \Gamma\ (\Gamma, x \utype B)$ can be constructed. Then by Lemma \ref{renaming}, the theorem can be proven.
+  Using \rname{agreeR-wk} and Lemma~\ref{agreerenrefl} a proof of the relation $agreeR\ id\ \Gamma\ (\Gamma, x \utype B)$ can be constructed. Then by Lemma~\ref{renaming}, the theorem can be proven.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $agreeS\ \sigma\ \Delta\ \Gamma$ and lemmas in Section \ref{agreeSprop}.
+  By induction on the derivation of $agreeS\ \sigma\ \Delta\ \Gamma$ and lemmas in Section~\ref{agreeSprop}.
 \end{proof}
@@
   The proof proceeds by induction on the derivation of $\Gamma \vdash m : A$.
-Similarly to the proof of Lemma \ref{renaming}, the interesting cases are the application cases where Lemma \ref{mergeagreesubstinv} must be utilized to split the $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ judgments for use in the induction hypothesis.
+Similarly to the proof of Lemma~\ref{renaming}, the interesting cases are the application cases where Lemma~\ref{mergeagreesubstinv} must be utilized to split the $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ judgments for use in the induction hypothesis.
 \end{proof}
@@
 \begin{proof}
-  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and the properties of $merge$ discussed in Section \ref{mergeprop}.
+  By induction on the derivation of $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and the properties of $merge$ discussed in Section~\ref{mergeprop}.
 \end{proof}
@@
   \begin{itemize}
-    \item For the \rname{App-U$\rightarrow$} case, from assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$ and Lemmas \ref{mergepure2}, \ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
-Applying Lemma \ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+    \item For the \rname{App-U$\rightarrow$} case, from assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\pure{\Gamma_2}$ and Lemmas~\ref{mergepure2}, \ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+Applying Lemma~\ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
 Now by the induction hypothesis, we can conclude there exist sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
-Applying Lemma \ref{uarrowinv} to assumption $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : U_{i'}$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$ where $s'$ is a sort and $i', j'$ are natural numbers.
-The goal can finally be proven by applying the substitution Lemma \ref{substitutionu} on assumptions $\Gamma_2 \vdash n : A$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$.
-    \item For the \rname{App-L$\rightarrow$} case, from assumption $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and Lemmas \ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
-Applying Lemma \ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
+Applying Lemma~\ref{uarrowinv} to assumption $\overline{\Gamma_1} \vdash (x \utype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : U_{i'}$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$ where $s'$ is a sort and $i', j'$ are natural numbers.
+The goal can finally be proven by applying the substitution Lemma~\ref{substitutionu} on assumptions $\Gamma_2 \vdash n : A$ and $\overline{\Gamma_1}, x \utype A \vdash B : s'_{j'}$.
+    \item For the \rname{App-L$\rightarrow$} case, from assumption $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and Lemmas~\ref{mergerere}, we can conclude that $\overline{\Gamma_1} = \overline{\Gamma}$ and $\overline{\Gamma_2} = \overline{\Gamma}$.
+Applying Lemma~\ref{mergecontextokinv} to assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $\Gamma \vdash$ obtains $\Gamma_1 \vdash$ and $\Gamma_2 \vdash$.
 Now by the induction hypothesis, we can conclude that there exists sorts $s, t$ and natural numbers $i, j$ such that there are $\overline{\Gamma_1} \vdash (x \ltype A) \rightarrow B : s_i$ and $\overline{\Gamma_2} \vdash A : t_j$.
-Applying Lemma \ref{larrowinv} to assumption $\overline{\Gamma_1} \vdash (x \ltype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : L_{i'}$ and $\overline{\Gamma_1} \vdash B : s'_{j'}$.
+Applying Lemma~\ref{larrowinv} to assumption $\overline{\Gamma_1} \vdash (x \ltype A) \rightarrow B : s_i$ allows us to derive $\overline{\Gamma_1} \vdash A : L_{i'}$ and $\overline{\Gamma_1} \vdash B : s'_{j'}$.
 Due to the fact that variable $x$ is not a free variable in $B$, the substitution occurring in goal $\exists s \in sort, i \in \mathbb{N}, \overline{\Gamma} \vdash B[n/x] : s_i$ is trivial, thus the judgment $\overline{\Gamma_1} \vdash B : s'_{j'}$ that we have proven shows the existence of the goal.
@@
   \begin{itemize}
-    \item For case \rname{App-U$\rightarrow$}, by assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $(x \ltype A) \in \Gamma$ and Lemma \ref{islmergeinv} we can conclude that $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
-In both cases, applying the induction hypothesis and Lemma \ref{narity} proves the goal.
+    \item For case \rname{App-U$\rightarrow$}, by assumptions $\mrg{\Gamma_1}{\Gamma_2}{\Gamma}$ and $(x \ltype A) \in \Gamma$ and Lemma~\ref{islmergeinv} we can conclude that $x \in \Gamma_1$ and $x \notin \Gamma_2$ or $x \in \Gamma_2$ and $x \notin \Gamma_1$.
+In both cases, applying the induction hypothesis and Lemma~\ref{narity} proves the goal.
     \item For case \rname{App-L$\rightarrow$}, the proof is the same as \rname{App-U$\rightarrow$}.
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
-Now applying Lemma \ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
-Now by \rname{U$\rightarrow$} and Lemma \ref{purere} the goal is proven.
-    \item If $s = L$, then we may apply Theorem \ref{valid} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
-Now applying Lemma \ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
-Now by \rname{L$\rightarrow$} and Lemma \ref{purere} the goal is proven.
+    \item If $s = U$, then we may apply Theorem~\ref{valid} to assumption $\Gamma \vdash m : (x \utype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$.
+Now applying Lemma~\ref{ulolliinv} to $\overline{\Gamma} \vdash (x \utype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+Now by \rname{U$\rightarrow$} and Lemma~\ref{purere} the goal is proven.
+    \item If $s = L$, then we may apply Theorem~\ref{valid} to assumption $\Gamma \vdash m : (x \ltype A) \multimap B$ to show that there exist sort $t$ and natural number $i$ such that $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$.
+Now applying Lemma~\ref{llolliinv} to $\overline{\Gamma} \vdash (x \ltype A) \multimap B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+Now by \rname{L$\rightarrow$} and Lemma~\ref{purere} the goal is proven.
   \end{itemize}
@@
   \begin{itemize}
-    \item If $s = U$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
-Now applying Lemma \ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
-By \rname{U$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \utype A) \multimap B : L_{i'}$.
+    \item If $s = U$, then we may apply Theorem~\ref{valid} to $\Gamma \vdash m : (x \utype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \utype A) \rightarrow B : t_i$.
+Now applying Lemma~\ref{uarrowinv} to $\overline{\Gamma} \vdash (x \utype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma}, x \utype A \vdash B : t'_{i'}$.
+By \rname{U$\multimap$} and Lemma~\ref{repure} we can prove $\overline{\Gamma} \vdash (x \utype A) \multimap B : L_{i'}$.
 By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
-    \item If $s = L$, then we may apply Theorem \ref{valid} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
-Now applying Lemma \ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
-By \rname{L$\multimap$} and Lemma \ref{repure} we can prove $\overline{\Gamma} \vdash (x \ltype A) \multimap B : L_{i'}$.
+    \item If $s = L$, then we may apply Theorem~\ref{valid} to $\Gamma \vdash m : (x \ltype A) \rightarrow B$ showing that there exist sort $t$ and natural number $i$ such that there is $\overline{\Gamma} \vdash (x \ltype A) \rightarrow B : t_i$.
+Now applying Lemma~\ref{larrowinv} to $\overline{\Gamma} \vdash (x \ltype A)\rightarrow B : t_i$ shows that there exist sort $t'$ and natural number $i'$ such that $\overline{\Gamma} \vdash A : U_{i'}$ and $\overline{\Gamma} \vdash B : t'_{i'}$.
+By \rname{L$\multimap$} and Lemma~\ref{repure} we can prove $\overline{\Gamma} \vdash (x \ltype A) \multimap B : L_{i'}$.
 By rule \rname{$\lambda$$\multimap$}, the rest of the goal can be proven in a straightforward manner.
@@
 
-The proof of the logical consistency of CLC proceeds by construction of a reduction preserving erasure from CLC to CC$\omega$. As CC$\omega$ is consistent, CLC must be consistent as well.
+The proof of the logical consistency of CLC proceeds by construction of a reduction preserving erasure from CLC to \CComg. As \CComg is consistent, CLC must be consistent as well.
 
@@
 \begin{lemma}
-  For CLC term $m$, map $\sigma$ from variables to CLC terms, map $\tau$ from variables to CC$\omega$ terms, if for all variables $x$ there is $\erase{\sigma\ x} = \tau\ x$, then $\erase{m[\sigma]} = \erase{m}[\tau]$.
+  For CLC term $m$, map $\sigma$ from variables to CLC terms, map $\tau$ from variables to \CComg terms, if for all variables $x$ there is $\erase{\sigma\ x} = \tau\ x$, then $\erase{m[\sigma]} = \erase{m}[\tau]$.
 \end{lemma}
@@
 
-For the following lemmas, we will index relations and judgments with subscript CLC or CC$\omega$ to emphasize the language it is defined over.
+For the following lemmas, we will index relations and judgments with subscript CLC or \CComg to emphasize the language it is defined over.
 
 \begin{lemma}\label{erasestep}
-  For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{\CComg}} \erase{n}$.
 \end{lemma}
@@
 \begin{lemma}
-  For any CLC terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \equiv_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{CLC}} n$ and Lemma \ref{erasestep}.
+  For any CLC terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \equiv_{\scriptscriptstyle \text{\CComg}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{CLC}} n$ and Lemma~\ref{erasestep}.
 \end{proof}
@@
 \begin{lemma}\label{erasesub1}
-  For any CLC terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \prec_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
+  For any CLC terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \prec_{\scriptscriptstyle \text{\CComg}} \erase{n}$.
 \end{lemma}
@@
 \begin{lemma}
-  For any CLC terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \preceq_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
-\end{lemma}
-\begin{proof}
-  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{CLC}} n$ and the properties of subtyping proven in Section \ref{subtyping}.
+  For any CLC terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \preceq_{\scriptscriptstyle \text{\CComg}} \erase{n}$.
+\end{lemma}
+\begin{proof}
+  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{CLC}} n$ and the properties of subtyping proven in Section~\ref{subtyping}.
 \end{proof}
@@
   Embedding.
-For any CLC context $\Gamma$ and CLC terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}} m : A$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{CC$\omega$}} \erase{m} : \erase{A}$.
+For any CLC context $\Gamma$ and CLC terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}} m : A$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{\CComg}} \erase{m} : \erase{A}$.
 \end{theorem}
@@
 \begin{corollary}\label{embeddingcontext}
-  For any CLC context $\Gamma$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}}$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{CC$\omega$}}$.
+  For any CLC context $\Gamma$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CLC}}$, then there is $\erase{\Gamma} \vdash_{\scriptscriptstyle \text{\CComg}}$.
 \end{corollary}
 \begin{proof}
-  Direct consequence of applying Theorem \ref{embed} to all types in context $\Gamma$.
+  Direct consequence of applying Theorem~\ref{embed} to all types in context $\Gamma$.
 \end{proof}
@@
   Suppose there exists a well-typed CLC term $m$ with an infinite sequence of reductions.
-Theorem \ref{embed} shows that there must exist some term $\erase{m}$ that is well-typed in CC$\omega$.
-Additionally, this infinite sequence of reductions on $m$ can be translated in CC$\omega$ step-wise by Lemma \ref{erasestep}.
-This shows that we have constructed a non-normalizing CC$\omega$ term $\erase{m}$, which a contradiction to the strong normalization property of CC$\omega$, thus CLC must be strongly normalizing as well.
+Theorem~\ref{embed} shows that there must exist some term $\erase{m}$ that is well-typed in \CComg.
+Additionally, this infinite sequence of reductions on $m$ can be translated in \CComg step-wise by Lemma~\ref{erasestep}.
+This shows that we have constructed a non-normalizing \CComg term $\erase{m}$, which a contradiction to the strong normalization property of \CComg, thus CLC must be strongly normalizing as well.
 \end{proof}
@@
 \subsection{Embedding}
-To show that CLC is compatible with the predicative fragment of CC$\omega$, we construct a lifting procedure that lifts CC$\omega$ terms into CLC in a straightforward way.
+To show that CLC is compatible with the predicative fragment of \CComg, we construct a lifting procedure that lifts \CComg terms into CLC in a straightforward way.
 \begin{align*}
@@
 
-With slight overloading of notation, we define lifting for CC$\omega$ recursively.
+With slight overloading of notation, we define lifting for \CComg recursively.
 \begin{align*}
@@
 \begin{lemma}
-  For CC$\omega$ context $\Gamma$, there is $\pure{\lift{\Gamma}}$.
+  For \CComg context $\Gamma$, there is $\pure{\lift{\Gamma}}$.
 \end{lemma}
@@
 \begin{lemma}
-  For CC$\omega$ term $m$, map $\sigma$ from variables to CC$\omega$ terms, map $\tau$ from variables to CLC terms, if for all variables $x$ there is $\lift{\sigma\ x} = \tau\ x$, then $\lift{m[\sigma]} = \lift{m}[\tau]$.
+  For \CComg term $m$, map $\sigma$ from variables to \CComg terms, map $\tau$ from variables to CLC terms, if for all variables $x$ there is $\lift{\sigma\ x} = \tau\ x$, then $\lift{m[\sigma]} = \lift{m}[\tau]$.
 \end{lemma}
@@
 \end{proof}
-For the following lemmas, we will index relations and judgments with subscript CLC or CC$\omega$ to emphasize the language it is defined over.
+For the following lemmas, we will index relations and judgments with subscript CLC or \CComg to emphasize the language it is defined over.
 
 \begin{lemma}\label{liftstep}
-  For any CC$\omega$ terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \step_{\scriptscriptstyle \text{CC$\omega$}} n$.
+  For any \CComg terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{\CComg}} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \step_{\scriptscriptstyle \text{\CComg}} n$.
 \end{proof}
@@
 \begin{lemma}
-  For any CC$\omega$ terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \equiv_{\scriptscriptstyle \text{CLC}} \lift{n}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{CC$\omega$}} n$ and Lemma \ref{liftstep}.
+  For any \CComg terms $m$ and $n$, if there is $m \equiv_{\scriptscriptstyle \text{\CComg}} n$, then there is $\lift{m} \equiv_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \equiv_{\scriptscriptstyle \text{\CComg}} n$ and Lemma~\ref{liftstep}.
 \end{proof}
@@
 \begin{lemma}\label{liftsub1}
-  For any CC$\omega$ terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \prec_{\scriptscriptstyle \text{CLC}} \lift{n}$.
-\end{lemma}
-\begin{proof}
-  By induction on the derivation of $m \prec_{\scriptscriptstyle \text{CC$\omega$}} n$.
+  For any \CComg terms $m$ and $n$, if there is $m \prec_{\scriptscriptstyle \text{\CComg}} n$, then there is $\lift{m} \prec_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By induction on the derivation of $m \prec_{\scriptscriptstyle \text{\CComg}} n$.
 \end{proof}
@@
 \begin{lemma}
-  For any CC$\omega$ terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{CC$\omega$}} n$, then there is $\lift{m} \preceq_{\scriptscriptstyle \text{CLC}} \lift{n}$.
-\end{lemma}
-\begin{proof}
-  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{CC$\omega$}} n$ and the properties of subtyping proven in Section \ref{subtyping}.
+  For any \CComg terms $m$ and $n$, if there is $m \preceq_{\scriptscriptstyle \text{\CComg}} n$, then there is $\lift{m} \preceq_{\scriptscriptstyle \text{CLC}} \lift{n}$.
+\end{lemma}
+\begin{proof}
+  By case analysis on the derivation of $m \preceq_{\scriptscriptstyle \text{\CComg}} n$ and the properties of subtyping proven in Section~\ref{subtyping}.
 \end{proof}
@@
   Lifting.
-For any CC$\omega$ context $\Gamma$ and CC$\omega$ terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{CC$\omega$}} m : A$, then there is $\lift{\Gamma} \vdash_{\scriptscriptstyle \text{CLC}} \lift{m} : \lift{A}$.
+For any \CComg context $\Gamma$ and \CComg terms $m, A$, if there is $\Gamma \vdash_{\scriptscriptstyle \text{\CComg}} m : A$, then there is $\lift{\Gamma} \vdash_{\scriptscriptstyle \text{CLC}} \lift{m} : \lift{A}$.
 \end{theorem}
 \begin{proof}
-  By induction on the derivation of $\Gamma \vdash_{\scriptscriptstyle \text{CC$\omega$}} m : A$.
+  By induction on the derivation of $\Gamma \vdash_{\scriptscriptstyle \text{\CComg}} m : A$.
 \end{proof}

### 2022-03-12 writing/rebuttal.md
rebuttal

# LICS rebuttal

## 1
Reviewer 1 has acutely observed that, given Γ = A :U L, B :U U, f :U (x :L A) → B, x :L A, the expression Γ ⊢ (λ(_ :U B).0) (f x) : 𝐍 by design cannot be typed. This is due to lack of knowledge at this point as to what resources it will take when constructing an object to substitute for x. So we conservatively forbid expressions of this form as it could lead to duplication/destruction of resources.

However, this expression can be refactored to work given a wrapper type

Inductive [_] : A :U U → U :=
| Just : A :U U → A → [A].

So with Γ = A :U L, B :U U, f' :U (x :L A) → [B], x :L A, the expression 
Γ ⊢ match (f' x) with Just y => (λ(_ :U B).0) y : 𝐍 is well typed. 
The sequencing effect imposed by constructor intro/elim ensures that y be free of linear resources when the branch is reached during reduction. This is very much the principle of call-by-push-value in effect system literature. Cbpv encoding is used extensively in our stateful reasoning example, where ΣL takes the role of [_].

Now when we have a linearly closed term in place of x, it is clear that this term can be freely generated without consuming any resources. From a linear logic perspective, there is an implicit ! in front of this term and it is linear "in name only".

## 2
Reviewers 1 and 2 have pointed out the lack of Prop in our system. We want to clarify that Prop is completely compatible in CLC, with meta-theory proven in Coq. Prop was removed during our Coq formalization of CILC simply as a means to reduce proof bureaucracy.

## 3
Reviewer 1 and 3 list Quantitative Type Theory and Krishnaswami et al.'s works as comparison points. We will comment on these works in this section.

QTT uses a semi-ring annotation system to control usage of function arguments within its body. It allows linear variables to occur in types by simply not counting their occurrences. This approach does not posses certain properties that we seek. Most notably, judgments of the form Γ, x :0 A ⊢ m : B cannot be strengthened to Γ ⊢ m : B as x may occur any number of times in B. The lack of this strengthening property proved to be fatal during our attempts to extend QTT with message passing concurrency.

Krishnaswami et al.'s work utilizes not just two adjoint operators G and F but also two different typing judgments of linear and non-linear. When writing programs in this system, the disparate judgments block linear and non-linear terms from direct interaction. All interactions between linear and non-linear terms require G or F to mediate, which lead to considerable complexity during back and forth interactions. Our system only uses a single typing judgment, making interaction between linear and non-linear much more straightforward.

The languages listed by reviewer 1 (Practical affine types and Linear Haskell) are not dependently typed, thus their linear inductive types permit no term dependency. The inductive types of CILC allow dependency on non-linear terms. This poses much more difficulty when defining the intro/elim rules as they must account for both dependency and linearity.

## 4
To reviewer 2's question on which major results have been verified in Coq, all explicitly stated theorems in the paper have been verified in Coq. All theorems restated for CILC apart from strong-normalization have been proven in Coq.

### 2022-03-12 writing/rebuttal.md
rebuttal

@@
 
-So with Γ = A :U L, B :U U, f' :U (x :L A) → [B], x :L A, the expression 
-Γ ⊢ match (f' x) with Just y => (λ(_ :U B).0) y : 𝐍 is well typed. 
+So with Γ = A :U L, B :U U, f' :U (x :L A) → [B], x :L A, the expression Γ ⊢ match (f' x) with Just y => (λ(_ :U B).0) y : 𝐍 is well typed.
+
 The sequencing effect imposed by constructor intro/elim ensures that y be free of linear resources when the branch is reached during reduction.
@@
 ## 3
-Reviewer 1 and 3 list Quantitative Type Theory and Krishnaswami et al.'s works as comparison points. We will comment on these works in this section.
+Reviewers 1 and 3 list Quantitative Type Theory and Krishnaswami et al.'s works as comparison points. We will comment on these works in this section.
 
@@
 The lack of this strengthening property proved to be fatal during our attempts to extend QTT with message passing concurrency.
+Strengthening is an easy property to prove for CLC and we are currently working towards concurrency.
 
-Krishnaswami et al.'s work utilizes not just two adjoint operators G and F but also two different typing judgments of linear and non-linear.
+Krishnaswami et al.'s work utilizes not just two adjoint operators G and F but also two different typing judgments of linear and non-linear types.
 When writing programs in this system, the disparate judgments block linear and non-linear terms from direct interaction.
 All interactions between linear and non-linear terms require G or F to mediate, which lead to considerable complexity during back and forth interactions.
-Our system only uses a single typing judgment, making interaction between linear and non-linear much more straightforward.
+Our system only uses a single typing judgment, making interaction between linear and non-linear term much more straightforward without needing any operators to mediate.
 
@@
 ## 4
-To reviewer 2's question on which major results have been verified in Coq, all explicitly stated theorems in the paper have been verified in Coq.
+To reviewer 2's question on which major results have been verified in Coq, all explicitly stated lemmas and theorems in the paper have been verified in Coq.
 All theorems restated for CILC apart from strong-normalization have been proven in Coq.

### 2022-03-14 writing/rebuttal.txt
rebuttal

--- 2022-03-12 writing/rebuttal.md
@@
 ## 1
-Reviewer 1 has acutely observed that, given Γ = A :U L, B :U U, f :U (x :L A) → B, x :L A, the expression Γ ⊢ (λ(_ :U B).0) (f x) : 𝐍 by design cannot be typed.
+Reviewer 1 has acutely observed that, given \Gamma = A :U L, B :U U, f :U (x :L A) -> B, x :L A, the expression \Gamma |- (\lambda(_ :U B).0) (f x) : N by design cannot be typed.
 This is due to lack of knowledge at this point as to what resources it will take when constructing an object to substitute for x.
@@
 
-Inductive [_] : A :U U → U :=
-| Just : A :U U → A → [A].
+Inductive [_] : A :U U -> U :=
+| Just : A :U U -> A -> [A].
 
-So with Γ = A :U L, B :U U, f' :U (x :L A) → [B], x :L A, the expression 
-Γ ⊢ match (f' x) with Just y => (λ(_ :U B).0) y : 𝐍 is well typed. 
+So with \Gamma = A :U L, B :U U, f' :U (x :L A) -> [B], x :L A, the expression \Gamma |- match (f' x) with Just y => (\lambda(_ :U B).0) y : N is well typed.
+
 The sequencing effect imposed by constructor intro/elim ensures that y be free of linear resources when the branch is reached during reduction.
@@
 From a linear logic perspective, there is an implicit !
-in front of this term and it is linear "in name only".
+in front of this term so it is linear "in name only".
 
@@
 This approach does not posses certain properties that we seek.
-Most notably, judgments of the form Γ, x :0 A ⊢ m : B cannot be strengthened to Γ ⊢ m : B as x may occur any number of times in B.
+Most notably, judgments of the form \Gamma, x :0 A |- m : B cannot be strengthened to \Gamma |- m : B as x may occur any number of times in B.
 The lack of this strengthening property proved to be fatal during our attempts to extend QTT with message passing concurrency.
+Strengthening is an easy property to prove for CLC and we are currently working towards concurrency.
 
-Krishnaswami et al.'s work utilizes not just two adjoint operators G and F but also two different typing judgments of linear and non-linear.
+Krishnaswami et al.'s work utilizes not just two adjoint operators G and F but also two different typing judgments of linear and non-linear types.
 When writing programs in this system, the disparate judgments block linear and non-linear terms from direct interaction.
 All interactions between linear and non-linear terms require G or F to mediate, which lead to considerable complexity during back and forth interactions.
-Our system only uses a single typing judgment, making interaction between linear and non-linear much more straightforward.
+Our system only uses a single typing judgment, making interaction between linear and non-linear term much more straightforward without needing any operators to mediate.
 
